### 2015-10-01 HW1.tex
[hw1] tex file and images.

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry} 
\usepackage{enumerate, xfrac, color, graphicx}
\usepackage{amsmath,amsthm,amssymb,amsfonts,mathabx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{pifont}
\usepackage{listings, courier}
\graphicspath{{/Users/mfzhao/Dropbox/}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\lstset{breaklines=true, basicstyle=\small\ttfamily, language=R, backgroundcolor=\color{highlight}, stepnumber=5}

\definecolor{highlight}{RGB}{248,248,248}

\begin{document}
	\title{6.867 Problem Set 1}
	\date{October 1, 2015}
	\maketitle

\subsubsection*{Problem 1}

We implemented gradient descent in Python using an algorithm similar to the pseudocode below. The function is written in such a way that it can accept arbitrary scalar functions $f$ and gradients, $\nabla f(x)$, where $f(x)$ here is a function of an arbitrary number of parameters, $n$, and return $\min{f(x)}$.

\begin{algorithm}
\caption{Gradient Descent}
\label{GradDescent}
\begin{algorithmic}[1]
\Procedure{Gradient Descent}{}
\State Specify initial guess, step size, and convergence criteria
\State Set current guess equal to initial guess
\While{Distance between last guess and current guess location is greater than convergence criteria}
\State Calculate function at current location
\State Calculate the gradient of function at current location
\State Updated guess = take a step from current guess in direction of gradient proportional to step size
\State Calculate distance between updated guess and current guess
\State Set current guess equal to updated guess
\EndWhile{}
\EndProcedure
\end{algorithmic}
\end{algorithm}

In order to understand the effect of various hyperparameters (our initial guess, the step size, and the convergence criteria) of the algorithm, we compared the results of our gradient descent routine given a number of different parameter combinations for two different functions: $f(x) = \sum_i^n x_i^2$ and $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$. The first of these is a convex function, whereas the second is quite non-convex, with many local minima. The results of our gradient descent algorithm for numerous values are found below.

We find that, in general, the convex function is much better behaved. Regardless of our initial guess or our learning rate, the algorithm converges to a guess in fewer than 1,000 steps, and these guess are consistently close to $x = (0,0,0,0)$, the ``true" value. The final guess' distance from the true value of $x$ is primarily a function of the convergence criteria, which makes sense - the less picky we are with our convergence criteria, the further from the true value of $x$ our final guess will be.

The non-convex, on the other hand, is much more sensitive to the hyperparameters we specify. The most notable departure from the convex function is the influence of our initial guess. It is much easier for the gradient descent algorithm to get trapped in local minima, and this is clearly illustrated in our results. While this is the most obvious difference, note that changes in both the learning rate and the convergence criteria also lead to non-trivially different guesses for $\min{f(x)}$.

Unfortunately, it might not always be convenient to specify the gradient of our scalar function, $f(x)$, explicitly. This might be because it is complicated to compute, or because no closed form solution exists. In order to get around this issue, we can approximate the gradient of the function using central differences. We approximate the $i$th element of the gradient, $\nabla_i$, as

\begin{equation}
\nabla_i \approx \frac{f(x_i+h) - f(x_i-h)}{2h}
\end{equation}

\noindent where $h$ is some small constant (we use $h = .001$).

We see that the performance in the calculation of the gradient is identical between the analytical approach and the central difference approach. That's great! Now let's see how our gradient descent algorithm compares to two other algorithms - Broyden?Fletcher?Goldfarb?Shanno (BFGS) and Conjugate Gradient (CG). We compare both the number of iterations and function evaluations. We'll consider our algorithm in two forms - one where we pass in an explicit gradient function, and another where we estimate the gradient using the central difference method. All benchmarking is done with the same starting point: [5, -3, 7, 8]. $f_1(x) =  \sum_i^n x_i^2$ and $f_2(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$. We see that when passed an analytical solution for $\nabla f(x)$, our algorithm requires about as many function calls, although many more iterations than both BFGS and CG. Given we do not have an analytical expression for $\nabla f(x)$, however, our algorithm requires many more iterations and function calls.

\begin{table}
\captionof{table}{Performance of gradient descent given various algorithm hyperparameters} 
\begin{tabular}{llllll}
\toprule
{} &                        $Guess_{i}$ &      $R_{learn}$ &   Thresh. &    $Guess_{f}$ ($f(x) = \sum_i^n x_i^2$) & $Guess_{f}$ ($f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$) \\
\midrule
&    [5.0, -3.0, 7.0, 8.0] &    0.3 &    0.1 &   [0.020, -0.012, 0.029, 0.033] &    [6.150, -0.314, 6.165, 6.176] \\
&    [5.0, -3.0, 7.0, 8.0] &    0.3 &  0.001 &   [0.001, -0.001, 0.002, 0.002] &    [6.159, -0.025, 6.160, 6.161] \\
&    [5.0, -3.0, 7.0, 8.0] &    0.3 &  1e-05 &   [0.000, -0.000, 0.000, 0.000] &    [6.160, -0.003, 6.160, 6.160] \\
&    [5.0, -3.0, 7.0, 8.0] &  0.003 &    0.1 &   [1.172, -0.703, 1.641, 1.876] &    [5.005, -2.999, 6.995, 7.993] \\
&    [5.0, -3.0, 7.0, 8.0] &  0.003 &  0.001 &   [0.118, -0.071, 0.165, 0.188] &    [6.092, -0.932, 6.200, 6.267] \\
&    [5.0, -3.0, 7.0, 8.0] &  0.003 &  1e-05 &   [0.012, -0.007, 0.017, 0.019] &    [6.092, -0.932, 6.200, 6.267] \\
 &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &    0.1 &  [-0.041, 0.016, 0.004, -0.020] &  [-12.059, 6.070, 0.014, -6.139] \\
&  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  0.001 &  [-0.003, 0.001, 0.000, -0.001] &  [-12.294, 6.152, 0.001, -6.158] \\
 &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  1e-05 &  [-0.000, 0.000, 0.000, -0.000] &  [-12.315, 6.159, 0.000, -6.159] \\
 &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &    0.1 &  [-2.388, 0.955, 0.239, -1.194] &  [-10.002, 4.004, 0.995, -5.005] \\
 &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  0.001 &  [-0.240, 0.096, 0.024, -0.120] &  [-11.748, 5.885, 0.063, -6.076] \\
 &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  1e-05 &  [-0.024, 0.010, 0.002, -0.012] &  [-11.848, 5.938, 0.051, -6.092] \\
 &   [1.0, -1.0, -3.0, 2.0] &    0.3 &    0.1 &  [0.010, -0.010, -0.031, 0.020] &   [0.007, -0.007, -0.314, 0.024] \\
 &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  0.001 &  [0.001, -0.001, -0.002, 0.001] &   [0.001, -0.001, -0.025, 0.002] \\
 &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  1e-05 &  [0.000, -0.000, -0.000, 0.000] &   [0.000, -0.000, -0.003, 0.000] \\
 &   [1.0, -1.0, -3.0, 2.0] &  0.003 &    0.1 &  [0.736, -0.736, -2.207, 1.471] &   [0.995, -0.995, -2.999, 1.994] \\
 &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  0.001 &  [0.073, -0.073, -0.220, 0.147] &   [0.051, -0.051, -0.932, 0.143] \\
 &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  1e-05 &  [0.007, -0.007, -0.022, 0.015] &   [0.051, -0.051, -0.932, 0.143] \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\captionof{table}{Performance in gradient calculation for analytical and central difference approaches} 
\begin{tabular}{llllll}
\toprule
{} &   $x_{i}$  &  function &   $\nabla_{analytical}$  & $\nabla_{approximate}$ \\
\midrule
&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [-0.8589 -0.2011  0.797   1.1494] &[-0.8589 -0.2011  0.797   1.1494] \\
&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^2$ &   [-20.   8.   2. -10.] &  [-20.   8.   2. -10.]\\
&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ &[ 0.344  -0.6768  0.8615  0.8589]&[ 0.344  -0.6768  0.8615  0.8589] \\
&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [ 0.8615 -0.8615 -0.2011  0.9493] & [ 0.8615 -0.8615 -0.2011  0.9493] \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\captionof{table}{Performance of authors' gradient descent algorithm compared to BFGS and CG}
\begin{tabular}{llllll}
\toprule
{} & algorithm & $f_1(x)$ function calls & $f_1(x)$ iterations &  $f_2(x)$ function calls &  $f_2(x)$ iterations \\
\midrule
  & BFGS & 24 & 2 & 78 & 10 \\
  & CG & 30 & 2 & 78 & 6 \\
  & Gradient Descent (analytical) & 20 & 10 & 48 & 24 \\ 
  & Gradient Descent (approximate) & 100 & 10 & 240 & 24 \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection*{Problem 2}

\subsubsection*{Problem 3}

To begin problem 3, we implemented the closed form solution of Ridge Regression in Python, making use of Scipy and Numpy. Note that if you want to test cases where the number of features is greater than the number of data points (something that a well-trained teacher of machine should be skeptical to do), it is useful to calculate the inverse of a matrix using the pseudo-inverse, rather than the inverse. We learned this the long way after debugging code for hours.

The closed form solution for ridge regression is given by the expression:

\begin{equation}
\hat{\theta} = (\mathbf{X}^T\mathbf{X} + \lambda \mathbf{I})^{-1}\mathbf{X}^t\mathbf{y}
\end{equation}

\noindent where $\mathbf{X}$ is your matrix of features, $y$ is your array of observed values, and $\hat{\theta}$ are your estimated feature weights. $\lambda$ is a regularization parameter, which represents how much you penalize the model for having large feature weights.

Given the data in Bishop 1.4 (which can be seen below), we attempt to fit polynomials of various orders, $M$.

\begin{center}
\includegraphics[scale=.3]{Bishop_14_Data.png}
\end{center}

The plot below shows the mean squared error (MSE) that our ridge regression fit gives for values of $M$ between 1 and 5, for values of $\lambda$ ranging from $10^0 -1$ to $10^2$. Note that in the case where $\lambda = 0$, we have $OLS$ regression, and as $M$ increases, the model is free to overfit our data. While the MSE here is low, if we evaluated this on a new data point, our model would likely perform very poorly. As $\lambda$ gets higher, we penalize the model for adding additional large weights to $\hat{\theta}$, but our $MSE$ stays relatively large. However, these ``large" MSE models are likely to perform better on a new data point!

\begin{center}
\includegraphics[scale=.4]{MSE_Lambda_Bishop.png}
\end{center}

In order to avoid overfitting, but in order to make sure that we have chosen values of $\lambda$ and $M$ (and in general, hyperparameters) that make sense, it is common to split the data into a training set and a test set. We will use the closed-form solution for ridge regression to calculate $\hat{\theta}$ on our training set. We will then use that $\hat{\theta}$ to attempt to predict values of $\mathbf{y}$ for the test data. By choosing the $\hat{\theta}$ that minimizes MSE on the test data, but is trained on the training data, we are ensuring that we have not overfit our model to the data it is trained on. As one final step in the model selection process, we can check the performance of our final $\hat{\theta}$ on a validation set, to make sure that performance is still good (and we have not overfit for performance on our test set).

\subsubsection*{Problem 4}

\end{document}

### 2015-10-01 HW1.tex
[problem3] adding new code, new latex, new image.

@@
 
+We use this methodology to find the best $\lambda$ and $M$ given three datasets similar to the bishop data.
+We calculate $\lambda$ and $M$ using the regress\_train.txt, data we validate using regress\_test.txt, and then we test using regress\_validate.txt.
+We search through a grid of various values for $\lambda$ and $M$, and find that $MSE$ is lowest on the validation set when $\lambda = 1$ and $M = 1$: $MSE = 1.776$ on the validation data set.
+These parameters gave us an $MSE$ of 1.460 on the training data and and an $MSE$ of 1.415 on the test data.
+The fact that our $MSE$ is on the same order of magnitude for the test data is a good sign - it indicates that in making sure we didn't overfit to the training data, we didn't accidentally overfit to the validation data.
+
+The $MSE$ for various combinations of $\lambda$ and $M$ on the three datasets (including our chosen set of hyperparameters) is found below.
+We see that with large $M$ and small $\lambda$, we do great on the training data, but bad on our validation data.
+For large values of both $\lambda$ and $M$, we do well, but not as well as our optimal values we found through the grid search.
+
+Let's now use this methodology on a real dataset, as opposed to a toy dataset.
+The dataset that we will use was compiled by Kirsztian Buza at Budapest University of Technology and Economics.
+It includes 280 numerical values describing a number of blog posts, such as day of week, number of parent pages, and number of comments received within 24 hours.
+We are hoping to predict the number of comments it will receive in the next 24 hours.
+We are given this data split into training, validation, and test datasets, so we will once again perform a grid search to find the optimal value of $\lambda$.
+In this particular case, we are not regressing a polynomial of arbitrary order, so we do not need to test different values of $M$.
+
+\begin{center}
+\includegraphics[scale=.4]{MSE_comparison.png}
+\end{center}
+
+We find that for very large $\lambda$ (as large as $10^6$, $MSE$ continues to decrease, although it is always close to 900.
+The MSE for a uniform predictor that predicts the average of $y$, $\hat{y}$, has an MSE of 1453, so the performance of our model is still better.
+However, we see that for values of $\lambda$ higher than $10^6$ or so, the value of the MSE increases, and eventually approaches (and eclipses) the MSE of our uniform estimator.
+
+This isn't surprising - the blog feedback dataset is large, with 10,480 observations in the test set.
+If we interpret the value of $\lambda$ as strength of our prior, it makes sense that it does not have a substantial effect until it is significantly larger than the size of our data.
+Furthermore, the inability of our model to get very low MSE for low $\lambda$ speaks both to the fact that the data has a much higher dimension than our training data, and also suggests that the simple linear model we are using here may not be a very good fit for the data.
+
+\begin{center}
+\includegraphics[scale=.4]{BlogFeedbackRegressionMSE.png}
+\end{center}
+
 \subsubsection*{Problem 4}

### 2015-10-01 problem_3.py

@@ -1,0 +1,248 @@
+import numpy as np
+import scipy.linalg as linalg 
+import pylab
+import homework1
+import sys
+import pandas as pd
+import matplotlib.pyplot as plt
+from ggplot import *
+
+def createFeatures(X=np.zeros(1), M=1):
+	features = X**0
+	for i in range(1, M+1):
+		features = np.hstack([features, X**i])
+	return features
+
+def ridgeRegression(X, Y, l=0, M=1, features=False):
+	# Closed form solution from http://www.hongliangjie.com/notes/lr.pdf
+	if not features:
+		features = createFeatures(X, M)
+	else:
+		features = X
+		M = features.shape[1]-1
+
+	firstTerm = linalg.pinv(np.dot(np.transpose(features), features) + l*np.identity(M+1))
+	secondTerm = np.transpose(features)
+	thirdTerm = Y
+
+	theta = np.dot(firstTerm, np.dot(secondTerm, thirdTerm))
+
+	return theta
+
+def predictRidge(X, theta, features=False):
+	if not features:
+		features = createFeatures(X, M=theta.size-1)
+	else:
+		features = X
+	
+	predictions = np.dot(features, theta)
+
+	return predictions
+
+def MSE(predictions, actuals):
+	mse = np.dot(np.transpose(predictions-actuals), (predictions-actuals))/float(predictions.size)
+
+	return mse 
+
+def gridSearch(lambdas, Ms, X_train, X_test, X_valid, Y_train, Y_test, Y_valid):
+	results = np.empty([4,1])
+	bestLambda = lambdas[0]
+	bestM = Ms[0]
+	bestTheta = ridgeRegression(X_train, Y_train, l=lambdas[0], M=Ms[0])
+	bestPredictions = predictRidge(X_test, bestTheta)
+	minMSE = MSE(bestPredictions, Y_test)
+	print Y_test 
+	print Y_train
+
+	for i in lambdas:
+		for j in Ms:
+			theta = ridgeRegression(X_train, Y_train, l=i, M=j)
+			predictions = predictRidge(X_test, theta)
+			mse = MSE(predictions, Y_test)
[…]

### 2015-10-01 problem4.py
[problem4] initial commit.

@@ -1,0 +1,78 @@
+import numpy as np
+import scipy.optimize as spo
+import pandas as pd 
+import scipy
+import scipy.io
+
+np.set_printoptions(precision=4)
+
+# Generic gradient descent function
+def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=1000, h=0.001):
+	count = 0
+	nIter = 0
+	fcall = 0
+	while count < 2 and nIter <= maxIter:
+		f_i = f(init)
+		# Calculates the gradient
+		grad = df(init, f, h)
+		# Update step
+		init = init - lr * grad.reshape(-1, 1)
+		f_u = f(init)
+		# Calculate the difference in f using initial and updated values
+		diff = abs((f_i - f_u))
+		nIter += 1
+		fcall += 2
+		if df == centdiff:
+			fcall += 2*len(init)
+		# Tracks successive number of times that difference is below the convergence criterion
+		if diff < crit:
+			count += 1
+		else:
+			count = 0
+
+	print "nIter: %d" % (nIter)
+	print "Fcall: %d" % (fcall)
+	return init
+
+# Central difference approximation of a gradient, returns a vector of length of the input vector x
+def centdiff(x, f, h=0.00001):
+	n = len(x)
+	out = np.zeros(n)
+	for i in range(0, n):
+		hplus = np.copy(x)
+		hminus = np.copy(x)
+		hplus[i] += h
+		hminus[i] -= h
+		# Calculates a better denominator to address potential problems with floating point arithmetic especially for small values of h
+		hfix = hplus[i] - hminus[i]
+		out[i] = (f(hplus) - f(hminus))/(hfix)
+	return out
+
+def ridgeRegressionLoss(theta):
+	xTtheta = np.dot(np.transpose(X_train), theta)
+	diff = Y_train - xTtheta
+	loss = np.sum(np.dot(np.transpose(diff), diff)) + Lambda * np.sum(np.dot(np.transpose(theta), theta))
+
+	return loss
+
+
+def LASSOLoss(theta):
+	XTtheta = np.dot(np.transpose(X_train), theta)
+	diff = Y_train - XTtheta
+	loss = np.dot(np.transpose((diff)), diff)*(1/Y_train.size) + Lambda * np.sum(np.absolute(theta))
+
+	return loss
+
+
[…]

### 2015-10-01 problem4.py
[problem4] get this working. seems to be good, but would be good to double check the loss functions. @mfzhao

@@ -3,11 +3,12 @@ import scipy.optimize as spo
 import pandas as pd 
 import scipy
 import scipy.io
+import pylab
 
 np.set_printoptions(precision=4)
 
 # Generic gradient descent function
-def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=1000, h=0.001):
+def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=100000, h=0.001):
 	count = 0
 	nIter = 0
 	fcall = 0
@@ -29,7 +30,9 @@ def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=1000, h=0.001):
 			count += 1
 		else:
 			count = 0
-
+		print 'Loss function: ', f_u
+		print 'Test MSE: ', MSE(predict(init, X_test), Y_test)
+		print 'Train MSE: ', MSE(predict(init, X_train), Y_train_val)
 	print "nIter: %d" % (nIter)
 	print "Fcall: %d" % (fcall)
 	return init
@@ -59,20 +62,45 @@ def ridgeRegressionLoss(theta):
 def LASSOLoss(theta):
 	XTtheta = np.dot(np.transpose(X_train), theta)
 	diff = Y_train - XTtheta
-	loss = np.dot(np.transpose((diff)), diff)*(1/Y_train.size) + Lambda * np.sum(np.absolute(theta))
+	loss = np.sum(np.dot(np.transpose((diff)), diff))*(1/Y_train.size) + Lambda * np.sum(np.absolute(theta))
 
 	return loss
+
+def predict(theta, X):
+	prediction = np.dot(np.transpose(X), theta)
+
+	return prediction 
+
+def MSE(predictions, actuals):
+	mse = np.dot((np.transpose(predictions)-actuals), np.transpose(np.transpose(predictions)-actuals))/float(predictions.size)
+
+	return mse 
 
 
 variables = scipy.io.loadmat('/Users/dholtz/Downloads/6867_hw1_data/regress-highdim.mat')
 X_train = variables['X_train']
 X_test = variables['X_test']
 Y_train = variables['Y_train']
+Y_train_val = Y_train
 Y_train = Y_train.reshape(-1, 1)	
 Y_test = variables['Y_test']
 Lambda = 5
 
-thetaInit = np.ones(X_train.shape[0]).reshape(-1, 1)
[…]

### 2015-10-01 problem4.py
[problem4] add comments.

@@ -88,8 +88,10 @@ Lambda = 5
 
 thetaInit = np.random.rand(12).reshape(-1, 1)
 
+## These don't always converge - might be because the actual function is sinusoidal?
 thetaLASSO = gradientDescent(f=LASSOLoss, df=centdiff, init=thetaInit, lr=0.00003, h=.00001, crit=.006, maxIter=1000000)
 thetaRidge = gradientDescent(f=ridgeRegressionLoss, df=centdiff, init=thetaInit, lr=0.00003, h=.00001, crit=.006, maxIter=1000000)
+
 LASSOPredictions = predict(thetaLASSO, X_test)
 ridgePredictions = predict(thetaRidge, X_test)
 LASSOPredictionsTrain = predict(thetaLASSO, X_train)
@@ -104,3 +106,4 @@ pylab.plot(X_test[:1,].flatten(), Y_test.flatten(), 'ro',
 	X_test[:1,].flatten(), LASSOPredictions.flatten(), 'bo',
 	X_test[:1,].flatten(), ridgePredictions.flatten(), 'go')
 pylab.show()
+

### 2015-10-01 problem_3.py
[problem3] this commit shouldnt go here but whatever. testing that our method for calculating theta in ridge regression is OK.

@@ -27,6 +27,17 @@ def ridgeRegression(X, Y, l=0, M=1, features=False):
 
 	theta = np.dot(firstTerm, np.dot(secondTerm, thirdTerm))
 
+	#Xprime = np.vstack([features, np.sqrt(l)*np.identity(features.shape[1])])
+	#yprime = np.vstack([Y, np.zeros(features.shape[1]).reshape(-1, 1)])
+    #
+	#thetaprime = np.dot(linalg.pinv(Xprime), yprime)
+	#print linalg.pinv(X).shape
+	#print Y.shape
+	#thetadoubleprime = np.dot(linalg.pinv(features), Y)
+	#print thetaprime
+	#print theta
+	#print thetadoubleprime
+
 	return theta
 
 def predictRidge(X, theta, features=False):

### 2015-10-01 problem4.py
[problem4] add intercept term, plots.

@@ -4,11 +4,13 @@ import pandas as pd
 import scipy
 import scipy.io
 import pylab
+import statsmodels.api as sm
+from ggplot import *
 
 np.set_printoptions(precision=4)
 
 # Generic gradient descent function
-def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=100000, h=0.001):
+def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=100000, h=0.001, verbose=False):
 	count = 0
 	nIter = 0
 	fcall = 0
@@ -33,6 +35,8 @@ def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=100000, h=0.001):
 		print 'Loss function: ', f_u
 		print 'Test MSE: ', MSE(predict(init, X_test), Y_test)
 		print 'Train MSE: ', MSE(predict(init, X_train), Y_train_val)
+		if verbose == True:
+			print init 
 	print "nIter: %d" % (nIter)
 	print "Fcall: %d" % (fcall)
 	return init
@@ -54,7 +58,7 @@ def centdiff(x, f, h=0.00001):
 def ridgeRegressionLoss(theta):
 	xTtheta = np.dot(np.transpose(X_train), theta)
 	diff = Y_train - xTtheta
-	loss = np.sum(np.dot(np.transpose(diff), diff)) + Lambda * np.sum(np.dot(np.transpose(theta), theta))
+	loss = np.sum(np.dot(np.transpose(diff), diff))*(1/Y_train.size) + Lambda * np.sum(np.dot(np.transpose(theta), theta))
 
 	return loss
 
@@ -76,17 +80,27 @@ def MSE(predictions, actuals):
 
 	return mse 
 
+def addInterceptTerm(X_array):
+	data = np.vstack([X_array, np.ones(X_array.shape[1]).reshape(1, -1)])
+
+	return data
 
 variables = scipy.io.loadmat('/Users/dholtz/Downloads/6867_hw1_data/regress-highdim.mat')
 X_train = variables['X_train']
+X_train = addInterceptTerm(X_train)
 X_test = variables['X_test']
+X_test = addInterceptTerm(X_test)
 Y_train = variables['Y_train']
 Y_train_val = Y_train
[…]

### 2015-10-01 problem4.py
[problem4] check against BFGS - problem persists. must be in the loss function.

@@ -95,10 +95,8 @@ Y_train_val = Y_train
 Y_train = Y_train.reshape(-1, 1)	
 Y_test = variables['Y_test']
 W_true = variables['W_true']
-print W_true.shape
-print np.array([0]).reshape(1,-1).shape
 W_true = np.hstack([np.array([0]).reshape(1,-1), W_true])
-Lambda = 7
+Lambda = .1
 
 thetaInit = np.random.rand(13).reshape(-1, 1)
 
@@ -116,14 +114,13 @@ X_lin = np.linspace(-1, 1, 1000)
 X_lin = np.array([X_lin, np.sin(0.4*np.pi*X_lin*1), np.sin(0.4*np.pi*X_lin*2), np.sin(0.4*np.pi*X_lin*3), np.sin(0.4*np.pi*X_lin*4), np.sin(0.4*np.pi*X_lin*5), np.sin(0.4*np.pi*X_lin*6), np.sin(0.4*np.pi*X_lin*7), np.sin(0.4*np.pi*X_lin*8), np.sin(0.4*np.pi*X_lin*9), np.sin(0.4*np.pi*X_lin*10), np.sin(0.4*np.pi*X_lin*11)])
 X_lin = addInterceptTerm(X_lin)
 
-print W_true.shape 
-print X_lin.shape
 ActualPredictions = predict(np.transpose(W_true), X_lin)
 
 YLasso_lin = predict(thetaLASSO, X_lin)
 YRidge_lin = predict(thetaRidge, X_lin)
 
-#print sm.OLS(Y_test, X_test).fit().summary()
+print spo.fmin_bfgs(LASSOLoss, thetaInit)
+print thetaLASSO
 
 Lambda = 0
 
@@ -158,8 +155,6 @@ print thetaOLS
 print thetaRidge 
 print thetaLASSO
 print W_true
-
-print np.concatenate([thetaOLS, thetaRidge, thetaLASSO, np.transpose(W_true)])
 
 d = {'weights' : pd.Series(np.concatenate([thetaOLS, thetaRidge, thetaLASSO, np.transpose(W_true)]).flatten()), \
 'Models': pd.Series(np.concatenate([np.repeat(np.array(['OLS']), 13),np.repeat(np.array(['Ridge Regression']), 13),\

### 2015-10-02 HW1.tex
WE ARE COMPLETED!i

@@
 
-We implemented gradient descent in Python using an algorithm similar to the pseudocode below.
-The function is written in such a way that it can accept arbitrary scalar functions $f$ and gradients, $\nabla f(x)$, where $f(x)$ here is a function of an arbitrary number of parameters, $n$, and return $\min{f(x)}$.
+We implemented a rather naive gradient descent in Python using an algorithm similar to the pseudocode below.
+The function is written in such a way that it can accept arbitrary scalar functions $f$ and their gradients, $\nabla f(x)$, where $x$ can be a vector of arbitrary length, $n$.
+The goal of this function is to try and return $\arg\min_x {f(x)}$.
 
@@
 
-In order to understand the effect of various hyperparameters (our initial guess, the step size, and the convergence criteria) of the algorithm, we compared the results of our gradient descent routine given a number of different parameter combinations for two different functions: $f(x) = \sum_i^n x_i^2$ and $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$.
+In order to understand the effect of various hyperparameters (our initial guess, the step size, and the convergence criteria) of the algorithm, we compared the results of our gradient descent routine given a number of different parameter combinations for two different functions: $f(x) = \sum_i^n x_i^2$ and $f(x) = \sum_i^n \frac{1}{100}x_i^2 - \cos(x)$.
 The first of these is a convex function, whereas the second is quite non-convex, with many local minima.
-The results of our gradient descent algorithm for numerous values are found below.
+The results of our gradient descent algorithm for numerous values are found below in Table 1.
 
-We find that, in general, the convex function is much better behaved.
-Regardless of our initial guess or our learning rate, the algorithm converges to a guess in fewer than 1,000 steps, and these guess are consistently close to $x = (0,0,0,0)$, the ``true" value.
+We find that, in general, the convex function behaves considerably better.
+Regardless of our initial guess or our learning rate, the algorithm returns a guess in fewer than 1,000 steps, and these guesses are consistently close to $x = (0,0,0,0)$, the "true" value.
 The final guess' distance from the true value of $x$ is primarily a function of the convergence criteria, which makes sense - the less picky we are with our convergence criteria, the further from the true value of $x$ our final guess will be.
 
-The non-convex, on the other hand, is much more sensitive to the hyperparameters we specify.
+The non-convex function, on the other hand, is much more sensitive to the hyperparameters we specify.
 The most notable departure from the convex function is the influence of our initial guess.
-It is much easier for the gradient descent algorithm to get trapped in local minima, and this is clearly illustrated in our results.
-While this is the most obvious difference, note that changes in both the learning rate and the convergence criteria also lead to non-trivially different guesses for $\min{f(x)}$.
+It is much easier for our gradient descent algorithm to get trapped in local minima, and this is clearly illustrated in our results.
+While this is the most obvious difference, note that changes in both the learning rate and the convergence criteria also lead to non-trivially different guesses for $\arg\min_x {f(x)}$.
+
+Generally speaking, choice of starting value is only in when were are dealing with non-convex functions as different starting values may lead to different local extremums.
+Though if we randomly initialized good starting values, it decreases the number of iterations necessary for convergence.
+Tuning the step size alters the performance of the algorithm, while tuning the convergence threshold alters the accuracy of the algorithm.
+
+To elaborate on this a little, for any given convergence threshold, increasing step size will decrease the number of iterations required for convergence.
+However, we need make sure that the step isn't so large that the algorithm gets caught in a loop where its constantly overshooting the minimum.
+On the other hand, for a given step size, decreasing the convergence threshold increases accuracy of the algorithm, but at the cost of performance as it'll take additional iterations to reach the higher level of precision.
+Here, we want to make sure that the convergence threshold isn't so small that we can never converge.
 
@@
 That's great!
-Now let's see how our gradient descent algorithm compares to two other algorithms - Broyden?Fletcher?Goldfarb?Shanno (BFGS) and Conjugate Gradient (CG).
+Now let's see how our gradient descent algorithm compares to two more sophisticated optimizers - BFGS and Conjugate Gradient (CG).
 We compare both the number of iterations and function evaluations.
@@
 $f_1(x) =  \sum_i^n x_i^2$ and $f_2(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$.
-We see that when passed an analytical solution for $\nabla f(x)$, our algorithm requires about as many function calls, although many more iterations than both BFGS and CG.
-Given we do not have an analytical expression for $\nabla f(x)$, however, our algorithm requires many more iterations and function calls.
+We see that when passed an analytical solution for $\nabla f(x)$, our algorithm requires about as many function calls, although many more iterations than either BFGS or CG.
+If we use the central difference approximation for $\nabla f(x)$ instead, our algorithm requires many more iterations and function calls.
 
@@
 \toprule
-{} &                        $Guess_{i}$ &      $R_{learn}$ &   Thresh. &    $Guess_{f}$ ($f(x) = \sum_i^n x_i^2$) & $Guess_{f}$ ($f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$) \\
+{} &                        $Guess_{i}$ &      step &   Thresh. &    $Guess_{f}$ ($f(x) = \sum_i^n x_i^2$) & $Guess_{f}$ ($f(x) = \sum_i^n \frac{1}{100} x_i^2 - \cos(x)$) \\
 \midrule
@@
 &    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
-&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [-0.8589 -0.2011  0.797   1.1494] &[-0.8589 -0.2011  0.797   1.1494] \\
+&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n \frac{1}{100} x_i^2 - \cos(x)$ & [-0.8589 -0.2011  0.797   1.1494] &[-0.8589 -0.2011  0.797   1.1494] \\
 &    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^2$ &   [-20.   8.   2. -10.] &  [-20.   8.   2. -10.]\\
-&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ &[ 0.344  -0.6768  0.8615  0.8589]&[ 0.344  -0.6768  0.8615  0.8589] \\
+&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n \frac{1}{100} x_i^2 - \cos(x)$ &[ 0.344  -0.6768  0.8615  0.8589]&[ 0.344  -0.6768  0.8615  0.8589] \\
 &    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
-&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [ 0.8615 -0.8615 -0.2011  0.9493] & [ 0.8615 -0.8615 -0.2011  0.9493] \\
+&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n \frac{1}{100} x_i^2 - \cos(x)$ & [ 0.8615 -0.8615 -0.2011  0.9493] & [ 0.8615 -0.8615 -0.2011  0.9493] \\
 \bottomrule
@@
 \begin{tabular}{llllll}
-\toprule
-{} & algorithm & $f_1(x)$ function calls & $f_1(x)$ iterations &  $f_2(x)$ function calls &  $f_2(x)$ iterations \\
+\toprule{} & algorithm & $f_1(x)$ function calls & $f_1(x)$ iterations &  $f_2(x)$ function calls &  $f_2(x)$ iterations \\
 \midrule
@@
 
+With problem 2, we first derive the maximum-likelihood estimator for the weight vector $\theta$.
+Note that if we assume Gaussian errors, that is $Y \sim N(X\theta, \sigma^2)$, then our maximum-likelihood estimator is equivalent to the OLS estimator.
+Hence:
+\begin{equation*}
+	\hat{\theta}_{ML}=\hat{\theta}_{OLS}=(\Phi^T \Phi)^{-1}\Phi^T Y
+\end{equation*}
+where $\Phi$ denotes the polynomial basis expansion of $X$.
+
+Using this closed form solution, we determined the maximum-likelihood weight vector for the polynomial basis expansions of $X$ of orders 0, 1, 3, and 9.
+Using these estimated weight vectors we obtained very similar plots to Bishop as seen below:
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq0.png}
+		\caption*{$M=0$}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq1.png}
+		\caption*{$M=1$}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq3.png}
+		\caption*{$M=3$}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq9.png}
+		\caption*{$M=9$}
+	\end{minipage}
+	\caption{Plots of polynomials having various orders $M$, in red, fitted to the the observed points. The green curve shows the DGF $\sin(2\pi x)$.}
+\end{figure}
+
+\begin{table}[ht]
+\centering
+\captionof{table}{Optimal Weight Vectors for Various Polynomial Basis Expansions of $\mathbf{X}$}
+\begin{tabular}{lrrrr}
+\toprule
+{} &     $M=0$ &     $M=1$ &      $M=3$ &           $M=9$ \\
+\midrule
+$\theta_0$ &  0.186299 &  0.820212 &   0.313703 &        0.349482 \\
+$\theta_1$ &        & -1.267826 &   7.985371 &      232.682435 \\
+$\theta_2$ &        &        & -25.426102 &    -5329.068232 \\
+$\theta_3$ &        &        &  17.374077 &    48633.377984 \\
+$\theta_4$ &        &        &         &  -231944.259392 \\
+$\theta_5$ &        &        &         &   640871.190655 \\
+$\theta_6$ &        &        &         & -1063154.903003 \\
+$\theta_7$ &        &        &         &  1043711.192742 \\
+$\theta_8$ &        &        &         &  -558375.328325 \\
+$\theta_9$ &        &        &         &   125355.027159 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The sum of squared errors of a linear polynomial basis model is defined by: 
+\begin{equation*}
+	SSE=(\Phi \theta-Y)^T(\Phi \theta-Y) 
+\end{equation*}
+which has gradient:
+\begin{equation*}
+	\nabla SSE = 2\Phi^T(\Phi \theta-Y)
+\end{equation*}
+We tested the accuracy to this closed form solution using our central differences approximator.
+The analytical and numerical gradients were essentially identical across many different points, even at order $M=9$.
+This is exactly what we want!
+
+We then applied gradient descent to the SSE function using the analytical gradient to see if we could replicate the plots in Bishop.
+Optimizing our hyperparameters got increasingly difficult as we increased the order $M$.
+For $M=O$, we merely needed to tune the step size and we could achieve consistent convergence to the Bishop plot across many different starting values.
+At $M=1$, we needed to change both the step size and increase the number of maximum iterations before we consistently converged to the optimum.
+At $M=3$, we also had to tune the convergence threshold to be several order of magnitudes smaller than it was before, and further increase the number of max interations to get consistent convergence.
+Finally $M=9$, regardless of how we tried to tune the hyperparameters, we couldn't manage to find a set of hyperparameters that consistently converged, in fact, even scipy's BFGS optimizer failed to consistently produce a plot similar to Bishop's.
+
+As we expected, our naive gradient descent implementation was vastly inferior the scipy's BFGS optimizer.
+While we could achieve the same level of accuracy as BFGS, our implementation took significantly more iterations.
+Furthermore, significant time and effort is needed to properly tune our algorithm's hyperparameters, while scipy's BFGS implementation didn't require any tuning.
+
+Alternatively, we can use a trigonometric basis expansion rather than a polynomial one and get somewhat similar results.
+We expect that our SSE would converge to 0 as we increased the order $M$.
+Also, we expect the fit to the "true" data generating function would deteriorate again with increasing order $M$.
+This is in fact exactly what we see:
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq0T.png}
+		\caption*{$M=0$}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq1T.png}
+		\caption*{$M=1$}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq3T.png}
+		\caption*{$M=3$}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=2in]{Meq9T.png}
+		\caption*{$M=9$}
+	\end{minipage}
+	\caption{Plots of trigonometric basis expansions with various orders $M$, in red, fitted to the the observed points. The green curve shows the DGF $\sin(2\pi x)$.}
+\end{figure}
+
+Note that orders $M=1$ and $M=3$ provide rather good fits to for the true DGF.
+We should expect as the the DGF is in fact $\sin(2\pi x)$.
+However, if we didn't know that this was the true DGF, it might be problematic using a trigonometric basis expansion as it imposes a structure of cyclicality to the fit which may lead to gross misspecification of the model.
+
 \subsubsection*{Problem 3}
@@
 To begin problem 3, we implemented the closed form solution of Ridge Regression in Python, making use of Scipy and Numpy.
-Note that if you want to test cases where the number of features is greater than the number of data points (something that a well-trained teacher of machine should be skeptical to do), it is useful to calculate the inverse of a matrix using the pseudo-inverse, rather than the inverse.
+Note that if you want to test cases where the number of features is greater than the number of data points (something that a well-trained teacher of machine should be skeptical to do), it is useful to use the pseudo-inverse, rather than the inverse to fit your parameters.
 We learned this the long way after debugging code for hours.
@@
 As $\lambda$ gets higher, we penalize the model for adding additional large weights to $\hat{\theta}$, but our $MSE$ stays relatively large.
-However, these ``large" MSE models are likely to perform better on a new data point!
+However, these "large" MSE models are likely to perform better on a new data point!
 
@@
 \subsubsection*{Problem 4}
+For problem 4, we are working with a highly sparse dataset so we can compare the differences between LASSO and ridge regression.
+We first start by determining the optimal penalty $\lambda$ that minimizes the MSE of test set.
+Our optimal $\lambda$, 0.1, results in an MSE of $0.4395$.
+When we disable the $L_2$ regularizer, our ridge regression produces the same estimated parameters as OLS which results in a MSE of $4.0979$.
+Separately, we determine that $\lambda = .9$ is optimal LASSO penalty.
+This resulted in a MSE of $0.4524$.
+We plotted our 3 obtained fits - OLS, Ridge with $\lambda=0.1$, and LASSO with $\lambda=0.9$ - against the the true DGF:
+
+\begin{center}
+\includegraphics[width=4.5in]{plot4}	
+\end{center}
+
+While its seems the ridge regression seems to fit the true DGF slightly better than LASSO, both these methods vastly outperform standard OLS.
+We further plotted the estimated optimal weight vectors from our 3 models against the true $w$:
+
+\begin{center}
+\includegraphics[width=4.5in]{Weight_Distributions.png}	
+\end{center}
+
+We see that the L1 and L2 regularizers are working as intended!
+Our ridge regression estimates are considerably smaller than our OLS estimates.
+Furthermore, our we see sparse estimation from LASSO.
+This sparsity results from the differences between the L1 and L2 penalty.
+In ridge regression, we are want to minimize the squared error, subject to the constraint that the sum of the weights squared is less than some constant.
+Geometrically, we can think of this constraint as a sphere of some distance around the origin.
+In LASSO, we are minimizing the squared error subject to the constraint the the sum of the absolute values of the weights is less than some constant.
+The L1 constraint can be thought of as a diamond (or regular octahedral) around the origin.
+
+Now consider sets weights that lead to the the same squared error.
+These sets which can represented as ellipses with the OLS estimates at their center, as that's the point the minimizes squared error.
+In Ridge regression, we are looking for the point where one of these ellipses touches the sphere around the origin, while in LASSO we're looking for the ellipse that touched the diamond.
+Since the diamond has shaper points, its fairly likely that the ellipse intersects the diamond at one of its points, we represents some weight begin equal to 0.
 
 \end{document}
+Status API Training Shop Blog About Pricing
+© 2015 GitHub, Inc. Terms Privacy Security Contact Help

### 2015-10-02 problem4.py

@@ -2,10 +2,10 @@ import numpy as np
 import matplotlib.patches as mpatches
 import matplotlib.pyplot as plt
 import scipy.optimize as spo
-import pandas as pd 
+import pandas as pd
 import scipy
 import scipy.io
-import pylab
+import matplotlib.pylab as pylab
 import statsmodels.api as sm
 from ggplot import *
 
@@ -39,17 +39,17 @@ def gradientDescent(f, df, init, lr=0.3, crit=0.0001, maxIter=100000, h=0.0001, 
 		diff = Y_train - XTtheta
 		#print 'Diff manual ', diff
 		#print 'Loss function: ', f_u
-		#print 'Squared Error: ', np.sum(np.dot(np.transpose(diff), diff))*(1./Y_train.size) 
+		#print 'Squared Error: ', np.sum(np.dot(np.transpose(diff), diff))*(1./Y_train.size)
 		#print 'Penalty: ', Lambda * np.sum(np.absolute(init))
 		#print 'Loss: ', np.sum(np.dot(np.transpose(diff), diff))*(1./Y_train.size) + Lambda * np.sum(np.absolute(init))
 		#print 'Lambda: ', Lambda
 		#print 'Theta term manual', np.sum(np.absolute(init))
 		#print 'Test MSE: ', MSE(predict(init, X_test), Y_test)
 		#print 'Train MSE: ', MSE(predict(init, X_train), Y_train_val)
-		#print 'theta', init 
+		#print 'theta', init
 		#print f(init)
 		if verbose == True:
-			print init 
+			print init
 	#print "nIter: %d" % (nIter)
 	#print "Fcall: %d" % (fcall)
 	return init
@@ -71,47 +71,47 @@ def centdiff(x, f, h=0.00001):
 def ridgeRegressionLoss(theta):
 	xTtheta = np.dot(np.transpose(X_train), theta)
 	diff = Y_train - xTtheta
-	loss = np.sum(np.dot(np.transpose(diff), diff))*(1./Y_train.size) + Lambda * np.sum(np.dot(np.transpose(theta), theta))
+	loss = np.sum(np.dot(np.transpose(diff), diff))*(1./Y_train.size) + Lambda2 * np.sum(np.dot(np.transpose(theta), theta))
 
 	return loss
 
 
 def LASSOLoss(theta):
[…]

### 2015-10-03 HW1/tex/HW1.tex
[homework1] reorganize stuff.

--- 2015-10-02 HW1.tex
@@
 	\title{6.867 Problem Set 1}
-	\date{October 1, 2015}
 	\maketitle
 
-\subsubsection*{Problem 1}
+\subsubsection*{Implementing Gradient Descent}
 
-We implemented a rather naive gradient descent in Python using an algorithm similar to the pseudocode below.
+We implemented a rather naive gradient descent algorithm in Python using an approach similar to the pseudocode below.
 The function is written in such a way that it can accept arbitrary scalar functions $f$ and their gradients, $\nabla f(x)$, where $x$ can be a vector of arbitrary length, $n$.
+If $\nabla f(x)$ is not specified, we approximate it using the central differences.
 The goal of this function is to try and return $\arg\min_x {f(x)}$.
@@
 The first of these is a convex function, whereas the second is quite non-convex, with many local minima.
-The results of our gradient descent algorithm for numerous values are found below in Table 1.
+A sample of the results of our gradient descent algorithm for numerous values are found below in Table 1.
 
@@
 While this is the most obvious difference, note that changes in both the learning rate and the convergence criteria also lead to non-trivially different guesses for $\arg\min_x {f(x)}$.
+Note that we do put a cap on the number of iterations allowable, so it is also possible that our non-convex function simply fails to converge, although this doesn't appear to happen in any of the cases illustrated in this text.
 
-Generally speaking, choice of starting value is only in when were are dealing with non-convex functions as different starting values may lead to different local extremums.
-Though if we randomly initialized good starting values, it decreases the number of iterations necessary for convergence.
+Generally speaking, choice of starting value is only important when dealing with non-convex functions, as different starting values may lead to different local extremums.
+That being said, even for convex functions, good starting values decrease the number of iterations necessary for convergence.
 Tuning the step size alters the performance of the algorithm, while tuning the convergence threshold alters the accuracy of the algorithm.
 
-To elaborate on this a little, for any given convergence threshold, increasing step size will decrease the number of iterations required for convergence.
+It's worth elaborating on the nuances of picking a step size and convergence threshold.
+For any given convergence threshold, increasing step size will decrease the number of iterations required for convergence.
 However, we need make sure that the step isn't so large that the algorithm gets caught in a loop where its constantly overshooting the minimum.
-On the other hand, for a given step size, decreasing the convergence threshold increases accuracy of the algorithm, but at the cost of performance as it'll take additional iterations to reach the higher level of precision.
-Here, we want to make sure that the convergence threshold isn't so small that we can never converge.
+On the other hand, for a given step size, decreasing the convergence threshold increases accuracy of the algorithm, but it will take additional iterations to reach this higher level of precision.
+We want to make sure that the convergence threshold isn't so small that we can never converge.
 
 Unfortunately, it might not always be convenient to specify the gradient of our scalar function, $f(x)$, explicitly.
-This might be because it is complicated to compute, or because no closed form solution exists.
+It may be complicated to compute, or there may not be a closed form solution.
 In order to get around this issue, we can approximate the gradient of the function using central differences.
@@
 
-\noindent where $h$ is some small constant (we use $h = .001$).
+\noindent where $h$ is some small constant (we typically use $h = .001$).
 
-We see that the performance in the calculation of the gradient is identical between the analytical approach and the central difference approach.
+The central difference approximation of the gradient is equivalent to the analytical solution for both the convex and non-convex function.
 That's great!
-Now let's see how our gradient descent algorithm compares to two more sophisticated optimizers - BFGS and Conjugate Gradient (CG).
-We compare both the number of iterations and function evaluations.
-We'll consider our algorithm in two forms - one where we pass in an explicit gradient function, and another where we estimate the gradient using the central difference method.
+We also want to know how our gradient descent algorithm compares to more sophisticated optimizers - BFGS and Conjugate Gradient (CG).
+We compare both the number of iterations and function evaluations required to calculate the minimum of the function.
+We'll check both the case where we provide an expression for the gradient, and the case where we estimate the gradient.
 All benchmarking is done with the same starting point: [5, -3, 7, 8].
@@
 If we use the central difference approximation for $\nabla f(x)$ instead, our algorithm requires many more iterations and function calls.
+Later on in this paper, we'll make use of this algorithm to solve some machine learning problems.
 
@@
 &    [5.0, -3.0, 7.0, 8.0] &    0.3 &    0.1 &   [0.020, -0.012, 0.029, 0.033] &    [6.150, -0.314, 6.165, 6.176] \\
-&    [5.0, -3.0, 7.0, 8.0] &    0.3 &  0.001 &   [0.001, -0.001, 0.002, 0.002] &    [6.159, -0.025, 6.160, 6.161] \\
 &    [5.0, -3.0, 7.0, 8.0] &    0.3 &  1e-05 &   [0.000, -0.000, 0.000, 0.000] &    [6.160, -0.003, 6.160, 6.160] \\
 &    [5.0, -3.0, 7.0, 8.0] &  0.003 &    0.1 &   [1.172, -0.703, 1.641, 1.876] &    [5.005, -2.999, 6.995, 7.993] \\
-&    [5.0, -3.0, 7.0, 8.0] &  0.003 &  0.001 &   [0.118, -0.071, 0.165, 0.188] &    [6.092, -0.932, 6.200, 6.267] \\
 &    [5.0, -3.0, 7.0, 8.0] &  0.003 &  1e-05 &   [0.012, -0.007, 0.017, 0.019] &    [6.092, -0.932, 6.200, 6.267] \\
- &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &    0.1 &  [-0.041, 0.016, 0.004, -0.020] &  [-12.059, 6.070, 0.014, -6.139] \\
-&  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  0.001 &  [-0.003, 0.001, 0.000, -0.001] &  [-12.294, 6.152, 0.001, -6.158] \\
- &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  1e-05 &  [-0.000, 0.000, 0.000, -0.000] &  [-12.315, 6.159, 0.000, -6.159] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &    0.1 &  [-2.388, 0.955, 0.239, -1.194] &  [-10.002, 4.004, 0.995, -5.005] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  0.001 &  [-0.240, 0.096, 0.024, -0.120] &  [-11.748, 5.885, 0.063, -6.076] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  1e-05 &  [-0.024, 0.010, 0.002, -0.012] &  [-11.848, 5.938, 0.051, -6.092] \\
  &   [1.0, -1.0, -3.0, 2.0] &    0.3 &    0.1 &  [0.010, -0.010, -0.031, 0.020] &   [0.007, -0.007, -0.314, 0.024] \\
- &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  0.001 &  [0.001, -0.001, -0.002, 0.001] &   [0.001, -0.001, -0.025, 0.002] \\
  &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  1e-05 &  [0.000, -0.000, -0.000, 0.000] &   [0.000, -0.000, -0.003, 0.000] \\
  &   [1.0, -1.0, -3.0, 2.0] &  0.003 &    0.1 &  [0.736, -0.736, -2.207, 1.471] &   [0.995, -0.995, -2.999, 1.994] \\
- &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  0.001 &  [0.073, -0.073, -0.220, 0.147] &   [0.051, -0.051, -0.932, 0.143] \\
  &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  1e-05 &  [0.007, -0.007, -0.022, 0.015] &   [0.051, -0.051, -0.932, 0.143] \\
@@
 &    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n \frac{1}{100} x_i^2 - \cos(x)$ & [-0.8589 -0.2011  0.797   1.1494] &[-0.8589 -0.2011  0.797   1.1494] \\
-&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^2$ &   [-20.   8.   2. -10.] &  [-20.   8.   2. -10.]\\
-&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n \frac{1}{100} x_i^2 - \cos(x)$ &[ 0.344  -0.6768  0.8615  0.8589]&[ 0.344  -0.6768  0.8615  0.8589] \\
 &    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
@@
 
-\subsubsection*{Problem 2}
+\subsubsection*{Linear Basis Function Regression}
 
-With problem 2, we first derive the maximum-likelihood estimator for the weight vector $\theta$.
+We now consider the linear combination of basis function class regression models.
+We first derive the maximum-likelihood estimator for the weight vector $\theta$.
 Note that if we assume Gaussian errors, that is $Y \sim N(X\theta, \sigma^2)$, then our maximum-likelihood estimator is equivalent to the OLS estimator.
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq0.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq0.png}
 		\caption*{$M=0$}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq1.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq1.png}
 		\caption*{$M=1$}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq3.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq3.png}
 		\caption*{$M=3$}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq9.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq9.png}
 		\caption*{$M=9$}
@@
 \begin{equation*}
-	SSE=(\Phi \theta-Y)^T(\Phi \theta-Y) 
+	SSE=(\Phi \theta-Y)^T(\Phi \theta-Y),         	\nabla SSE = 2\Phi^T(\Phi \theta-Y)
 \end{equation*}
-which has gradient:
-\begin{equation*}
-	\nabla SSE = 2\Phi^T(\Phi \theta-Y)
-\end{equation*}
-We tested the accuracy to this closed form solution using our central differences approximator.
-The analytical and numerical gradients were essentially identical across many different points, even at order $M=9$.
+We tested the accuracy of this closed form solution using central differences.
+The analytical and numerical gradients were essentially identical across many points, even at order $M=9$.
 This is exactly what we want!
@@
 Optimizing our hyperparameters got increasingly difficult as we increased the order $M$.
-For $M=O$, we merely needed to tune the step size and we could achieve consistent convergence to the Bishop plot across many different starting values.
+For $M=0$, we merely needed to tune the step size and we could achieve consistent convergence to the Bishop plot across many different starting values.
 At $M=1$, we needed to change both the step size and increase the number of maximum iterations before we consistently converged to the optimum.
 At $M=3$, we also had to tune the convergence threshold to be several order of magnitudes smaller than it was before, and further increase the number of max interations to get consistent convergence.
-Finally $M=9$, regardless of how we tried to tune the hyperparameters, we couldn't manage to find a set of hyperparameters that consistently converged, in fact, even scipy's BFGS optimizer failed to consistently produce a plot similar to Bishop's.
+For $M=9$, regardless of how we tried to tune the hyperparameters, we couldn't manage to find a set of hyperparameters that consistently converged, in fact, even scipy's BFGS optimizer failed to consistently produce a plot similar to Bishop's.
 
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq0T.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq0T.png}
 		\caption*{$M=0$}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq1T.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq1T.png}
 		\caption*{$M=1$}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq3T.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq3T.png}
 		\caption*{$M=3$}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=2in]{Meq9T.png}
+		\includegraphics[width=1\linewidth, height=1in]{Meq9T.png}
 		\caption*{$M=9$}
@@
 
-\subsubsection*{Problem 3}
+\subsubsection*{Implementing Ridge Regression}
 
-To begin problem 3, we implemented the closed form solution of Ridge Regression in Python, making use of Scipy and Numpy.
+Let's now see how regression behaves when we impose a regularization term.
+The regularization term, $\lambda$, serves to make sure we don't overfit to our training data.
+We implemented the closed form solution of Ridge Regression in Python, making use of Scipy and Numpy.
 Note that if you want to test cases where the number of features is greater than the number of data points (something that a well-trained teacher of machine should be skeptical to do), it is useful to use the pseudo-inverse, rather than the inverse to fit your parameters.
@@
 Given the data in Bishop 1.4 (which can be seen below), we attempt to fit polynomials of various orders, $M$.
-
-\begin{center}
-\includegraphics[scale=.3]{Bishop_14_Data.png}
-\end{center}
-
-The plot below shows the mean squared error (MSE) that our ridge regression fit gives for values of $M$ between 1 and 5, for values of $\lambda$ ranging from $10^0 -1$ to $10^2$.
+The plot below shows the mean squared error (MSE) that our ridge regression fit gives for values of $M$ between 1 and 5, for values of $\lambda$ ranging from $(10^0 -1)$ to $10^2$.
 Note that in the case where $\lambda = 0$, we have $OLS$ regression, and as $M$ increases, the model is free to overfit our data.
-While the MSE here is low, if we evaluated this on a new data point, our model would likely perform very poorly.
-As $\lambda$ gets higher, we penalize the model for adding additional large weights to $\hat{\theta}$, but our $MSE$ stays relatively large.
+While the MSE on training data is low, if we evaluated this on a new data point, our model would likely perform very poorly.
+As $\lambda$ gets higher, we penalize the model for having large $\hat{\theta}$ components.
+Our $MSE$s for these models are relatively large.
 However, these "large" MSE models are likely to perform better on a new data point!
 
-\begin{center}
-\includegraphics[scale=.4]{MSE_Lambda_Bishop.png}
-\end{center}
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.48\linewidth}
+                \includegraphics[width=1\linewidth]{Bishop_14_Data.png}
+		\caption*{Original Bishop Data}
+	\end{minipage}
+	\begin{minipage}[b]{.48\linewidth}
+                \includegraphics[width=1\linewidth]{MSE_Lambda_Bishop.png}
+		\caption*{$MSE$($M$,$\lambda$)}
+	\end{minipage}
+\end{figure}
 
-In order to avoid overfitting, but in order to make sure that we have chosen values of $\lambda$ and $M$ (and in general, hyperparameters) that make sense, it is common to split the data into a training set and a test set.
+In order to avoid overfitting, while making sure that we have chosen values of $\lambda$ and $M$ that make sense, it is common to split the data into a training set and a test set.
 We will use the closed-form solution for ridge regression to calculate $\hat{\theta}$ on our training set.
 We will then use that $\hat{\theta}$ to attempt to predict values of $\mathbf{y}$ for the test data.
-By choosing the $\hat{\theta}$ that minimizes MSE on the test data, but is trained on the training data, we are ensuring that we have not overfit our model to the data it is trained on.
-As one final step in the model selection process, we can check the performance of our final $\hat{\theta}$ on a validation set, to make sure that performance is still good (and we have not overfit for performance on our test set).
+By choosing the $\hat{\theta}$ that is trained on the training data and minimizes MSE on the test data, we are ensuring that we have not overfit our model to the data it is trained on.
+As one final step in the model selection process, we can check the performance of our final $\hat{\theta}$ on a validation set, to make sure that performance is still good (and we have not overfit to the data in on our test set).
 
@@
 We are hoping to predict the number of comments it will receive in the next 24 hours.
-We are given this data split into training, validation, and test datasets, so we will once again perform a grid search to find the optimal value of $\lambda$.
+This data is split into training, validation, and test datasets, so we will once again perform a grid search to find the value of $\lambda$ that minimizes $MSE$ on our validation set.
 In this particular case, we are not regressing a polynomial of arbitrary order, so we do not need to test different values of $M$.
 
-\begin{center}
-\includegraphics[scale=.4]{MSE_comparison.png}
-\end{center}
-
 We find that for very large $\lambda$ (as large as $10^6$, $MSE$ continues to decrease, although it is always close to 900.
-The MSE for a uniform predictor that predicts the average of $y$, $\hat{y}$, has an MSE of 1453, so the performance of our model is still better.
+A predictor that uniformly predicts the average value of $y$, $\hat{y}$, has an $MSE$ of 1453, so the performance of our model is still better.
 However, we see that for values of $\lambda$ higher than $10^6$ or so, the value of the MSE increases, and eventually approaches (and eclipses) the MSE of our uniform estimator.
@@
 This isn't surprising - the blog feedback dataset is large, with 10,480 observations in the test set.
-If we interpret the value of $\lambda$ as strength of our prior, it makes sense that it does not have a substantial effect until it is significantly larger than the size of our data.
+If we interpret the value of $\lambda$ as the strength of our Bayesian prior on our model parameters, it makes sense that our choice of $\lambda$ does not have a substantial effect unless it is significantly larger than the size of our data.
 Furthermore, the inability of our model to get very low MSE for low $\lambda$ speaks both to the fact that the data has a much higher dimension than our training data, and also suggests that the simple linear model we are using here may not be a very good fit for the data.
 
-\begin{center}
-\includegraphics[scale=.4]{BlogFeedbackRegressionMSE.png}
-\end{center}
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.48\linewidth}
+                \includegraphics[width=1\linewidth]{MSE_comparison.png}
+                \caption*{$MSE$ across different subsets of toy data}                
+	\end{minipage}
+	\begin{minipage}[b]{.48\linewidth}
+                \includegraphics[width=1\linewidth]{BlogFeedbackRegressionMSE.png}
+                \caption*{$MSE$($\lambda$) for blog data}
+	\end{minipage}
+\end{figure}
 
-\subsubsection*{Problem 4}
-For problem 4, we are working with a highly sparse dataset so we can compare the differences between LASSO and ridge regression.
-We first start by determining the optimal penalty $\lambda$ that minimizes the MSE of test set.
+\subsubsection*{Sparse Regression with LASSO}
+
+Let's now consider the opposite of a 10,480 point dataset - the case where we regress with only 5  training points.
+For this case, we might consider using LASSO, which penalizes our model with the magnitude of $\lambda$ (as opposed to its squared magnitude).
+Let's compare LASSO and ridge regression.
+We first start by determining the optimal penalty $\lambda$ that minimizes the ridge regression $MSE$ on the test set.
 Our optimal $\lambda$, 0.1, results in an MSE of $0.4395$.
-When we disable the $L_2$ regularizer, our ridge regression produces the same estimated parameters as OLS which results in a MSE of $4.0979$.
-Separately, we determine that $\lambda = .9$ is optimal LASSO penalty.
-This resulted in a MSE of $0.4524$.
-We plotted our 3 obtained fits - OLS, Ridge with $\lambda=0.1$, and LASSO with $\lambda=0.9$ - against the the true DGF:
+When we disable the $L_2$ regularizer, our ridge regression produces the same estimated parameters as OLS, which results in a MSE of $4.0979$.
+Separately, we determine that $\lambda = .9$ is the optimal LASSO penalty.
+This results in a $MSE$ of $0.4524$.
+We plot our 3 obtained fits - OLS, Ridge with $\lambda=0.1$, and LASSO with $\lambda=0.9$ - against the the true DGF.
 
-\begin{center}
-\includegraphics[width=4.5in]{plot4}	
-\end{center}
+While its seems the ridge regression fits the true DGF slightly better than LASSO, both these methods vastly outperform standard OLS.
+We also plot the estimated optimal weight vectors from our 3 models against the true $w$:
 
-While its seems the ridge regression seems to fit the true DGF slightly better than LASSO, both these methods vastly outperform standard OLS.
-We further plotted the estimated optimal weight vectors from our 3 models against the true $w$:
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.48\linewidth}
+                \includegraphics[width=1\linewidth]{plot4}
+	\end{minipage}
+	\begin{minipage}[b]{.48\linewidth}
+                \includegraphics[width=1\linewidth]{Weight_Distributions.png}
+	\end{minipage}
+\end{figure}
 
-\begin{center}
-\includegraphics[width=4.5in]{Weight_Distributions.png}	
-\end{center}
-
-We see that the L1 and L2 regularizers are working as intended!
+We see that the $L_1$ and $L_2$ regularizers are working as intended!
 Our ridge regression estimates are considerably smaller than our OLS estimates.
 Furthermore, our we see sparse estimation from LASSO.
-This sparsity results from the differences between the L1 and L2 penalty.
-In ridge regression, we are want to minimize the squared error, subject to the constraint that the sum of the weights squared is less than some constant.
+This sparsity results from the differences between the $L_1$ and $L_2$ penalty.
+In ridge regression, we want to minimize the squared error, subject to the constraint that the sum of the weights squared is less than some constant.
 Geometrically, we can think of this constraint as a sphere of some distance around the origin.
 In LASSO, we are minimizing the squared error subject to the constraint the the sum of the absolute values of the weights is less than some constant.
-The L1 constraint can be thought of as a diamond (or regular octahedral) around the origin.
+The $L_1$ constraint can be thought of as a diamond (or regular octahedral) around the origin.
 
-Now consider sets weights that lead to the the same squared error.
-These sets which can represented as ellipses with the OLS estimates at their center, as that's the point the minimizes squared error.
-In Ridge regression, we are looking for the point where one of these ellipses touches the sphere around the origin, while in LASSO we're looking for the ellipse that touched the diamond.
-Since the diamond has shaper points, its fairly likely that the ellipse intersects the diamond at one of its points, we represents some weight begin equal to 0.
+Now consider sets of weights that lead to the the same squared error.
+These sets can represented as ellipses with the OLS estimates at their center, as that's the point the minimizes squared error.
+In ridge regression, we're looking for the point where one of these ellipses touches the sphere around the origin, while in LASSO we're looking for the ellipse that touches the diamond.
+Since the diamond has sharper points, its fairly likely that the ellipse intersects the diamond at one of its points, which represents some weights = 0.
+It seems that if you expect your weights are sparse, LASSO is the correct regression method to use!
 

### 2015-10-03 HW1/code/problem2.py
Added problem 2

@@ -1,0 +1,146 @@
+import numpy as np
+import pandas as pd
+import random
+import scipy as sp
+import homework1
+import sys
+import matplotlib as mpl
+from matplotlib import pylab as pl
+from ggplot import *
+
+np.set_printoptions(precision=4)
+# Maximum Likelihood Weight Vector Function assuming Gaussian Errors (aka OLS)
+def OLS(X, Y, M=1):
+	phi = createTFeatures(X, M)
+	XtX = np.dot(np.transpose(phi), phi)
+	XtY = np.dot(np.transpose(phi), Y)
+	theta = sp.linalg.solve(XtX, XtY)
+	return theta
+
+def yPred(X, Y, M=1):
+    pl.plot(X.T.tolist()[0],Y.T.tolist()[0], 'gs')
+    phi = createTFeatures(X, M)
+    w = OLS(X, Y, M)
+    pts = np.array([[p] for p in pl.linspace(min(X), max(X), 100)])
+    yhat = np.dot(createTFeatures(pts, M), w)
+    return yhat
+
+def plot(yhat):
+	pts = np.array([[p] for p in pl.linspace(min(X), max(X), 100)])
+	true = np.sin(2*np.pi*pts)
+	DF = pd.DataFrame({'X': pd.Series(X.flatten()), 'Y': pd.Series(Y.flatten()), 'axis': pd.Series(pts.flatten()), 'yhat': pd.Series(yhat.flatten()), 'true': pd.Series(true.flatten())})
+	mpl.rcParams["figure.figsize"] = "2, 4"
+	print ggplot(DF, aes(x='axis', y='yhat')) + geom_line(color="red", size=4) + geom_line(aes(y='true'), color='green', size=2) + geom_point(aes(x='X', y='Y'), size = 200, alpha=0.5) + xlim(-0.1,1.1) + xlab('X') + ylab('Y')
+
+def createTFeatures(X=np.zeros(1), M=1):
+    features = X**0
+    for i in range(1, M+1):
+        features = np.hstack([features, np.sin(2*np.pi*X*i)])
+    return features
+
+def predict(X, theta):
+	features = createTFeatures(X, M=theta.size-1)
+	yhat = np.dot(features, theta)
+	return yhat
+
+def SSE(theta):
+	yhat = predict(X, theta)
+	res = yhat - Y.flatten()
[…]

### 2015-10-19 HW2/problem2/plotBoundary.py
[hw2] problem 2 wont work.

@@ -1,0 +1,28 @@
+import pdb
+from numpy import *
+import pylab as pl
+
+# X is data matrix (each row is a data point)
+# Y is desired output (1 or -1)
+# scoreFn is a function of a data point
+# values is a list of values to plot
+
+def plotDecisionBoundary(X, Y, scoreFn, values, title = ""):
+    # Plot the decision boundary. For that, we will asign a score to
+    # each point in the mesh [x_min, m_max]x[y_min, y_max].
+    x_min, x_max = X[:, 0].min() - 1, X[:, 0].max() + 1
+    y_min, y_max = X[:, 1].min() - 1, X[:, 1].max() + 1
+    h = max((x_max-x_min)/200., (y_max-y_min)/200.)
+    xx, yy = meshgrid(arange(x_min, x_max, h),
+                      arange(y_min, y_max, h))
+    zz = array([scoreFn(x) for x in c_[xx.ravel(), yy.ravel()]])
+    zz = zz.reshape(xx.shape)
+    print xx, yy, zz
+    pl.figure()
+    CS = pl.contour(xx, yy, zz, values, colors = 'green', linestyles = 'solid', linewidths = 2)
+    pl.clabel(CS, fontsize=9, inline=1)
+    # Plot the training points
+    pl.scatter(X[:, 0], X[:, 1], c=(1.-Y), s=50, cmap = pl.cm.cool)
+    pl.title(title)
+    pl.axis('tight')
+    pl.show()

### 2015-10-19 HW2/problem2/problem2.py

@@ -1,0 +1,115 @@
+import numpy as np
+from plotBoundary import *
+import cvxopt
+# import your SVM training code
+
+# parameters
+name = 'nonsep'
+print '======Training======'
+# load data from csv files
+train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
+# use deep copy here to make cvxopt happy
+X = train[:, 0:2].copy()
+Y = train[:, 2:3].copy()
+
+def linear_kernel(x1, x2):
+    return np.dot(x1, x2)
+
+def polynomial_kernel(x, y, p=3):
+    return (1 + np.dot(x, y)) ** p
+
+def gaussian_kernel(x, y, sigma=1):
+    return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
+
+#X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
+#y = np.array([1., 1., -1., -1.])
+C = 1
+kernel = gaussian_kernel
+
+
+# Carry out training, primal and/or dual
+def trainSVM(X, y, kernel, C):
+	n_samples, n_features = X.shape
+	K = np.zeros((n_samples, n_samples))
+	for i in range(n_samples):
+		for j in range(n_samples):
+			K[i, j] = kernel(X[i], X[j])
+
+	P = cvxopt.matrix(np.outer(y,y) * K)
+	q = cvxopt.matrix(np.ones(n_samples) * -1.)
+	A = cvxopt.matrix(y, (1, n_samples))
+	b = cvxopt.matrix(0.0)
+
+	G = cvxopt.matrix(np.vstack([np.diag(np.ones(n_samples) * -1.), np.diag(np.ones(n_samples))]))
+	h = cvxopt.matrix(np.hstack([np.zeros(n_samples), C*np.ones(n_samples)])) 
+ 
+	# Solve QP problem
+	solution = cvxopt.solvers.qp(P, q, G, h, A, b)
+	alpha = np.ravel(solution['x'])
+
+	sv_bool = alpha > .00001
+	ind = np.arange(len(alpha))[sv_bool]
+	alpha = alpha[sv_bool]
+	sv_y = y[sv_bool]
+	sv = X[sv_bool]
+	print "%d support vectors out of %d points" % (len(alpha), n_samples)
+
+	return n_features, K, alpha, sv, sv_y, sv_bool, ind
+
[…]

### 2015-10-19 HW2/problem2/problem2.py
[hw2] ahhh.

@@ -105,7 +105,7 @@ print '======Plot Training======'
 plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'SVM Train')
 
 
-print '======Validation======'
+print '======Validation======='
 # load data from csv files
 validate = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_validate.csv')
 X = validate[:, 0:2]

### 2015-10-20 HW2/problem2/plotBoundary.py
[hw2] fixing bugs of unknown provenance.

@@ -15,7 +15,8 @@ def plotDecisionBoundary(X, Y, scoreFn, values, title = ""):
     h = max((x_max-x_min)/200., (y_max-y_min)/200.)
     xx, yy = meshgrid(arange(x_min, x_max, h),
                       arange(y_min, y_max, h))
-    zz = array([scoreFn(x) for x in c_[xx.ravel(), yy.ravel()]])
+    X_tmp = array([[x1, x2] for x1, x2 in zip(ravel(xx), ravel(yy))])
+    zz = array([scoreFn(X_tmp)])
     zz = zz.reshape(xx.shape)
     print xx, yy, zz
     pl.figure()

### 2015-10-20 HW2/problem2/problem2.py

@@ -24,7 +24,7 @@ def gaussian_kernel(x, y, sigma=1):
 #X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
 #y = np.array([1., 1., -1., -1.])
 C = 1
-kernel = gaussian_kernel
+kernel = linear_kernel
 
 
 # Carry out training, primal and/or dual
@@ -75,12 +75,12 @@ def predictSVM(X):
 		theta_0 /= len(alpha)
 
 		# Weight vector
-	#if kernel == linear_kernel:
-	weight = np.zeros(n_features)
-	for n in range(len(alpha)):
-		weight += alpha[n] * sv_y[n] * sv[n]
-	#else:
-	#	weight = None
+	if kernel == linear_kernel:
+		weight = np.zeros(n_features)
+		for n in range(len(alpha)):
+			weight += alpha[n] * sv_y[n] * sv[n]
+	else:
+		weight = None
 
 	if weight != None:
 		print [X, np.sign(np.dot(X, weight) + theta_0)]
@@ -90,14 +90,13 @@ def predictSVM(X):
 		if len(np.array(X).shape) == 1:
 			n_points = 1 
 		else:
-			n_points = np.array(X).shape[2]
+			n_points = np.array(X).shape[0]
 		y_predict = np.zeros(n_points)
 		for i in range(n_points):
 			s = 0 
 			for a_0, sv_y_0, sv_0 in zip(alpha, sv_y, sv):
 				s += a_0*sv_y_0*kernel(X[i], sv_0)
 			y_predict[i] = s
-		print [X, np.sign(y_predict + theta_0)]
 		return np.sign(y_predict + theta_0)
 
 # plot training results
@@ -105,7 +104,7 @@ print '======Plot Training======'
 plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'SVM Train')
 
 
-print '======Validation======='
+print '======Validation=======	'
 # load data from csv files
 validate = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_validate.csv')
 X = validate[:, 0:2]

### 2015-10-20 HW2/tex/HW2.tex
[hw2] more reorganization.

--- 2015-10-01 HW1.tex
@@
 \begin{document}
-	\title{6.867 Problem Set 1}
-	\date{October 1, 2015}
+	\title{6.867 Problem Set 2}
 	\maketitle
 
-\subsubsection*{Problem 1}
-
-We implemented gradient descent in Python using an algorithm similar to the pseudocode below.
-The function is written in such a way that it can accept arbitrary scalar functions $f$ and gradients, $\nabla f(x)$, where $f(x)$ here is a function of an arbitrary number of parameters, $n$, and return $\min{f(x)}$.
-
-\begin{algorithm}
-\caption{Gradient Descent}
-\label{GradDescent}
-\begin{algorithmic}[1]
-\Procedure{Gradient Descent}{}
-\State Specify initial guess, step size, and convergence criteria
-\State Set current guess equal to initial guess
-\While{Distance between last guess and current guess location is greater than convergence criteria}
-\State Calculate function at current location
-\State Calculate the gradient of function at current location
-\State Updated guess = take a step from current guess in direction of gradient proportional to step size
-\State Calculate distance between updated guess and current guess
-\State Set current guess equal to updated guess
-\EndWhile{}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-In order to understand the effect of various hyperparameters (our initial guess, the step size, and the convergence criteria) of the algorithm, we compared the results of our gradient descent routine given a number of different parameter combinations for two different functions: $f(x) = \sum_i^n x_i^2$ and $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$.
-The first of these is a convex function, whereas the second is quite non-convex, with many local minima.
-The results of our gradient descent algorithm for numerous values are found below.
-
-We find that, in general, the convex function is much better behaved.
-Regardless of our initial guess or our learning rate, the algorithm converges to a guess in fewer than 1,000 steps, and these guess are consistently close to $x = (0,0,0,0)$, the ``true" value.
-The final guess' distance from the true value of $x$ is primarily a function of the convergence criteria, which makes sense - the less picky we are with our convergence criteria, the further from the true value of $x$ our final guess will be.
-
-The non-convex, on the other hand, is much more sensitive to the hyperparameters we specify.
-The most notable departure from the convex function is the influence of our initial guess.
-It is much easier for the gradient descent algorithm to get trapped in local minima, and this is clearly illustrated in our results.
-While this is the most obvious difference, note that changes in both the learning rate and the convergence criteria also lead to non-trivially different guesses for $\min{f(x)}$.
-
-Unfortunately, it might not always be convenient to specify the gradient of our scalar function, $f(x)$, explicitly.
-This might be because it is complicated to compute, or because no closed form solution exists.
-In order to get around this issue, we can approximate the gradient of the function using central differences.
-We approximate the $i$th element of the gradient, $\nabla_i$, as
-
-\begin{equation}
-\nabla_i \approx \frac{f(x_i+h) - f(x_i-h)}{2h}
-\end{equation}
-
-\noindent where $h$ is some small constant (we use $h = .001$).
-
-We see that the performance in the calculation of the gradient is identical between the analytical approach and the central difference approach.
-That's great!
-Now let's see how our gradient descent algorithm compares to two other algorithms - Broyden?Fletcher?Goldfarb?Shanno (BFGS) and Conjugate Gradient (CG).
-We compare both the number of iterations and function evaluations.
-We'll consider our algorithm in two forms - one where we pass in an explicit gradient function, and another where we estimate the gradient using the central difference method.
-All benchmarking is done with the same starting point: [5, -3, 7, 8].
-$f_1(x) =  \sum_i^n x_i^2$ and $f_2(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$.
-We see that when passed an analytical solution for $\nabla f(x)$, our algorithm requires about as many function calls, although many more iterations than both BFGS and CG.
-Given we do not have an analytical expression for $\nabla f(x)$, however, our algorithm requires many more iterations and function calls.
-
-\begin{table}
-\captionof{table}{Performance of gradient descent given various algorithm hyperparameters} 
-\begin{tabular}{llllll}
-\toprule
-{} &                        $Guess_{i}$ &      $R_{learn}$ &   Thresh. &    $Guess_{f}$ ($f(x) = \sum_i^n x_i^2$) & $Guess_{f}$ ($f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$) \\
-\midrule
-&    [5.0, -3.0, 7.0, 8.0] &    0.3 &    0.1 &   [0.020, -0.012, 0.029, 0.033] &    [6.150, -0.314, 6.165, 6.176] \\
-&    [5.0, -3.0, 7.0, 8.0] &    0.3 &  0.001 &   [0.001, -0.001, 0.002, 0.002] &    [6.159, -0.025, 6.160, 6.161] \\
-&    [5.0, -3.0, 7.0, 8.0] &    0.3 &  1e-05 &   [0.000, -0.000, 0.000, 0.000] &    [6.160, -0.003, 6.160, 6.160] \\
-&    [5.0, -3.0, 7.0, 8.0] &  0.003 &    0.1 &   [1.172, -0.703, 1.641, 1.876] &    [5.005, -2.999, 6.995, 7.993] \\
-&    [5.0, -3.0, 7.0, 8.0] &  0.003 &  0.001 &   [0.118, -0.071, 0.165, 0.188] &    [6.092, -0.932, 6.200, 6.267] \\
-&    [5.0, -3.0, 7.0, 8.0] &  0.003 &  1e-05 &   [0.012, -0.007, 0.017, 0.019] &    [6.092, -0.932, 6.200, 6.267] \\
- &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &    0.1 &  [-0.041, 0.016, 0.004, -0.020] &  [-12.059, 6.070, 0.014, -6.139] \\
-&  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  0.001 &  [-0.003, 0.001, 0.000, -0.001] &  [-12.294, 6.152, 0.001, -6.158] \\
- &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  1e-05 &  [-0.000, 0.000, 0.000, -0.000] &  [-12.315, 6.159, 0.000, -6.159] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &    0.1 &  [-2.388, 0.955, 0.239, -1.194] &  [-10.002, 4.004, 0.995, -5.005] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  0.001 &  [-0.240, 0.096, 0.024, -0.120] &  [-11.748, 5.885, 0.063, -6.076] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  1e-05 &  [-0.024, 0.010, 0.002, -0.012] &  [-11.848, 5.938, 0.051, -6.092] \\
- &   [1.0, -1.0, -3.0, 2.0] &    0.3 &    0.1 &  [0.010, -0.010, -0.031, 0.020] &   [0.007, -0.007, -0.314, 0.024] \\
- &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  0.001 &  [0.001, -0.001, -0.002, 0.001] &   [0.001, -0.001, -0.025, 0.002] \\
- &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  1e-05 &  [0.000, -0.000, -0.000, 0.000] &   [0.000, -0.000, -0.003, 0.000] \\
- &   [1.0, -1.0, -3.0, 2.0] &  0.003 &    0.1 &  [0.736, -0.736, -2.207, 1.471] &   [0.995, -0.995, -2.999, 1.994] \\
- &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  0.001 &  [0.073, -0.073, -0.220, 0.147] &   [0.051, -0.051, -0.932, 0.143] \\
- &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  1e-05 &  [0.007, -0.007, -0.022, 0.015] &   [0.051, -0.051, -0.932, 0.143] \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-\captionof{table}{Performance in gradient calculation for analytical and central difference approaches} 
-\begin{tabular}{llllll}
-\toprule
-{} &   $x_{i}$  &  function &   $\nabla_{analytical}$  & $\nabla_{approximate}$ \\
-\midrule
-&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
-&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [-0.8589 -0.2011  0.797   1.1494] &[-0.8589 -0.2011  0.797   1.1494] \\
-&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^2$ &   [-20.   8.   2. -10.] &  [-20.   8.   2. -10.]\\
-&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ &[ 0.344  -0.6768  0.8615  0.8589]&[ 0.344  -0.6768  0.8615  0.8589] \\
-&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
-&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [ 0.8615 -0.8615 -0.2011  0.9493] & [ 0.8615 -0.8615 -0.2011  0.9493] \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-\captionof{table}{Performance of authors' gradient descent algorithm compared to BFGS and CG}
-\begin{tabular}{llllll}
-\toprule
-{} & algorithm & $f_1(x)$ function calls & $f_1(x)$ iterations &  $f_2(x)$ function calls &  $f_2(x)$ iterations \\
-\midrule
-  & BFGS & 24 & 2 & 78 & 10 \\
-  & CG & 30 & 2 & 78 & 6 \\
-  & Gradient Descent (analytical) & 20 & 10 & 48 & 24 \\ 
-  & Gradient Descent (approximate) & 100 & 10 & 240 & 24 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection*{Problem 2}
-
-\subsubsection*{Problem 3}
-
-To begin problem 3, we implemented the closed form solution of Ridge Regression in Python, making use of Scipy and Numpy.
-Note that if you want to test cases where the number of features is greater than the number of data points (something that a well-trained teacher of machine should be skeptical to do), it is useful to calculate the inverse of a matrix using the pseudo-inverse, rather than the inverse.
-We learned this the long way after debugging code for hours.
-
-The closed form solution for ridge regression is given by the expression:
-
-\begin{equation}
-\hat{\theta} = (\mathbf{X}^T\mathbf{X} + \lambda \mathbf{I})^{-1}\mathbf{X}^t\mathbf{y}
-\end{equation}
-
-\noindent where $\mathbf{X}$ is your matrix of features, $y$ is your array of observed values, and $\hat{\theta}$ are your estimated feature weights.
-$\lambda$ is a regularization parameter, which represents how much you penalize the model for having large feature weights.
-
-Given the data in Bishop 1.4 (which can be seen below), we attempt to fit polynomials of various orders, $M$.
-
-\begin{center}
-\includegraphics[scale=.3]{Bishop_14_Data.png}
-\end{center}
-
-The plot below shows the mean squared error (MSE) that our ridge regression fit gives for values of $M$ between 1 and 5, for values of $\lambda$ ranging from $10^0 -1$ to $10^2$.
-Note that in the case where $\lambda = 0$, we have $OLS$ regression, and as $M$ increases, the model is free to overfit our data.
-While the MSE here is low, if we evaluated this on a new data point, our model would likely perform very poorly.
-As $\lambda$ gets higher, we penalize the model for adding additional large weights to $\hat{\theta}$, but our $MSE$ stays relatively large.
-However, these ``large" MSE models are likely to perform better on a new data point!
-
-\begin{center}
-\includegraphics[scale=.4]{MSE_Lambda_Bishop.png}
-\end{center}
-
-In order to avoid overfitting, but in order to make sure that we have chosen values of $\lambda$ and $M$ (and in general, hyperparameters) that make sense, it is common to split the data into a training set and a test set.
-We will use the closed-form solution for ridge regression to calculate $\hat{\theta}$ on our training set.
-We will then use that $\hat{\theta}$ to attempt to predict values of $\mathbf{y}$ for the test data.
-By choosing the $\hat{\theta}$ that minimizes MSE on the test data, but is trained on the training data, we are ensuring that we have not overfit our model to the data it is trained on.
-As one final step in the model selection process, we can check the performance of our final $\hat{\theta}$ on a validation set, to make sure that performance is still good (and we have not overfit for performance on our test set).
-
-We use this methodology to find the best $\lambda$ and $M$ given three datasets similar to the bishop data.
-We calculate $\lambda$ and $M$ using the regress\_train.txt, data we validate using regress\_test.txt, and then we test using regress\_validate.txt.
-We search through a grid of various values for $\lambda$ and $M$, and find that $MSE$ is lowest on the validation set when $\lambda = 1$ and $M = 1$: $MSE = 1.776$ on the validation data set.
-These parameters gave us an $MSE$ of 1.460 on the training data and and an $MSE$ of 1.415 on the test data.
-The fact that our $MSE$ is on the same order of magnitude for the test data is a good sign - it indicates that in making sure we didn't overfit to the training data, we didn't accidentally overfit to the validation data.
-
-The $MSE$ for various combinations of $\lambda$ and $M$ on the three datasets (including our chosen set of hyperparameters) is found below.
-We see that with large $M$ and small $\lambda$, we do great on the training data, but bad on our validation data.
-For large values of both $\lambda$ and $M$, we do well, but not as well as our optimal values we found through the grid search.
-
-Let's now use this methodology on a real dataset, as opposed to a toy dataset.
-The dataset that we will use was compiled by Kirsztian Buza at Budapest University of Technology and Economics.
-It includes 280 numerical values describing a number of blog posts, such as day of week, number of parent pages, and number of comments received within 24 hours.
-We are hoping to predict the number of comments it will receive in the next 24 hours.
-We are given this data split into training, validation, and test datasets, so we will once again perform a grid search to find the optimal value of $\lambda$.
-In this particular case, we are not regressing a polynomial of arbitrary order, so we do not need to test different values of $M$.
-
-\begin{center}
-\includegraphics[scale=.4]{MSE_comparison.png}
-\end{center}
-
-We find that for very large $\lambda$ (as large as $10^6$, $MSE$ continues to decrease, although it is always close to 900.
-The MSE for a uniform predictor that predicts the average of $y$, $\hat{y}$, has an MSE of 1453, so the performance of our model is still better.
-However, we see that for values of $\lambda$ higher than $10^6$ or so, the value of the MSE increases, and eventually approaches (and eclipses) the MSE of our uniform estimator.
-
-This isn't surprising - the blog feedback dataset is large, with 10,480 observations in the test set.
-If we interpret the value of $\lambda$ as strength of our prior, it makes sense that it does not have a substantial effect until it is significantly larger than the size of our data.
-Furthermore, the inability of our model to get very low MSE for low $\lambda$ speaks both to the fact that the data has a much higher dimension than our training data, and also suggests that the simple linear model we are using here may not be a very good fit for the data.
-
-\begin{center}
-\includegraphics[scale=.4]{BlogFeedbackRegressionMSE.png}
-\end{center}
-
-\subsubsection*{Problem 4}
+\subsubsection*{Support Vector Machines}
 
 \end{document}
+Status API Training Shop Blog About Pricing
+© 2015 GitHub, Inc. Terms Privacy Security Contact Help

### 2015-10-20 HW2/tex/HW2.tex
[hw2] starting the tex file, plots.

@@
 
+Let's now compare the performance of SVM to that of logistic regression for classification problems.
+To illustrate the objective and constraints of the support vector machine, we have included below the explicit objective we would optimize over, as well as the constraints, for the dual form of a linear SVM with slack variables.
+The equations below correspond to the 2D problem where we have positive examples (1, 2), (2, 2) and negative examples (0, 0), (-2, 3).
+
+\[
+\min_{\alpha_1, \alpha_2, \alpha_3, \alpha_4}
+\frac{1}{2}
+\begin{bmatrix}
+    \alpha_1 & \alpha_2 & \alpha_3 & \alpha_4 \\
+\end{bmatrix}
+\begin{bmatrix}
+    5       & 6 & 0 & -4 \\
+    6       & 8 & 0 & -2 \\
+    0       & 0 & 0 & 0 \\
+    -4       & -2 & 0 & 13 \\
+\end{bmatrix}
+\begin{bmatrix}
+    \alpha_1 \\
+    \alpha_2 \\
+    \alpha_3 \\
+    \alpha_4 \\
+\end{bmatrix} 
++
+\begin{bmatrix}
+    -1       & -1 & -1 & -1 \\
+\end{bmatrix}
+\begin{bmatrix}
+    \alpha_1 \\
+    \alpha_2 \\
+    \alpha_3 \\
+    \alpha_4 \\
+\end{bmatrix} 
+\]
+
+\begin{center}
+s.t.
+\end{center}
+
+\[
+\begin{bmatrix}
+    -1 & 0 & 0 & 0 \\
+    0 & -1 & 0 & 0 \\
+    0 & 0 & -1 & 0 \\
+    0 & 0 & 0 & -1 \\
+    1 & 0 & 0 &0 \\
+    0 & 1 & 0 & 0 \\
+    0 & 0 & 1 & 0 \\
+    0 & 0 & 0 & 1 \\
+\end{bmatrix}
+\begin{bmatrix}
+    \alpha_1 \\
+    \alpha_2 \\
+    \alpha_3 \\
+    \alpha_4 \\
+\end{bmatrix} 
+\leq
+\begin{bmatrix}
+0 \\
+0 \\ 
+0 \\ 
+0 \\ 
+C \\
+C \\ 
+C \\ 
+C \\
+\end{bmatrix},
+\]
+
+\[
+\begin{bmatrix}
+    1 & 1 & -1 & -1 \\
+\end{bmatrix}
+\begin{bmatrix}
+    \alpha_1 \\
+    \alpha_2 \\
+    \alpha_3 \\
+    \alpha_4 \\
+\end{bmatrix} 
+= 0
+\]
+
+\begin{table}
+\captionof{table}{Performance of Linear SVM on provided data sets}
+\begin{tabular}{llllll}
+\toprule
+{} & dataset & classification error rate (training set) & classification error rate (validation set) \\
+\midrule
+  & data\_stdev1 & 0.00\% & 0.00\% \\
+  & data\_stdev2 & 9.50\% & 7.50\% \\
+  & data\_stdev4 & 26.00\% & 23.50\% \\
+  & data\_nonsep & 49.50\% & 51.25\% \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev1_train.png}
+		\caption*{stdev1 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev2_train.png}
+		\caption*{stdev2 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev4_train.png}
+		\caption*{stdev4 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_nonsep_train.png}
+		\caption*{nonsep (Training)}
+	\end{minipage}
+		\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev1_validation.png}
+		\caption*{stdev1 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev2_validation.png}
+		\caption*{stdev2 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev4_validation.png}
+		\caption*{stdev4 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{linear_svm_nonsep_validation.png}
+		\caption*{nonsep (Validation)}
+	\end{minipage}
+	\caption{The decision boundaries generated by SVM plotted against the stdev1, stdev2, stdev4, and nonsep (training and validation) datasets}
+\end{figure}
+
 \end{document}

### 2015-10-20 HW2/code/problem2/problem_2/plotBoundary.py

@@ -18,7 +18,6 @@ def plotDecisionBoundary(X, Y, scoreFn, values, title = ""):
     X_tmp = array([[x1, x2] for x1, x2 in zip(ravel(xx), ravel(yy))])
     zz = array([scoreFn(X_tmp)])
     zz = zz.reshape(xx.shape)
-    print xx, yy, zz
     pl.figure()
     CS = pl.contour(xx, yy, zz, values, colors = 'green', linestyles = 'solid', linewidths = 2)
     pl.clabel(CS, fontsize=9, inline=1)

### 2015-10-20 HW2/code/problem2/problem_2/problem2.py

@@ -1,10 +1,11 @@
 import numpy as np
 from plotBoundary import *
 import cvxopt
+import sys
 # import your SVM training code
 
 # parameters
-name = 'nonsep'
+name = 'stdev4'
 print '======Training======'
 # load data from csv files
 train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
@@ -22,7 +23,7 @@ def gaussian_kernel(x, y, sigma=1):
     return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
 
 #X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
-#y = np.array([1., 1., -1., -1.])
+#Y = np.array([1., 1., -1., -1.])
 C = 1
 kernel = linear_kernel
 
@@ -83,7 +84,6 @@ def predictSVM(X):
 		weight = None
 
 	if weight != None:
-		print [X, np.sign(np.dot(X, weight) + theta_0)]
 		return np.sign(np.dot(X, weight) + theta_0)
 
 	else:
@@ -101,7 +101,11 @@ def predictSVM(X):
 
 # plot training results
 print '======Plot Training======'
-plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'SVM Train')
+plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'Linear SVM, stdev4 Training')
+y_predict = predictSVM(X)
+y_predict = np.reshape(y_predict, (len(y_predict), -1))
+correct = float(np.sum(Y == y_predict))/len(Y)
+print correct
 
 
 print '======Validation=======	'
@@ -110,5 +114,9 @@ validate = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_val
 X = validate[:, 0:2]
 Y = validate[:, 2:3]
 # plot validation results
-plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'SVM Validate')
+plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'Linear SVM, stdev4 Validation')
+y_predict = predictSVM(X)
+y_predict = np.reshape(y_predict, (len(y_predict), -1))
+correct = float(np.sum(Y == y_predict))/len(Y)
+print correct
 

### 2015-10-20 HW2/code/problem2/problem2.py
[hw2] rename some stuff, move some stuff.

@@ -5,7 +5,7 @@ import sys
 # import your SVM training code
 
 # parameters
-name = 'stdev4'
+name = 'nonsep'
 print '======Training======'
 # load data from csv files
 train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
@@ -98,6 +98,32 @@ def predictSVM(X):
 				s += a_0*sv_y_0*kernel(X[i], sv_0)
 			y_predict[i] = s
 		return np.sign(y_predict + theta_0)
+
+def geometricMarginSVM():
+	global alpha 
+	global sv 
+	global sv_y
+	global sv_bool 
+	global ind 
+	global K
+	global n_features
+
+	theta_0 = 0
+	for n in range(len(alpha)):
+		theta_0 += sv_y[n]
+		theta_0 -= np.sum(alpha * sv_y * K[ind[n],sv_bool])
+		theta_0 /= len(alpha)
+
+	weight = np.zeros(n_features)
+	for n in range(len(alpha)):
+		weight += alpha[n] + sv_y[n] + sv[n]
+
+	weights = np.append(weight, theta_0)
+	#print np.linalg.norm(weights)
+	return 1./np.linalg.norm(weights)
+
+
+print 'geometric margin', geometricMarginSVM()
 
 # plot training results
 print '======Plot Training======'

### 2015-10-20 HW2/tex/HW2.tex
Updated Tex File with problem 1 tables and plots

@@
 	\maketitle
+
+\subsubsection*{Logistic Regression}
+
+\begin{table}
+\captionof{table}{Performance of Logistic Regression on provided data sets with a decision boundary of 0.5}
+\begin{tabular}{llllll}
+\toprule
+{} & dataset & classification error (training set) & classification error (validation set) \\
+\midrule
+  & data\_stdev1 & 0.00\% & 0.00\% \\
+  & data\_stdev2 & 9.25\% & 8.00\% \\
+  & data\_stdev4 & 26.00\% & 24.75\% \\
+  & data\_nonsep & 48.50\% & 50.75\% \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_stdev1_train.png}
+		\caption*{stdev1 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_stdev2_train.png}
+		\caption*{stdev2 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_stdev4_train.png}
+		\caption*{stdev4 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_nonsep_train.png}
+		\caption*{nonsep (Training)}
+	\end{minipage}
+		\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_stdev1_validation.png}
+		\caption*{stdev1 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_stdev2_validation.png}
+		\caption*{stdev2 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_stdev4_validation.png}
+		\caption*{stdev4 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CEDB_nonsep_validation.png}
+		\caption*{nonsep (Validation)}
+	\end{minipage}
+	\caption{The classification error of Logistic Regression as a function of our choice of a decision boundary for the stdev1, stdev2, stdev4, and nonsep (training and validation) datasets}
+\end{figure}
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_train.png}
+		\caption*{stdev1 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_train.png}
+		\caption*{stdev2 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_train.png}
+		\caption*{stdev4 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_train.png}
+		\caption*{nonsep (Training)}
+	\end{minipage}
+		\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_validation.png}
+		\caption*{stdev1 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_validation.png}
+		\caption*{stdev2 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_validation.png}
+		\caption*{stdev4 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_validation.png}
+		\caption*{nonsep (Validation)}
+	\end{minipage}
+	\caption{The decision boundaries generated by Logistic Regression plotted against the stdev1, stdev2, stdev4, and nonsep (training and validation) datasets}
+\end{figure}
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev1.png}
+		\caption*{Classification Error - stdev1}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev2.png}
+		\caption*{Classification Error - stdev2}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev4.png}
+		\caption*{Classification Error - stdev4}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_nonsep.png}
+		\caption*{Classification Error - nonsep}
+	\end{minipage}
+		\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_stdev1.png}
+		\caption*{Logistic Loss - stdev1}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_stdev2.png}
+		\caption*{Logistic Loss - stdev2}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_stdev4.png}
+		\caption*{Logistic Loss - stdev4}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_nonsep.png}
+		\caption*{Logistic Loss - nonsep}
+	\end{minipage}
+	\caption{Classification Error (for a decision boundary of 0.5) and Logistic Loss of Logistic Regression as a function of the L2 penalty $\lambda$}
+\end{figure}
 

### 2015-10-23 HW2/tex/HW2.tex
[hw2] adding table of soft SVM performance with params.

@@
 
+\begin{table}
+\captionof{table}{Soft SVM Performance for Different Parameters and Kernels} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+{} &                        Kernel &      $C$ &   $\sigma$ &    Number of Support Vectors & $1/||\mathbf{w}||$ \\
+\midrule
+&    Gaussian &    .01 &    .01 &   400 &    .0286 \\
+&    Gaussian &    .01 &    .1 &   400 &    .0286 \\
+&    Gaussian &    .01 &    1 &   400 &    .0286 \\
+&    Gaussian &    .01 &    10 &   400 &    .0286 \\
+&    Gaussian &    .01 &    100 &   400 &    .0286 \\
+&    Gaussian &    .1 &    .01 &   400 &    .0141 \\
+&    Gaussian &    .1 &    .1 &   400 &    .0141 \\
+&    Gaussian &    .1 &    1 &   400 &    .0207 \\
+&    Gaussian &    .1 &    10 &   400 &    .0141 \\
+&    Gaussian &    .1 &    100 &   400 &    .0141 \\
+&    Gaussian &    1 &    .01 &   400 &    .0017 \\
+&    Gaussian &    1 &    .1 &   391 &    .0022 \\
+&    Gaussian &    1 &    1 &   100 &    .0076 \\
+&    Gaussian &    1 &    10 &   368 &    .0018 \\
+&    Gaussian &    1 &    100 &   400 &    .0017 \\
+&    Gaussian &    10 &    .01 &   400 &    .0017 \\
+&    Gaussian &    10 &    .1 &   390 &    .0018 \\
+&    Gaussian &    10 &    1 &   71 &    .0017 \\
+&    Gaussian &    10 &    10 &   209 &    .0003 \\
+&    Gaussian &    10 &    100 &   400 &    .0002 \\
+&    Gaussian &    100 &    .01 &   400 &    .0017 \\
+&    Gaussian &    100 &    .1 &   390 &    .0018 \\
+&    Gaussian &    100 &    1 &   59 &    .0002 \\
+&    Gaussian &    100 &    10 &   159 &    7.51e-5 \\
+&    Gaussian &    100 &    100 &   399 &    1.77e-5 \\
+&    Linear &    .01 &    N/A &   399 &    .0282 \\
+&    Linear &    .1 &    N/A &   393 &    .0133 \\
+&    Linear &    1 &    N/A &   392 &    .0017 \\
+&    Linear &    10 &    N/A &   393 &    .0002 \\
+&    Linear &    100 &    N/A &   397 &    1.8e-5 \\
+\midrule
+\bottomrule
+\end{tabular}
+\end{table}
+
 \end{document}

### 2015-10-23 HW2/code/problem2/problem2.py

@@ -19,13 +19,14 @@ def linear_kernel(x1, x2):
 def polynomial_kernel(x, y, p=3):
     return (1 + np.dot(x, y)) ** p
 
-def gaussian_kernel(x, y, sigma=1):
+def gaussian_kernel(x, y):
+    global sigma
     return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
 
 #X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
 #Y = np.array([1., 1., -1., -1.])
 C = 1
-kernel = linear_kernel
+kernel = gaussian_kernel
 
 
 # Carry out training, primal and/or dual
@@ -57,7 +58,7 @@ def trainSVM(X, y, kernel, C):
 
 	return n_features, K, alpha, sv, sv_y, sv_bool, ind
 
-n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
+#n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
 
 # Define the predictSVM(x) function, which uses trained parameters
 def predictSVM(X):
@@ -123,26 +124,36 @@ def geometricMarginSVM():
 	return 1./np.linalg.norm(weights)
 
 
-print 'geometric margin', geometricMarginSVM()
+#print 'geometric margin', geometricMarginSVM()
 
 # plot training results
-print '======Plot Training======'
-plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'Linear SVM, stdev4 Training')
-y_predict = predictSVM(X)
-y_predict = np.reshape(y_predict, (len(y_predict), -1))
-correct = float(np.sum(Y == y_predict))/len(Y)
-print correct
+#print '======Plot Training======'
+#plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'Linear SVM, stdev4 Training')
+#y_predict = predictSVM(X)
+#y_predict = np.reshape(y_predict, (len(y_predict), -1))
+#correct = float(np.sum(Y == y_predict))/len(Y)
+#print correct
+#
+#
+#print '======Validation=======	'
+## load data from csv files
+#validate = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_validate.csv')
+#X = validate[:, 0:2]
+#Y = validate[:, 2:3]
[…]

### 2015-10-23 HW2/code/problem3/problem2.py
Couldn't get SVM to work, kept getting a bug TypeError: buffer format not supported, all the code is there, maybe you can try to see if it works?

@@ -1,0 +1,128 @@
+import numpy as np
+from plotBoundary import *
+import cvxopt
+import sys
+# import your SVM training code
+
+# parameters
+name = 'nonsep'
+print '======Training======'
+# load data from csv files
+try:
+    train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
+except:
+    train = loadtxt('/Users/mfzhao/Downloads/hw2_resources/data/data_'+name+'_train.csv')
+    
+# use deep copy here to make cvxopt happy
+X = train[:, 0:2].copy()
+Y = train[:, 2:3].copy()
+
+def linear_kernel(x1, x2):
+    return np.dot(x1, x2)
+
+def polynomial_kernel(x, y, p=3):
+    return (1 + np.dot(x, y)) ** p
+
+def gaussian_kernel(x, y, sigma=1):
+    return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
+
+#X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
+#Y = np.array([1., 1., -1., -1.])
+C = 1
+kernel = linear_kernel
+
+
+# Carry out training, primal and/or dual
+def trainSVM(X, y, kernel, C):
+	n_samples, n_features = X.shape
+	K = np.zeros((n_samples, n_samples))
+	for i in range(n_samples):
+		for j in range(n_samples):
+			K[i, j] = kernel(X[i], X[j])
+
+	P = cvxopt.matrix(np.outer(y,y) * K)
+	q = cvxopt.matrix(np.ones(n_samples) * -1.)
+	A = cvxopt.matrix(y, (1, n_samples))
+	b = cvxopt.matrix(0.0)
+
+	G = cvxopt.matrix(np.vstack([np.diag(np.ones(n_samples) * -1.), np.diag(np.ones(n_samples))]))
+	h = cvxopt.matrix(np.hstack([np.zeros(n_samples), C*np.ones(n_samples)])) 
+ 
+	# Solve QP problem
+	solution = cvxopt.solvers.qp(P, q, G, h, A, b)
+	alpha = np.ravel(solution['x'])
+
+	sv_bool = alpha > .00001
+	ind = np.arange(len(alpha))[sv_bool]
+	alpha = alpha[sv_bool]
+	sv_y = y[sv_bool]
+	sv = X[sv_bool]
[…]

### 2015-10-23 HW2/code/problem3/titanic.py

@@ -1,0 +1,136 @@
+from numpy import *
+from plotBoundary import *
+import scipy.optimize
+import matplotlib as mpl
+from ggplot import *
+import pandas as pd
+import scipy.io
+import cvxopt
+from cvxopt import matrix
+from problem2 import *
+
+def scaling(X):
+    list = (4,5,6,7)
+    for i in list:
+        xmax = max(X[:,i])
+        xmin = min(X[:,i])
+        X[:,i] =  (X[:,i] - xmin)/(xmax - xmin)
+
+# Logistic Regression Loss Function
+def LRLoss(w, X, Y, l=0):
+    return sum(log(1+exp(-hstack(Y)*(dot(X,w[1:len(w)])+w[0]))))+l*dot(w[1:len(w)], w[1:len(w)])
+
+# Prediction Function
+def LRPredict(w, X):
+    return 1/(1+exp(-(dot(X,w[1:len(w)])+w[0])))
+    
+# Training Function
+def Train(X, Y, l=0):
+    w0 = random.randn(12)
+    w_opt = scipy.optimize.fmin(LRLoss, w0, args=(X,Y,l))
+    return w_opt
+
+# Classification Error Function
+def classifyErr(w, X, Y, db):
+    py = LRPredict(w,X)
+    yhat = ones(len(py))
+    for i in range(len(py)):
+        if py[i] < db:
+            yhat[i] = -1
+    return 1-mean(Y.flatten() == yhat)
+
+def loadTitan(name, typ):
+    file = 'data_'+name+'_'+typ+'.csv'
+    try:
+        data = scipy.io.loadmat('/Users/dholtz/Downloads/hw2_resources/data/'+file)['data']
+    except:
+        data = scipy.io.loadmat('/Users/mfzhao/Downloads/hw2_resources/data/'+file)['data']
+    X = data[:,0:11]
+    Y = data[:,11:12]
+    return array(X), array(Y)
+
+def GridLR(Xt, Yt, Xv, Yv, l, db=0.5):
+    l_opt = 0
+    bestLoss = 10000
+    w0 = random.randn(12)
+    w_opt = zeros(12)
+    vErr = []
+    for i in range(len(l)):
+        w = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(Xt,Yt,l[i]))
+        Loss = LRLoss(w, Xv, Yv, 0)/len(Yv)
+        vErr.append(Loss)
[…]

### 2015-10-23 HW2/code/problem2/problem2.py
Fixed some things

@@ -8,7 +8,10 @@ import sys
 name = 'nonsep'
 print '======Training======'
 # load data from csv files
-train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
+try:
+    train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
+except:
+    train = loadtxt('/Users/mfzhao/Downloads/hw2_resources/data/data_'+name+'_train.csv')
 # use deep copy here to make cvxopt happy
 X = train[:, 0:2].copy()
 Y = train[:, 2:3].copy()
@@ -19,13 +22,14 @@ def linear_kernel(x1, x2):
 def polynomial_kernel(x, y, p=3):
     return (1 + np.dot(x, y)) ** p
 
-def gaussian_kernel(x, y, sigma=1):
+def gaussian_kernel(x, y):
+    global sigma
     return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
 
 #X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
 #Y = np.array([1., 1., -1., -1.])
 C = 1
-kernel = linear_kernel
+kernel = gaussian_kernel
 
 
 # Carry out training, primal and/or dual
@@ -57,7 +61,7 @@ def trainSVM(X, y, kernel, C):
 
 	return n_features, K, alpha, sv, sv_y, sv_bool, ind
 
-n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
+#n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
 
 # Define the predictSVM(x) function, which uses trained parameters
 def predictSVM(X):
@@ -123,26 +127,36 @@ def geometricMarginSVM():
 	return 1./np.linalg.norm(weights)
 
 
-print 'geometric margin', geometricMarginSVM()
+#print 'geometric margin', geometricMarginSVM()
 
 # plot training results
-print '======Plot Training======'
-plotDecisionBoundary(X, Y, predictSVM, [-1, 0, 1], title = 'Linear SVM, stdev4 Training')
-y_predict = predictSVM(X)
-y_predict = np.reshape(y_predict, (len(y_predict), -1))
-correct = float(np.sum(Y == y_predict))/len(Y)
-print correct
[…]

### 2015-10-23 HW2/code/problem3/plotBoundary.py

@@ -1,0 +1,26 @@
+import pdb
+from numpy import *
+from matplotlib import pylab as pl
+
+# X is data matrix (each row is a data point)
+# Y is desired output (1 or -1)
+# scoreFn is a function of a data point
+# values is a list of values to plot
+
+def plotDecisionBoundary(w, X, Y, scoreFn, values, title = ''):
+    # Plot the decision boundary. For that, we will asign a score to
+    # each point in the mesh [x_min, m_max]x[y_min, y_max].
+    x_min, x_max = X[:, 0].min() - 1, X[:, 0].max() + 1
+    y_min, y_max = X[:, 1].min() - 1, X[:, 1].max() + 1
+    h = max((x_max-x_min)/200., (y_max-y_min)/200.)
+    xx, yy = meshgrid(arange(x_min, x_max, h),
+                      arange(y_min, y_max, h))
+    zz = array([scoreFn(w, x) for x in c_[xx.ravel(), yy.ravel()]])
+    zz = zz.reshape(xx.shape)
+    pl.figure()
+    CS = pl.contour(xx, yy, zz, values, colors = 'green', linestyles = 'solid', linewidths = 2)
+    pl.clabel(CS, fontsize=9, inline=1)
+    # Plot the training points
+    pl.scatter(X[:, 0], X[:, 1], c=(1.-Y), s=50, cmap = pl.cm.cool)
+    pl.title(title)
+    pl.axis('tight')

### 2015-10-23 HW2/code/problem3/problem2.py

@@ -2,6 +2,7 @@ import numpy as np
 from plotBoundary import *
 import cvxopt
 import sys
+from cvxopt import matrix
 # import your SVM training code
 
 # parameters
@@ -12,8 +13,11 @@ try:
     train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
 except:
     train = loadtxt('/Users/mfzhao/Downloads/hw2_resources/data/data_'+name+'_train.csv')
-    
-# use deep copy here to make cvxopt happy
+import numpy as np
+from plotBoundary import *
+import cvxopt
+import sys
+# import your SVM training code
 X = train[:, 0:2].copy()
 Y = train[:, 2:3].copy()
 
@@ -23,13 +27,15 @@ def linear_kernel(x1, x2):
 def polynomial_kernel(x, y, p=3):
     return (1 + np.dot(x, y)) ** p
 
-def gaussian_kernel(x, y, sigma=1):
+sigma =  1
+def gaussian_kernel(x, y):
+    global sigma
     return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
 
 #X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
 #Y = np.array([1., 1., -1., -1.])
 C = 1
-kernel = linear_kernel
+kernel = gaussian_kernel
 
 
 # Carry out training, primal and/or dual
@@ -61,7 +67,7 @@ def trainSVM(X, y, kernel, C):
 
 	return n_features, K, alpha, sv, sv_y, sv_bool, ind
 
-n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
+#n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
 
 # Define the predictSVM(x) function, which uses trained parameters
 def predictSVM(X):
@@ -79,7 +85,7 @@ def predictSVM(X):
 		theta_0 -= np.sum(alpha * sv_y * K[ind[n],sv_bool])
 		theta_0 /= len(alpha)
 
-	# Weight vector
+		# Weight vector
 	if kernel == linear_kernel:
 		weight = np.zeros(n_features)
 		for n in range(len(alpha)):
@@ -126,3 +132,26 @@ def geometricMarginSVM():
 	#print np.linalg.norm(weights)
 	return 1./np.linalg.norm(weights)
 
+
[…]

### 2015-10-23 HW2/code/problem3/titanic.py

@@ -1,5 +1,4 @@
 from numpy import *
-from plotBoundary import *
 import scipy.optimize
 import matplotlib as mpl
 from ggplot import *
@@ -10,8 +9,7 @@ from cvxopt import matrix
 from problem2 import *
 
 def scaling(X):
-    list = (4,5,6,7)
-    for i in list:
+    for i in (4,5,6,7):
         xmax = max(X[:,i])
         xmin = min(X[:,i])
         X[:,i] =  (X[:,i] - xmin)/(xmax - xmin)
@@ -26,8 +24,8 @@ def LRPredict(w, X):
     
 # Training Function
 def Train(X, Y, l=0):
-    w0 = random.randn(12)
-    w_opt = scipy.optimize.fmin(LRLoss, w0, args=(X,Y,l))
+    w0 = random.randn(13)
+    w_opt = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(X,Y,l))
     return w_opt
 
 # Classification Error Function
@@ -45,15 +43,18 @@ def loadTitan(name, typ):
         data = scipy.io.loadmat('/Users/dholtz/Downloads/hw2_resources/data/'+file)['data']
     except:
         data = scipy.io.loadmat('/Users/mfzhao/Downloads/hw2_resources/data/'+file)['data']
-    X = data[:,0:11]
+    temp = data[:,0:11]
+    scaling(temp)
     Y = data[:,11:12]
-    return array(X), array(Y)
+    X = ones([len(temp), 12])
+    X[:,1:12] = temp
+    return X.copy(), Y.copy()
 
 def GridLR(Xt, Yt, Xv, Yv, l, db=0.5):
     l_opt = 0
     bestLoss = 10000
-    w0 = random.randn(12)
-    w_opt = zeros(12)
+    w0 = random.randn(13)
+    w_opt = zeros(13)
     vErr = []
     for i in range(len(l)):
         w = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(Xt,Yt,l[i]))
@@ -86,22 +87,12 @@ def SVMGrid(Xtr, Ytr, Xv, Yv):
 Xtr, Ytr= loadTitan('titanic', 'train')
 Xv, Yv= loadTitan('titanic', 'validate')
 Xtest, Ytest = loadTitan('titanic', 'test')
-scaling(Xtr)
-scaling(Xv)
-scaling(Xtest)
 
[…]

### 2015-10-23 HW2/code/problem3/titanic.py
Made a seperate file for just the SVM stuff, still not working

@@ -4,12 +4,9 @@ import matplotlib as mpl
 from ggplot import *
 import pandas as pd
 import scipy.io
-import cvxopt
-from cvxopt import matrix
-from problem2 import *
 
 def scaling(X):
-    for i in (4,5,6,7):
+    for i in range(11):
         xmax = max(X[:,i])
         xmin = min(X[:,i])
         X[:,i] =  (X[:,i] - xmin)/(xmax - xmin)
@@ -24,7 +21,7 @@ def LRPredict(w, X):
     
 # Training Function
 def Train(X, Y, l=0):
-    w0 = random.randn(13)
+    w0 = random.randn(12)
     w_opt = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(X,Y,l))
     return w_opt
 
@@ -43,12 +40,10 @@ def loadTitan(name, typ):
         data = scipy.io.loadmat('/Users/dholtz/Downloads/hw2_resources/data/'+file)['data']
     except:
         data = scipy.io.loadmat('/Users/mfzhao/Downloads/hw2_resources/data/'+file)['data']
-    temp = data[:,0:11]
-    scaling(temp)
+    X = data[:,0:11]
+    scaling(X)
     Y = data[:,11:12]
-    X = ones([len(temp), 12])
-    X[:,1:12] = temp
-    return X.copy(), Y.copy()
+    return X, Y
 
 def GridLR(Xt, Yt, Xv, Yv, l, db=0.5):
     l_opt = 0
@@ -69,21 +64,6 @@ def GridLR(Xt, Yt, Xv, Yv, l, db=0.5):
 def SVMErr(yhat, y):
     return 1 - mean((yhat == y))
 
-def SVMGrid(Xtr, Ytr, Xv, Yv):
-    C = linspace(0,100,101)
-    bestErr = 1
-    kernel = gaussian_kernel
-    bestC = 0
-    for i in range(len(C)):
-        n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(Xtr, Ytr, kernel, C=C[i])
-        yhat = predictSVM(Xv)
-        cErr = SVMErr(yhat, Yv)
-        if cErr < bestErr:
-            bestErr = cErr
-            bestC = C[i]
-            n_featuresb, Kb, alphab, svb, sv_yb, sv_boolb, indb = n_features, K, alpha, sv, sv_y, sv_bool, ind
[…]

### 2015-10-23 HW2/code/problem3/titanicSVM.py

@@ -1,0 +1,171 @@
+import numpy as np
+from plotBoundary import *
+import cvxopt
+import sys
+from cvxopt import matrix
+import scipy.io
+
+def scaling(X):
+    for i in range(11):
+        xmax = max(X[:,i])
+        xmin = min(X[:,i])
+        X[:,i] =  (X[:,i] - xmin)/(xmax - xmin)
+
+def loadTitan(name, typ):
+    file = 'data_'+name+'_'+typ+'.csv'
+    try:
+        data = scipy.io.loadmat('/Users/dholtz/Downloads/hw2_resources/data/'+file)['data']
+    except:
+        data = scipy.io.loadmat('/Users/mfzhao/Downloads/hw2_resources/data/'+file)['data']
+    X = np.array(data[:,0:11]).copy()
+    scaling(X)
+    Y = np.array(data[:,11:12]).copy()
+    return X, Y
+
+def linear_kernel(x1, x2):
+    return np.dot(x1, x2)
+
+def polynomial_kernel(x, y, p=3):
+    return (1 + np.dot(x, y)) ** p
+
+def gaussian_kernel(x, y):
+    global sigma
+    return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
+
+# Carry out training, primal and/or dual
+def trainSVM(X, y, kernel, C):
+	n_samples, n_features = X.shape
+	K = np.zeros((n_samples, n_samples))
+	for i in range(n_samples):
+		for j in range(n_samples):
+			K[i, j] = kernel(X[i], X[j])
+
+	P = cvxopt.matrix(np.outer(y,y) * K)
+	q = cvxopt.matrix(np.ones(n_samples) * -1.)
+	A = cvxopt.matrix(y, (1, n_samples))
+	b = cvxopt.matrix(0.0)
+
+	G = cvxopt.matrix(np.vstack([np.diag(np.ones(n_samples) * -1.), np.diag(np.ones(n_samples))]))
+	h = cvxopt.matrix(np.hstack([np.zeros(n_samples), C*np.ones(n_samples)])) 
+ 
+	# Solve QP problem
+	solution = cvxopt.solvers.qp(P, q, G, h, A, b)
+	alpha = np.ravel(solution['x'])
+
+	sv_bool = alpha > .00001
+	ind = np.arange(len(alpha))[sv_bool]
+	alpha = alpha[sv_bool]
+	sv_y = y[sv_bool]
[…]

### 2015-10-26 HW2/code/problem1/lr_test.py
Updated Latex File

@@ -5,10 +5,6 @@ import matplotlib as mpl
 from ggplot import *
 import pandas as pd
 
-# Plotting Params
-mpl.rcParams["figure.figsize"] = "2.25, 2"
-pltsize = {"figure.figsize": "2.25, 2"}
-
 # Logistic Regression Loss Function
 def LRLoss(w, X, Y, l=0):
     return sum(log(1+exp(-hstack(Y)*(dot(X,w[1:len(w)])+w[0]))))+l*dot(w[1:len(w)], w[1:len(w)])
@@ -21,7 +17,7 @@ def LRPredict(w, X):
 def loadData(name, typ):
     try:
         data = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_'+typ+'.csv')
-    try:
+    except:
         data = loadtxt('/Users/mfzhao/Downloads/hw2_resources/data/data_'+name+'_'+typ+'.csv')
     X = data[:,0:2]
     Y = data[:,2:3]
@@ -69,7 +65,6 @@ def GridL(Xt, Yt, Xv, Yv, l, db=0.5):
 
 # Wrapper to do everything necessary for each dataset
 def wrapper(name):
-    global pltsize
     Xt, Yt=loadData(name, 'train')
     Xv, Yv=loadData(name, 'validate')
     w = Train(Xt, Yt, 0)
@@ -91,10 +86,22 @@ def wrapper(name):
     DF2 = pd.melt(DF2,id_vars=['Lambda'])
     title1 = 'Classification Error vs Lambda - ' + name
     title2 = 'Logisitic Loss vs Lambda - ' + name
-    print p1 = ggplot(DF1, aes(x='Lambda', y='value', color='variable')) + geom_line(size=4) + ggtitle('') + ylab('Error') + theme_matplotlib(rc=pltsize, matplotlib_defaults=False)
-    print p2 = ggplot(DF2, aes(x='Lambda', y='value', color='variable')) + geom_line(size=4) + ggtitle('') + ylab('Error') + theme_matplotlib(rc=pltsize, matplotlib_defaults=False)
-    
-wrapper('stdev1')
-wrapper('stdev2')
-wrapper('stdev4')
-wrapper('nonsep')
+    #print p1 = ggplot(DF1, aes(x='Lambda', y='value', color='variable')) + geom_line(size=4) + ggtitle('') + ylab('Error')
[…]

### 2015-10-26 HW2/code/problem3/titanicSVM.py
Problem 3 updated

@@ -20,6 +20,7 @@ def loadTitan(name, typ):
     X = np.array(data[:,0:11]).copy()
     scaling(X)
     Y = np.array(data[:,11:12]).copy()
+    Y = matrix(Y.tolist())
     return X, Y
 
 def linear_kernel(x1, x2):
@@ -53,9 +54,10 @@ def trainSVM(X, y, kernel, C):
 	alpha = np.ravel(solution['x'])
 
 	sv_bool = alpha > .00001
+	sv_bool = np.array(sv_bool)
 	ind = np.arange(len(alpha))[sv_bool]
 	alpha = alpha[sv_bool]
-	sv_y = y[sv_bool]
+	sv_y = np.ravel(y)[sv_bool]
 	sv = X[sv_bool]
 	print "%d support vectors out of %d points" % (len(alpha), n_samples)
 
@@ -69,24 +71,20 @@ def predictSVM(X):
 	global ind 
 	global K
 	global n_features
-
+	global kernel
+	
 	theta_0 = 0
 	for n in range(len(alpha)):
 		theta_0 += sv_y[n]
 		theta_0 -= np.sum(alpha * sv_y * K[ind[n],sv_bool])
 		theta_0 /= len(alpha)
 
-		# Weight vector
+	# Weight vector
 	if kernel == linear_kernel:
 		weight = np.zeros(n_features)
 		for n in range(len(alpha)):
 			weight += alpha[n] * sv_y[n] * sv[n]
-	else:
-		weight = None
-
-	if weight != None:
 		return np.sign(np.dot(X, weight) + theta_0)
-
 	else:
 		if len(np.array(X).shape) == 1:
 			n_points = 1 
@@ -127,34 +125,86 @@ def SVMErr(yhat, y):
     return 1 - mean((yhat == y))
 
 def SVMGrid(Xtr, Ytr, Xv, Yv):
-    C = linspace(0,100,101)
-    bestErr = 1
-    kernel = gaussian_kernel
+    global alpha
+    global sv 
+    global sv_y
+    global sv_bool 
+    global ind 
+    global K
+    global n_features
+    global kernel
+    global sigma
+
+    C = linspace(0,1,101)
+    bestErr = 10
     bestC = 0
-    for i in range(len(C)):
-        n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(Xtr, Ytr, kernel, C=C[i])
-        yhat = predictSVM(Xv)
-        cErr = SVMErr(yhat, Yv)
-        if cErr < bestErr:
[…]

### 2015-10-26 HW2/tex/HW2.tex
Latex updated

--- 2015-10-20 HW2/tex/HW2.tex
@@
 \subsubsection*{Logistic Regression}
+Here we want to predict the class of a discrete (or categorical) variable $Y$ given some some features $X$.
+While we can formulate this classification problem in the context of the linear regression framework we're already familiar with (these are known as linear probability models), these classical linear models are generally unbounded which may not be the best choice for modeling probabilities.
+Rather, we consider a slightly modified version of linear regression where we take the standard linear prediction $w_o + \sum_i w_i x_i$ and transform it such that it is bounded on $[0,1]$.
+In our case, our choice of a transformation function is the sigmoid function:
+\begin{equation*}
+	\sigma(z) = \frac{1}{1+e^{z}}
+\end{equation*}
 
-\begin{table}
-\captionof{table}{Performance of Logistic Regression on provided data sets with a decision boundary of 0.5}
-\begin{tabular}{llllll}
+This transformed model is known as logistic or logit regression. Using this formulation, our the negative log-likelihood becomes:
+\begin{equation*}
+	\textnormal{NLL}=\sum_i \log \left(1+e^{-y^{(i)}(w_0 + x^{(i)}\cdot w)} \right)
+\end{equation*}
+We think that overfitting might be problem in logistic regression especially in the case of linearly separable data which will drive our weights $w$ to very large values to minimize the the logistic loss.
+Hence, we add a ridge penalty to the weights on the features to shrink them:
+\begin{equation*}
+	\min_w \sum_i \log \left(1+e^{-y^{(i)}(w_0 + x^{(i)}\cdot w)} \right) + \lambda w^T w
+\end{equation*}
+Using this, we trained Logistic Regressions for each of our 4 datasets using gradient descent. The optimal weights are reported below:
+
+\begin{table}[ht]
+\centering
+\captionof{table}{Optimal Weights Trained on Various Datasets}
+\begin{tabular}{lrrrr}
 \toprule
-{} & dataset & classification error (training set) & classification error (validation set) \\
+{}    &      stdev1 &    stdev2 &    stdev4 & nonsep\\
 \midrule
-  & data\_stdev1 & 0.00\% & 0.00\% \\
-  & data\_stdev2 & 9.25\% & 8.00\% \\
-  & data\_stdev4 & 26.00\% & 24.75\% \\
-  & data\_nonsep & 48.50\% & 50.75\% \\
+$w_0$ &  -63.930971 & -0.009267 & -0.046649 & 0.000597\\
+$w_1$ &  266.198033 &  0.236279 &  0.763589 & -0.024737\\
+$w_2$ &  160.999118 &  0.203428 &  1.114811 & -0.023726\\
 \bottomrule
@@
 \end{table}
+
+Logistic regression produces a predicted probability of a class rather than an actual classification.
+This means that we need to decide a decision boundary to minimizes our classifcation error.
+Considering we have symmetric loss, we intuitively might want our decision boundary to be 0.5 which produces the following error rates:
+
+\begin{table}[ht]
+\centering
+\captionof{table}{Performance of Logistic Regression with Decision boundary of 0.5}
+\begin{tabular}{lrrrr}
+\toprule
+{} & stdev1 & stdev2 & stdev4 & nonsep\\
+\midrule
+  Training Error   & 0.00\%  & 9.25\% &  26.00\% & 48.50\% \\
+  Validation Error & 0.00\%  & 8.00\% &  24.75\% & 50.75\% \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+These results make a lot of sense as we should expect the classification error to be 0 for the perfectly linearly separable case and get progressively worse as as the two classes overlap.
+In fact, in the nonseparable case, our classifier is no better (or even slightly worse) than random guessing!
+We can see this as we plot our decision boundary in X-space below:
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_train.png}
+		\caption*{stdev1 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_train.png}
+		\caption*{stdev2 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_train.png}
+		\caption*{stdev4 (Training)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_train.png}
+		\caption*{nonsep (Training)}
+	\end{minipage}
+		\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_validation.png}
+		\caption*{stdev1 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_validation.png}
+		\caption*{stdev2 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_validation.png}
+		\caption*{stdev4 (Validation)}
+	\end{minipage}
+	\begin{minipage}[b]{.24\linewidth}
+		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_validation.png}
+		\caption*{nonsep (Validation)}
+	\end{minipage}
+	\caption{The decision boundaries generated by Logistic Regression plotted against the stdev1, stdev2, stdev4, and nonsep (training and validation) datasets}
+\end{figure}
+
+We are also interested in examining classification error as we vary our choice of the decision boundary. This is depicted in in figure 2 below:
 
@@
 
-\begin{figure}[ht]
-	\centering
-	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_train.png}
-		\caption*{stdev1 (Training)}
-	\end{minipage}
-	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_train.png}
-		\caption*{stdev2 (Training)}
-	\end{minipage}
-	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_train.png}
-		\caption*{stdev4 (Training)}
-	\end{minipage}
-	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_train.png}
-		\caption*{nonsep (Training)}
-	\end{minipage}
-		\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_validation.png}
-		\caption*{stdev1 (Validation)}
-	\end{minipage}
-	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_validation.png}
-		\caption*{stdev2 (Validation)}
-	\end{minipage}
-	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_validation.png}
-		\caption*{stdev4 (Validation)}
-	\end{minipage}
-	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_validation.png}
-		\caption*{nonsep (Validation)}
-	\end{minipage}
-	\caption{The decision boundaries generated by Logistic Regression plotted against the stdev1, stdev2, stdev4, and nonsep (training and validation) datasets}
-\end{figure}
+In the linearly separable case (stdev1), we see that just about any value greater than 0 gives us perfect classification, which makes intuitive sense since the optimal weights will try to push everything to either 0 or 1 resulting in the graph we see: the error starts out at .5 and almost immediately drops to 0.
+However if the data is still generally but not perfectly linearly separable (stdev2 and stdev4), we get what we would expect--classification error starts high and decreases until around decision boundary of 0.5 and then starts to increase again.
+In the linearly non-seperable case (nonsep), the probabilities are all relatively close to 0.5.
+Hence we don't see any effect of changing the decision boundary until we get relatively close to 0.5 in this case, we see that we get a spike since we start to misclassify the green points while all the purple points are still incorrectly classified.
+The spike drops back to 0.5 at around decision boundary of 0.5 and decreases as we start to correctly classify all the purple points.
+As we continue to increase the decision boundary, we get to the case were we classify everything as purple leading us with a misclassification rate of 0.5.
+
+We also explore the effect of adding a ridge penalty to logistic regression. We plot the classification error and loss of our regression as a function of the penalty $lambda$:
 
@@
 		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev1.png}
-		\caption*{Classification Error - stdev1}
+		\caption*{Error - stdev1}
 	\end{minipage}
@@
 		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev2.png}
-		\caption*{Classification Error - stdev2}
+		\caption*{Error - stdev2}
 	\end{minipage}
@@
 		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev4.png}
-		\caption*{Classification Error - stdev4}
+		\caption*{Error - stdev4}
 	\end{minipage}
@@
 		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_nonsep.png}
-		\caption*{Classification Error - nonsep}
+		\caption*{Error - nonsep}
 	\end{minipage}
@@
 	\end{minipage}
-	\caption{Classification Error (for a decision boundary of 0.5) and Logistic Loss of Logistic Regression as a function of the L2 penalty $\lambda$}
+	\caption{Classification Error (for a decision boundary of 0.5) and Loss of Logistic Regression as a function of the L2 penalty $\lambda$}
 \end{figure}
+
+We see that generally speaking, the classification error on the validation set decreases (except in the linearly separable case, where it stays at 0) as we increase our ridge penalty which is something we might expect if we think there's possible overfitting.
+We also see that generally speaking, loss monotonically increases with lambda, something we should expect since the ridge penalty is pushing the weights toward 0 causing logistic regression to output probabilities closer to 0.5.
 
@@
 
+\subsubsection*{Titanic Data}
+We're interesting in now testing the performance of our classifiers on a real dataset.
+In this case, we are using the Titanic dataset where we're doing the rather macabre task of trying to predict passenger survival given his or her features.
+So this problem we first scale all the features so they lie between $[0,1]$.
+WE accomplish by taking each observation and transforming it via:
+\begin{equation*}
+	x_{i,j}^* = \frac{x_{i,j}-\min(x_j)}{\max(x_j) - \min(x_j)}
+\end{equation*}
+We scale our data in such a way so our logistic regression behaves nicely. Otherwise, we encounter precision loss errors that end up significantly impacting our predictive accuracy.
+
+Using this transformed data, we train a logistic regression, a linear SVM, a polynomial SVM ($K(x,z) = (1+x\cdot z)^3$), and a gaussian SVM.
+For each classifier, we perform a grid search to choose the optimal hyperparameters that minimize the classification error of the validation set.
+We find that for:
+\begin{itemize}
+	\item Logistic Regression, $\lambda = 0.38$
+	\item Linear SVM, $C = 0.07$
+	\item Polynomial SVM, $C=0.03$
+	\item Gaussian SVM, $\sigma = 1$ and $C=0.04$
+\end{itemize}
+
+With these settings we obtain the following results:
+\begin{table}[ht]
+\centering
+\captionof{table}{Titanic Data Classification Performance}
+\begin{tabular}{lrrrr}
+\toprule
+{} & Logit & L-SVM & P-SVM & G-SVM \\
+\midrule
+Penalty & $\lambda = 0.38$ & $C = 0.07$ & $C = 0.03$ & $C = 0.04$\\
+\midrule
+Training Error    &  0.165 &  0.225 & 0.140 &  0.195 \\
+Validation Error  &  0.208 &  0.208 & 0.222 &  0.210 \\
+Test Error        &  0.249 &  0.233 & 0.265 &  0.243 \\
+Geometric Margin  & {}     &  0.0055& 0.0112& 0.0063 \\
+\bottomrule
+\end{tabular}	
+\end{table}
+
+\begin{table}[ht]
+\centering
+\captionof{table}{Optimal Weight Table}
+\begin{tabular}{lrrrr}
+\toprule
+{} & Logit & L-SVM & P-SVM & G-SVM \\
+\midrule
+Penalty & $\lambda = 0.38$ & $C = 0.07$ & $C = 0.03$ & $C = 0.04$\\
+\midrule
+$w_1$  &  0.603693 &  0.031302 & -0.012892 &  0.120059 \\
+$w_2$  &  0.455286 &  0.033952 & -0.010501 &  0.255103 \\
+$w_3$  & -1.058978 & -0.065245 &  0.023511 & -0.375139 \\
+$w_4$  &  2.475588 &  1.901134 &  0.101932 &  1.440000 \\
+$w_5$  & -0.914359 & -0.078502 & -0.009049 & -0.068269 \\
+$w_6$  & -0.251056 & -0.011503 &  0.000163 & -0.036648 \\
+$w_7$  &  0.538155 &  0.131051 &  0.010683 &  0.104903 \\
+$w_8$  &  0.330644 &  0.057787 & -0.001549 &  0.117515 \\
+$w_9$  & -0.233114 & -0.038255 & -0.009864 & -0.159978 \\
+$w_{10}$ &  0.291611 &  0.029324 &  0.024844 &  0.240000 \\
+$w_{11}$ & -0.058496 &  0.008941 & -0.014863 & -0.080000 \\
+\bottomrule
+\end{tabular}	
+\end{table}
+
 \end{document}
-Status API Training Shop Blog About Pricing
-© 2015 GitHub, Inc. Terms Privacy Security Contact Help

### 2015-10-26 HW2/tex/HW2.tex
[hw2] getting up to date.

--- 2015-10-23 HW2/tex/HW2.tex
@@
 \captionof{table}{Performance of Linear SVM on provided data sets}
-\begin{tabular}{llllll}
+\begin{tabular}{c c|c|c|c|c|c}
 \toprule
-{} & dataset & classification error rate (training set) & classification error rate (validation set) \\
+{} & dataset & $w_0$ & $w_1$ & $w_2$ & classification error rate (training) & classification error rate (validation) \\
 \midrule
-  & data\_stdev1 & 0.00\% & 0.00\% \\
-  & data\_stdev2 & 9.50\% & 7.50\% \\
-  & data\_stdev4 & 26.00\% & 23.50\% \\
-  & data\_nonsep & 49.50\% & 51.25\% \\
+  & data\_stdev1 & 0.31 & 5.42 & 0.87 & 0.00\% & 0.00\% \\
+  & data\_stdev2 & 9.09e-2 & 9.83e1 & 9.11e1 & 9.50\% & 7.50\% \\
+  & data\_stdev4 & -2.05e-2 & 2.58e2 & 2.40e2 & 26.00\% & 23.50\% \\
+  & data\_nonsep & 2.56e-3 & 4.23e2 & 3.84e2 & 49.50\% & 51.25\% \\
 \bottomrule

### 2015-10-26 HW2/code/problem2/problem2.py

@@ -5,7 +5,7 @@ import sys
 # import your SVM training code
 
 # parameters
-name = 'nonsep'
+name = 'stdev1'
 print '======Training======'
 # load data from csv files
 try:
@@ -29,13 +29,11 @@ def gaussian_kernel(x, y):
 #X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
 #Y = np.array([1., 1., -1., -1.])
 C = 1
-kernel = gaussian_kernel
+kernel = linear_kernel
 
 
 # Carry out training, primal and/or dual
 def trainSVM(X, y, kernel, C):
-	print y.shape 
-	print X.shape
 	n_samples, n_features = X.shape
 	K = np.zeros((n_samples, n_samples))
 	for i in range(n_samples):
@@ -44,11 +42,6 @@ def trainSVM(X, y, kernel, C):
 
 	P = cvxopt.matrix(np.outer(y,y) * K)
 	q = cvxopt.matrix(np.ones(n_samples) * -1.)
-	print y.shape 
-	print y.dtype
-	print (1, n_samples)
-	print y
-	print 
 	A = cvxopt.matrix(np.array(y), (1, n_samples))
 	b = cvxopt.matrix(0.0)
 
@@ -68,7 +61,7 @@ def trainSVM(X, y, kernel, C):
 
 	return n_features, K, alpha, sv, sv_y, sv_bool, ind
 
-#n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
+n_features, K, alpha, sv, sv_y, sv_bool, ind = trainSVM(X, Y, kernel, C=C)
 
 # Define the predictSVM(x) function, which uses trained parameters
 def predictSVM(X):
@@ -133,38 +126,70 @@ def geometricMarginSVM():
 	#print np.linalg.norm(weights)
 	return 1./np.linalg.norm(weights)
 
+def weightsSVM():
+	global alpha 
+	global sv 
+	global sv_y
+	global sv_bool 
+	global ind 
+	global K
+	global n_features
+
+	theta_0 = 0
+	for n in range(len(alpha)):
+		theta_0 += sv_y[n]
+		theta_0 -= np.sum(alpha * sv_y * K[ind[n],sv_bool])
+		theta_0 /= len(alpha)
+
+	weight = np.zeros(n_features)
+	for n in range(len(alpha)):
+		weight += alpha[n] + sv_y[n] + sv[n]
+
+	weights = np.append(weight, theta_0)
[…]

### 2015-10-26 HW2/code/problem3/titanic_weight_comparison.R

@@ -1,0 +1,41 @@
+feature_weights = data.frame(`Model` = vector(mode='character', length=44))
+
+feature_weights[['Model']] = c(rep('Logistic Regression', 11), 
+                               rep('Linear SVM', 11), 
+                               rep('Polynomial SVM', 11), 
+                               rep('RBF SVM', 11))
+
+feature_names = c('Passenger Class = 1st', 'Passenger Class = 2nd', 
+                  'Passenger Class = 3rd', 'Gender', 'Age', '# Siblings/Spouses', 
+                  '# Parents/Children', 'Passenger Fare', 'Embarked at Southhampton', 
+                  'Embarked at Cherbourg', 'Embarked at Queenstown')
+feature_weights[['Feature Name']] = rep(feature_names, 4)
+
+weights = c(0.6036931,0.45528582,-1.05897834,2.47558843,-0.91435904,-0.251056,
+            0.53815509,0.33064367,-0.2331139,0.29161072,-0.05849566, 0.0313021,
+            0.0339519,-0.06524453,1.90113447,-0.07850197,-0.01150343,0.13105129,
+            0.0577874,-0.03825494,0.02932354,0.00894087, -0.01289241,-0.01050105,
+            0.02351099,0.10193199,-0.00904929,0.00016298,0.01068262,-0.00154886,
+            -0.00986365,0.0248439,-0.01486273, 0.12005924,0.25510333,-0.37513945,
+            1.44000021,-0.06826854,-0.03664755,0.1049032,0.11751473,-0.15997783,
+            0.24000049,-0.07999953)
+
+feature_weights[['Feature Weight']] = weights
+
+feature_weights %>%
+  group_by(`Model`) %>%
+  mutate(max_weight = max(abs(`Feature Weight`))) %>%
+  ungroup() %>% 
+  mutate(`Feature Weight (Normalized)` = `Feature Weight`/max_weight) %>%
+  ggplot(., aes(x=as.factor(`Feature Name`), y=`Feature Weight (Normalized)`, fill=`Model`)) + 
+  geom_bar(stat="identity", position="dodge") + xlab('Feature') + 
+  ggtitle('Feature Weights (Normalized by Model)') + theme(axis.text.x = element_text(angle = 90, hjust = 1)) -> p
+ggsave(p, file='~/Desktop/mit_6867/HW2/tex/titanic_weight_comparison.png')
+
+feature_weights %>%
[…]

### 2015-10-26 HW2/tex/HW2.tex

--- 2015-10-26 HW2/tex/HW2.tex
@@
 \end{figure}
+
+\begin{table}
+\captionof{table}{Soft SVM Performance for Different Parameters and Kernels} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+{} &                        Kernel &      $C$ &   $\sigma$ &    Number of Support Vectors & $1/||\mathbf{w}||$ \\
+\midrule
+&    Gaussian &    .01 &    .1 &   400 &    .0286 \\
+&    Gaussian &    .01 &    1 &   400 &    .0286 \\
+&    Gaussian &    .01 &    10 &   400 &    .0286 \\
+&    Gaussian &    .1 &    .1 &   400 &    .0141 \\
+&    Gaussian &    .1 &    1 &   400 &    .0207 \\
+&    Gaussian &    .1 &    10 &   400 &    .0141 \\
+&    Gaussian &    1 &    .1 &   391 &    .0022 \\
+&    Gaussian &    1 &    1 &   100 &    .0076 \\
+&    Gaussian &    1 &    10 &   368 &    .0018 \\
+&    Gaussian &    10 &    .1 &   390 &    .0018 \\
+&    Gaussian &    10 &    1 &   71 &    .0017 \\
+&    Gaussian &    10 &    10 &   209 &    .0003 \\
+&    Gaussian &    100 &    .1 &   390 &    .0018 \\
+&    Gaussian &    100 &    1 &   59 &    .0002 \\
+&    Gaussian &    100 &    10 &   159 &    7.51e-5 \\
+&    Linear &    .01 &    N/A &   399 &    .0282 \\
+&    Linear &    .1 &    N/A &   393 &    .0133 \\
+&    Linear &    1 &    N/A &   392 &    .0017 \\
+&    Linear &    10 &    N/A &   393 &    .0002 \\
+&    Linear &    100 &    N/A &   397 &    1.8e-5 \\
+\midrule
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[height=3in]{titanic_weight_comparison.png}
+	\caption*{Comparison of Feature Weights Across Models (Normalized per Model)}
+\end{figure}
+
 \end{document}

### 2015-10-27 HW2/tex/HW2.tex
[hw2] first draft of finished writeup.

@@
 We think that overfitting might be problem in logistic regression especially in the case of linearly separable data which will drive our weights $w$ to very large values to minimize the the logistic loss.
-Hence, we add a ridge penalty to the weights on the features to shrink them:
+Hence, we add a ridge penalty to the weights:
 \begin{equation*}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_stdev1_train.png}
 		\caption*{stdev1 (Training)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_stdev2_train.png}
 		\caption*{stdev2 (Training)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_stdev4_train.png}
 		\caption*{stdev4 (Training)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_nonsep_train.png}
 		\caption*{nonsep (Training)}
@@
 		\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev1_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_stdev1_validation.png}
 		\caption*{stdev1 (Validation)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev2_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_stdev2_validation.png}
 		\caption*{stdev2 (Validation)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_stdev4_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_stdev4_validation.png}
 		\caption*{stdev4 (Validation)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{LR_nonsep_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{LR_nonsep_validation.png}
 		\caption*{nonsep (Validation)}
@@
 
-We also explore the effect of adding a ridge penalty to logistic regression. We plot the classification error and loss of our regression as a function of the penalty $lambda$:
+We also explore the effect of adding a ridge penalty to logistic regression. We plot the classification error and loss of our regression as a function of the penalty $\lambda$:
 
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev1.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{CErr_lambda_stdev1.png}
 		\caption*{Error - stdev1}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev2.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{CErr_lambda_stdev2.png}
 		\caption*{Error - stdev2}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_stdev4.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{CErr_lambda_stdev4.png}
 		\caption*{Error - stdev4}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{CErr_lambda_nonsep.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{CErr_lambda_nonsep.png}
 		\caption*{Error - nonsep}
@@
 		\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_stdev1.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{Loss_lambda_stdev1.png}
 		\caption*{Logistic Loss - stdev1}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_stdev2.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{Loss_lambda_stdev2.png}
 		\caption*{Logistic Loss - stdev2}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_stdev4.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{Loss_lambda_stdev4.png}
 		\caption*{Logistic Loss - stdev4}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{Loss_lambda_nonsep.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{Loss_lambda_nonsep.png}
 		\caption*{Logistic Loss - nonsep}
@@
 
+Table 3 shows the weights generated by our soft SVM algorithm, as well as its performance on the training and validation sets, given the four datasets we've been provided for testing.
+The soft SVM, in this case, uses a simple linear kernel, so the decision boundaries that are generated are still straight lines in $\mathbb{R}^2$.
+However, the slack provided by the soft SVM allows the algorithm to classify data and create a ``margin", even when the data is not linearly separable.
+As expected, in general the algorithm does slightly better on the training set than the validation set.
+Also, however, note that as the data becomes less and less linearly separable, our soft SVM with a linear kernel does increasingly poorly.
+This motivates our decision to explore the use of a Gaussian (or RBF) kernel.
+
+Table 4 includes the performance of our soft SVM algorithm on various training sets when we change kernels (we provide data on both linear and Gaussian kernels), and also when we modify the bandwidth of our Gaussian kernel (where applicable) and change our regularization parameter, $C$.
+In general, our geometric margin increases as we decrease $C$.
+This is to be expected - as we provide the SVM with more slack, it is able to allow more points within its margin, and can consequently expand its geometric margin.
+In general, the larger the geometric margin, the more support vectors - again, this is intuitive.
+The more slack we allow, the more data points we will observe on or inside of our margin.
+It is not, however, necessarily true that the geometric margin will continue to increase as we decrease $C$ (and also, geometric margin will not always decrease as we increase $C$).
+As C gets very large, our decision boundary will approach the hard SVM decision boundary, and will not move at all.
+As $C$ gets increasingly small, the additional slack will lead our decision boundary to encompass all of the data, at which point our geometric margin will stop increasing.
+
+Choosing $C$ by maximizing the geometric margin on the training data not a mistake that a practitioner should make.
+This is because maximizing the geometric margin on the training data is, in a sense, overfitting on the training data.
+Smaller $C$ would almost always give us larger geometric margins, and the practitioner would feel compelled to choose a very small $C$.
+However, this choice of $C$ may lead to very poor performance on validation sets, or future data.
+The best method for selecting $C$ would be to perform some sort of cross-validation.
+In the most basic sense, we might perform a grid search with various values of C, train our soft SVM, and see which value of C provides the best performance on a held-out validation set.
+
 \begin{table}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev1_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_stdev1_train.png}
 		\caption*{stdev1 (Training)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev2_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_stdev2_train.png}
 		\caption*{stdev2 (Training)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev4_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_stdev4_train.png}
 		\caption*{stdev4 (Training)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_nonsep_train.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_nonsep_train.png}
 		\caption*{nonsep (Training)}
@@
 		\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev1_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_stdev1_validation.png}
 		\caption*{stdev1 (Validation)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev2_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_stdev2_validation.png}
 		\caption*{stdev2 (Validation)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_stdev4_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_stdev4_validation.png}
 		\caption*{stdev4 (Validation)}
@@
 	\begin{minipage}[b]{.24\linewidth}
-		\includegraphics[width=1\linewidth, height=1in]{linear_svm_nonsep_validation.png}
+		\centering
+		\includegraphics[width=.5\linewidth, height=.5in]{linear_svm_nonsep_validation.png}
 		\caption*{nonsep (Validation)}
@@
 We're interesting in now testing the performance of our classifiers on a real dataset.
-In this case, we are using the Titanic dataset where we're doing the rather macabre task of trying to predict passenger survival given his or her features.
-So this problem we first scale all the features so they lie between $[0,1]$.
-WE accomplish by taking each observation and transforming it via:
+In this case, we are using the Titanic dataset, where we're performing the rather macabre task of predicting passenger survival given his or her features.
+For this problem we first scale all the features so they lie between $[0,1]$.
+We accomplish this by taking each observation and transforming it via:
 \begin{equation*}
@@
 \end{equation*}
-We scale our data in such a way so our logistic regression behaves nicely. Otherwise, we encounter precision loss errors that end up significantly impacting our predictive accuracy.
+This scaling of our data ensures that logistic regression behaves nicely. Otherwise, we encounter precision loss errors that end up significantly impacting our predictive accuracy.
 
@@
 For each classifier, we perform a grid search to choose the optimal hyperparameters that minimize the classification error of the validation set.
-We find that for:
-\begin{itemize}
-	\item Logistic Regression, $\lambda = 0.38$
-	\item Linear SVM, $C = 0.07$
-	\item Polynomial SVM, $C=0.03$
-	\item Gaussian SVM, $\sigma = 1$ and $C=0.04$
-\end{itemize}
+We find that for logistic regression,  $\lambda = 0.38$.
+For linear SVM, $C=0.07$.
+For polynomial SVM, $C=0.03$.
+And for Gaussian SVM, $\sigma = 1$ and $C=0.04$.
 
-With these settings we obtain the following results:
+The results generated by these settings, as well as the weights generated by these settings, are found in Tables 5 and 6.
+Note that we omit our bias term, $w_0$, although it is included in the model, since it does not have any immediate physical interpretation (other than perhaps the baseline rate at which people survived the sinking of the Titanic).
+
+As can be seen in Table 5, logistic regression outperforms polynomial SVM, with a test error rate of 24.9\% (as opposed to 26.5\% the SVM).
+Gaussian SVM performs outperforms the logistic regression, with a test error rate of 24.3\%, and the linear SVM performs the best, with a test error rate of 23.3\%.
+This is an interesting, and somewhat surprising result, as we might expect Gaussian SVM to perform as well (if not better) than linear SVM.
+
+To better understand the differences in our models, one can look at Figure 5, which compares the weights of different variables for our different models.
+Note that each model's array of weights is normalized so that its maximum weight (in this case, always gender) is equal to 1.
+One can see that the linear SVM places almost all of its weight on gender, with a small amount on \# of parents/children.
+Logistic regression also places the most weight on gender, but also puts significant weight on (in order of decreasing magnitude) 3rd class passengers, age, \# of parents and children, 1st class passengers, and 2nd class passengers.
+The polynomial SVM puts most weight on passenger class and passenger embarkation locations.
+And the Gaussian SVM, after age, places the most weight on 3rd class passengers, 2nd class passengers, and passengers who embarked in Cherbourg.
+
 \begin{table}[ht]
@@
 	\includegraphics[height=3in]{titanic_weight_comparison.png}
-	\caption*{Comparison of Feature Weights Across Models (Normalized per Model)}
+	\caption*{Figure 5: Comparison of Feature Weights Across Models (Normalized per Model)}
 \end{figure}

### 2015-11-05 HW2/tex/HW2.tex
Final Draft

@@
 As expected, in general the algorithm does slightly better on the training set than the validation set.
-Also, however, note that as the data becomes less and less linearly separable, our soft SVM with a linear kernel does increasingly poorly.
-This motivates our decision to explore the use of a Gaussian (or RBF) kernel.
-
-Table 4 includes the performance of our soft SVM algorithm on various training sets when we change kernels (we provide data on both linear and Gaussian kernels), and also when we modify the bandwidth of our Gaussian kernel (where applicable) and change our regularization parameter, $C$.
-In general, our geometric margin increases as we decrease $C$.
-This is to be expected - as we provide the SVM with more slack, it is able to allow more points within its margin, and can consequently expand its geometric margin.
-In general, the larger the geometric margin, the more support vectors - again, this is intuitive.
-The more slack we allow, the more data points we will observe on or inside of our margin.
-It is not, however, necessarily true that the geometric margin will continue to increase as we decrease $C$ (and also, geometric margin will not always decrease as we increase $C$).
-As C gets very large, our decision boundary will approach the hard SVM decision boundary, and will not move at all.
-As $C$ gets increasingly small, the additional slack will lead our decision boundary to encompass all of the data, at which point our geometric margin will stop increasing.
-
-Choosing $C$ by maximizing the geometric margin on the training data not a mistake that a practitioner should make.
-This is because maximizing the geometric margin on the training data is, in a sense, overfitting on the training data.
-Smaller $C$ would almost always give us larger geometric margins, and the practitioner would feel compelled to choose a very small $C$.
-However, this choice of $C$ may lead to very poor performance on validation sets, or future data.
-The best method for selecting $C$ would be to perform some sort of cross-validation.
-In the most basic sense, we might perform a grid search with various values of C, train our soft SVM, and see which value of C provides the best performance on a held-out validation set.
-
-\begin{table}
-\captionof{table}{Performance of Linear SVM on provided data sets}
-\begin{tabular}{c c|c|c|c|c|c}
-\toprule
-{} & dataset & $w_0$ & $w_1$ & $w_2$ & classification error rate (training) & classification error rate (validation) \\
-\midrule
-  & data\_stdev1 & 0.31 & 5.42 & 0.87 & 0.00\% & 0.00\% \\
-  & data\_stdev2 & 9.09e-2 & 9.83e1 & 9.11e1 & 9.50\% & 7.50\% \\
-  & data\_stdev4 & -2.05e-2 & 2.58e2 & 2.40e2 & 26.00\% & 23.50\% \\
-  & data\_nonsep & 2.56e-3 & 4.23e2 & 3.84e2 & 49.50\% & 51.25\% \\
-\bottomrule
-\end{tabular}
-\end{table}
+We plot our linear SVM classifier below:
 
@@
 
+However, note that as the data becomes less and less linearly separable, our soft SVM with a linear kernel does increasingly poorly.
+This motivates our decision to explore the use of a Gaussian (or RBF) kernel.
+Table 4 includes the performance of our soft SVM algorithm on various training sets when we change kernels (for both linear and Gaussian kernels), and also when we modify the bandwidth of our Gaussian kernel (where applicable) and change our regularization parameter, $C$.
+
+In general, our geometric margin increases as we decrease $C$.
+This is to be expected - as we provide the SVM with more slack, it is able to allow more points within its margin, and can consequently expand its geometric margin.
+In general, the larger the geometric margin, the more support vectors - again, this is intuitive.
+The more slack we allow, the more data points we will observe on or inside of our margin.
+It is not, however, necessarily true that the geometric margin will continue to increase as we decrease $C$ (and also, geometric margin will not always decrease as we increase $C$).
+As C gets very large, our decision boundary will approach the hard SVM decision boundary, and will not move at all.
+As $C$ gets increasingly small, the additional slack will lead our decision boundary to encompass all of the data, at which point our geometric margin will stop increasing.
+
+Choosing $C$ by maximizing the geometric margin on the training data not a mistake that a practitioner should make.
+This is because maximizing the geometric margin on the training data is, in a sense, overfitting on the training data.
+Smaller $C$ would almost always give us larger geometric margins, and the practitioner would feel compelled to choose a very small $C$.
+However, this choice of $C$ may lead to very poor performance on validation sets, or future data.
+The best method for selecting $C$ would be to perform some sort of cross-validation.
+In the most basic sense, we might perform a grid search with various values of C, train our soft SVM, and see which value of C provides the best performance on a held-out validation set.
+
+\begin{table}
+\centering
+\captionof{table}{Classification Performance of Linear SVM}
+\begin{tabular}{c c|c|c|c|c|c}
+\toprule
+{} & dataset & $w_0$ & $w_1$ & $w_2$ & Training Error & Validation Error \\
+\midrule
+  & stdev1 & 0.31 & 5.42 & 0.87 & 0.00\% & 0.00\% \\
+  & stdev2 & 9.09e-2 & 9.83e1 & 9.11e1 & 9.50\% & 7.50\% \\
+  & stdev4 & -2.05e-2 & 2.58e2 & 2.40e2 & 26.00\% & 23.50\% \\
+  & nonsep & 2.56e-3 & 4.23e2 & 3.84e2 & 49.50\% & 51.25\% \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \begin{table}
@@
 \toprule
-{} &                        Kernel &      $C$ &   $\sigma$ &    Number of Support Vectors & $1/||\mathbf{w}||$ \\
+{} &                        Kernel &      $C$ &   $\sigma$ &    \# SV & $1/||\mathbf{w}||$ \\
 \midrule
@@
 \subsubsection*{Titanic Data}
-We're interesting in now testing the performance of our classifiers on a real dataset.
-In this case, we are using the Titanic dataset, where we're performing the rather macabre task of predicting passenger survival given his or her features.
+We're interested in now testing the performance of our classifiers on a real dataset.
+In this case, we are using the Titanic dataset, where we're assigned with the rather macabre task of predicting passenger survival given his or her features.
 For this problem we first scale all the features so they lie between $[0,1]$.
@@
 
-As can be seen in Table 5, logistic regression outperforms polynomial SVM, with a test error rate of 24.9\% (as opposed to 26.5\% the SVM).
-Gaussian SVM performs outperforms the logistic regression, with a test error rate of 24.3\%, and the linear SVM performs the best, with a test error rate of 23.3\%.
-This is an interesting, and somewhat surprising result, as we might expect Gaussian SVM to perform as well (if not better) than linear SVM.
-
-To better understand the differences in our models, one can look at Figure 5, which compares the weights of different variables for our different models.
-Note that each model's array of weights is normalized so that its maximum weight (in this case, always gender) is equal to 1.
-One can see that the linear SVM places almost all of its weight on gender, with a small amount on \# of parents/children.
-Logistic regression also places the most weight on gender, but also puts significant weight on (in order of decreasing magnitude) 3rd class passengers, age, \# of parents and children, 1st class passengers, and 2nd class passengers.
-The polynomial SVM puts most weight on passenger class and passenger embarkation locations.
-And the Gaussian SVM, after age, places the most weight on 3rd class passengers, 2nd class passengers, and passengers who embarked in Cherbourg.
-
 \begin{table}[ht]
@@
 \end{table}
+
+As can be seen in Table 5, logistic regression outperforms polynomial SVM, with a test error rate of 24.9\% (as opposed to 26.5\% the SVM).
+Gaussian SVM performs outperforms the logistic regression, with a test error rate of 24.3\%, and the linear SVM performs the best, with a test error rate of 23.3\%.
+This is an interesting, and somewhat surprising result, as we might expect Gaussian SVM to perform as well (if not better) than linear SVM.
+
+To better understand the differences in our models, one can look at Table 6, which reports the calculated weights for each of our trained models.
+We also depict this visually in Figure 5.
+Note that in Figure 5, each model's array of weights is normalized so that its maximum weight (in this case, always gender) is equal to 1.
+One can see that the linear SVM places almost all of its weight on gender, with a small amount on \# of parents/children.
+Logistic regression also places the most weight on gender, but also puts significant weight on (in order of decreasing magnitude) 3rd class passengers, age, \# of parents and children, 1st class passengers, and 2nd class passengers.
+The polynomial SVM puts most weight on passenger class and passenger embarkation locations.
+And the Gaussian SVM, after age, places the most weight on 3rd class passengers, 2nd class passengers, and passengers who embarked in Cherbourg.
 

### 2015-11-05 HW2/code/problem1/lr_test.py

@@ -1,107 +1,0 @@
-from numpy import *
-from plotBoundary import *
-import scipy.optimize
-import matplotlib as mpl
-from ggplot import *
-import pandas as pd
-
-# Logistic Regression Loss Function
-def LRLoss(w, X, Y, l=0):
-    return sum(log(1+exp(-hstack(Y)*(dot(X,w[1:len(w)])+w[0]))))+l*dot(w[1:len(w)], w[1:len(w)])
-
-# Prediction Function
-def LRPredict(w, X):
-    return 1/(1+exp(-(dot(X,w[1:len(w)])+w[0])))
-
-# Data Loading Function
-def loadData(name, typ):
-    try:
-        data = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_'+typ+'.csv')
-    except:
-        data = loadtxt('/Users/mfzhao/Downloads/hw2_resources/data/data_'+name+'_'+typ+'.csv')
-    X = data[:,0:2]
-    Y = data[:,2:3]
-    return X, Y
-
-# Training Function
-def Train(X, Y, l=0):
-    w0 = random.randn(3)
-    w_opt = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(X,Y,l))
-    return w_opt
-
-# Classification Error Function
-def classifyErr(py, Y, db):
-    yhat = ones(len(py))
-    for i in range(len(py)):
-        if py[i] < db:
-            yhat[i] = -1
-    return 1-mean(Y.flatten() == yhat)
-
-# Classification Error vs Decision Boundary Plotting
-def plotCEDB(w, X, Y, title):
-    global pltsize
-    axis = array(linspace(0,1,101))
-    out = zeros(len(axis))
-    py = LRPredict(w,X)
-    for i in range(len(axis)):
-        out[i] = classifyErr(py, Y, axis[i])
-    DF = pd.DataFrame({'Decision Boundary': pd.Series(axis),'Classification Error': pd.Series(out)})
-    print ggplot(DF, aes(x='Decision Boundary', y='Classification Error')) + geom_line(size=4) + ggtitle(title) + theme_matplotlib(rc=pltsize, matplotlib_defaults=False) 
-
-# Grid Search Lambda Function
-def GridL(Xt, Yt, Xv, Yv, l, db=0.5):
-    tErr = []
[…]

### 2015-11-05 HW2/code/problem2/plotBoundary.py

@@ -1,28 +1,0 @@
-import pdb
-from numpy import *
-import pylab as pl
-
-# X is data matrix (each row is a data point)
-# Y is desired output (1 or -1)
-# scoreFn is a function of a data point
-# values is a list of values to plot
-
-def plotDecisionBoundary(X, Y, scoreFn, values, title = ""):
-    # Plot the decision boundary. For that, we will asign a score to
-    # each point in the mesh [x_min, m_max]x[y_min, y_max].
-    x_min, x_max = X[:, 0].min() - 1, X[:, 0].max() + 1
-    y_min, y_max = X[:, 1].min() - 1, X[:, 1].max() + 1
-    h = max((x_max-x_min)/200., (y_max-y_min)/200.)
-    xx, yy = meshgrid(arange(x_min, x_max, h),
-                      arange(y_min, y_max, h))
-    X_tmp = array([[x1, x2] for x1, x2 in zip(ravel(xx), ravel(yy))])
-    zz = array([scoreFn(X_tmp)])
-    zz = zz.reshape(xx.shape)
-    pl.figure()
-    CS = pl.contour(xx, yy, zz, values, colors = 'green', linestyles = 'solid', linewidths = 2)
-    pl.clabel(CS, fontsize=9, inline=1)
-    # Plot the training points
-    pl.scatter(X[:, 0], X[:, 1], c=(1.-Y), s=50, cmap = pl.cm.cool)
-    pl.title(title)
-    pl.axis('tight')
-    pl.show()

### 2015-11-05 HW2/code/problem3/plotBoundary.py

@@ -1,26 +1,0 @@
-import pdb
-from numpy import *
-from matplotlib import pylab as pl
-
-# X is data matrix (each row is a data point)
-# Y is desired output (1 or -1)
-# scoreFn is a function of a data point
-# values is a list of values to plot
-
-def plotDecisionBoundary(w, X, Y, scoreFn, values, title = ''):
-    # Plot the decision boundary. For that, we will asign a score to
-    # each point in the mesh [x_min, m_max]x[y_min, y_max].
-    x_min, x_max = X[:, 0].min() - 1, X[:, 0].max() + 1
-    y_min, y_max = X[:, 1].min() - 1, X[:, 1].max() + 1
-    h = max((x_max-x_min)/200., (y_max-y_min)/200.)
-    xx, yy = meshgrid(arange(x_min, x_max, h),
-                      arange(y_min, y_max, h))
-    zz = array([scoreFn(w, x) for x in c_[xx.ravel(), yy.ravel()]])
-    zz = zz.reshape(xx.shape)
-    pl.figure()
-    CS = pl.contour(xx, yy, zz, values, colors = 'green', linestyles = 'solid', linewidths = 2)
-    pl.clabel(CS, fontsize=9, inline=1)
-    # Plot the training points
-    pl.scatter(X[:, 0], X[:, 1], c=(1.-Y), s=50, cmap = pl.cm.cool)
-    pl.title(title)
-    pl.axis('tight')

### 2015-11-05 HW2/code/problem3/problem2.py

@@ -1,157 +1,0 @@
-import numpy as np
-from plotBoundary import *
-import cvxopt
-import sys
-from cvxopt import matrix
-# import your SVM training code
-
-# parameters
-name = 'nonsep'
-print '======Training======'
-# load data from csv files
-try:
-    train = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_train.csv')
-except:
-    train = loadtxt('/Users/mfzhao/Downloads/hw2_resources/data/data_'+name+'_train.csv')
-import numpy as np
-from plotBoundary import *
-import cvxopt
-import sys
-# import your SVM training code
-X = train[:, 0:2].copy()
-Y = train[:, 2:3].copy()
-
-def linear_kernel(x1, x2):
-    return np.dot(x1, x2)
-
-def polynomial_kernel(x, y, p=3):
-    return (1 + np.dot(x, y)) ** p
-
-sigma =  1
-def gaussian_kernel(x, y):
-    global sigma
-    return np.exp(-linalg.norm(x-y)**2 / (2 * (sigma ** 2)))
-
-#X = np.array([[1., 2.], [2., 2.], [0., 0.], [-2., 3.]])
-#Y = np.array([1., 1., -1., -1.])
-C = 1
-kernel = gaussian_kernel
-
-
-# Carry out training, primal and/or dual
-def trainSVM(X, y, kernel, C):
-	n_samples, n_features = X.shape
-	K = np.zeros((n_samples, n_samples))
-	for i in range(n_samples):
-		for j in range(n_samples):
-			K[i, j] = kernel(X[i], X[j])
-
-	P = cvxopt.matrix(np.outer(y,y) * K)
-	q = cvxopt.matrix(np.ones(n_samples) * -1.)
-	A = cvxopt.matrix(y, (1, n_samples))
-	b = cvxopt.matrix(0.0)
-
-	G = cvxopt.matrix(np.vstack([np.diag(np.ones(n_samples) * -1.), np.diag(np.ones(n_samples))]))
-	h = cvxopt.matrix(np.hstack([np.zeros(n_samples), C*np.ones(n_samples)])) 
- 
-	# Solve QP problem
-	solution = cvxopt.solvers.qp(P, q, G, h, A, b)
-	alpha = np.ravel(solution['x'])
-
-	sv_bool = alpha > .00001
-	ind = np.arange(len(alpha))[sv_bool]
-	alpha = alpha[sv_bool]
[…]

### 2015-11-05 HW2/code/problem3/titanic.py

@@ -1,84 +1,0 @@
-from numpy import *
-import scipy.optimize
-import matplotlib as mpl
-from ggplot import *
-import pandas as pd
-import scipy.io
-
-def scaling(X):
-    for i in range(11):
-        xmax = max(X[:,i])
-        xmin = min(X[:,i])
-        X[:,i] =  (X[:,i] - xmin)/(xmax - xmin)
-
-# Logistic Regression Loss Function
-def LRLoss(w, X, Y, l=0):
-    return sum(log(1+exp(-hstack(Y)*(dot(X,w[1:len(w)])+w[0]))))+l*dot(w[1:len(w)], w[1:len(w)])
-
-# Prediction Function
-def LRPredict(w, X):
-    return 1/(1+exp(-(dot(X,w[1:len(w)])+w[0])))
-    
-# Training Function
-def Train(X, Y, l=0):
-    w0 = random.randn(12)
-    w_opt = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(X,Y,l))
-    return w_opt
-
-# Classification Error Function
-def classifyErr(w, X, Y, db):
-    py = LRPredict(w,X)
-    yhat = ones(len(py))
-    for i in range(len(py)):
-        if py[i] < db:
-            yhat[i] = -1
-    return 1-mean(Y.flatten() == yhat)
-
-def loadTitan(name, typ):
-    file = 'data_'+name+'_'+typ+'.csv'
-    try:
-        data = scipy.io.loadmat('/Users/dholtz/Downloads/hw2_resources/data/'+file)['data']
-    except:
-        data = scipy.io.loadmat('/Users/mfzhao/Downloads/hw2_resources/data/'+file)['data']
-    X = data[:,0:11]
-    scaling(X)
-    Y = data[:,11:12]
-    return X, Y
-
-def GridLR(Xt, Yt, Xv, Yv, l, db=0.5):
-    l_opt = 0
-    bestLoss = 10000
-    w0 = random.randn(12)
-    w_opt = zeros(12)
-    vErr = []
-    for i in range(len(l)):
-        w = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(Xt,Yt,l[i]))
-        Loss = LRLoss(w, Xv, Yv, 0)/len(Yv)
-        vErr.append(Loss)
-        if Loss < bestLoss:
-            bestLoss = Loss
-            l_opt = l[i]
-            w_opt = w
-    return w_opt, l_opt
-
[…]

### 2015-11-09 HW3/tex/HW3.tex
[hw3] initial commit hw3.

--- 2015-10-01 HW1.tex
@@
 \begin{document}
-	\title{6.867 Problem Set 1}
-	\date{October 1, 2015}
+	\title{6.867 Problem Set 3}
 	\maketitle
 
-\subsubsection*{Problem 1}
-
-We implemented gradient descent in Python using an algorithm similar to the pseudocode below.
-The function is written in such a way that it can accept arbitrary scalar functions $f$ and gradients, $\nabla f(x)$, where $f(x)$ here is a function of an arbitrary number of parameters, $n$, and return $\min{f(x)}$.
-
-\begin{algorithm}
-\caption{Gradient Descent}
-\label{GradDescent}
-\begin{algorithmic}[1]
-\Procedure{Gradient Descent}{}
-\State Specify initial guess, step size, and convergence criteria
-\State Set current guess equal to initial guess
-\While{Distance between last guess and current guess location is greater than convergence criteria}
-\State Calculate function at current location
-\State Calculate the gradient of function at current location
-\State Updated guess = take a step from current guess in direction of gradient proportional to step size
-\State Calculate distance between updated guess and current guess
-\State Set current guess equal to updated guess
-\EndWhile{}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-In order to understand the effect of various hyperparameters (our initial guess, the step size, and the convergence criteria) of the algorithm, we compared the results of our gradient descent routine given a number of different parameter combinations for two different functions: $f(x) = \sum_i^n x_i^2$ and $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$.
-The first of these is a convex function, whereas the second is quite non-convex, with many local minima.
-The results of our gradient descent algorithm for numerous values are found below.
-
-We find that, in general, the convex function is much better behaved.
-Regardless of our initial guess or our learning rate, the algorithm converges to a guess in fewer than 1,000 steps, and these guess are consistently close to $x = (0,0,0,0)$, the ``true" value.
-The final guess' distance from the true value of $x$ is primarily a function of the convergence criteria, which makes sense - the less picky we are with our convergence criteria, the further from the true value of $x$ our final guess will be.
-
-The non-convex, on the other hand, is much more sensitive to the hyperparameters we specify.
-The most notable departure from the convex function is the influence of our initial guess.
-It is much easier for the gradient descent algorithm to get trapped in local minima, and this is clearly illustrated in our results.
-While this is the most obvious difference, note that changes in both the learning rate and the convergence criteria also lead to non-trivially different guesses for $\min{f(x)}$.
-
-Unfortunately, it might not always be convenient to specify the gradient of our scalar function, $f(x)$, explicitly.
-This might be because it is complicated to compute, or because no closed form solution exists.
-In order to get around this issue, we can approximate the gradient of the function using central differences.
-We approximate the $i$th element of the gradient, $\nabla_i$, as
+\subsubsection*{Neural Networks}
 
 \begin{equation}
-\nabla_i \approx \frac{f(x_i+h) - f(x_i-h)}{2h}
+\frac{\partial J(w)}{\partial w_{kj}^{(2)}} = \sum_{i=1}^N z_j \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) + 2 \lambda w_{kj}^{(2)}
+\end{equation}
+\begin{equation}
+\frac{\partial J(w)}{\partial w_{jd}^{(1)}} = \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right) + 2 \lambda w_{jd}^{(1)}
 \end{equation}
 
-\noindent where $h$ is some small constant (we use $h = .001$).
-
-We see that the performance in the calculation of the gradient is identical between the analytical approach and the central difference approach.
-That's great!
-Now let's see how our gradient descent algorithm compares to two other algorithms - Broyden?Fletcher?Goldfarb?Shanno (BFGS) and Conjugate Gradient (CG).
-We compare both the number of iterations and function evaluations.
-We'll consider our algorithm in two forms - one where we pass in an explicit gradient function, and another where we estimate the gradient using the central difference method.
-All benchmarking is done with the same starting point: [5, -3, 7, 8].
-$f_1(x) =  \sum_i^n x_i^2$ and $f_2(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$.
-We see that when passed an analytical solution for $\nabla f(x)$, our algorithm requires about as many function calls, although many more iterations than both BFGS and CG.
-Given we do not have an analytical expression for $\nabla f(x)$, however, our algorithm requires many more iterations and function calls.
-
-\begin{table}
-\captionof{table}{Performance of gradient descent given various algorithm hyperparameters} 
-\begin{tabular}{llllll}
-\toprule
-{} &                        $Guess_{i}$ &      $R_{learn}$ &   Thresh. &    $Guess_{f}$ ($f(x) = \sum_i^n x_i^2$) & $Guess_{f}$ ($f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$) \\
-\midrule
-&    [5.0, -3.0, 7.0, 8.0] &    0.3 &    0.1 &   [0.020, -0.012, 0.029, 0.033] &    [6.150, -0.314, 6.165, 6.176] \\
-&    [5.0, -3.0, 7.0, 8.0] &    0.3 &  0.001 &   [0.001, -0.001, 0.002, 0.002] &    [6.159, -0.025, 6.160, 6.161] \\
-&    [5.0, -3.0, 7.0, 8.0] &    0.3 &  1e-05 &   [0.000, -0.000, 0.000, 0.000] &    [6.160, -0.003, 6.160, 6.160] \\
-&    [5.0, -3.0, 7.0, 8.0] &  0.003 &    0.1 &   [1.172, -0.703, 1.641, 1.876] &    [5.005, -2.999, 6.995, 7.993] \\
-&    [5.0, -3.0, 7.0, 8.0] &  0.003 &  0.001 &   [0.118, -0.071, 0.165, 0.188] &    [6.092, -0.932, 6.200, 6.267] \\
-&    [5.0, -3.0, 7.0, 8.0] &  0.003 &  1e-05 &   [0.012, -0.007, 0.017, 0.019] &    [6.092, -0.932, 6.200, 6.267] \\
- &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &    0.1 &  [-0.041, 0.016, 0.004, -0.020] &  [-12.059, 6.070, 0.014, -6.139] \\
-&  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  0.001 &  [-0.003, 0.001, 0.000, -0.001] &  [-12.294, 6.152, 0.001, -6.158] \\
- &  [-10.0, 4.0, 1.0, -5.0] &    0.3 &  1e-05 &  [-0.000, 0.000, 0.000, -0.000] &  [-12.315, 6.159, 0.000, -6.159] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &    0.1 &  [-2.388, 0.955, 0.239, -1.194] &  [-10.002, 4.004, 0.995, -5.005] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  0.001 &  [-0.240, 0.096, 0.024, -0.120] &  [-11.748, 5.885, 0.063, -6.076] \\
- &  [-10.0, 4.0, 1.0, -5.0] &  0.003 &  1e-05 &  [-0.024, 0.010, 0.002, -0.012] &  [-11.848, 5.938, 0.051, -6.092] \\
- &   [1.0, -1.0, -3.0, 2.0] &    0.3 &    0.1 &  [0.010, -0.010, -0.031, 0.020] &   [0.007, -0.007, -0.314, 0.024] \\
- &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  0.001 &  [0.001, -0.001, -0.002, 0.001] &   [0.001, -0.001, -0.025, 0.002] \\
- &   [1.0, -1.0, -3.0, 2.0] &    0.3 &  1e-05 &  [0.000, -0.000, -0.000, 0.000] &   [0.000, -0.000, -0.003, 0.000] \\
- &   [1.0, -1.0, -3.0, 2.0] &  0.003 &    0.1 &  [0.736, -0.736, -2.207, 1.471] &   [0.995, -0.995, -2.999, 1.994] \\
- &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  0.001 &  [0.073, -0.073, -0.220, 0.147] &   [0.051, -0.051, -0.932, 0.143] \\
- &   [1.0, -1.0, -3.0, 2.0] &  0.003 &  1e-05 &  [0.007, -0.007, -0.022, 0.015] &   [0.051, -0.051, -0.932, 0.143] \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-\captionof{table}{Performance in gradient calculation for analytical and central difference approaches} 
-\begin{tabular}{llllll}
-\toprule
-{} &   $x_{i}$  &  function &   $\nabla_{analytical}$  & $\nabla_{approximate}$ \\
-\midrule
-&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
-&    [5.0, -3.0, 7.0, 8.0] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [-0.8589 -0.2011  0.797   1.1494] &[-0.8589 -0.2011  0.797   1.1494] \\
-&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^2$ &   [-20.   8.   2. -10.] &  [-20.   8.   2. -10.]\\
-&    [-10.   4.   1.  -5.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ &[ 0.344  -0.6768  0.8615  0.8589]&[ 0.344  -0.6768  0.8615  0.8589] \\
-&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^2$ &  [ 10.  -6.  14.  16.] & [ 10.  -6.  14.  16.]\\
-&    [ 1. -1. -3.  2.] & $f(x) = \sum_i^n x_i^{\frac{2}{100}} + \cos(x)$ & [ 0.8615 -0.8615 -0.2011  0.9493] & [ 0.8615 -0.8615 -0.2011  0.9493] \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-\captionof{table}{Performance of authors' gradient descent algorithm compared to BFGS and CG}
-\begin{tabular}{llllll}
-\toprule
-{} & algorithm & $f_1(x)$ function calls & $f_1(x)$ iterations &  $f_2(x)$ function calls &  $f_2(x)$ iterations \\
-\midrule
-  & BFGS & 24 & 2 & 78 & 10 \\
-  & CG & 30 & 2 & 78 & 6 \\
-  & Gradient Descent (analytical) & 20 & 10 & 48 & 24 \\ 
-  & Gradient Descent (approximate) & 100 & 10 & 240 & 24 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection*{Problem 2}
-
-\subsubsection*{Problem 3}
-
-To begin problem 3, we implemented the closed form solution of Ridge Regression in Python, making use of Scipy and Numpy.
-Note that if you want to test cases where the number of features is greater than the number of data points (something that a well-trained teacher of machine should be skeptical to do), it is useful to calculate the inverse of a matrix using the pseudo-inverse, rather than the inverse.
-We learned this the long way after debugging code for hours.
-
-The closed form solution for ridge regression is given by the expression:
-
-\begin{equation}
-\hat{\theta} = (\mathbf{X}^T\mathbf{X} + \lambda \mathbf{I})^{-1}\mathbf{X}^t\mathbf{y}
-\end{equation}
-
-\noindent where $\mathbf{X}$ is your matrix of features, $y$ is your array of observed values, and $\hat{\theta}$ are your estimated feature weights.
-$\lambda$ is a regularization parameter, which represents how much you penalize the model for having large feature weights.
-
-Given the data in Bishop 1.4 (which can be seen below), we attempt to fit polynomials of various orders, $M$.
-
-\begin{center}
-\includegraphics[scale=.3]{Bishop_14_Data.png}
-\end{center}
-
-The plot below shows the mean squared error (MSE) that our ridge regression fit gives for values of $M$ between 1 and 5, for values of $\lambda$ ranging from $10^0 -1$ to $10^2$.
-Note that in the case where $\lambda = 0$, we have $OLS$ regression, and as $M$ increases, the model is free to overfit our data.
-While the MSE here is low, if we evaluated this on a new data point, our model would likely perform very poorly.
-As $\lambda$ gets higher, we penalize the model for adding additional large weights to $\hat{\theta}$, but our $MSE$ stays relatively large.
-However, these ``large" MSE models are likely to perform better on a new data point!
-
-\begin{center}
-\includegraphics[scale=.4]{MSE_Lambda_Bishop.png}
-\end{center}
-
-In order to avoid overfitting, but in order to make sure that we have chosen values of $\lambda$ and $M$ (and in general, hyperparameters) that make sense, it is common to split the data into a training set and a test set.
-We will use the closed-form solution for ridge regression to calculate $\hat{\theta}$ on our training set.
-We will then use that $\hat{\theta}$ to attempt to predict values of $\mathbf{y}$ for the test data.
-By choosing the $\hat{\theta}$ that minimizes MSE on the test data, but is trained on the training data, we are ensuring that we have not overfit our model to the data it is trained on.
-As one final step in the model selection process, we can check the performance of our final $\hat{\theta}$ on a validation set, to make sure that performance is still good (and we have not overfit for performance on our test set).
-
-We use this methodology to find the best $\lambda$ and $M$ given three datasets similar to the bishop data.
-We calculate $\lambda$ and $M$ using the regress\_train.txt, data we validate using regress\_test.txt, and then we test using regress\_validate.txt.
-We search through a grid of various values for $\lambda$ and $M$, and find that $MSE$ is lowest on the validation set when $\lambda = 1$ and $M = 1$: $MSE = 1.776$ on the validation data set.
-These parameters gave us an $MSE$ of 1.460 on the training data and and an $MSE$ of 1.415 on the test data.
-The fact that our $MSE$ is on the same order of magnitude for the test data is a good sign - it indicates that in making sure we didn't overfit to the training data, we didn't accidentally overfit to the validation data.
-
-The $MSE$ for various combinations of $\lambda$ and $M$ on the three datasets (including our chosen set of hyperparameters) is found below.
-We see that with large $M$ and small $\lambda$, we do great on the training data, but bad on our validation data.
-For large values of both $\lambda$ and $M$, we do well, but not as well as our optimal values we found through the grid search.
-
-Let's now use this methodology on a real dataset, as opposed to a toy dataset.
-The dataset that we will use was compiled by Kirsztian Buza at Budapest University of Technology and Economics.
-It includes 280 numerical values describing a number of blog posts, such as day of week, number of parent pages, and number of comments received within 24 hours.
-We are hoping to predict the number of comments it will receive in the next 24 hours.
-We are given this data split into training, validation, and test datasets, so we will once again perform a grid search to find the optimal value of $\lambda$.
-In this particular case, we are not regressing a polynomial of arbitrary order, so we do not need to test different values of $M$.
-
-\begin{center}
-\includegraphics[scale=.4]{MSE_comparison.png}
-\end{center}
-
-We find that for very large $\lambda$ (as large as $10^6$, $MSE$ continues to decrease, although it is always close to 900.
-The MSE for a uniform predictor that predicts the average of $y$, $\hat{y}$, has an MSE of 1453, so the performance of our model is still better.
-However, we see that for values of $\lambda$ higher than $10^6$ or so, the value of the MSE increases, and eventually approaches (and eclipses) the MSE of our uniform estimator.
-
-This isn't surprising - the blog feedback dataset is large, with 10,480 observations in the test set.
-If we interpret the value of $\lambda$ as strength of our prior, it makes sense that it does not have a substantial effect until it is significantly larger than the size of our data.
-Furthermore, the inability of our model to get very low MSE for low $\lambda$ speaks both to the fact that the data has a much higher dimension than our training data, and also suggests that the simple linear model we are using here may not be a very good fit for the data.
-
-\begin{center}
-\includegraphics[scale=.4]{BlogFeedbackRegressionMSE.png}
-\end{center}
-
-\subsubsection*{Problem 4}
-
 \end{document}

### 2015-11-09 HW3/code/hw3.py

@@ -1,0 +1,50 @@
+import numpy as np 
+import scipy.io 
+import sys
+from scipy.special import expit
+
+filename = "/Users/dholtz/Downloads/hw3_resources/toy_multiclass_1_train.csv"
+T = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+
+X_train = T[:,0:2]
+Y_train = T[:,2:3]
+print len(np.unique(Y_train))
+print Y_train.shape[0]
+
+def sigmoidFunction( signal, derivative=False ):
+    signal = np.clip( signal, -500, 500 )
+    
+    signal = expit( signal )
+    
+    if derivative:
+        return np.multiply(signal, 1-signal)
+    else:
+        return signal
+
+def forwardProp(X_train, w1, w2, activation_functions):
+    output = np.dot(X_train, w1[1:,:]) + w1[0:1,:]
+    activations = activation_functions[0](output)
+    output2 = np.dot(activations, w2[1:,:]) + w2[0:1,:]
+    activations2 = activation_functions[1](output2)
+    return activations2
+
+def lossFunction(output, Y_train, l, w1, w2):
+	Y_train_matrix = np.zeros((Y_train.shape[0], len(np.unique(Y_train))))
+	Y_train_locations = np.hstack([np.arange(Y_train.shape[0]).reshape(-1, 1), Y_train-1])
+	Y_train_matrix[Y_train_locations[:,0:1].astype(int), Y_train_locations[:,1:].astype(int)] = 1
+	loss = np.sum(-Y_train_matrix*np.log(output) - (1-Y_train_matrix)*np.log(1-output)) + l*(np.sum(w1**2) + np.sum(w2**2))
+	print loss
+	return loss
+
+def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_rate):
+    w1 = 2*np.random.random((X_train.shape[1]+1,n1)) - 1
+    w2 = 2*np.random.random((w1.shape[1]+1, n2)) - 1 
+    
+    MSE = () 
+    error_limit = 1e-3
+    while MSE > error_limit:
+    	output = forwardProp(X_train, w1, w2, activation_functions)
+    	error = lossFunction(output, Y_train, l, w1, w2)
+	
+
+neuralNetwork(X_train, Y_train, 4, 3, .1, [sigmoidFunction, sigmoidFunction])

### 2015-11-09 HW3/code/hw3.py
[nn] ugh.

@@ -26,15 +26,15 @@ def forwardProp(X_train, w1, w2, activation_functions):
     activations = activation_functions[0](output)
     output2 = np.dot(activations, w2[1:,:]) + w2[0:1,:]
     activations2 = activation_functions[1](output2)
-    return activations2
+    return output, activations, output2, activations2
 
 def lossFunction(output, Y_train, l, w1, w2):
 	Y_train_matrix = np.zeros((Y_train.shape[0], len(np.unique(Y_train))))
 	Y_train_locations = np.hstack([np.arange(Y_train.shape[0]).reshape(-1, 1), Y_train-1])
 	Y_train_matrix[Y_train_locations[:,0:1].astype(int), Y_train_locations[:,1:].astype(int)] = 1
-	loss = np.sum(-Y_train_matrix*np.log(output) - (1-Y_train_matrix)*np.log(1-output)) + l*(np.sum(w1**2) + np.sum(w2**2))
+	loss = np.sum(-Y_train_matrix*np.log(output) - (1-Y_train_matrix)*np.log(1-output)) + l*(np.sum(w1[1:,:]**2) + np.sum(w2[1:,:]**2))
 	print loss
-	return loss
+	return loss, Y_train_matrix
 
 def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_rate):
     w1 = 2*np.random.random((X_train.shape[1]+1,n1)) - 1
@@ -43,8 +43,29 @@ def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_ra
     MSE = () 
     error_limit = 1e-3
     while MSE > error_limit:
-    	output = forwardProp(X_train, w1, w2, activation_functions)
-    	error = lossFunction(output, Y_train, l, w1, w2)
+    	output, activations, output2, activations2 = forwardProp(X_train, w1, w2, activation_functions)
+    	error, Y_train_matrix = lossFunction(activations2, Y_train, l, w1, w2)
+
+    	corrections2 = Y_train_matrix*(1-activations2) + (1-Y_train_matrix)*activations2
+    	activationsBias2 = np.hstack([activations, np.ones((activations.shape[0], 1)).reshape(-1, 1)])
+    	errorTerm2 = np.dot(np.transpose(activationsBias2), corrections2)
+    	errorTermReg2 = errorTerm2 + 2*l*w2
+        print errorTermReg2
+
+
[…]

### 2015-11-09 HW3/code/hw3.py
[hw3] nn works, does poorly.

@@ -41,16 +41,18 @@ def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_ra
     w2 = 2*np.random.random((w1.shape[1]+1, n2)) - 1 
     
     MSE = () 
-    error_limit = 1e-3
-    while MSE > error_limit:
+    c = 1e-5
+    error = 10000
+    old_error = 1000000
+    while np.abs(old_error-error) > c:
+    	old_error = error
     	output, activations, output2, activations2 = forwardProp(X_train, w1, w2, activation_functions)
     	error, Y_train_matrix = lossFunction(activations2, Y_train, l, w1, w2)
 
     	corrections2 = Y_train_matrix*(1-activations2) + (1-Y_train_matrix)*activations2
     	activationsBias2 = np.hstack([activations, np.ones((activations.shape[0], 1)).reshape(-1, 1)])
     	errorTerm2 = np.dot(np.transpose(activationsBias2), corrections2)
-    	errorTermReg2 = errorTerm2 + 2*l*w2
-        print errorTermReg2
+    	errorTermReg2 = errorTerm2/float(X_train.shape[0]) + 2*l*w2
 
 
         X_trainBias = np.hstack([X_train, np.ones((X_train.shape[0], 1)).reshape(-1, 1)])
@@ -58,14 +60,28 @@ def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_ra
         activationDeriv = sigmoidFunction(output, derivative=True)
         activationDerivBias = np.hstack([activationDeriv, np.ones((activationDeriv.shape[0], 1)).reshape(-1,1)])
 
-        print X_trainBias.shape
-        print X_train.shape
-        print corrections2.shape
-        print w2.shape
-        print np.dot(np.dot(np.transpose(np.dot(corrections2, np.transpose(w2))), activations), np.transpose(activationDeriv)).shape
-        print activationDeriv.shape
-        print w1.shape
+        XtimesCorrections = np.dot(np.transpose(X_trainBias), corrections2)
+        productTimesW2 = np.dot(XtimesCorrections, np.transpose(w2[1:,:]))
+        errorTermReg1 = np.dot(np.dot(productTimesW2, np.transpose(activations)), activationDeriv)/float(X_train.shape[0]) + 2*l*w1
 
-	
+        w1 = w1 - learning_rate*errorTermReg1
[…]

### 2015-11-09 HW3/code/hw3.py
[hw3] separate out bias.

@@ -3,7 +3,7 @@ import scipy.io
 import sys
 from scipy.special import expit
 
-filename = "/Users/dholtz/Downloads/hw3_resources/toy_multiclass_1_train.csv"
+filename = "/Users/dholtz/Downloads/hw3_resources/toy_multiclass_2_train.csv"
 T = scipy.io.loadmat(filename, appendmat=False)['toy_data']
 
 X_train = T[:,0:2]
@@ -13,7 +13,6 @@ print Y_train.shape[0]
 
 def sigmoidFunction( signal, derivative=False ):
     signal = np.clip( signal, -500, 500 )
-    
     signal = expit( signal )
     
     if derivative:
@@ -21,10 +20,10 @@ def sigmoidFunction( signal, derivative=False ):
     else:
         return signal
 
-def forwardProp(X_train, w1, w2, activation_functions):
-    output = np.dot(X_train, w1[1:,:]) + w1[0:1,:]
+def forwardProp(X_train, w1, w2, b1, b2, activation_functions):
+    output = np.dot(X_train, w1) + b1
     activations = activation_functions[0](output)
-    output2 = np.dot(activations, w2[1:,:]) + w2[0:1,:]
+    output2 = np.dot(activations, w2) + b2
     activations2 = activation_functions[1](output2)
     return output, activations, output2, activations2
 
@@ -32,13 +31,21 @@ def lossFunction(output, Y_train, l, w1, w2):
 	Y_train_matrix = np.zeros((Y_train.shape[0], len(np.unique(Y_train))))
 	Y_train_locations = np.hstack([np.arange(Y_train.shape[0]).reshape(-1, 1), Y_train-1])
 	Y_train_matrix[Y_train_locations[:,0:1].astype(int), Y_train_locations[:,1:].astype(int)] = 1
-	loss = np.sum(-Y_train_matrix*np.log(output) - (1-Y_train_matrix)*np.log(1-output)) + l*(np.sum(w1[1:,:]**2) + np.sum(w2[1:,:]**2))
+	loss = np.sum(-Y_train_matrix*np.log(output) - (1-Y_train_matrix)*np.log(1-output)) + l*(np.sum(w1**2) + np.sum(w2**2))
 	print loss
 	return loss, Y_train_matrix
 
 def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_rate):
-    w1 = 2*np.random.random((X_train.shape[1]+1,n1)) - 1
[…]

### 2015-11-09 HW3/code/hw3.py
[hw3] clean up some y stuff.

@@ -6,7 +6,7 @@ from scipy.special import expit
 filename = "/Users/dholtz/Downloads/hw3_resources/toy_multiclass_2_train.csv"
 T = scipy.io.loadmat(filename, appendmat=False)['toy_data']
 
-X_train = T[:,0:2]
+X_train = T[:,0:2]	
 Y_train = T[:,2:3]
 print len(np.unique(Y_train))
 print Y_train.shape[0]
@@ -28,8 +28,13 @@ def forwardProp(X_train, w1, w2, b1, b2, activation_functions):
     return output, activations, output2, activations2
 
 def lossFunction(output, Y_train, l, w1, w2):
-	Y_train_matrix = np.zeros((Y_train.shape[0], len(np.unique(Y_train))))
-	Y_train_locations = np.hstack([np.arange(Y_train.shape[0]).reshape(-1, 1), Y_train-1])
+	# m = number of data points
+	# k = number of outputs
+    m = Y_train.shape[0]
+    k = len(np.unique(Y_train))
+
+	Y_train_matrix = np.zeros((m, k))
+	Y_train_locations = np.hstack([np.arange(m).reshape(-1, 1), Y_train-1])
 	Y_train_matrix[Y_train_locations[:,0:1].astype(int), Y_train_locations[:,1:].astype(int)] = 1
 	loss = np.sum(-Y_train_matrix*np.log(output) - (1-Y_train_matrix)*np.log(1-output)) + l*(np.sum(w1**2) + np.sum(w2**2))
 	print loss

### 2015-11-10 HW3/code/hw3.py
HW3mk2 working

@@ -3,11 +3,18 @@ import scipy.io
 import sys
 from scipy.special import expit
 
-filename = "/Users/dholtz/Downloads/hw3_resources/toy_multiclass_2_train.csv"
-T = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+try:
+    filename = "/Users/dholtz/Downloads/hw3_resources/toy_multiclass_1_train.csv"
+    T = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+except:
+    filename = "/Users/mfzhao/Downloads/hw3_resources/toy_multiclass_2_train.csv"
+    T = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+
 
 X_train = T[:,0:2]
 Y_train = T[:,2:3]
+print np.max(X_train[0])
+print np.max(X_train[1])
 print len(np.unique(Y_train))
 print Y_train.shape[0]
 
@@ -42,13 +49,13 @@ def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_ra
     # n1 = number of hidden units
     m = Y_train.shape[0]
     d = X_train.shape[1]
-    w1 = (2*np.random.random((d,n1)) - 1)*.001
-    w2 = (2*np.random.random((n1, n2)) - 1)*.001
+    w1 = (2*np.random.random((d,n1)) - 1)*.1
+    w2 = (2*np.random.random((n1, n2)) - 1)*.1
     b1 =  np.zeros((1, n1))
     b2 =  np.zeros((1, n2))
     
     MSE = () 
-    c = 1e-5
+    c = 1e-8
     error = 10000
     old_error = 1000000
     while np.abs(old_error-error) > c:
@@ -56,7 +63,7 @@ def neuralNetwork(X_train, Y_train, n1, n2, l, activation_functions, learning_ra
     	output, activations, output2, activations2 = forwardProp(X_train, w1, w2, b1, b2, activation_functions)
     	error, Y_train_matrix = lossFunction(activations2, Y_train, l, w1, w2)
 
-    	corrections2 = Y_train_matrix*(1-activations2) + (1-Y_train_matrix)*activations2
+    	corrections2 = -Y_train_matrix*(1-activations2) + (1-Y_train_matrix)*activations2
     	errorTerm2 = np.dot(np.transpose(activations), corrections2)
     	errorTermReg2 = errorTerm2 + 2*l*w2
 
[…]

### 2015-11-10 HW3/code/hw3mk2.py

@@ -1,0 +1,86 @@
+import numpy as np 
+import scipy.io 
+import sys
+from scipy.special import expit
+
+
+path = "/Downloads/hw3_resources/mnist_train.csv"
+try:
+    filename = "/Users/dholtz/"+path
+    T = scipy.io.loadmat(filename, appendmat=False)['tr']
+except:
+    filename = "/Users/mfzhao/"+path
+    T = scipy.io.loadmat(filename, appendmat=False)['tr']
+
+def sigm(x):
+    x = np.clip(x, -500, 500)
+    return expit(x)
+
+def dsigm(x):
+    return sigm(x)*(1-sigm(x))
+
+def fprop(w1, w2, b1, b2, x):
+    z2 = np.dot(x,w1) + b1
+    a2 = sigm(z2)
+    z3 = np.dot(a2, w2) + b2
+    a3 = sigm(z3)
+    return z2, a2, z3, a3
+
+def oneHot(Y):
+	Y_1H = np.zeros((Y.shape[0], len(np.unique(Y))))
+	Y_train_locations = np.hstack([np.arange(Y.shape[0]).reshape(-1, 1), Y-1])
+	Y_1H[Y_train_locations[:,0:1].astype(int), Y_train_locations[:,1:].astype(int)] = 1
+	return Y_1H
+
+def LossFn(py, Y):
+    return -Y*np.log(py)-(1-Y)*np.log(1-py)
+    
+def NNet(X, Y, m, l, a, crit=1e-5):
+    n = Y.shape[0]
+    d = X.shape[1] 
+    k = Y.shape[1]
+    w1 = (2*np.random.random((d,m))-1)*0.5
+    w2 = (2*np.random.random((m,k))-1)*0.5
+    b1 = np.zeros((1,m))
+    b2 = np.zeros((1,k))
+    count = 0
+    lossPrev = 10000000
+    lossNow = 100000
+    
+    while count < 2:
+        lossPrev = lossNow
+        z2, a2, z3, a3 = fprop(w1, w2, b1, b2, X)
+        lossNow = np.sum(LossFn(a3, Y))
+        print lossNow
+        if np.abs(lossPrev - lossNow) < crit:
+            count += 1
+        delta3 = a3 - Y
+        w2update = np.dot(np.transpose(a2), delta3)/n + 2*l*w2
+        delta2 = np.dot(delta3, np.transpose(w2))*dsigm(z2)
[…]

### 2015-11-10 HW3/tex/HW3.tex

@@
 
+In some cases, we might want to use a neural network rather than another algorithm, such as SVM or logistic regression, in order to be somewhat more agnostic about the parametric form of the data-generating process.
+In this case, we investigate the use of neural networks, first on some toy training data sets, and subsequently on a subset of the popular MNIST dataset.
+
+Let us first briefly review the way that neural network works.
+A neural network consists of a variable number of ``hidden layers``.
+A weight matrix maps each of your data points to a value for each node in the hidden layer.
+We then evaluate some ``activation function`` on each node value (typically a sigmoid function or tanh function) to get a new set of ``inputs`` for the next layer of the neural network.
+We continue this process until we finally arrive at our output predictions.
+A schematic sketch of a neural net can be found below.
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[height=3in]{neuralnetdiagram.png}
+	\caption*{Figure 1: A schematic diagram of a neural network (taken from the homework assignment document)}
+\end{figure}
+
+Alternatively, a neural network may be specified as follows:
+
 \begin{equation}
-\frac{\partial J(w)}{\partial w_{kj}^{(2)}} = \sum_{i=1}^N z_j \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) + 2 \lambda w_{kj}^{(2)}
+a_{n}^{(1)} = \sum_d w_{dn}^{(1)}x_d
 \end{equation}
+
 \begin{equation}
-\frac{\partial J(w)}{\partial w_{jd}^{(1)}} = \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right) + 2 \lambda w_{jd}^{(1)}
+z_{n}^{(1)} = \sigma_1(a_n^{(1)})
+\end{equation}
+
+\begin{equation}
+a_{k}^{(2)} = \sum_k w_{nk}^{(2)}z_n
+\end{equation}
+
+\begin{equation}
+z_{k}^{(2)} = \sigma_2(a_k^{(2)})
+\end{equation}
+
+The most trivial neural network has only one output layer with one node, and no hidden layers. This would correspond to vanilla logistic regression.
+
+For the purposes of this paper, we will try to solve simple multi-class prediction problems using a neural network with one hidden layer.
+In order to determine the weight matrices that give us the best predictive accuracy, we will specify a loss function and perform (batch or stochastic) gradient descent.
+We specify our objective function as the regularized negative log-likelihood function:
+
+\begin{equation}
+J(w) = l(w) +\lambda(||w^{(1)}||^2_F + ||w^{(2)}||^2_F
+\end{equation}
+
+\noindent where $l(w)$ is
+
+\begin{equation}
+l(w) = \sum_{i=1}^N \sum_{k=1}^{K} [-y_k^{(i)}\log(h_k(x^{(i)},w)) - (1-y_k^{(i)})\log(1-(h_k(x^{(i)},w))].
+\end{equation}
+
+In order to calculate the gradient as a function of $w_1$ and $w_2$, we use the two gradient calculation expressions below:
+
+\begin{equation}
+\frac{\partial J(w)}{\partial w_{kj}^{(2)}} = -\left ( \sum_{i=1}^N z_j \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) \right ) + 2 \lambda w_{kj}^{(2)}
+\end{equation}
+
+\noindent and
+
+\begin{equation}
+\frac{\partial J(w)}{\partial w_{jd}^{(1)}} = -\left ( \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\right ) + 2 \lambda w_{jd}^{(1)}.
 \end{equation}

### 2015-11-10 HW3/code/hw3mk2.py
[hw3] small change.

@@ -80,7 +80,7 @@ Ytr = T[:,T.shape[1]-1:T.shape[1]]
 YtrM = oneHot(Ytr)
 
 np.random.seed(11210)
-w1, w2, b1, b2 = NNet(Xtr, YtrM, 200, 0.01, .1, crit=1e-4)
+w1, w2, b1, b2 = NNet(Xtr, YtrM, 200, 0.01, .05, crit=1e-4)
 yhat = nnPredict(w1, w2, b1, b2, Xtr)
 
 classificationError(yhat, Ytr)

### 2015-11-10 HW3/code/hw3mk2.py
SGD implemented/ Toy grid serach done.

@@ -2,15 +2,7 @@ import numpy as np
 import scipy.io 
 import sys
 from scipy.special import expit
-
-
-path = "/Downloads/hw3_resources/mnist_train.csv"
-try:
-    filename = "/Users/dholtz/"+path
-    T = scipy.io.loadmat(filename, appendmat=False)['tr']
-except:
-    filename = "/Users/mfzhao/"+path
-    T = scipy.io.loadmat(filename, appendmat=False)['tr']
+import random
 
 def sigm(x):
     x = np.clip(x, -500, 500)
@@ -35,7 +27,7 @@ def oneHot(Y):
 def LossFn(py, Y):
     return -Y*np.log(py)-(1-Y)*np.log(1-py)
     
-def NNet(X, Y, m, l, a, crit=1e-5):
+def NNet(X, Y, m, l, a, crit=1e-5, maxiter = 100000):
     n = Y.shape[0]
     d = X.shape[1] 
     k = Y.shape[1]
@@ -43,10 +35,10 @@ def NNet(X, Y, m, l, a, crit=1e-5):
     w2 = (2*np.random.random((m,k))-1)*0.5
     b1 = np.zeros((1,m))
     b2 = np.zeros((1,k))
+    it = 0
     count = 0
     lossPrev = 10000000
     lossNow = 100000
-    
     while count < 2:
         lossPrev = lossNow
         z2, a2, z3, a3 = fprop(w1, w2, b1, b2, X)
@@ -54,6 +46,8 @@ def NNet(X, Y, m, l, a, crit=1e-5):
         print lossNow
         if np.abs(lossPrev - lossNow) < crit:
             count += 1
+        else:
+            count = 0
         delta3 = a3 - Y
         w2update = np.dot(np.transpose(a2), delta3)/n + 2*l*w2
         delta2 = np.dot(delta3, np.transpose(w2))*dsigm(z2)
@@ -63,6 +57,47 @@ def NNet(X, Y, m, l, a, crit=1e-5):
         w1 -= a*w1update
         b2 -= delta3.sum(axis=0)/n
         b1 -= delta2.sum(axis=0)/n
+        it += 1
+        if it == maxiter:
+            print 'Maximum number of iterations reached. Convergence Criterion not met.'
+            break
+    return w1, w2, b1, b2
+
+def sgdNNet(X, Y, m, l, a, n, crit=1e-5, maxiter = 100000):
+    d = X.shape[1] 
+    k = Y.shape[1]
[…]

### 2015-11-10 HW3/code/mnist.py

@@ -1,0 +1,19 @@
+import hw3mk2.py
+
+path = "/Downloads/hw3_resources/mnist_train.csv"
+try:
+    filename = "/Users/dholtz/"+path
+    T = scipy.io.loadmat(filename, appendmat=False)['tr']
+except:
+    filename = "/Users/mfzhao/"+path
+    T = scipy.io.loadmat(filename, appendmat=False)['tr']
+    
+Xtr = T[:,0:T.shape[1]-1]
+Ytr = T[:,T.shape[1]-1:T.shape[1]]
+YtrM = oneHot(Ytr)
+
+np.random.seed(11210)
+w1, w2, b1, b2 = NNet(Xtr, YtrM, 200, 0.001, .01, crit=1e-4)
+yhat = nnPredict(w1, w2, b1, b2, Xtr)
+
+classificationError(yhat, Ytr)

### 2015-11-10 HW3/code/toy1.py

@@ -1,0 +1,96 @@
+from hw3mk2 import *
+import scipy
+
+pathTr = "/Downloads/hw3_resources/toy_multiclass_1_train.csv"
+pathV = "/Downloads/hw3_resources/toy_multiclass_1_validate.csv"
+pathTe = "/Downloads/hw3_resources/toy_multiclass_1_test.csv"
+try:
+    filename = "/Users/mfzhao/"+pathTr
+    Tr = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+    filename = "/Users/mfzhao/"+pathV
+    V = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+    filename = "/Users/mfzhao/"+pathTe
+    Te = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+except:
+    filename = "/Users/mfzhao/"+pathTr
+    Tr = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+    filename = "/Users/mfzhao/"+pathV
+    V = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+    filename = "/Users/mfzhao/"+pathTe
+    Te = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+    
+    
+Xtr = Tr[:,0:Tr.shape[1]-1]
+Ytr = Tr[:,Tr.shape[1]-1:Tr.shape[1]]
+YtrM = oneHot(Ytr)
+Xv = V[:,0:V.shape[1]-1]
+Yv = V[:,V.shape[1]-1:V.shape[1]]
+YvM = oneHot(Yv)
+Xte = Te[:,0:Te.shape[1]-1]
+Yte = Te[:,Te.shape[1]-1:Te.shape[1]]
+YteM = oneHot(Yte)
+
+def gridSearch(Xtr, YtrM, Xv, YvM):
+    lamb = [0, 1e-6, 1e-5, 1e-4, 1e-3, 1e-2, 1e-1, 1]
+    h = [1,2,3,4,5,6,7,8,9,10]
+    besttr = 1
+    besterr = 1
+    bestl = 0
+    bestm = 1
+    trE = np.zeros((len(lamb), len(h)))
+    vE = np.zeros((len(lamb), len(h)))
+    i=0
+    for l in lamb:
+        j=0
+        for m in h:
+            w1, w2, b1, b2 = NNet(Xtr, YtrM, m, l, 15, crit=1e-3)
[…]

### 2015-11-10 HW3/tex/HW3.tex
[hw3] adding toy data set performance stuff.

@@
 \begin{equation}
-\frac{\partial J(w)}{\partial w_{kj}^{(2)}} = -\left ( \sum_{i=1}^N z_j \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) \right ) + 2 \lambda w_{kj}^{(2)}
+\frac{\partial J(w)}{\partial w_{kj}^{(2)}} = \sum_{i=1}^N z_j \left(\sigma(a_k^{(2)}) -y_k^{(i)} \right) + 2 \lambda w_{kj}^{(2)}
 \end{equation}
@@
 \begin{equation}
-\frac{\partial J(w)}{\partial w_{jd}^{(1)}} = -\left ( \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\right ) + 2 \lambda w_{jd}^{(1)}.
+\frac{\partial J(w)}{\partial w_{jd}^{(1)}} =  \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(\sigma(a_k^{(2)}) -y_k^{(i)}\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right) + 2 \lambda w_{jd}^{(1)}.
 \end{equation}
 
+Let's now see how a neural network performs on two different toy data sets.
+We vary both the number of hidden nodes and the regularization parameter in our objective function, and use both full gradient descent and batch gradient descent.
+The two algorithms perform comparably - in the case of toy data set number 2, we find that the full gradient descent algorithm does a slightly better job at finding the optimal values of our parameter than stochastic gradient descent (classification error of 6.7\% as opposed to 9\%).
+Our changes in parameters have a much larger effect on the performance for toy dataset 2 than for our features in toy dataset 1.
+This is likely due to the lack of an easy way to separate examples of different classes in dataset 2.
+
+\begin{table}
+\captionof{table}{Performance of neural network with full gradient descent on Toy Data Set 1 (training)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &  0.007 & \bf{0}     & 0     & 0     & 0     \\
+$\lambda$ = \bf{1e-5} & 0.007 & 0     & 0     & 0     & 0     \\
+$\lambda$ = \bf{1e-3} & 0.093 & 0.003 & 0.003 & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-1} & 0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Performance of neural network with full gradient descent on Toy Data Set 1 (validation)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &      0.01  & \bf{0.003} & 0.003 & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-5} &  0.01  & 0.003 & 0.003 & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-3} &  0.063 & 0.003 & 0.003 & 0.003 & 0.007 \\
+$\lambda$ = \bf{1e-1} &  0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Performance of neural network with stochastic gradient descent on Toy Data Set 1 (training)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &       0.003 & \bf{0.003} & 0.003 & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-5} &   0.037 & 0.003 & 0.003 & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-3} &   0.333 & 0.003 & 0     & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-1} &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Performance of neural network with stochastic gradient descent on Toy Data Set 1 (validation)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &        0.01  & \bf{0.003} & 0.003 & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-5} &    0.03  & 0.003 & 0.007 & 0.003 & 0.007 \\
+$\lambda$ = \bf{1e-3} &    0.333 & 0.007 & 0.01  & 0.01  & 0.017 \\
+$\lambda$ = \bf{1e-1} &    0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Performance of neural network with full gradient descent on Toy Data Set 2 (training)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &  0.36  & 0.063 & 0.047 & 0.05  & 0.027 \\
+$\lambda$ = \bf{1e-5} & 0.397 & 0.063 & 0.067 & 0.063 & 0.067 \\
+$\lambda$ = \bf{1e-3} & 0.413 & \bf{0.06}  & 0.06  & 0.06  & 0.06  \\
+$\lambda$ = \bf{1e-1} & 0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Performance of neural network with full gradient descent on Toy Data Set 2 (validation)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &      0.343 & 0.07  & 0.093 & 0.1   & 0.107 \\
+$\lambda$ = \bf{1e-5} &   0.39  & 0.073 & 0.07  & 0.073 & 0.073 \\
+$\lambda$ = \bf{1e-3} &  0.38  & \bf{0.067} & 0.067 & 0.067 & 0.067 \\
+$\lambda$ = \bf{1e-1} &  0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Performance of neural network with stochastic gradient descent on Toy Data Set 2 (training)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &       0.367 & 0.077 & 0.063 & 0.057 & 0.05  \\
+$\lambda$ = \bf{1e-5} &   0.363 & 0.073 & \bf{0.067} & 0.063 & 0.063 \\
+$\lambda$ = \bf{1e-3} &   0.357 & 0.087 & 0.083 & 0.083 & 0.09  \\
+$\lambda$ = \bf{1e-1} &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Performance of neural network with stochastic gradient descent on Toy Data Set 2 (validation)} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+$\lambda$ = \bf{0} &        0.373 & 0.097 & 0.11  & 0.117 & 0.113 \\
+$\lambda$ = \bf{1e-5} &    0.373 & 0.103 & \bf{0.09}  & 0.09  & 0.09  \\
+$\lambda$ = \bf{1e-3} &    0.37  & 0.107 & 0.103 & 0.103 & 0.103 \\
+$\lambda$ = \bf{1e-1} &    0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \end{document}

### 2015-11-10 HW3/code/hw3mk2.py

@@ -43,7 +43,6 @@ def NNet(X, Y, m, l, a, crit=1e-5, maxiter = 100000):
         lossPrev = lossNow
         z2, a2, z3, a3 = fprop(w1, w2, b1, b2, X)
         lossNow = np.sum(LossFn(a3, Y))
-        print lossNow
         if np.abs(lossPrev - lossNow) < crit:
             count += 1
         else:
@@ -80,7 +79,6 @@ def sgdNNet(X, Y, m, l, a, n, crit=1e-5, maxiter = 100000):
         lossPrev = lossNow
         z2, a2, z3, a3 = fprop(w1, w2, b1, b2, X[ind])
         lossNow = np.sum(LossFn(a3, Y[ind]))
-        print lossNow
         if np.abs(lossPrev - lossNow) < crit:
             count += 1
         else:
@@ -88,7 +86,7 @@ def sgdNNet(X, Y, m, l, a, n, crit=1e-5, maxiter = 100000):
         delta3 = a3 - Y[ind]
         w2update = np.dot(np.transpose(a2), delta3)/n + 2*l*w2
         delta2 = np.dot(delta3, np.transpose(w2))*dsigm(z2)
-        w1update = np.dot(np.transpose(X), delta2)/n + 2*l*w1
+        w1update = np.dot(np.transpose(X[ind]), delta2)/n + 2*l*w1
         
         w2 -= a*w2update
         w1 -= a*w1update
@@ -107,6 +105,5 @@ def nnPredict(w1, w2, b1, b2, Xv):
 
 def classificationError(yPredict, Y):
     classificationError = 1-np.sum(np.array(yPredict.reshape(-1, 1) == Y.reshape(-1, 1), dtype=bool))/float(Y.shape[0])
-    print classificationError
     return classificationError
 

### 2015-11-10 HW3/code/toy1.py

@@ -1,9 +1,10 @@
 from hw3mk2 import *
 import scipy
+from tabulate import tabulate
 
-pathTr = "/Downloads/hw3_resources/toy_multiclass_1_train.csv"
-pathV = "/Downloads/hw3_resources/toy_multiclass_1_validate.csv"
-pathTe = "/Downloads/hw3_resources/toy_multiclass_1_test.csv"
+pathTr = "/Downloads/hw3_resources/toy_multiclass_2_train.csv"
+pathV = "/Downloads/hw3_resources/toy_multiclass_2_validate.csv"
+pathTe = "/Downloads/hw3_resources/toy_multiclass_2_test.csv"
 try:
     filename = "/Users/mfzhao/"+pathTr
     Tr = scipy.io.loadmat(filename, appendmat=False)['toy_data']
@@ -12,11 +13,11 @@ try:
     filename = "/Users/mfzhao/"+pathTe
     Te = scipy.io.loadmat(filename, appendmat=False)['toy_data']
 except:
-    filename = "/Users/mfzhao/"+pathTr
+    filename = "/Users/dholtz/"+pathTr
     Tr = scipy.io.loadmat(filename, appendmat=False)['toy_data']
-    filename = "/Users/mfzhao/"+pathV
+    filename = "/Users/dholtz/"+pathV
     V = scipy.io.loadmat(filename, appendmat=False)['toy_data']
-    filename = "/Users/mfzhao/"+pathTe
+    filename = "/Users/dholtz/"+pathTe
     Te = scipy.io.loadmat(filename, appendmat=False)['toy_data']
     
     
@@ -31,8 +32,8 @@ Yte = Te[:,Te.shape[1]-1:Te.shape[1]]
 YteM = oneHot(Yte)
 
 def gridSearch(Xtr, YtrM, Xv, YvM):
-    lamb = [0, 1e-6, 1e-5, 1e-4, 1e-3, 1e-2, 1e-1, 1]
-    h = [1,2,3,4,5,6,7,8,9,10]
+    lamb = [0, 1e-5, 1e-3, 1e-1]
+    h = [1,3,5,7,9]
     besttr = 1
     besterr = 1
     bestl = 0
@@ -43,17 +44,18 @@ def gridSearch(Xtr, YtrM, Xv, YvM):
     for l in lamb:
         j=0
         for m in h:
+            print l, m
[…]

### 2015-11-10 HW3/code/toyplot.R
[hw3] toy data set plots.

@@ -1,0 +1,41 @@
+library(ggplot2)
+library(dplyr)
+
+toy_test_1 <- read.csv('~/Downloads/toy_multiclass_1_testCLEAN.csv', head=T)
+toy_validate_1 <- read.csv('~/Downloads/toy_multiclass_1_validateCLEAN.csv', head=T)
+toy_train_1 <- read.csv('~/Downloads/toy_multiclass_1_trainCLEAN.csv', head=T)
+
+toy_test_2 <- read.csv('~/Downloads/toy_multiclass_2_testCLEAN.csv', head=T)
+toy_validate_2 <- read.csv('~/Downloads/toy_multiclass_2_validateCLEAN.csv', head=T)
+toy_train_2 <- read.csv('~/Downloads/toy_multiclass_2_trainCLEAN.csv', head=T)
+
+toy_1_complete <- rbind(toy_validate_1 %>%
+  mutate(type = 'Validation'),
+  toy_test_1 %>%
+    mutate(type = 'Test'),
+  toy_train_1 %>%
+  mutate(type = 'Training')
+  ) %>%
+  mutate(Dataset = as.factor(type),
+         `Class` = as.factor(toy.data.3))
+
+toy_2_complete <- rbind(toy_validate_2 %>%
+                          mutate(type = 'Validation'),
+                        toy_test_2 %>%
+                          mutate(type = 'Test'),
+                        toy_train_2 %>%
+                          mutate(type = 'Training')
+) %>%
+  mutate(Dataset = as.factor(type),
+         `Class` = as.factor(toy.data.3))
+
+t1 <- ggplot(toy_1_complete, aes(x=as.numeric(toy.data.1), y=as.numeric(toy.data.2), 
+                       color=Class, shape=Dataset)) + geom_point() + xlab('X1') + ylab('X2') + 
+  ggtitle('Distribution of Toy Data Set 1 in 2-dimensions')
+ggsave(t1, file='../tex/t1dist.jpg')
+
+t2 <- ggplot(toy_2_complete, aes(x=as.numeric(toy.data.1), y=as.numeric(toy.data.2), 
+                           color=Class, shape=Dataset)) + geom_point() + xlab('X1') + ylab('X2') +
+  ggtitle('Distribution of Toy Data Set 2 in 2-dimensions')
+ggsave(t2, file='t2dist.jpg')
+

### 2015-11-10 HW3/code/hw3mk2.py
Adagrad + fixed SGD

@@ -35,6 +35,8 @@ def NNet(X, Y, m, l, a, crit=1e-5, maxiter = 100000):
     w2 = (2*np.random.random((m,k))-1)*0.5
     b1 = np.zeros((1,m))
     b2 = np.zeros((1,k))
+    g1 = np.ones((d,m))
+    g2 = np.ones((m,k))
     it = 0
     count = 0
     lossPrev = 10000000
@@ -43,23 +45,26 @@ def NNet(X, Y, m, l, a, crit=1e-5, maxiter = 100000):
         lossPrev = lossNow
         z2, a2, z3, a3 = fprop(w1, w2, b1, b2, X)
         lossNow = np.sum(LossFn(a3, Y))
+        #print it, ':', lossNow
         if np.abs(lossPrev - lossNow) < crit:
             count += 1
         else:
             count = 0
         delta3 = a3 - Y
-        w2update = np.dot(np.transpose(a2), delta3)/n + 2*l*w2
+        w2update = np.dot(np.transpose(a2), delta3)/n
         delta2 = np.dot(delta3, np.transpose(w2))*dsigm(z2)
-        w1update = np.dot(np.transpose(X), delta2)/n + 2*l*w1
+        w1update = np.dot(np.transpose(X), delta2)/n
         
-        w2 -= a*w2update
-        w1 -= a*w1update
+        g2 += w2update**2
+        g1 += w1update**2
+        w2 -= a*(w2update + 2*l*w2)/np.sqrt(g2)
+        w1 -= a*(w1update + 2*l*w1)/np.sqrt(g1)
         b2 -= delta3.sum(axis=0)/n
         b1 -= delta2.sum(axis=0)/n
         it += 1
         if it == maxiter:
             print 'Maximum number of iterations reached. Convergence Criterion not met.'
-            break
+            return w1, w2, b1, b2
     return w1, w2, b1, b2
 
 def sgdNNet(X, Y, m, l, a, n, crit=1e-5, maxiter = 100000):
@@ -69,6 +74,8 @@ def sgdNNet(X, Y, m, l, a, n, crit=1e-5, maxiter = 100000):
     w2 = (2*np.random.random((m,k))-1)*0.5
     b1 = np.zeros((1,m))
     b2 = np.zeros((1,k))
+    g1 = np.ones((d,m))
+    g2 = np.ones((m,k))
     it = 0
     count = 0
     lossPrev = 10000000
[…]

### 2015-11-10 HW3/code/mnist.py

@@ -1,19 +1,136 @@
-import hw3mk2.py
+from hw3mk2 import *
+import scipy
+from tabulate import tabulate
 
-path = "/Downloads/hw3_resources/mnist_train.csv"
+pathTr = "Downloads/hw3_resources/mnist_train.csv"
+pathV = "Downloads/hw3_resources/mnist_validate.csv"
+pathTe = "Downloads/hw3_resources/mnist_test.csv"
 try:
-    filename = "/Users/dholtz/"+path
-    T = scipy.io.loadmat(filename, appendmat=False)['tr']
+    filename = "/Users/mfzhao/"+pathTr
+    Tr = scipy.io.loadmat(filename, appendmat=False)['tr']
+    filename = "/Users/mfzhao/"+pathV
+    V = scipy.io.loadmat(filename, appendmat=False)['va']
+    filename = "/Users/mfzhao/"+pathTe
+    Te = scipy.io.loadmat(filename, appendmat=False)['te']
 except:
-    filename = "/Users/mfzhao/"+path
-    T = scipy.io.loadmat(filename, appendmat=False)['tr']
+    filename = "/Users/dholtz/"+pathTr
+    Tr = scipy.io.loadmat(filename, appendmat=False)['tr']
+    filename = "/Users/dholtz/"+pathV
+    V = scipy.io.loadmat(filename, appendmat=False)['va']
+    filename = "/Users/dholtz/"+pathTe
+    Te = scipy.io.loadmat(filename, appendmat=False)['te']
+
+def gridSearch(Xtr, YtrM, Xv, YvM, a):
+    lamb = [0, .001, .01, .1]
+    h = [50,100,150]
+    besttr = 1
+    besterr = 1
+    bestl = 0
+    bestm = 1
+    trE = np.zeros((len(lamb), len(h)))
+    vE = np.zeros((len(lamb), len(h)))
+    i=0
+    for l in lamb:
+        j=0
+        for m in h:
+            print l, m
+            w1, w2, b1, b2 = NNet(Xtr, YtrM, m, l, a, crit=1e-3, maxiter=5000)
+            yhattr = nnPredict(w1, w2, b1, b2, Xtr)
+            trE[i,j] = classificationError(yhattr, Ytr)
+            yhatv = nnPredict(w1, w2, b1, b2, Xv)
[…]

### 2015-11-10 HW3/code/toy.py

@@ -1,0 +1,30 @@
+from hw3mk2 import *
+import scipy
+
+pathTr = "/Downloads/hw3_resources/toy_multiclass_1_train.csv"
+pathV = "/Downloads/hw3_resources/toy_multiclass_1_validate.csv"
+pathTe = "/Downloads/hw3_resources/toy_multiclass_1_test.csv"
+
+filename = "/Users/mfzhao/"+pathTr
+Tr = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+filename = "/Users/mfzhao/"+pathV
+V = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+filename = "/Users/mfzhao/"+pathTe
+Te = scipy.io.loadmat(filename, appendmat=False)['toy_data']
+
+Xtr = Tr[:,0:Tr.shape[1]-1]
+Ytr = Tr[:,Tr.shape[1]-1:Tr.shape[1]]
+YtrM = oneHot(Ytr)
+Xv = V[:,0:V.shape[1]-1]
+Yv = V[:,V.shape[1]-1:V.shape[1]]
+YvM = oneHot(Yv)
+Xte = Te[:,0:Te.shape[1]-1]
+Yte = Te[:,Te.shape[1]-1:Te.shape[1]]
+YteM = oneHot(Yte)
+
+w1, w2, b1, b2 = NNet(Xtr, YtrM, 4, 0.001, 15, crit=1e-3)
+yhattr = nnPredict(w1, w2, b1, b2, Xtr)
+print yhattr
+print classificationError(yhattr, Ytr)
+yhatv = nnPredict(w1, w2, b1, b2, Xv)
+print classificationError(yhatv, Yv)

### 2015-11-10 HW3/code/toy1.py
[hw3] fix SGD numbers.

@@ -2,9 +2,9 @@ from hw3mk2 import *
 import scipy
 from tabulate import tabulate
 
-pathTr = "/Downloads/hw3_resources/toy_multiclass_2_train.csv"
-pathV = "/Downloads/hw3_resources/toy_multiclass_2_validate.csv"
-pathTe = "/Downloads/hw3_resources/toy_multiclass_2_test.csv"
+pathTr = "/Downloads/hw3_resources/toy_multiclass_1_train.csv"
+pathV = "/Downloads/hw3_resources/toy_multiclass_1_validate.csv"
+pathTe = "/Downloads/hw3_resources/toy_multiclass_1_test.csv"
 try:
     filename = "/Users/mfzhao/"+pathTr
     Tr = scipy.io.loadmat(filename, appendmat=False)['toy_data']
@@ -95,10 +95,10 @@ def SGDgrid(Xtr, YtrM, Xv, YvM):
     print 'SGD - Validation Error:     ', besterr
     return trE, vE
     
-TrErr_M, VErr_M = gridSearch(Xtr, YtrM, Xv, YvM)
+#TrErr_M, VErr_M = gridSearch(Xtr, YtrM, Xv, YvM)
 TrErr_SGD, VErr_SGD = SGDgrid(Xtr, YtrM, Xv, YvM)
 
-print tabulate(np.round(TrErr_M, 3), tablefmt='latex')
-print tabulate(np.round(VErr_M, 3), tablefmt='latex')
+#print tabulate(np.round(TrErr_M, 3), tablefmt='latex')
+#print tabulate(np.round(VErr_M, 3), tablefmt='latex')
 print tabulate(np.round(TrErr_SGD, 3), tablefmt='latex')
 print tabulate(np.round(VErr_SGD, 3), tablefmt='latex')

### 2015-11-11 HW3/code/mnist.py
mnist fix

@@ -22,7 +22,7 @@ except:
 
 def gridSearch(Xtr, YtrM, Xv, YvM, a):
     lamb = [0, .001, .01, .1]
-    h = [50,100,150]
+    h = [200, 250]
     besttr = 1
     besterr = 1
     bestl = 0
@@ -54,7 +54,7 @@ def gridSearch(Xtr, YtrM, Xv, YvM, a):
     
 def SGDgrid(Xtr, YtrM, Xv, YvM, a):
     lamb = [0, .001, .01, .1]
-    h = [50,100,150]
+    h = [50, 100, 150]
     besttr = 1
     besterr = 1
     bestl = 0
@@ -120,17 +120,18 @@ np.random.seed(11210)
 #print classificationError(sgdyhatv1, Yv)
 #print classificationError(sgdyhattr2, Ytr)
 #print classificationError(sgdyhatv2, Yv)
-
+print '--------------------------step size: .1--------------------------' 
 TrErr_M1, VErr_M1 = gridSearch(Xtr, YtrM, Xv, YvM, .1)
 TrErr_SGD1, VErr_SGD1 = SGDgrid(Xtr, YtrM, Xv, YvM, .1)
-TrErr_M2, VErr_M2 = gridSearch(Xtr, YtrM, Xv, YvM, .1)
-TrErr_SGD2, VErr_SGD2 = SGDgrid(Xtr, YtrM, Xv, YvM, .1)
+print '--------------------------step size: .01--------------------------'
+TrErr_M2, VErr_M2 = gridSearch(Xtr, YtrM, Xv, YvM, .01)
+TrErr_SGD2, VErr_SGD2 = SGDgrid(Xtr, YtrM, Xv, YvM, .01)
 
 print tabulate(np.round(TrErr_M1, 3), tablefmt='latex')
 print tabulate(np.round(VErr_M1, 3), tablefmt='latex')
 print tabulate(np.round(TrErr_SGD1, 3), tablefmt='latex')
 print tabulate(np.round(VErr_SGD1, 3), tablefmt='latex')
-print tabulate(np.round(TrErr_M1, 3), tablefmt='latex')
-print tabulate(np.round(VErr_M1, 3), tablefmt='latex')
-print tabulate(np.round(TrErr_SGD1, 3), tablefmt='latex')
-print tabulate(np.round(VErr_SGD1, 3), tablefmt='latex')
+print tabulate(np.round(TrErr_M2, 3), tablefmt='latex')
[…]

### 2015-11-11 HW3/tex/HW3.tex
[hw3] toy data set plots.
[hw3] table setup.

@@
 
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.48\linewidth}
+		\includegraphics[width=1\linewidth, height=2.7in]{t1dist.jpg}
+		\caption*{Toy Dataset \#1}
+	\end{minipage}
+	\begin{minipage}[b]{.48\linewidth}
+		\includegraphics[width=1\linewidth, height=2.7in]{t2dist.jpg}
+		\caption*{Toy Dataset \#2}
+	\end{minipage}
+	\caption{Distributions of different classes in $X1$, $X2$ for toy datasets}
+\end{figure}
+
 \begin{table}
@@
 \midrule
-$\lambda$ = \bf{0} &       0.003 & \bf{0.003} & 0.003 & 0.003 & 0.003 \\
-$\lambda$ = \bf{1e-5} &   0.037 & 0.003 & 0.003 & 0.003 & 0.003 \\
-$\lambda$ = \bf{1e-3} &   0.333 & 0.003 & 0     & 0.003 & 0.003 \\
+$\lambda$ = \bf{0} &       \bf{0.003} & 0.003 & 0.003 & 0.003 & 0.003 \\
+$\lambda$ = \bf{1e-5} &   0.043 & 0     & 0.003 & 0.003 & 0     \\
+$\lambda$ = \bf{1e-3} &   0.117 & 0     & 0.003 & 0.003 & 0.003 \\
 $\lambda$ = \bf{1e-1} &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
@@
 \midrule
-$\lambda$ = \bf{0} &        0.01  & \bf{0.003} & 0.003 & 0.003 & 0.003 \\
-$\lambda$ = \bf{1e-5} &    0.03  & 0.003 & 0.007 & 0.003 & 0.007 \\
-$\lambda$ = \bf{1e-3} &    0.333 & 0.007 & 0.01  & 0.01  & 0.017 \\
+$\lambda$ = \bf{0} &        \bf{0.003} & 0.003 & 0.01  & 0.007 & 0.003 \\
+$\lambda$ = \bf{1e-5} &    0.033 & 0.007 & 0.003 & 0.007 & 0.003 \\
+$\lambda$ = \bf{1e-3} &    0.097 & 0.013 & 0.013 & 0.013 & 0.013 \\
 $\lambda$ = \bf{1e-1} &    0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
@@
 \midrule
-$\lambda$ = \bf{0} &       0.367 & 0.077 & 0.063 & 0.057 & 0.05  \\
-$\lambda$ = \bf{1e-5} &   0.363 & 0.073 & \bf{0.067} & 0.063 & 0.063 \\
-$\lambda$ = \bf{1e-3} &   0.357 & 0.087 & 0.083 & 0.083 & 0.09  \\
+$\lambda$ = \bf{0} &       0.363 & 0.073 & 0.053 & 0.05  & 0.06  \\
+$\lambda$ = \bf{1e-5} &   0.363 & 0.07  & \bf{0.067} & 0.057 & 0.067 \\
+$\lambda$ = \bf{1e-3} &   0.363 & 0.07  & 0.07  & 0.07  & 0.07  \\
 $\lambda$ = \bf{1e-1} &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
@@
 \midrule
-$\lambda$ = \bf{0} &        0.373 & 0.097 & 0.11  & 0.117 & 0.113 \\
-$\lambda$ = \bf{1e-5} &    0.373 & 0.103 & \bf{0.09}  & 0.09  & 0.09  \\
-$\lambda$ = \bf{1e-3} &    0.37  & 0.107 & 0.103 & 0.103 & 0.103 \\
+$\lambda$ = \bf{0} &        0.373 & 0.083 & 0.11  & 0.103 & 0.113 \\
+$\lambda$ = \bf{1e-5} &    0.373 & 0.087 & \bf{0.077} & 0.08  & 0.09  \\
+$\lambda$ = \bf{1e-3} &    0.377 & 0.097 & 0.097 & 0.093 & 0.093 \\
 $\lambda$ = \bf{1e-1} &    0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
@@
 
+\begin{table}
+\captionof{table}{Neural Net Performance on MNIST dataset for different choices of parameters} 
+\centering
+\begin{tabular}{llllll}
+\toprule
+& $n_1$ = 50 & $n_1$ = 100 & $n_1 = 150$ \\
+\midrule
+Training, $lr = .1$, full gradient descent, $\lambda$ = 0 & && \\
+Training, $lr = .1$, full gradient descent, $\lambda$= 1e-3 & && \\
+Training, $lr = .1$, full gradient descent, $\lambda$= 1e-2 & && \\
+Training, $lr = .1$, full gradient descent, $\lambda$= 1e-1 & && \\
+Validation, $lr = .1$, full gradient descent, $\lambda$ = 0 & && \\
+Validation, $lr = .1$, full gradient descent, $\lambda$= 1e-3 & && \\
+Validation, $lr = .1$, full gradient descent, $\lambda$= 1e-2 & && \\
+Validation, $lr = .1$, full gradient descent, $\lambda$= 1e-1 & && \\
+Training, $lr = .1$, SGD, $\lambda$ = 0 & && \\
+Training, $lr = .1$, SGD, $\lambda$= 1e-3 & && \\
+Training, $lr = .1$, SGD, $\lambda$= 1e-2 & && \\
+Training, $lr = .1$, SGD, $\lambda$= 1e-1 & && \\
+Validation, $lr = .1$, SGD, $\lambda$ = 0 & && \\
+Validation, $lr = .1$, SGD, $\lambda$= 1e-3 & && \\
+Validation, $lr = .1$, SGD, $\lambda$= 1e-2 & && \\
+Validation, $lr = .1$, SGD, $\lambda$= 1e-1 & && \\
+Training, $lr = .01$, full gradient descent, $\lambda$ = 0 & && \\
+Training, $lr = .01$, full gradient descent, $\lambda$= 1e-3 & && \\
+Training, $lr = .01$, full gradient descent, $\lambda$= 1e-2 & && \\
+Training, $lr = .01$, full gradient descent, $\lambda$= 1e-1 & && \\
+Validation, $lr = .01$, full gradient descent, $\lambda$ = 0 & && \\
+Validation, $lr = .01$, full gradient descent, $\lambda$= 1e-3 & && \\
+Validation, $lr = .01$, full gradient descent, $\lambda$= 1e-2 & && \\
+Validation, $lr = .01$, full gradient descent, $\lambda$= 1e-1 & && \\
+Training, $lr = .01$, SGD, $\lambda$ = 0 & && \\
+Training, $lr = .01$, SGD, $\lambda$= 1e-3 & && \\
+Training, $lr = .01$, SGD, $\lambda$= 1e-2 & && \\
+Training, $lr = .01$, SGD, $\lambda$= 1e-1 & && \\
+Validation, $lr = .01$, SGD, $\lambda$ = 0 & && \\
+Validation, $lr = .01$, SGD, $\lambda$= 1e-3 & && \\
+Validation, $lr = .01$, SGD, $\lambda$= 1e-2 & && \\
+Validation, $lr = .01$, SGD, $\lambda$= 1e-1 & && \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \end{document}

### 2015-11-11 HW3/tex/HW3.tex
[hw3] prettify tables.

--- 2015-11-10 HW3/tex/HW3.tex
@@
 \begin{equation}
-\frac{\partial J(w)}{\partial w_{jd}^{(1)}} = -\left ( \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(y_k^{(i)}(1-\sigma(a_k^{(2)}) + (1-y_k^{(i)})(\sigma(a_k^{(2)})\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\right ) + 2 \lambda w_{jd}^{(1)}.
+\frac{\partial J(w)}{\partial w_{jd}^{(1)}} =  \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(\sigma(a_k^{(2)}) -y_k^{(i)}\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right) + 2 \lambda w_{jd}^{(1)}.
 \end{equation}
 
+Let's now see how a neural network performs on two different toy data sets.
+We vary both the number of hidden nodes and the regularization parameter in our objective function, and use both full gradient descent and batch gradient descent.
+The two algorithms perform comparably - in the case of toy data set number 2, we find that the full gradient descent algorithm does a slightly better job at finding the optimal values of our parameter than stochastic gradient descent (classification error of 6.7\% as opposed to 9\%).
+Our changes in parameters have a much larger effect on the performance for toy dataset 2 than for our features in toy dataset 1.
+This is likely due to the lack of an easy way to separate examples of different classes in dataset 2.
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.48\linewidth}
+		\includegraphics[width=1\linewidth, height=2.7in]{t1dist.jpg}
+		\caption*{Toy Dataset \#1}
+	\end{minipage}
+	\begin{minipage}[b]{.48\linewidth}
+		\includegraphics[width=1\linewidth, height=2.7in]{t2dist.jpg}
+		\caption*{Toy Dataset \#2}
+	\end{minipage}
+	\caption{Distributions of different classes in $X1$, $X2$ for toy datasets}
+\end{figure}
+
+\begin{table}
+\captionof{table}{Neural Network Classification Error Rates With Various Parameters on Toy Datasets} 
+\centering
+\begin{tabular}{llllllll}
+\toprule
+Dataset & Method & $\lambda$ & $n_1$ = \bf{1} & $n_1$ =  \bf{3} & $n_1$ =  \bf{5} & $n_1$ =  \bf{7} & $n_1$ =  \bf{9} \\
+\midrule
+Toy Dataset 1 (Training) & full gradient descent & 0 &     0.007 & \bf{0}     & 0     & 0     & 0     \\
+Toy Dataset 1 (Training) & full gradient descent & 1e-5 &   0.007 & 0     & 0     & 0     & 0     \\
+Toy Dataset 1 (Training) & full gradient descent & 1e-3 &  0.093 & 0.003 & 0.003 & 0.003 & 0.003 \\
+Toy Dataset 1 (Training) & full gradient descent & 1e-1 &  0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\midrule
+Toy Dataset 1 (Validation) & full gradient descent & 0 &      0.01  & \bf{0.003} & 0.003 & 0.003 & 0.003 \\
+Toy Dataset 1 (Validation) & full gradient descent & 1e-5 &   0.01  & 0.003 & 0.003 & 0.003 & 0.003 \\
+Toy Dataset 1 (Validation) & full gradient descent & 1e-3 &  0.063 & 0.003 & 0.003 & 0.003 & 0.007 \\
+Toy Dataset 1 (Validation) & full gradient descent & 1e-1 &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\midrule
+Toy Dataset 1 (Training) & SGD & 0 &       \bf{0.003} & 0.003 & 0.003 & 0.003 & 0.003 \\
+Toy Dataset 1 (Training) & SGD & 1e-5 &   0.043 & 0     & 0.003 & 0.003 & 0     \\
+Toy Dataset 1 (Training) & SGD & 1e-3 &   0.117 & 0     & 0.003 & 0.003 & 0.003 \\
+Toy Dataset 1 (Training) & SGD & 1e-1 &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\midrule
+Toy Dataset 1 (Validation) & SGD & 0 &       \bf{0.003} & 0.003 & 0.01  & 0.007 & 0.003 \\
+Toy Dataset 1 (Validation) & SGD & 1e-5 &   0.033 & 0.007 & 0.003 & 0.007 & 0.003 \\
+Toy Dataset 1 (Validation) & SGD & 1e-3 &   0.097 & 0.013 & 0.013 & 0.013 & 0.013 \\
+Toy Dataset 1 (Validation) & SGD & 1e-1 &  0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\midrule
+Toy Dataset 2 (Training) & full gradient descent & 0 &      0.36  & 0.063 & 0.047 & 0.05  & 0.027 \\
+Toy Dataset 2 (Training) & full gradient descent & 1e-5 &   0.397 & 0.063 & 0.067 & 0.063 & 0.067 \\
+Toy Dataset 2 (Training) & full gradient descent & 1e-3 &  0.413 & \bf{0.06}  & 0.06  & 0.06  & 0.06  \\
+Toy Dataset 2 (Training) & full gradient descent & 1e-1 &  0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\midrule
+Toy Dataset 2 (Validation) & full gradient descent & 0 &      0.343 & 0.07  & 0.093 & 0.1   & 0.107 \\
+Toy Dataset 2 (Validation) & full gradient descent & 1e-5 &   0.39  & 0.073 & 0.07  & 0.073 & 0.073 \\
+Toy Dataset 2 (Validation) & full gradient descent & 1e-3 &  0.38  & \bf{0.067} & 0.067 & 0.067 & 0.067 \\
+Toy Dataset 2 (Validation) & full gradient descent & 1e-1 &  0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\midrule
+Toy Dataset 2 (Training) & SGD & 0 &       0.363 & 0.073 & 0.053 & 0.05  & 0.06  \\
+Toy Dataset 2 (Training) & SGD & 1e-5 &   0.363 & 0.07  & \bf{0.067} & 0.057 & 0.067 \\
+Toy Dataset 2 (Training) & SGD & 1e-3 &   0.363 & 0.07  & 0.07  & 0.07  & 0.07  \\
+Toy Dataset 2 (Training) & SGD & 1e-1 &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\midrule
+Toy Dataset 2 (Validation) & SGD & 0 &       0.373 & 0.083 & 0.11  & 0.103 & 0.113 \\
+Toy Dataset 2 (Validation) & SGD & 1e-5 &   0.373 & 0.087 & \bf{0.077} & 0.08  & 0.09  \\
+Toy Dataset 2 (Validation) & SGD & 1e-3 &   0.377 & 0.097 & 0.097 & 0.093 & 0.093 \\
+Toy Dataset 2 (Validation) & SGD & 1e-1 &   0.667 & 0.667 & 0.667 & 0.667 & 0.667 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\captionof{table}{Neural Net Classification Error Rates on MNIST dataset for different choices of parameters} 
+\centering
+\begin{tabular}{llllllllll}
+\toprule
+Dataset & $lr$ & method & $\lambda$ & $n_1$ = 50 & $n_1$ = 100 & $n_1 = 150$ \\
+\midrule
+Training & .1 & full gradient descent &  0 & 0     & 0     & \bf{0}     \\
+Training & .1 & full gradient descent & 1e-3 & 0.007 & 0.005 & 0.006 \\
+Training & .1 & full gradient descent & 1e-2 & 0.069 & 0.069 & 0.074 \\
+Training & .1 & full gradient descent & 1e-1 & 0.812 & 0.812 & 0.812 \\
+\midrule
+Validation & .1 & full gradient descent &  0 & 0.068 & 0.078 & \bf{0.062} \\
+Validation & .1 & full gradient descent & 1e-3 &  0.064 & 0.07  & 0.062 \\
+Validation & .1 & full gradient descent & 1e-2 & 0.072 & 0.078 & 0.078 \\
+Validation & .1 & full gradient descent & 1e-1 & 0.822 & 0.822 & 0.822 \\
+\midrule
+Training & .1 & SGD &  0 & 0.112 & 0.109 & \bf{0.106} \\
+Training & .1 & SGD & 1e-3 & 0.109 & 0.119 & 0.114 \\
+Training & .1 & SGD & 1e-2 & 0.13  & 0.132 & 0.131 \\
+Training & .1 & SGD & 1e-1 & 0.683 & 0.703 & 0.78  \\
+\midrule
+Validation & .1 & SGD &  0 & 0.09  & 0.1   & \bf{0.086} \\
+Validation & .1 & SGD & 1e-3 & 0.092 & 0.09  & 0.098 \\
+Validation & .1 & SGD & 1e-2 & 0.122 & 0.122 & 0.12  \\
+Validation & .1 & SGD & 1e-1 & 0.68  & 0.698 & 0.762 \\
+\midrule
+Training & .01 & full gradient descent &  0 & 0.084 & 0.064 & 0.057 \\
+Training & .01 & full gradient descent &  1e-3 & 0.076 & 0.067 & \bf{0.056} \\
+Training & .01 & full gradient descent &  1e-2 & 0.094 & 0.087 & 0.081 \\
+Training & .01 & full gradient descent & 1e-1 & 0.663 & 0.66  & 0.66  \\
+\midrule
+Validation & .01 & full gradient descent &  0 & 0.09  & 0.076 & 0.08  \\
+Validation & .01 & full gradient descent & 1e-3 & 0.102 & 0.074 & \bf{0.07}  \\
+Validation & .01 & full gradient descent & 1e-2 & 0.088 & 0.08  & 0.076 \\
+Validation & .01 & full gradient descent & 1e-1 & 0.674 & 0.672 & 0.672 \\
+\midrule
+Training &.01 & SGD &  0 & 0.14  & 0.122 & 0.115 \\
+Training &.01 & SGD & 1e-3 & 0.153 & 0.132 & \bf{0.124} \\
+Training &.01 & SGD & 1e-2 & 0.153 & 0.152 & 0.138 \\
+Training &.01 & SGD & 1e-1 & 0.629 & 0.545 & 0.511 \\
+\midrule
+Validation & .01 & SGD &  0 & 0.14  & 0.116 & 0.118 \\
+Validation & .01 & SGD & 1e-3 & 0.136 & 0.126 & \bf{0.108} \\
+Validation & .01 & SGD & 1e-2 & 0.128 & 0.142 & 0.124 \\
+Validation & .01 & SGD & 1e-1 & 0.622 & 0.548 & 0.504 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \end{document}

### 2015-11-11 HW3/code/mnist.py
more writing

@@ -22,7 +22,7 @@ except:
 
 def gridSearch(Xtr, YtrM, Xv, YvM, a):
     lamb = [0, .001, .01, .1]
-    h = [200, 250]
+    h = [50,100,150]
     besttr = 1
     besterr = 1
     bestl = 0

### 2015-11-11 HW3/tex/HW3.tex
tex updates.

--- 2015-11-11 HW3/tex/HW3.tex
@@
 Let us first briefly review the way that neural network works.
-A neural network consists of a variable number of ``hidden layers``.
-A weight matrix maps each of your data points to a value for each node in the hidden layer.
-We then evaluate some ``activation function`` on each node value (typically a sigmoid function or tanh function) to get a new set of ``inputs`` for the next layer of the neural network.
-We continue this process until we finally arrive at our output predictions.
-A schematic sketch of a neural net can be found below.
+Neural networks are inspired by how brains process information.
+In the brain we have interconnected neurons that produce electrochemical impulses that feed into other neurons.
+In a neural network, we try to mimic this architecture by using interconnected ``perceptrons`` or nodes.
+These nodes are organized into ``layers`` which are connected to nodes in other layers, but not to nodes in the same layer.
+
+In a feed-forward neural network (the specific type of neural network architecture we are working with).
+The outputs of each layer are fully connected to the next layer.
+These layer to layer connections are represented by a weight matrix.
+Functionally, the weight matrix produces a weighted linear combination of the outputs of a layer to use as the input for the next layer.
+This weighted linear combination is then fed into some nonlinear ``activation function`` on each node (typically a sigmoid function or tanh function) which produces the output of a node.
+Its important to note that the activation function needs to be nonlinear otherwise we would end up with a standard linear model (a linear combination of linear function is still a linear function).
+These outputs can then be fed into a new layer via the same procedure as above.
+We can arbitrarily choose how many of these interconnected layers we want to include in a neural network model.
 
@@
 
-Alternatively, a neural network may be specified as follows:
+However, a neural network model needs at the very least two layers, an input layer and an output layer.
+These layers are our data, where the input layer is composed of our features and the output layer is what we want to predict.
+In this sense, the simplest neural network  with a single node in the output layer is just a logistic regression.
+For networks with more than just the input layer and output layer, the intermediate layers are called ``hidden layers`` since they're not directly observed but are learned (much like a hidden state in a Hidden Markov Model).
+A schematic sketch of a neural net with a single hidden layer can be found above.
+Alternatively, this neural network may be mathematically specified as follows:
 
@@
 \begin{equation}
-z_{n}^{(1)} = \sigma_1(a_n^{(1)})
+z_{n}^{(1)} = f_1(a_n^{(1)})
 \end{equation}
@@
 \begin{equation}
-z_{k}^{(2)} = \sigma_2(a_k^{(2)})
+y_k = z_{k}^{(2)} = f_2(a_k^{(2)})
 \end{equation}
 
-The most trivial neural network has only one output layer with one node, and no hidden layers. This would correspond to vanilla logistic regression.
+Or in matrix notation:
+\begin{equation}
+a^{(1)} = X\cdot W^{(1)}	
+\end{equation}
+
+\begin{equation}
+z^{(1)} = f_1(a^{(1)})
+\end{equation}
+
+\begin{equation}
+a^{(2)} = z^{(1)}\cdot W^{(2)}	
+\end{equation}
+
+\begin{equation}
+y = z^{(2)} = f_1(a^{(2)})
+\end{equation}
 
@@
 \begin{equation}
-\frac{\partial J(w)}{\partial w_{jd}^{(1)}} =  \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(\sigma(a_k^{(2)}) -y_k^{(i)}\right) w_{kj}^{(2)} \left(\frac{1}{1+e^{-a_{jd}^{(1)}}}\right)\left(1-\frac{1}{1+e^{-a_{jd}^{(1)}}}\right) + 2 \lambda w_{jd}^{(1)}.
+\frac{\partial J(w)}{\partial w_{jd}^{(1)}} =  \sum_{i=1}^N x_i \sum_{k=1}^{K} \left(\sigma(a_k^{(2)}) -y_k^{(i)}\right) w_{kj}^{(2)} \left(\sigma(a_{jd}^{(1)})\right)\left(1-\sigma(a_{jd}^{(1)})\right) + 2 \lambda w_{jd}^{(1)}.
 \end{equation}
@@
 Let's now see how a neural network performs on two different toy data sets.
-We vary both the number of hidden nodes and the regularization parameter in our objective function, and use both full gradient descent and batch gradient descent.
+We vary both the number of hidden nodes and the regularization parameter in our objective function, and use both full gradient descent and mini-batch stochastic gradient descent (with a batch size of 10).
+In particular, we use mini-batch SGD as opposed to full SGD (where batch size is 1) since mini-batch generally has better convergence properties.
 The two algorithms perform comparably - in the case of toy data set number 2, we find that the full gradient descent algorithm does a slightly better job at finding the optimal values of our parameter than stochastic gradient descent (classification error of 6.7\% as opposed to 9\%).
-Our changes in parameters have a much larger effect on the performance for toy dataset 2 than for our features in toy dataset 1.
+While full gradient descent does perform better, its important to note the SGD was significantly faster.
+This should be expected since full-batch uses the entire dataset to perform an update while SGD only uses a fraction, meaning that when we use SGD as opposed to full gradient descent we lose some of the nice convergence properties to gain faster updates.
+
+It's important to note that our changes in parameters have a much larger effect on the performance for toy dataset 2 than for our features in toy dataset 1.
 This is likely due to the lack of an easy way to separate examples of different classes in dataset 2.
@@
 
-\begin{table}
-\captionof{table}{Neural Net Performance on MNIST dataset for different choices of parameters} 
-\centering
-\begin{tabular}{llllll}
-\toprule
-& $n_1$ = 50 & $n_1$ = 100 & $n_1 = 150$ \\
-\midrule
-Training, $lr = .1$, full gradient descent, $\lambda$ = 0 & && \\
-Training, $lr = .1$, full gradient descent, $\lambda$= 1e-3 & && \\
-Training, $lr = .1$, full gradient descent, $\lambda$= 1e-2 & && \\
-Training, $lr = .1$, full gradient descent, $\lambda$= 1e-1 & && \\
-Validation, $lr = .1$, full gradient descent, $\lambda$ = 0 & && \\
-Validation, $lr = .1$, full gradient descent, $\lambda$= 1e-3 & && \\
-Validation, $lr = .1$, full gradient descent, $\lambda$= 1e-2 & && \\
-Validation, $lr = .1$, full gradient descent, $\lambda$= 1e-1 & && \\
-Training, $lr = .1$, SGD, $\lambda$ = 0 & && \\
-Training, $lr = .1$, SGD, $\lambda$= 1e-3 & && \\
-Training, $lr = .1$, SGD, $\lambda$= 1e-2 & && \\
-Training, $lr = .1$, SGD, $\lambda$= 1e-1 & && \\
-Validation, $lr = .1$, SGD, $\lambda$ = 0 & && \\
-Validation, $lr = .1$, SGD, $\lambda$= 1e-3 & && \\
-Validation, $lr = .1$, SGD, $\lambda$= 1e-2 & && \\
-Validation, $lr = .1$, SGD, $\lambda$= 1e-1 & && \\
-Training, $lr = .01$, full gradient descent, $\lambda$ = 0 & && \\
-Training, $lr = .01$, full gradient descent, $\lambda$= 1e-3 & && \\
-Training, $lr = .01$, full gradient descent, $\lambda$= 1e-2 & && \\
-Training, $lr = .01$, full gradient descent, $\lambda$= 1e-1 & && \\
-Validation, $lr = .01$, full gradient descent, $\lambda$ = 0 & && \\
-Validation, $lr = .01$, full gradient descent, $\lambda$= 1e-3 & && \\
-Validation, $lr = .01$, full gradient descent, $\lambda$= 1e-2 & && \\
-Validation, $lr = .01$, full gradient descent, $\lambda$= 1e-1 & && \\
-Training, $lr = .01$, SGD, $\lambda$ = 0 & && \\
-Training, $lr = .01$, SGD, $\lambda$= 1e-3 & && \\
-Training, $lr = .01$, SGD, $\lambda$= 1e-2 & && \\
-Training, $lr = .01$, SGD, $\lambda$= 1e-1 & && \\
-Validation, $lr = .01$, SGD, $\lambda$ = 0 & && \\
-Validation, $lr = .01$, SGD, $\lambda$= 1e-3 & && \\
-Validation, $lr = .01$, SGD, $\lambda$= 1e-2 & && \\
-Validation, $lr = .01$, SGD, $\lambda$= 1e-1 & && \\
-\bottomrule
-\end{tabular}
-\end{table}
+We are now interested in testing out neural network code on a real dataset.
+In our case, we're using a subset the very popular MNIST data collected by Yann Lecunn.
+This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent.
+For our features, each 28x28 images into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
+
+When training
 

### 2015-11-11 HW3/tex/HW3.tex

--- 2015-11-11 HW3/tex/HW3.tex
@@
 
+We are now interested in testing out neural network code on a real dataset.
+In our case, we're using a subset the very popular MNIST data collected by Yann Lecunn.
+This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent.
+For our features, each 28x28 images into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
+
+When training
+
 \end{document}

### 2015-11-11 HW3/tex/HW3.tex
[hw3] boo.

@@
 
-When training
+When training on the MNIST data, we capped the number of iterations of our gradient descent (for both full batch and SGD) so that to convergence wouldn't take forever.
+Furthermore, in practice, this may serves as a way of preventing overfitting as long as we are still getting reasonably close.
+
+Again, we observe that full gradient descent performs better than SGD.
+However, in this case, we see the performance difference to be much more substantial than what we saw in the toy data.
+This is mostly the result of the 5000 iteration cap we imposed on the gradient descent procedure.
+Since we often hit the cap in our training, full-batch was considerably closer to convergence compared to the SGD case as the each full-batch update more reliably and steadily minimizes the loss.
+Again, however each SGD update was considerably faster than a full-batch update.
+If we allowed the SGD to continue a little longer, we probably the results from our SGD probably would've considerably closer to the full-batch case.
 

### 2015-11-11 HW3/tex/HW3.tex
[hw3] nn plot.

--- 2015-11-11 HW3/tex/HW3.tex
@@
 It's important to note that our changes in parameters have a much larger effect on the performance for toy dataset 2 than for our features in toy dataset 1.
-This is likely due to the lack of an easy way to separate examples of different classes in dataset 2.
+This is likely due to the fact that the second toy dataset is not as easily separable, as seen in the plots below.
+While a researcher can essentially achieve 100\% classification accuracy by drawing vertical lines to classify toy dataset \#1, there is overlap in the locations of our classes in toy dataset \#2.
+Consequently, the regularization constants have more influence on performance.
 
@@
 
+We are now interested in testing out our neural network code on a real dataset.
+In our case, we're using a subset of the very popular MNIST data collected by Yann Lecunn.
+This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent.
+For our features, each 28x28 image is transformed into a feature vector of 784 where each of these features represents the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
+Our dataset only consists of 6 distinct digits.
+
+We again train both full-batch and stochastic gradient descent (batch size = 25) neural networks on the MNIST data, with a variety of different learning rates, regularization parameters and numbers of hidden nodes.
+The classification errors on both training and validation sets are found in Table 2, with best results in bold.
+
+As expected, we see that SGD performs slightly worse than full-batch gradient descent, since it does not overfit as much.
+Our data seems to show that increasing the number of hidden nodes strictly increasing performance - this is not surprising, although we imagine that as you continue to increase $n_1$, you begin to get diminishing returns for increasing computational complexity.
+Decreasing the step size (i.e., smaller learning rate) also seems to lead us to achieve slightly better learning rates, although this causes the algorithm to require many more iterations to converge.
+A chart of the neural network classification error on the MNIST validation set can be found in Figure 2.
+
 \begin{table}
@@
 
-We are now interested in testing out neural network code on a real dataset.
-In our case, we're using a subset the very popular MNIST data collected by Yann Lecunn.
-This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent.
-For our features, each 28x28 images into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
-
-When training
+\begin{figure}
+\center
+\includegraphics[height=4in, width=5in]{nnperf.jpg}
+\caption*{Figure 2: Neural Network Classification Error on MNIST Validation Sets}
+\end{figure}
 

### 2015-11-11 HW3/tex/HW3.tex

--- 2015-11-11 HW3/tex/HW3.tex
@@
 
+We are now interested in testing out neural network code on a real dataset.
+In our case, we're using a subset the very popular MNIST data collected by Yann Lecunn.
+This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent.
+For our features, each 28x28 images into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
+Our dataset only consists of 6 distinct digits.
+
+We again train both full-batch and stochastic gradient descent (batch size = 25) neural networks on the MNIST data, with a variety of different learning rates, regularization parameters and numbers of hidden nodes.
+The classification errors on both training and validation sets are found in Table 2, with best results in bold.
+
+As expected, we see that SGD performs slightly worse than full-batch gradient descent, since it does not overfit as much.
+Our data seems to show that increasing the number of hidden nodes strictly increasing performance - this is not surprising, although we imagine that as you continue to increase $n_1$, you begin to get diminishing returns for increasing computational complexity.
+Decreasing the step size (i.e., smaller learning rate) also seems to lead us to achieve slightly better learning rates, although this causes the algorithm to require many more iterations to converge.
+A chart of the neural network classification error on the MNIST validation set can be found in Figure 2.
+
+When training on the MNIST data, we capped the number of iterations of our gradient descent (for both full batch and SGD) so that to convergence wouldn't take forever.
+Furthermore, in practice, this may serves as a way of preventing overfitting as long as we are still getting reasonably close.
+
+Again, we observe that full gradient descent performs better than SGD.
+However, in this case, we see the performance difference to be much more substantial than what we saw in the toy data.
+This is mostly the result of the 5000 iteration cap we imposed on the gradient descent procedure.
+Since we often hit the cap in our training, full-batch was considerably closer to convergence compared to the SGD case as the each full-batch update more reliably and steadily minimizes the loss.
+Again, however each SGD update was considerably faster than a full-batch update.
+If we allowed the SGD to continue a little longer, we probably the results from our SGD probably would've considerably closer to the full-batch case.
+
 \begin{table}
@@
 
-We are now interested in testing out neural network code on a real dataset.
-In our case, we're using a subset the very popular MNIST data collected by Yann Lecunn.
-This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent.
-For our features, each 28x28 images into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
-
-When training on the MNIST data, we capped the number of iterations of our gradient descent (for both full batch and SGD) so that to convergence wouldn't take forever.
-Furthermore, in practice, this may serves as a way of preventing overfitting as long as we are still getting reasonably close.
-
-Again, we observe that full gradient descent performs better than SGD.
-However, in this case, we see the performance difference to be much more substantial than what we saw in the toy data.
-This is mostly the result of the 5000 iteration cap we imposed on the gradient descent procedure.
-Since we often hit the cap in our training, full-batch was considerably closer to convergence compared to the SGD case as the each full-batch update more reliably and steadily minimizes the loss.
-Again, however each SGD update was considerably faster than a full-batch update.
-If we allowed the SGD to continue a little longer, we probably the results from our SGD probably would've considerably closer to the full-batch case.
+\begin{figure}
+\center
+\includegraphics[height=4in, width=5in]{nnperf.jpg}
+\caption*{Figure 2: Neural Network Classification Error on MNIST Validation Sets}
+\end{figure}
 

### 2015-11-11 HW3/code/toyplot.R
[hw3] final draft.

@@ -1,5 +1,6 @@
 library(ggplot2)
 library(dplyr)
+library(scales)
 
 toy_test_1 <- read.csv('~/Downloads/toy_multiclass_1_testCLEAN.csv', head=T)
 toy_validate_1 <- read.csv('~/Downloads/toy_multiclass_1_validateCLEAN.csv', head=T)
@@ -31,11 +32,22 @@ toy_2_complete <- rbind(toy_validate_2 %>%
 
 t1 <- ggplot(toy_1_complete, aes(x=as.numeric(toy.data.1), y=as.numeric(toy.data.2), 
                        color=Class, shape=Dataset)) + geom_point() + xlab('X1') + ylab('X2') + 
-  ggtitle('Distribution of Toy Data Set 1 in 2-dimensions')
+  ggtitle('Scatterplot of Toy Data Set 1 in 2-dimensions')
 ggsave(t1, file='../tex/t1dist.jpg')
 
 t2 <- ggplot(toy_2_complete, aes(x=as.numeric(toy.data.1), y=as.numeric(toy.data.2), 
                            color=Class, shape=Dataset)) + geom_point() + xlab('X1') + ylab('X2') +
-  ggtitle('Distribution of Toy Data Set 2 in 2-dimensions')
-ggsave(t2, file='t2dist.jpg')
+  ggtitle('Scatterplot of Toy Data Set 2 in 2-dimensions')
+ggsave(t2, file='../tex/t2dist.jpg')
 
+nn_performance <- read.csv('nnperformance.csv', head=T)
+head(nn_performance)
+
+nnperf <- nn_performance %>% 
+  mutate(`Learning Rate` = as.factor(lr),
+         `Regularization` = as.factor(lambda)) %>%
+  ggplot(aes(x=n, y=error, color=`Regularization`)) + 
+  geom_line(aes(linetype=`Learning Rate`), size=1) + scale_y_continuous(labels=percent) + facet_wrap(~method) + 
+  xlab('Hidden Units') + ylab('Classification Error Rate') + 
+  ggtitle('Neural Network Classification Error on MNIST Validation Dataset')
+ggsave(nnperf, file='../tex/nnperf.jpg')

### 2015-11-11 HW3/tex/HW3.tex

--- 2015-11-11 HW3/tex/HW3.tex
@@
 
-In some cases, we might want to use a neural network rather than another algorithm, such as SVM or logistic regression, in order to be somewhat more agnostic about the parametric form of the data-generating process.
+We might want to use a neural network rather than another algorithm, such as SVM or logistic regression, in order to be somewhat more agnostic about the parametric form of the data-generating process.
 In this case, we investigate the use of neural networks, first on some toy training data sets, and subsequently on a subset of the popular MNIST dataset.
 
-Let us first briefly review the way that neural network works.
+Let us first briefly review the way that neural networks works.
 Neural networks are inspired by how brains process information.
 In the brain we have interconnected neurons that produce electrochemical impulses that feed into other neurons.
-In a neural network, we try to mimic this architecture by using interconnected ``perceptrons`` or nodes.
-These nodes are organized into ``layers`` which are connected to nodes in other layers, but not to nodes in the same layer.
+In a neural network, we try to mimic this architecture by using interconnected ``perceptrons" or nodes.
+These nodes are organized into ``layers" which are connected to nodes in other layers, but not to nodes in the same layer.
 
-In a feed-forward neural network (the specific type of neural network architecture we are working with).
-The outputs of each layer are fully connected to the next layer.
+In a feed-forward neural network (the specific type of neural network architecture we are working with), the outputs of each layer are fully connected to the next layer.
 These layer to layer connections are represented by a weight matrix.
 Functionally, the weight matrix produces a weighted linear combination of the outputs of a layer to use as the input for the next layer.
-This weighted linear combination is then fed into some nonlinear ``activation function`` on each node (typically a sigmoid function or tanh function) which produces the output of a node.
-Its important to note that the activation function needs to be nonlinear otherwise we would end up with a standard linear model (a linear combination of linear function is still a linear function).
+This weighted linear combination is then fed into some nonlinear ``activation function" on each node (typically a sigmoid function or tanh function) which produces the output of a node.
+Its important to note that the activation function needs to be nonlinear, otherwise we would end up with a standard linear model (a linear combination of linear functions is still a linear function).
 These outputs can then be fed into a new layer via the same procedure as above.
@@
 
-However, a neural network model needs at the very least two layers, an input layer and an output layer.
-These layers are our data, where the input layer is composed of our features and the output layer is what we want to predict.
+Neural network models at the very least need two layers, an input layer and an output layer.
+These layers are constructed from  data, where the input layer is composed of features and the output layer is the label.
 In this sense, the simplest neural network  with a single node in the output layer is just a logistic regression.
-For networks with more than just the input layer and output layer, the intermediate layers are called ``hidden layers`` since they're not directly observed but are learned (much like a hidden state in a Hidden Markov Model).
+For networks with more than just the input layer and output layer, the intermediate layers are called ``hidden layers," since they're not directly observed (much like a hidden state in a Hidden Markov Model).
 A schematic sketch of a neural net with a single hidden layer can be found above.
@@
 
-Or in matrix notation:
+\noindent Or in matrix notation:
+
 \begin{equation}
@@
 
-\noindent where $l(w)$ is
+\noindent where $l(w)$ is cross entropy loss:
 
@@
 
-In order to calculate the gradient as a function of $w_1$ and $w_2$, we use the two gradient calculation expressions below:
+To make our gradient descent faster, we want to use the analytical gradients of $w_1$ and $w_2$ for our gradient descent procedure. These gradients are calculated below.
 
@@
 \end{equation}
-
-Let's now see how a neural network performs on two different toy data sets.
-We vary both the number of hidden nodes and the regularization parameter in our objective function, and use both full gradient descent and mini-batch stochastic gradient descent (with a batch size of 10).
-In particular, we use mini-batch SGD as opposed to full SGD (where batch size is 1) since mini-batch generally has better convergence properties.
-The two algorithms perform comparably - in the case of toy data set number 2, we find that the full gradient descent algorithm does a slightly better job at finding the optimal values of our parameter than stochastic gradient descent (classification error of 6.7\% as opposed to 9\%).
-While full gradient descent does perform better, its important to note the SGD was significantly faster.
-This should be expected since full-batch uses the entire dataset to perform an update while SGD only uses a fraction, meaning that when we use SGD as opposed to full gradient descent we lose some of the nice convergence properties to gain faster updates.
-
-It's important to note that our changes in parameters have a much larger effect on the performance for toy dataset 2 than for our features in toy dataset 1.
-This is likely due to the fact that the second toy dataset is not as easily separable, as seen in the plots below.
-While a researcher can essentially achieve 100\% classification accuracy by drawing vertical lines to classify toy dataset \#1, there is overlap in the locations of our classes in toy dataset \#2.
-Consequently, the regularization constants have more influence on performance.
-
-\begin{figure}[ht]
-	\centering
-	\begin{minipage}[b]{.48\linewidth}
-		\includegraphics[width=1\linewidth, height=2.7in]{t1dist.jpg}
-		\caption*{Toy Dataset \#1}
-	\end{minipage}
-	\begin{minipage}[b]{.48\linewidth}
-		\includegraphics[width=1\linewidth, height=2.7in]{t2dist.jpg}
-		\caption*{Toy Dataset \#2}
-	\end{minipage}
-	\caption{Distributions of different classes in $X1$, $X2$ for toy datasets}
-\end{figure}
 
@@
 
-We are now interested in testing out our neural network code on a real dataset.
+Let's now see how a neural network performs on two different toy data sets.
+We vary both the number of hidden nodes and the regularization parameter in our objective function, and use both full gradient descent and mini-batch stochastic gradient descent (with a batch size of 10).
+In particular, we use mini-batch SGD as opposed to full SGD (where batch size is 1) since mini-batch  has better convergence properties.
+The two algorithms perform comparably - in the case of toy data set number 2, we find that the full gradient descent algorithm does a slightly better job at finding the optimal values of our parameter than stochastic gradient descent (classification error of 6.7\% as opposed to 9\%).
+While full gradient descent does perform better, its important to note the SGD was significantly faster.
+This should be expected since full-batch uses the entire dataset to perform an update while SGD only uses a fraction, meaning that when we use SGD as opposed to full gradient descent we lose some of the nice convergence properties to gain faster updates.
+
+It's important to note that our changes in parameters have a much larger effect on the performance for toy dataset 2 than for our features in toy dataset 1.
+This is likely due to the fact that the second toy dataset is not linearly separable, as seen in the plots below.
+While a researcher can essentially achieve 100\% classification accuracy by drawing vertical lines to classify toy dataset \#1, there is overlap in the locations of our classes in toy dataset \#2.
+Consequently, the regularization constants have more influence on performance.
+
+\begin{figure}[ht]
+	\centering
+	\begin{minipage}[b]{.48\linewidth}
+		\includegraphics[width=1\linewidth, height=2.7in]{t1dist.jpg}
+		\caption*{Toy Dataset \#1}
+	\end{minipage}
+	\begin{minipage}[b]{.48\linewidth}
+		\includegraphics[width=1\linewidth, height=2.7in]{t2dist.jpg}
+		\caption*{Toy Dataset \#2}
+	\end{minipage}
+	\caption{Scatterplots of different classes in $X1$, $X2$ for toy datasets}
+\end{figure}
+
+We are now interested in testing out neural network code on a real dataset.
 In our case, we're using a subset of the very popular MNIST data collected by Yann Lecunn.
 This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent.
-For our features, each 28x28 image is transformed into a feature vector of 784 where each of these features represents the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
+For our features, each 28x28 image is flattened into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).
 Our dataset only consists of 6 distinct digits.
 
+When training on the MNIST data, we capped the number of iterations of our gradient descent (for both full batch and SGD) so that to convergence wouldn't take forever.
+Furthermore, in practice, this may serves as a way of preventing overfitting as long as we are still getting reasonably close.
 We again train both full-batch and stochastic gradient descent (batch size = 25) neural networks on the MNIST data, with a variety of different learning rates, regularization parameters and numbers of hidden nodes.
 The classification errors on both training and validation sets are found in Table 2, with best results in bold.
-
-As expected, we see that SGD performs slightly worse than full-batch gradient descent, since it does not overfit as much.
-Our data seems to show that increasing the number of hidden nodes strictly increasing performance - this is not surprising, although we imagine that as you continue to increase $n_1$, you begin to get diminishing returns for increasing computational complexity.
-Decreasing the step size (i.e., smaller learning rate) also seems to lead us to achieve slightly better learning rates, although this causes the algorithm to require many more iterations to converge.
-A chart of the neural network classification error on the MNIST validation set can be found in Figure 2.
 
@@
 
-\begin{figure}
+As expected, we see that SGD performs slightly worse than full-batch gradient descent, since it does not overfit as much.
+Our data seems to show that increasing the number of hidden nodes strictly increasing performance - this is not surprising, although we imagine that as you continue to increase $n_1$, you begin to get diminishing returns for increasing computational complexity.
+Decreasing the step size (i.e., smaller learning rate) also seems to lead us to achieve slightly better learning rates, although this causes the algorithm to require many more iterations to converge.
+A chart of the neural network classification error on the MNIST validation set can be found in Figure 2.
+
+Note, however that we see the performance difference to be much more substantial than what we saw in the toy data.
+This is mostly the result of the 5000 iteration cap we imposed on the gradient descent procedure.
+Since we often hit the cap in while training, full-batch was considerably closer to convergence compared to the SGD as the each full-batch update more reliably and steadily decreases the loss.
+Again, however each SGD update was considerably faster than a full-batch update.
+If we allowed the SGD to continue a little longer, the results from our SGD probably would've considerably closer to the full-batch case.
+
+\begin{figure}[ht]
 \center

### 2015-11-11 HW2/code/problem1/problem1.py
[hw2] problem 1.

@@ -1,0 +1,100 @@
+from numpy import *
+from plotBoundary import *
+import scipy.optimize
+import matplotlib as mpl
+from ggplot import *
+import pandas as pd
+
+# Plotting Params
+mpl.rcParams["figure.figsize"] = "2.25, 2"
+pltsize = {"figure.figsize": "2.25, 2"}
+
+# Logistic Regression Loss Function
+def LRLoss(w, X, Y, l=0):
+    return sum(log(1+exp(-hstack(Y)*(dot(X,w[1:len(w)])+w[0]))))+l*dot(w[1:len(w)], w[1:len(w)])
+
+# Prediction Function
+def LRPredict(w, X):
+    return 1/(1+exp(-(dot(X,w[1:len(w)])+w[0])))
+
+# Data Loading Function
+def loadData(name, typ):
+    try:
+        data = loadtxt('/Users/dholtz/Downloads/hw2_resources/data/data_'+name+'_'+typ+'.csv')
+    try:
+        data = loadtxt('/Users/mfzhao/Downloads/hw2_resources/data/data_'+name+'_'+typ+'.csv')
+    X = data[:,0:2]
+    Y = data[:,2:3]
+    return X, Y
+
+# Training Function
+def Train(X, Y, l=0):
+    w0 = random.randn(3)
+    w_opt = scipy.optimize.fmin_bfgs(LRLoss, w0, args=(X,Y,l))
+    return w_opt
+
+# Classification Error Function
+def classifyErr(py, Y, db):
+    yhat = ones(len(py))
+    for i in range(len(py)):
+        if py[i] < db:
+            yhat[i] = -1
+    return 1-mean(Y.flatten() == yhat)
+
+# Classification Error vs Decision Boundary Plotting
+def plotCEDB(w, X, Y, title):
+    global pltsize
+    axis = array(linspace(0,1,101))
+    out = zeros(len(axis))
+    py = LRPredict(w,X)
+    for i in range(len(axis)):
+        out[i] = classifyErr(py, Y, axis[i])
+    DF = pd.DataFrame({'Decision Boundary': pd.Series(axis),'Classification Error': pd.Series(out)})
+    print ggplot(DF, aes(x='Decision Boundary', y='Classification Error')) + geom_line(size=4) + ggtitle(title) + theme_matplotlib(rc=pltsize, matplotlib_defaults=False) 
+
[…]

### 2015-11-12 Project/Proposal/proposal.tex

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry} 
\usepackage{enumerate, xfrac, color, graphicx}
\usepackage{amsmath,amsthm,amssymb,amsfonts,mathabx}
\usepackage{booktabs}
\usepackage{bbm}
\usepackage{caption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{pifont}
\usepackage{listings, courier}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\lstset{breaklines=true, basicstyle=\small\ttfamily, language=R, backgroundcolor=\color{highlight}, stepnumber=5}

\definecolor{highlight}{RGB}{248,248,248}

\begin{document}
	\title{Proposal: Predicting New  York Times Article Readership}
	\date{November 11, 2015}
	\author{Dave Holtz, Jeremy Yang, Michael Zhao}
	\maketitle

\subsubsection*{Part I: Description of Project}

Our group has access to micro-level readership and sharing data from the New York Times from April 3, 2013 to October 31st, 2013. We are interested in using content and contextual data to predict how many page views a given article will receive. This task itself should be achievable using relatively straightforward regression techniques. However, the real challenge of this project lies in the feature extraction. A 2012 paper by Berger and Milkman \cite{berger} found that the virality of online content is related to the positive or negative valence of the content of the article. They framed the question as a classification problem, wherein they extract numerous features from an article and trained a model to predict whether or not that article would appear on the New York Times most e-mailed list. The feature extraction methods they use provide a great baseline for comparison, and also a great starting point for our research. We expect that with more granular data and more sophisticated feature extraction techniques, we can achieve even better predictive accuracy for a very similar (and arguably more important) task.

We are currently considering the following features (and methodologies to extract them) for inclusion in our model:
\begin{itemize}
\item Sentiment analysis (Naive Bayes text classification and tf-idf weighting)
\item Contextual data (day-of-week and time of publication, article topic, article author, Flesch-Kincaid Reading Ease, etc.)
\item Article uniqueness (determined using a Neural Network based language model with word vectors)
\end{itemize}

Note: Michael is also working on this project for Natural Language Processing (6.864). Dave and Jeremy are only enrolled in Machine Learning.

\subsubsection*{Part II: Timeline and Division of Labor}

Timeline:
\begin{itemize}
\item Nov. 12 - Nov. 19: Scrape New York Times article text, Aggregate readership data and collect non-text contextual data
\item Nov. 19 - Nov. 25: Perform sentiment analysis on text, begin training of neural network language model
\item Nov. 25 - Dec. 2: Complete training of neural network language model, train readership regression model
\item Dec. 3 - Dec. 8: Tweak readership model, write report, do additional fixes and optimizations as necessary
\end{itemize}

Division of Labor:
\begin{itemize}
\item Dave: Scrape New York Times article text, neural network language model, sentiment analysis, contribute to report, last minute optimizations
\item Jeremy: Scrape New York Times article text, sentiment analysis, readership model, contribute to report, last minute optimizations
\item Michael: Non-text contextual data and readership data aggregation, neural network language model (primary contributor), readership model, contribute to report, last minute optimizations
\end{itemize}

\subsubsection*{Part III: Risks}

This project brings with it a few risks. They are listed below.

\begin{itemize}
\item The neural network language model proves very difficult to train, both in that it can be complicated to train for programmers, and require significant computational resources / time. Although we are confident we can get code to function, it may not work well given the amount of time necessary to both train the neural network and find appropriate parameter values.
\item Data sanitization could more time-consuming than anticipated. We are using data that comes directly from the New York Times database, as well as text data scraped from the internet. This data is real and messy. It may take longer than expected to extract useful features from it, and certain features may prove infeasible to extract.
\item Although we plan on using Berger and Milkman's result as a rough guideline, their prediction task was fundamentally different than ours, as their labels are categorical rather than numerical. We may find it difficult to draw direct comparisons between their results and ours, and may need to resort to a different (and simpler) baseline of comparison.
\end{itemize}

 \begin{thebibliography}{1}

 \bibitem{berger} Berger, Jonah, and Katherine L. Milkman. "What makes online content viral?." {\em Journal of marketing research} 49.2 (2012): 192-205.
APA

 \end{thebibliography}

\end{document}

### 2015-11-12 Project/Proposal/proposal.tex
edits

@@
 \item Sentiment analysis (Naive Bayes text classification and tf-idf weighting)
-\item Contextual data (day-of-week and time of publication, article topic, article author, Flesch-Kincaid Reading Ease, etc.)
-\item Article uniqueness (determined using a Neural Network based language model with word vectors)
+\item Contextual data (day-of-week and time of publication, article topic, author characteristics, Flesch-Kincaid Reading Ease, etc.)
+\item Article uniqueness (determined by a Neural Network based language model using word vectors)
 \end{itemize}
@@
 \begin{itemize}
-\item Nov. 12 - Nov. 19: Scrape New York Times article text, Aggregate readership data and collect non-text contextual data
+\item Nov. 12 - Nov. 19: Scrape and clean New York Times article text, Aggregate readership data and collect non-text contextual data
 \item Nov. 19 - Nov. 25: Perform sentiment analysis on text, begin training of neural network language model
@@
 \item Jeremy: Scrape New York Times article text, sentiment analysis, readership model, contribute to report, last minute optimizations
-\item Michael: Non-text contextual data and readership data aggregation, neural network language model (primary contributor), readership model, contribute to report, last minute optimizations
+\item Michael: Non-text contextual data extraction and readership data aggregation, neural network language model (primary contributor), readership model, contribute to report, last minute optimizations
 \end{itemize}
@@
 Although we are confident we can get code to function, it may not work well given the amount of time necessary to both train the neural network and find appropriate parameter values.
-\item Data sanitization could more time-consuming than anticipated.
+\item Data sanitization could be more time-consuming than anticipated.
 We are using data that comes directly from the New York Times database, as well as text data scraped from the internet.
@@
 \item Although we plan on using Berger and Milkman's result as a rough guideline, their prediction task was fundamentally different than ours, as their labels are categorical rather than numerical.
+Furthermore, they include some features that we are unable to replicate.
 We may find it difficult to draw direct comparisons between their results and ours, and may need to resort to a different (and simpler) baseline of comparison.

### 2015-12-05 Project/gender_guessing/add_genders.R
[project] add code to guess genders.

@@ -1,0 +1,29 @@
+library(gender)
+
+first_name <- function(x) {
+  split <- strsplit(x, ' ')
+  first_name <- split[[1]][1]
+  return (first_name)
+}
+
+male_prop_for_df <- function(x) {
+  predicted_gender <- gender(x[['first_name']], years = c(1935, 1997), method="ssa")
+  return(predicted_gender$proportion_male)
+}
+
+female_prop_for_df <- function(x) {
+  predicted_gender <- gender(x[['first_name']], years = c(1935, 1997), method="ssa")
+  return(predicted_gender$proportion_female)
+}
+
+authors <- read.delim('author_list.tsv', head=T)
+authors$author_name <- as.character(authors$author_name)
+authors$first_name <- apply(authors, 1, FUN=first_name)
+authors$prop_male <- apply(authors, 1, FUN=male_prop_for_df)
+authors$prop_female <- apply(authors, 1, FUN=female_prop_for_df)
+
+authors$prop_male <- as.numeric(authors$prop_male)
+authors$prop_female <- as.numeric(authors$prop_female)
+
+write.csv(authors, file='authors_with_genders.csv')
+

### 2015-12-05 Project/gender_guessing/parse_names.py

@@ -1,0 +1,35 @@
+import pickle
+import pandas
+import itertools
+import requests
+from bs4 import BeautifulSoup
+import time
+
+def get_popularity(author):
+	print author
+	r = requests.get('http://www.bing.com/search',
+	                 params={'q':'"'+author+'"',
+	                         "tbs":"li:1"}
+	                )
+	
+	soup = BeautifulSoup(r.text, "html.parser")
+	raw_results_text = soup.find('div',{'id':'b_tween'}).text
+	time.sleep(30)
+	return raw_results_text.replace(' resultsAny time', '').replace(',', '').replace(' ', '')
+	
+def remove_quotes(string):
+	return string.replace('"', '')
+
+df = pandas.read_pickle('/Users/dholtz/Downloads/articleData.pkl')
+
+authors = list(itertools.chain(*df['authors']))
+unique_authors = set(authors)
+
+authors_df = pandas.DataFrame(list(unique_authors), columns = ['author_name'])
+authors_df['author_name'] = authors_df.author_name.apply(remove_quotes)
+authors_df['author_name'] = authors_df.author_name.apply(remove_quotes)
+authors_df.to_csv('author_list.csv', index=False, encoding='utf-8')
+
+popularities = []
+for i in authors_df['author_name']:
+	popularities += get_popularity(i)

### 2015-12-05 Project/nytData_parsing/consolidate.py
data parser

@@ -1,0 +1,58 @@
+import pandas as pd
+import numpy as np
+
+def addToCell(cell, alist):
+    for i in alist:
+        cell.append(i)
+
+path = '/Users/mfzhao/NYT NLP/Data/'
+df1 = pd.read_pickle(path + '1.pkl')
+df2 = pd.read_pickle(path + '2.pkl')
+df3 = pd.read_pickle(path + '3.pkl')
+
+url_list = open(path+'article_list.txt','r').read().splitlines()
+n = len(url_list)
+col_list = ['id', 'count', 'publishedDate', 'headline', 'authors', 'typeOfMaterial', 'desk', 'type', 'section', 'desFacet', 'perFacet', 'orgFacet', 'geoFacet']
+df = pd.DataFrame({'id': range(n), 'count': np.zeros(n), 'publishedDate': np.zeros(n), 'headline': ['None'] * n, 'authors': [[] for _ in range(n)], 'typeOfMaterial': ['None'] * n, 'desk': ['None'] * n, 'type': ['None'] * n, 'section' : ['None'] * n, 'desFacet': [[] for _ in range(n)], 'perFacet': [[] for _ in range(n)], 'orgFacet': [[] for _ in range(n)], 'geoFacet': [[] for _ in range(n)]}, index = url_list, columns = col_list)
+
+for url in url_list:
+    df.loc[url,'count'] = df1.loc[url,'count'] + df2.loc[url,'count'] + df3.loc[url,'count']
+    if df1.loc[url,'count'] != 0:
+        df.loc[url,'headline']       = df1.loc[url,'headline']
+        df.loc[url,'publishedDate']  = df1.loc[url,'publishedDate']
+        df.loc[url,'desk']           = df1.loc[url,'desk']
+        df.loc[url,'type']           = df1.loc[url,'type']
+        df.loc[url,'section']        = df1.loc[url,'section']
+        df.loc[url,'typeOfMaterial'] = df1.loc[url,'typeOfMaterial']
+        addToCell(df.loc[url,'authors'], df1.loc[url,'authors'])
+        addToCell(df.loc[url,'desFacet'], df1.loc[url,'desFacet'])
+        addToCell(df.loc[url,'perFacet'], df1.loc[url,'perFacet'])
+        addToCell(df.loc[url,'orgFacet'], df1.loc[url,'orgFacet'])
+        addToCell(df.loc[url,'geoFacet'], df1.loc[url,'geoFacet'])
[…]

### 2015-12-05 Project/nytData_parsing/xArticleData.py

@@ -1,0 +1,76 @@
+import numpy as np
+import re
+import os
+import time
+import simplejson as json
+import sys
+import pandas as pd
+reload(sys)
+sys.setdefaultencoding('utf-8')
+
+month = sys.argv[1]
+half = int(sys.argv[2])
+
+if half == 1:
+    path = '/mnt/nfs6/sinan_data.proj/NYT/' + month + '/days1-15/'
+if half == 2:
+    path = '/mnt/nfs6/sinan_data.proj/NYT/' + month + '/days16-30/'
+
+listpath = '/srv/ml_project/article_list/08-list'
+url_list = open(listpath,'r').read().splitlines()
+n = len(url_list)
+col_list = ['id', 'count', 'publishedDate', 'headline', 'authors', 'typeOfMaterial', 'desk', 'type', 'section', 'desFacet', 'perFacet', 'orgFacet', 'geoFacet']
+df = pd.DataFrame({'id': range(n), 'count': np.zeros(n), 'publishedDate': np.zeros(n), 'headline': ['None'] * n, 'authors': [[] for _ in range(n)], 'typeOfMaterial': ['None'] * n, 'desk': ['None'] * n, 'type': ['None'] * n, 'section' : ['None'] * n, 'desFacet': [[] for _ in range(n)], 'perFacet': [[] for _ in range(n)], 'orgFacet': [[] for _ in range(n)], 'geoFacet': [[] for _ in range(n)]}, index = url_list, columns = col_list)
+pattern = '%m-%d-%Y %H:%M'
+
+def addToCell(cell, alist):
+    for i in alist:
+        cell.append(i)
+
+for root, dirs, files in os.walk(path):
+	name = os.path.join(root)
+	name = re.sub('/mnt.*NYT/', '', name)
+	name = re.sub('(\d\d)/days1-15', '\\1', name)
+	name = re.sub('(\d\d)/days16-30', '\\1', name)
+	name = re.sub('/(\d\d)/', '-\\1', name)
+	name = re.sub('/', '-', name)
+	date = name + '-2013 00:00'
+
+	try:
+		epoch = max((int(time.mktime(time.strptime(date, pattern)))- 7 * 86400) * 1000, 1375329600000)
+	except:
+		epoch = 3000000000000
+
+	for file in files:
+		f = open(os.path.join(root, file), 'r')
+		print os.path.join(root, file).encode('utf-8')
+		for line in f:
+			try:
+				obj = json.loads(line)
[…]

### 2015-12-06 Project/feature_extraction/sentiment.R
[project] reorganize and rename some stuff, new parser, etc.

@@ -1,35 +1,0 @@
-library('tm')
-library('SnowballC')
-library('slam')
-library("tm.lexicon.GeneralInquirer")
-
-normalize <- function(corpus){ # conversion to lower case, other function: removeNumbers,removePunctuation,removeWords, 
-  wd <- tm_map(corpus,content_transformer(removePunctuation))
-  wd <- tm_map(wd,content_transformer(tolower)) 
-  wd <- tm_map(wd,content_transformer(removeNumbers))
-  #  wd <- tm_map(wd,content_transformer(removeWords),stopwords("english"))
-  #  wd <- tm_map(wd,content_transformer(removeWords),c('may','will'))
-  #  wd <- tm_map(wd,content_transformer(stripWhitespace))
-  #  wd <- tm_map(wd,content_transformer(stemDocument))
-  return(wd)}
-
-path <- 'C:/Users/Jeremy/Desktop/Data/R/articles'
-c <- Corpus(DirSource(path,encoding='UTF-8'))
-#c <- Corpus(VectorSource(text))
-c <- normalize(c)
-dtm <- DocumentTermMatrix(c)
-
-# tf-idf weighted dtm
-dtm_weighted <- DocumentTermMatrix(c, control = list(weighting = weightTfIdf))
-  
-pos <- as.numeric(tm_term_score(dtm,terms_in_General_Inquirer_categories("Positiv"))) 
-neg <- as.numeric(tm_term_score(dtm,terms_in_General_Inquirer_categories("Negativ"))) 
-len <- rollup(dtm, 2, na.rm=TRUE, FUN = sum)
-len <- as.vector(as.matrix(len))
-
-index <- seq(1,length(len))
-data <- cbind(index,pos,neg,len)
-head(data)
-dim(data)
-
-write.table(data,'sentiment.txt',row.names = F,sep=',')

### 2015-12-06 Project/feature_extraction/sentiment/make_training_data.R

@@ -1,0 +1,31 @@
+library(stringr)
+library(dplyr)
+
+text <- data.frame(index = c(), text=c(), stringsAsFactors=FALSE)
+for (i in 1:7368){
+  print(i)
+  name <- paste('../../text_scraping/article_text/', i,'.txt',sep='')
+  t <- tryCatch({
+    readChar(name, file.info(name)$size)
+  }, warning=function(w) {
+    return('NA')
+  }, error=function(e) {
+    return('NA')
+  }, finally = {
+  })
+  t <- str_replace_all(t, '\n', '')
+  text = rbind(text, c(i, t))
+  names(text) <- c('index', 'text')
+  text$index <- as.numeric(text$index)
+  text$text <- as.character(text$text)
+}
+
+write.csv(text,'articles.csv',row.names=FALSE)
+
+nonempty_articles <- filter(text,text!='')
+set.seed(111)
+s <- sample(nonempty_articles$index,200)
+sampled_articles <- nonempty_articles[s,]
+
+head(sampled_articles)
+write.csv(sampled_articles,'sample_articles.csv',row.names=FALSE, fileEncoding='utf-8')

### 2015-12-06 Project/feature_extraction/sentiment/toascii.py

@@ -1,0 +1,193 @@
+import unicodedata
+import types
+
+def toascii(line, errfilename = None, in_encoding = 'utf-8'):
+    """
+    Convert a line to plain ascii, making reasonable substitutions
+    for accented characters, curly quotes, emdashes etc.
+    Unknown characters will be backslash replaced, and
+    can also be appended, with context, to an error file.
+    Characters in the error file can then be added to the
+    xlate table so they will be handled next time.
+
+    Arguments:
+    line:        a string or unicode string to be converted to ASCII.
+    errfilename: a place to log unknown characters.
+    in_encoding: the encoding used if line is a string
+    (not used for unicode input).
+    """
+
+    # Define the error file. In Python 2.6, the sub-function
+    # log_error can see a list but not a scalar variable.
+    # So the file pointer has to be errfile[0], not just errfile.
+    errfile = [ None, errfilename ]
+
+    # Log an error, giving some context around the problematic area:
+    def log_error(uni, start, end, msg="error"):
+        contextsize = 15
+        if errfile[0] == None:
+            if errfilename == None:
+                return
+            errfile[0] = open(errfile[1], "a")
+        unilen = len(uni)
+        strstart = start -contextsize
+        if strstart < 0 : strstart = 0
+        strend = end + contextsize
+        if strend > unilen : strend = unilen
+        print >>errfile[0], msg, ":", \
+            uni[strstart:strend].encode('ascii', 'backslashreplace')
+# [ u for u in uni[start:end]]
+
+    output = ''
+
+    # If it's a string, decode it to Unicode.
+    # If it's already unicode, no need to do that.
+    if type(line) == types.StringType:
+        if in_encoding == '':
+            in_encoding = 'utf-8'
+            # Slower but safer: try decoding with utf-8 then iso8859-15
+        line = line.decode(in_encoding, 'replace')
+    elif type(line) != types.UnicodeType:
+        return "toascii needs either string or unicode, not" + str(type(line))
+
+    normalized = unicodedata.normalize('NFKD', line)
+    while normalized != None:
+        try:
+            output += normalized.encode('ascii', 'strict')
+            # or ignore, replace, etc.
+            normalized = None
+            break
+
+        except UnicodeEncodeError, e:
+            # At this point, e has these useful attributes:
+            # e.encoding: 'ascii'
+            # e.args: (encoding, unicode, start, end?, message)
[…]

### 2015-12-06 Project/gender_guessing/parse_names.py

@@ -4,6 +4,7 @@ import itertools
 import requests
 from bs4 import BeautifulSoup
 import time
+import csv
 
 def get_popularity(author):
 	print author
@@ -15,7 +16,7 @@ def get_popularity(author):
 	soup = BeautifulSoup(r.text, "html.parser")
 	raw_results_text = soup.find('div',{'id':'b_tween'}).text
 	time.sleep(30)
-	return raw_results_text.replace(' resultsAny time', '').replace(',', '').replace(' ', '')
+	return [int(raw_results_text.split()[0].replace(',', ''))]
 	
 def remove_quotes(string):
 	return string.replace('"', '')
@@ -28,8 +29,11 @@ unique_authors = set(authors)
 authors_df = pandas.DataFrame(list(unique_authors), columns = ['author_name'])
 authors_df['author_name'] = authors_df.author_name.apply(remove_quotes)
 authors_df['author_name'] = authors_df.author_name.apply(remove_quotes)
-authors_df.to_csv('author_list.csv', index=False, encoding='utf-8')
+authors_df.to_csv('author_list.tsv', index=False, encoding='utf-8', quotechar='', quoting=csv.QUOTE_NONE, escapechar=' ', sep='\t')
+
+authors_df = pandas.read_csv('author_list.tsv', sep='\t')
 
 popularities = []
 for i in authors_df['author_name']:
 	popularities += get_popularity(i)
+	print popularities

### 2015-12-06 Project/text_scraping/scraper.R

@@ -1,74 +1,0 @@
-# install packages and define parse function
-
-if (!require('XML')) {install.packages('XML')}
-if (!require('httr')) {install.packages('httr')}
-
-# old parse function, article only
-
-parseArticleBody <- function(artHTML) {
-  xpath2try <- c('//div[@class="articleBody"]//p',
-                 '//p[@class="story-body-text story-content"]',
-                 '//p[@class="story-body-text"]'
-  )
-  for(xp in xpath2try) {
-    bodyi <- paste(xpathSApply(htmlParse(artHTML), xp, xmlValue), collapse='')
-    if(nchar(bodyi)>0) break
-  }
-  return(bodyi)
-}
-
-# new parse function, dummy for video and interactive graph
-
-parseArticleBody <- function(artHTML) {
-  xpath2try <- c('//div[@class="articleBody"]//p',
-                 '//p[@class="story-body-text story-content"]',
-                 '//p[@class="story-body-text"]'
-  )
-  
-  for(xp in xpath2try) {
-    bodyi <- paste(xpathSApply(htmlParse(artHTML), xp, xmlValue), collapse='')
-    if(nchar(bodyi)>0) break
-  }
-  
-  xpath <- c('//a[@class="video-link"]')
-  video <- paste(xpathSApply(htmlParse(artHTML), xpath, xmlValue), collapse='')
-  if (nchar(video)!= 0) {v <- 1} else {v <- 0}
-  
-  ypath <- c('//div[@class="interactive-graphic"]')
-  graph <- paste(xpathSApply(htmlParse(artHTML), ypath, xmlValue), collapse='')
-  if (nchar(graph)!= 0) {g <- 1} else {g <- 0}
-  
-  return(list(bodyi,v,g))
-}
-
-path_to_article_list <- 'article_list.txt'
-#path_to_article_list <- '/srv/ml_project/article_list/9-list'
-
-
-article_list <- read.csv(path_to_article_list, header=F)
-names(article_list) <- c('article_url')
-
-article_dummies <- data.frame(index = c(), 
-  has_video = c(),
-  has_graph = c()
-  )
-
-for (i in 1:length(article_list$article_url)) {
-  print(i)
-  url <- GET(as.character(article_list$article_url[i]))
-  html <- content(url, 'text')
-  artBody <- parseArticleBody(html)
-  write(artBody[[1]], file=paste0('article_text/', as.character(i), '.txt', sep="", collapse=""))
[…]

### 2015-12-06 Project/text_scraping/textParser.py

@@ -1,0 +1,20 @@
+from newspaper import Article
+import pandas as pd 
+
+article_urls = pd.read_csv('article_list.txt', header=None)
+url = 'http://www.nytimes.com/2013/08/31/sports/ncaafootball/lewan-staying-at-michigan-with-renewed-focus.html'
+
+for i in range(len(article_urls)):
+	print i 
+	article = Article(url=article_urls[0][i])
+	article.download()
+	article.parse()
+	text = article.text
+	text = text.replace('Photo\n', '')
+	text = text.replace('Photo \n', '')
+	text = text.replace('Advertisement Continue reading the main story\n', '')
+	text = text.replace('\n\n', ' ')
+	text = text[1:]
+	f = open('article_text/' + str(i + 1) + '.txt', 'w')
+	f.write(text.encode('utf8'))
+	f.close()

### 2015-12-07 Project/text_scraping/scraper.R
[project] some updates to scraper and thing to make sentiment training data.

@@ -1,0 +1,64 @@
+# install packages and define parse function
+
+if (!require('XML')) {install.packages('XML')}
+if (!require('httr')) {install.packages('httr')}
+
+# old parse function, article only
+
+parseArticleBody <- function(artHTML) {
+  xpath2try <- c('//div[@class="articleBody"]//p',
+                 '//p[@class="story-body-text story-content"]',
+                 '//p[@class="story-body-text"]'
+  )
+  for(xp in xpath2try) {
+    bodyi <- paste(xpathSApply(htmlParse(artHTML, asText=TRUE), xp, xmlValue), collapse='')
+    if(nchar(bodyi)>0) break
+  }
+  return(bodyi)
+}
+
+# new parse function, dummy for video and interactive graph
+
+parseArticleBody <- function(artHTML) {
+  xpath2try <- c('//div[@class="articleBody"]//p',
+                 '//p[@class="story-body-text story-content"]',
+                 '//p[@class="story-body-text"]'
+  )
+  
+  for(xp in xpath2try) {
+    bodyi <- paste(xpathSApply(htmlParse(artHTML), xp, xmlValue), collapse='')
+    if(nchar(bodyi)>0) break
+  }
+  
+  xpath <- c('//a[@class="video-link"]')
+  video <- paste(xpathSApply(htmlParse(artHTML), xpath, xmlValue), collapse='')
+  if (nchar(video)!= 0) {v <- 1} else {v <- 0}
+  
+  ypath <- c('//div[@class="interactive-graphic"]')
+  graph <- paste(xpathSApply(htmlParse(artHTML), ypath, xmlValue), collapse='')
+  if (nchar(graph)!= 0) {g <- 1} else {g <- 0}
+  
+  return(list(bodyi,v,g))
+}
+
+path_to_article_list <- 'article_list.txt'
+#path_to_article_list <- '/srv/ml_project/article_list/9-list'
+
+
+article_list <- read.csv(path_to_article_list, header=F)
+names(article_list) <- c('article_url')
+
+article_dummies <- data.frame(index = c(), 
+  has_video = c(),
+  has_graph = c()
+  )
+
+for (i in 1:length(article_list$article_url)) {
+  print(i)
+  url <- GET(as.character(article_list$article_url[i]))
+  html <- content(url, 'text')
+  artBody <- parseArticleBody(html)
+  article_dummies <- rbind(article_dummies, c(i, artBody[[2]], artBody[[3]]))
[…]

### 2015-12-08 Project/text_scraping/textParser.py
textParser fixes

@@ -1,20 +1,37 @@
 from newspaper import Article
-import pandas as pd 
+import sys
+import re
+import codecs
 
-article_urls = pd.read_csv('article_list.txt', header=None)
-url = 'http://www.nytimes.com/2013/08/31/sports/ncaafootball/lewan-staying-at-michigan-with-renewed-focus.html'
+blk = int(sys.argv[1])
 
-for i in range(len(article_urls)):
-	print i 
-	article = Article(url=article_urls[0][i])
-	article.download()
-	article.parse()
-	text = article.text
-	text = text.replace('Photo\n', '')
-	text = text.replace('Photo \n', '')
-	text = text.replace('Advertisement Continue reading the main story\n', '')
-	text = text.replace('\n\n', ' ')
-	text = text[1:]
-	f = open('article_text/' + str(i + 1) + '.txt', 'w')
-	f.write(text.encode('utf8'))
-	f.close()
+article_urls = codecs.open('list_fixed.txt', encoding = 'utf-8').read().splitlines()
+
+def wrapper(block):
+    for i in block:
+        print i 
+        article = Article(url=article_urls[i])
+        article.download()
+        article.parse()
+        text = article.text
+        text = text.replace('Photo\n', '')
+        text = text.replace('Photo \n', '')
+        text = text.replace('Advertisement Continue reading the main story\n', '')
+        text = text.replace('Photo Advertisement Continue reading the main story\n', '')
+        text = re.sub("\n+" , " ", text)
+        text = re.sub('^ ', '', text)
+        f = open('article_text/' + str(i + 1) + '.txt', 'w')
+        f.write(text.encode('utf-8'))
+        f.close()
+
+if blk == 1:
+    wrapper(xrange(1670))
+    
+elif blk == 2:
+    wrapper(xrange(1670, 3341))
+    
+elif blk == 3:
+    wrapper(xrange(3341,5011))
+    
+elif blk == 4:
+    wrapper(xrange(5011,6682))

### 2015-12-08 Project/gender_guessing/add_genders.R
[project] popularities, new name parser, new gender stuff, start to do regression script.

@@ -25,5 +25,5 @@ authors$prop_female <- apply(authors, 1, FUN=female_prop_for_df)
 authors$prop_male <- as.numeric(authors$prop_male)
 authors$prop_female <- as.numeric(authors$prop_female)
 
-write.csv(authors, file='authors_with_genders.csv')
-
+write.table(authors, file='authors_with_genders.tsv', row.names=FALSE, sep="\t", 
+            quote=FALSE, fileEncoding='UTF-8')

### 2015-12-08 Project/regression/regression.py

@@ -1,0 +1,56 @@
+import pandas as pd 
+import pickle 
+
+
+def create_dummy_variables(df, columns):
+	dummies = pd.get_dummies(articleData[columns], prefix=columns)
+	df_reduced = df.ix[:, df.columns - columns]
+	new_df = pd.concat([df_reduced, dummies], axis=1)
+	return new_df
+
+def clean_up_whitespace(string):
+	while '  ' in string:
+		string = string.replace('  ', ' ')
+	return string	.encode('utf-8', 'ignore')
+
+
+
+articleData = pickle.load(open('articleData.pkl', 'rb'))
+dummyColumns = ['typeOfMaterial', 'desk', 'type', 'section']
+
+articleDataDummies = create_dummy_variables(articleData, dummyColumns)
+
+authorData = pd.DataFrame.from_csv('../gender_guessing/authors_with_genders.tsv', sep='\t', index_col=None, encoding='utf-8')
+authorData['author_name'] = authorData.author_name.apply(clean_up_whitespace)
+
+def get_author_gender(author_array):
+	genders = []
+
+	if len(author_array) > 0:
+		for i in author_array:
+			print i 
+			j = clean_up_whitespace(i)
+			print type(j)
+			prop_male = authorData[authorData['author_name'] == j]['prop_male']
+			print prop_male
+			print min(prop_male)
+			print max(prop_male)
+			if min(prop_male) >= .9:
+				genders += [1.] 
+			elif max(prop_male) <= .1:
+				genders += [0.] 
+			else: 
+				genders += [.5]
+	else: 
+		genders += [.5]
+
+	if max(genders) == 1. and min(genders) == 1.:
+		return 'male'
+	elif min(genders) == 0. and max(genders) == 0.:
+		return 'female'
+	else:
+		return 'ambiguous / unknown'
+
+articleData['author_gender'] = articleData.authors.apply(get_author_gender)
+
+print articleData

### 2015-12-08 Project/regression/regression.py
[project] add popularity features.

@@ -28,19 +28,15 @@ def get_author_gender(author_array):
 
 	if len(author_array) > 0:
 		for i in author_array:
-			print i 
 			j = clean_up_whitespace(i)
-			print type(j)
 			prop_male = authorData[authorData['author_name'] == j]['prop_male']
-			print prop_male
-			print min(prop_male)
-			print max(prop_male)
 			if min(prop_male) >= .9:
 				genders += [1.] 
 			elif max(prop_male) <= .1:
 				genders += [0.] 
 			else: 
 				genders += [.5]
+
 	else: 
 		genders += [.5]
 
@@ -51,6 +47,22 @@ def get_author_gender(author_array):
 	else:
 		return 'ambiguous / unknown'
 
+def get_author_popularity(author_array):
+	popularities = []
+
+	if len(author_array) > 0:
+		for i in author_array:
+			j = clean_up_whitespace(i)
+			popularity = authorData[authorData['author_name'] == j]['popularity'].tolist()
+			popularities += popularity
+
+	else: 
+		popularities += [int(authorData['popularity'].median())]
+
+	return sum(popularities) / float(len(popularities))
+
+
 articleData['author_gender'] = articleData.authors.apply(get_author_gender)
+articleData['popularity_pre_log'] = articleData.authors.apply(get_author_popularity)
 
 print articleData

### 2015-12-08 Project/feature_extraction/readability/flesch_kincaid_score.py
[project finish feature extraction, etc.

@@ -6,9 +6,9 @@ import pandas as pd
 
 flesch_kincaid_scores = pd.DataFrame(columns = ('index_value', 'flesch-kincaid_score'))
 
-for i in os.listdir('../text_scraping/article_text'):
+for i in os.listdir('../../text_scraping/article_text'):
 	if i.endswith(".txt"):
-		with open ('../text_scraping/article_text/' + i, "r") as myfile:
+		with open ('../../text_scraping/article_text/' + i, "r") as myfile:
 			data=myfile.read().replace('\n', '')
 			if data != '' and '.' in data:
 				data_append = [{'index_value': i.replace('.txt', ''), 'flesch-kincaid_score': textstat.flesch_reading_ease(data)}]

### 2015-12-08 Project/regression/regression.py

@@ -15,13 +15,17 @@ def clean_up_whitespace(string):
 
 
 
-articleData = pickle.load(open('articleData.pkl', 'rb'))
-dummyColumns = ['typeOfMaterial', 'desk', 'type', 'section']
-
-articleDataDummies = create_dummy_variables(articleData, dummyColumns)
+articleData = pickle.load(open('articleData-withWords.pkl', 'rb'))
 
 authorData = pd.DataFrame.from_csv('../gender_guessing/authors_with_genders.tsv', sep='\t', index_col=None, encoding='utf-8')
 authorData['author_name'] = authorData.author_name.apply(clean_up_whitespace)
+
+fleschKincaid = pd.DataFrame.from_csv('../feature_extraction/readability/flesch_kincaid.txt', sep=',', index_col=None)
+fleschKincaid['id'] = fleschKincaid['index_value']
+fleschKincaid = fleschKincaid.ix[:, fleschKincaid.columns - ['index_value']]
+
+bigramPerplexity = pickle.load(open('bigram-perplexity.pkl', 'rb'))
+articleData = pd.concat([articleData, bigramPerplexity], axis=1)
 
 def get_author_gender(author_array):
 	genders = []
@@ -65,4 +69,8 @@ def get_author_popularity(author_array):
 articleData['author_gender'] = articleData.authors.apply(get_author_gender)
 articleData['popularity_pre_log'] = articleData.authors.apply(get_author_popularity)
 
-print articleData
+dummyColumns = ['typeOfMaterial', 'desk', 'type', 'section', 'author_gender']
+articleDataDummies = create_dummy_variables(articleData, dummyColumns)
+articleDataDummies = articleDataDummies.merge(fleschKincaid, on=['id'])
+
+print articleDataDummies

### 2015-12-08 Project/regression/regression.py
[project] regression, MSE

@@ -1,17 +1,17 @@
 import pandas as pd 
 import pickle 
-
+import numpy as np 
 
 def create_dummy_variables(df, columns):
-	dummies = pd.get_dummies(articleData[columns], prefix=columns)
-	df_reduced = df.ix[:, df.columns - columns]
-	new_df = pd.concat([df_reduced, dummies], axis=1)
-	return new_df
+    dummies = pd.get_dummies(articleData[columns], prefix=columns)
+    df_reduced = df.ix[:, df.columns - columns]
+    new_df = pd.concat([df_reduced, dummies], axis=1)
+    return new_df
 
 def clean_up_whitespace(string):
-	while '  ' in string:
-		string = string.replace('  ', ' ')
-	return string	.encode('utf-8', 'ignore')
+    while '  ' in string:
+        string = string.replace('  ', ' ')
+    return string    .encode('utf-8', 'ignore')
 
 
 
@@ -28,42 +28,42 @@ bigramPerplexity = pickle.load(open('bigram-perplexity.pkl', 'rb'))
 articleData = pd.concat([articleData, bigramPerplexity], axis=1)
 
 def get_author_gender(author_array):
-	genders = []
+    genders = []
 
-	if len(author_array) > 0:
-		for i in author_array:
-			j = clean_up_whitespace(i)
-			prop_male = authorData[authorData['author_name'] == j]['prop_male']
-			if min(prop_male) >= .9:
-				genders += [1.] 
-			elif max(prop_male) <= .1:
-				genders += [0.] 
-			else: 
-				genders += [.5]
+    if len(author_array) > 0:
+        for i in author_array:
+            j = clean_up_whitespace(i)
+            prop_male = authorData[authorData['author_name'] == j]['prop_male']
+            if min(prop_male) >= .9:
+                genders += [1.] 
+            elif max(prop_male) <= .1:
+                genders += [0.] 
+            else: 
+                genders += [.5]
 
-	else: 
-		genders += [.5]
+    else: 
+        genders += [.5]
 
-	if max(genders) == 1. and min(genders) == 1.:
-		return 'male'
-	elif min(genders) == 0. and max(genders) == 0.:
-		return 'female'
-	else:
-		return 'ambiguous / unknown'
+    if max(genders) == 1. and min(genders) == 1.:
+        return 'male'
+    elif min(genders) == 0. and max(genders) == 0.:
+        return 'female'
+    else:
+        return 'ambiguous / unknown'
 
[…]

### 2015-12-08 Latex/Paper.tex
Latex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Stylish Article
% LaTeX Template
% Version 2.1 (1/10/15)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Mathias Legrand (legrand.mathias@gmail.com) 
% With extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fleqn,12pt]{SelfArx} % Document font size and equations flushed left

\usepackage[english]{babel} % Specify a different language here - english by default

\usepackage{lipsum} % Required to insert dummy text. To be removed otherwise

%----------------------------------------------------------------------------------------
%	COLUMNS
%----------------------------------------------------------------------------------------

\setlength{\columnsep}{0.55cm} % Distance between the two columns of text
\setlength{\fboxrule}{0.75pt} % Width of the border around the abstract

%----------------------------------------------------------------------------------------
%	COLORS
%----------------------------------------------------------------------------------------

\definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
\definecolor{color2}{RGB}{0,20,20} % Color of the boxes behind the abstract and headings

%----------------------------------------------------------------------------------------
%	HYPERLINKS
%----------------------------------------------------------------------------------------

\usepackage{hyperref} % Required for hyperlinks
\hypersetup{hidelinks,colorlinks,breaklinks=true,urlcolor=color2,citecolor=color1,linkcolor=color1,bookmarksopen=false,pdftitle={Title},pdfauthor={Author}}

%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------

\JournalInfo{6.867 Final Paper} % Journal information
\Archive{} % Additional notes (e.g. copyright, DOI, review/research article)

\PaperTitle{How Does Content Drive Viewership?} % Article title

\Authors{Dave Holtz\textsuperscript{1}, Jeremy Yang\textsuperscript{2}, Michael Zhao\textsuperscript{3}} % Authors
\affiliation{\textsuperscript{1}\textit{dholtz@mit.edu}}
\affiliation{\textsuperscript{2}\textit{zheny@mit.edu}} 
\affiliation{\textsuperscript{3}\textit{mfzhao@mit.edu}}

\Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{Why do some webpages receive massive numbers of pageviews? To determine how content drives viewership, we construct a unique dataset of all articles published by the New York Times (NYT) in August 2013. Our dataset is built from 2 major components, the NYT's internal web traffic data and article content data parsed from the NYT website. We use the internal web traffic data to accurately track the number of page views of each article as well as construct a set of robust control variables such as the desk and section of each article. To build content features, we use various machine learning and statistical natural language processing techniques on our parsed article content data and construct features such as article perplexity, sentiment, reading difficulty, and indicators that denote the presence of pictures, videos, etc. Additionally, we have access to the NYT's internal website traffic data. We feed all of our constructed features to into a predictive regression model. We find [MAJOR RESULTS HERE].
}

%----------------------------------------------------------------------------------------

\begin{document}

\flushbottom % Makes all text pages the same height

\maketitle % Print the title and abstract box

\tableofcontents % Print the contents section

\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction} % The \section*{} command stops section numbering

In today's digital economy, many companies are very interested in attracting users to visit their websites in order to earn ad revenue. While many factors might motivate a user to visit a particular page, certainly one important factor is the content in that webpage. This paper explores the relationship between the content of a webpage and the number of page views it ultimately ends up receiving by constructing a unique dataset of all articles published by the New York Times (NYT) during August 2013. This dataset is built from two major components: the NYT's internal web traffic data and parsed NYT article content data.

Typically, a study such as ours tends to be very difficult to conduct as either accurate measures of viewership are unavailable\footnote{While oftentimes precise viewership data tends to be not available openly, oftentimes researchers use related observables, such as Facebook likes} or the feature extraction of the content is too challenging (for example Youtube), or or both. Fortunately, our access to the the NYT's internal web traffic data allows us to exactly measure the number of page views an article receives. The web traffic data is rather rich and also includes internal meta-data that we use to build various control features. Moreover, since we are working with mostly textual data, we are able to take advantage of recent advancements in machine learning and statistical NLP to do feature extraction on article text.

A similar study by Berger and Milkman (2012) \cite{BergerMilkman12} examines the relationship between content and word-of-mouth virality. They find that the emotional content of a NYT article is predictive of its virality. Using simple measures of an article's sentiment and emotionality, Berger and Milkman show that positive articles are more likely to show up on the New York Times "Most-Emailed" list. They also show that articles that evoke high physiological positive or negative arousal (such as awe or anger) tend to be more viral than articles that evoke deactivating emotions (sadness). We build on this study in two ways: first, we relate an article's content back to the number of page views it receives rather than its virality\footnote{Which companies arguably care more about since word-of-mouth virality is usually a means to increase page views}. Second, we employ more sophisticated machine learning feature extraction techniques to see if they work any better over their simple measures.
%------------------------------------------------

\section{Data}
\subsection{NYT Internal Web Traffic Data}
Our NYT internal web traffic dataset is a record of all individual user activity on the NYT website covering the period of April 3rd, 2013 to October 31st, 2013. This activity data is stored as individual lines of json and includes who (if available) accessed what page at what time. Overall, it is over 20 terabytes in size and contains over 3 billion page views\footnote{Not all page views are article views, for example, some events that are also tracked are searches, or user account settings.} Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013.

Looking at only pages that only contain articles or blogposts,

\subsection{Parsed NYT Article Content Data}

%\begin{figure*}[ht]\centering % Using \begin{figure*} makes the figure take up the entire width of the page
%\includegraphics[width=\linewidth]{view}
%\caption{Wide Picture}
%\label{fig:view}
%\end{figure*}

\begin{figure}[ht]\centering
\includegraphics[width=\linewidth]{results}
\caption{In-text Picture}
\label{fig:results}
\end{figure}

Reference to Figure \ref{fig:results}.

%------------------------------------------------

\section{}

\subsection{Subsection}

\begin{table}[hbt]
\caption{Table of Grades}
\centering
\begin{tabular}{llr}
\toprule
\multicolumn{2}{c}{Name} \\
\cmidrule(r){1-2}
First name & Last Name & Grade \\
\midrule
John & Doe & $7.5$ \\
Richard & Miles & $2$ \\
\bottomrule
\end{tabular}
\label{tab:label}
\end{table}

\begin{description}
\item[Word] Definition
\item[Concept] Explanation
\item[Idea] Text
\end{description}

\subsubsection{Subsubsection}

\begin{itemize}[noitemsep] % [noitemsep] removes whitespace between the items for a compact look
\item First item in a list
\item Second item in a list
\item Third item in a list
\end{itemize}

\subsubsection{Subsubsection}

\subsection{Subsection}

%------------------------------------------------
\phantomsection
\section*{Acknowledgments} % The \section*{} command stops section numbering

\addcontentsline{toc}{section}{Acknowledgments} % Adds this section to the table of contents

So long and thanks for all the fish \cite{Figueredo:2009dg}.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\phantomsection
\bibliographystyle{unsrt}
\bibliography{sample}

%----------------------------------------------------------------------------------------

\end{document}

### 2015-12-08 Latex/Paper.tex

@@
 
-\Abstract{Why do some webpages receive massive numbers of pageviews?
+\Abstract{Why do some webpages receive massive numbers of page views?
 To determine how content drives viewership, we construct a unique dataset of all articles published by the New York Times (NYT) in August 2013.
@@
 This activity data is stored as individual lines of json and includes who (if available) accessed what page at what time.
-Overall, it is over 20 terabytes in size and contains over 3 billion page views\footnote{Not all page views are article views, for example, some events that are also tracked are searches, or user account settings.} Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013.
+Overall, it is over 20 terabytes in size and contains over 3 billion page views\footnote{Not all page views are content views, for example, some events that are also tracked are searches, or user account settings.} Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013.
 
-Looking at only pages that only contain articles or blogposts,
+We limit our dataset to consist of pages that only contain articles or blogposts published during the month of August.
+We parse the data to obtain a list of urls, which need to be stripped of potential garbage.
+After cleaning up the url data, we are left with 6682 unique pieces of content.
+We then parse our dataset and aggregate the number of counts each url receives.
+In order to make the comparison between articles fairer since an article that's been out longer will have more page views on average, we only count the number of page views received up to 7 days after publication\footnote{Given that page views tend to sharply drop off soon after publication since recency is quite important to the News, the number of page views obtained during the 7 days after an article is published represents the vast majority (usually well over 90\%) of total page views an article receives.}.
+In total, our data consists of over 250 million page views.
+As seen in Figure 1 below, the distribution of page views is highly skewed with very have tables.
+After applying a log transformation (as seen in Figure 2), our distribution looks considerably more normal.
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{pageviews_hist}
+\caption{Histogram of Articles by Number of Page Views}
+\label{fig:pv_hist}
+\end{figure}
+% TWEAK TABLE LOOK
+\begin{table}[hbt]
+\caption{Page Views Distribution Summary Statistics}
+\centering
+\begin{tabular}{rr}
+\toprule
+Total Page Views  &  248161455\\
+\midrule
+Min               &  1\\
+Max               &  2545288\\  
+Mean              &  37138.8\\
+Median            &  10298.5\\
+Std. Dev.         &  88972.9\\
+Skewness          &  9.52191\\
+Kurtosis          &  173.061\\
+\midrule
+Observations &  6682\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{log_pageviews_hist}
+\caption{Histogram of Articles by Log of Page Views}
+\label{fig:lpv_hist}
+\end{figure}
+
+\begin{table}[hbt]
+\caption{Log Page Views Distribution Summary Statistics}
+\centering
+\begin{tabular}{rr}
+\toprule
+Min               &  0\\
+Max               &  14.74975\\  
+Mean              &  9.122868\\
+Median            &  9.239754\\
+Std. Dev.         &  2.028668\\
+Skewness          &  -1.270368\\
+Kurtosis          &  3.800911\\
+\midrule
+Observations &  6682\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+In addition to aggregating the counts, when
 
 \subsection{Parsed NYT Article Content Data}
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{wordcount_hist}
+\caption{Histogram of Articles by Word Count}
+\label{fig:pv_hist}
+\end{figure}
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{log_wordcount_hist}
+\caption{Histogram of Articles by Log Word Count}
+\label{fig:lpv_hist}
+\end{figure}
 
@@
 
-\begin{figure}[ht]\centering
-\includegraphics[width=\linewidth]{results}
-\caption{In-text Picture}
-\label{fig:results}
-\end{figure}
-
-Reference to Figure \ref{fig:results}.
-
 %------------------------------------------------
 
-\section{}
+\section{Constructed Features}
 
@@
 
+\section{Predictive Regression Model}
+
 %------------------------------------------------

### 2015-12-08 Latex/Paper.tex
[project] write a little, add a gitignore.

@@
 
-A similar study by Berger and Milkman (2012) \cite{BergerMilkman12} examines the relationship between content and word-of-mouth virality.
+A similar study by Berger and Milkman (2012) \cite{berger2012makes} examines the relationship between content and word-of-mouth virality.
 They find that the emotional content of a NYT article is predictive of its virality.
@@
 In total, our data consists of over 250 million page views.
-As seen in Figure 1 below, the distribution of page views is highly skewed with very have tables.
+As seen in Figure 1 below, the distribution of page views is highly skewed with very heavy tails.
 After applying a log transformation (as seen in Figure 2), our distribution looks considerably more normal.
@@
 
-\subsection{Subsection}
-
-\begin{table}[hbt]
-\caption{Table of Grades}
-\centering
-\begin{tabular}{llr}
-\toprule
-\multicolumn{2}{c}{Name} \\
-\cmidrule(r){1-2}
-First name & Last Name & Grade \\
-\midrule
-John & Doe & $7.5$ \\
-Richard & Miles & $2$ \\
-\bottomrule
-\end{tabular}
-\label{tab:label}
-\end{table}
+In preparing to perform our regression, we construct a number of features, using both the parsed NYT article data and secondary sources.
+These features include the Flesch reading ease, the (guessed) gender of the author(s), the popularity of the author(s), the sentiment of the article text, the perplexity of the article text, and variables indicating the section the article appeared in and the article's content type.
+Found below is a full list of these features, as well as the methodology used to construct them and validation of the features, where appropriate.
 
 \begin{description}
-\item[Word] Definition
-\item[Concept] Explanation
-\item[Idea] Text
+	\item[Flesch Reading Ease] The Flesch reading ease is a metric developed by Flesch in 1948 \cite{flesch1948new}.
+The score indicates how difficult a piece of English text is to understand.
+Lower scores correspond to more difficult passages, and the highest score attainable is 120.0.
+The formula for calculating a passage's Flesch reading ease is
+
+	\begin{equation}
+	206.835 - 1.015 \left ( \frac{\textrm{\# words}}{\textrm{\# sentences}} \right ) - 84.6 \left ( \frac{\textrm{\# syllables}}{\textrm{\# words}} \right )
+	\end{equation}
+
+	To calculate the Flesch reading ease, we use the python library ``textstat." Despite the fact that the above formula is relatively straightforward, the task of counting the number of syllables in a block of text is non-trivial, so we rely on ``textstat" to do so accurately.
+In cases where the Flesch reading ease was for some reason null (e.g., a blog post containing only a picture), we assign the Flesch reading ease its median value.
+
+	\item[Author Popularity] We attempt to include some measure of a particular author's popularity.
+It stands to reason that a new article by Paul Krugman or A.O.
+Scott should garner more readership than a new article by an unknown graduate student enrolled in 6.867 at MIT!
+
+	In order to measure something that will serve as a decent proxy for popularity, we programmatically searched for every distinct author in our dataset using Bing, and recorded the number of search results that were returned by the query.
+In cases where a particular article has more than one distinct author, we calculate an ``effective" popularity by simply calculating the average number of search results over all article authors.
+
+	\item[Author Gender] We also attempt to construct a feature that indicates the most likely gender of the article author(s).
+In cases where the gender of the author is unclear (e.g., Robin) or there are likely multiple authors with different genders (e.g., The New York Times Staff), we record a third gender value, ``ambiguous / unknown."
+
+	We construct our gender data by cross-referencing the first names of all of the authors in our dataset against U.S.
+Social Security Administration baby name data from 1935 to 1997.
+If over 90\% of the babies with a given name have been male, we assume a given author is male.
+If over 90\% of the babies with a given name have been female, we assume a given author is female.
+Otherwise, we record ``ambiguous / unknown."
+
+	\item[Material Type, Section, Desk, and Article Type] We also include dummy variables including the material type (e.g., `News' or `Obituary'), desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), and section (e.g., `Movies' or `World').
+
 \end{description}
 
-\subsubsection{Subsubsection}
+We also include features that attempt to capture the article sentiment and the article text perplexity.
+Since the construction of these features was more involved and involved validation of our algorithms, we discuss these two features in separate subsections.
 
-\begin{itemize}[noitemsep] % [noitemsep] removes whitespace between the items for a compact look
-\item First item in a list
-\item Second item in a list
-\item Third item in a list
-\end{itemize}
+\subsection{Article Sentiment}
 
-\subsubsection{Subsubsection}
-
-\subsection{Subsection}
+\subsubsection{Article Perplexity}
 
@@
 
-So long and thanks for all the fish \cite{Figueredo:2009dg}.
-
 %----------------------------------------------------------------------------------------
@@
 \bibliographystyle{unsrt}
-\bibliography{sample}
+\bibliography{Paper}
 

### 2015-12-08 Project/Latex/Paper.tex
[project] some linear regression eval and plots.

--- 2015-12-08 Latex/Paper.tex
@@
 
-\subsubsection{Article Perplexity}
+\subsection{Article Perplexity}
 
@@
 
+\begin{figure*}[ht]\centering
+\includegraphics[width=\linewidth]{feature_weights.png}
+\caption{Linear Regression Feature Weights (excluding intercept term)}
+\label{fig:lin_reg_weights}
+\end{figure*}
+
+\begin{tabular}{lr}
+\toprule
+Feature  &  Weight\\
+\midrule
+intercept & 9.047\\
+log\_wcount & 0.729\\
+desk\_Foreign & 0.657\\
+section\_World & -0.553\\
+desk\_Travel & -0.472\\
+typeOfMaterial\_Schedule & -0.291\\
+section\_Movies & -0.277\\
+section\_Opinion & 0.261\\
+typeOfMaterial\_Review & -0.255\\
+desk\_National & 0.253\\
+typeOfMaterial\_Letter & -0.243\\
+section\_Sports & -0.215\\
+section\_Education & -0.199\\
+section\_Books & 0.192\\
+section\_Corrections & -0.179\\
+desk\_Weekend & -0.175\\
+section\_Health & 0.172\\
+type\_BlogPost & -0.170\\
+desk\_BookReview & -0.169\\
+typeOfMaterial\_Op-Ed & 0.167\\
+\end{tabular}
 %------------------------------------------------

### 2015-12-08 Project/feature_extraction/sentiment/make_training_data.R

@@ -22,13 +22,13 @@ for (i in 1:7368){
 
 write.csv(text,'articles.csv',row.names=FALSE)
 
-nonempty_articles <- filter(text,text!='')
+nonempty_articles <- filter(text,text !='' & text!= 'NA' & !is.na(text))
 set.seed(111)
 s <- sample(nonempty_articles$index,200)
 sampled_articles <- nonempty_articles[s,]
 
-head(sampled_articles)
 sampled_articles %>%
-  mutate(text = str_replace_all(text, ',', '')) -> sampled_articles
+  mutate(text = str_replace_all(str_replace_all(str_replace_all(text, ',', ''), '\"', '”'), '\u0096', '')) -> sampled_articles
 
 write.csv(dplyr::select(sampled_articles, text),'sample_articles.csv',row.names=FALSE, fileEncoding='utf-8', quote=FALSE)
+write.csv(sampled_articles,'sample_articles_with_indices.csv',row.names=FALSE, fileEncoding='utf-8', quote=FALSE)

### 2015-12-08 Project/regression/plots.R

@@ -1,0 +1,14 @@
+library(ggplot2)
+library(knitr)
+
+weights <- c(-0.0155673547965, -0.032512577208, -0.0183928993822, 0.0436023817917, -0.168735260358, -0.0548502431781, 0.0400864539071, 0.0477745936425, 0.0482044874247, 0.656851318914, 0.101863103466, 0.0787925572846, 0.0772113800894, -0.109357780179, 0.00259863923261, 0.253263211162, 0.00423288154355, -0.151785209635, -0.086932735094, 0.0642537764487, 0.0594296289471, -0.0784431104261, 0.0100914683538, -0.0455614111585, -0.150498327579, 0.0370920476908, -0.471571091079, 0.00575529953804, -0.174645309934, -0.00122143569646, -0.120772053981, 0.729336915393, 0.0191174119751, -0.0296600316436, 0.0414784388088, 0.192408869835, 0.0525969201364, -0.0430805415879, -0.179258076024, -0.13951006772, 0.0400864539072, -0.199291009302, 0.125170423372, -0.00107423696482, 0.1722575493, -0.103366129896, -0.0142565167554, -0.0481040689734, -0.277144753309, -0.00275229839806, 0.123937726235, 0.111792685733, 0.261123875719, -0.0202035470324, -0.0667582303632, 0.0161341447383, -0.214593015947, 0.0525950326542, 0.0457459855028, 0.0809072398456, 0.0841238574083, -0.0947854358489, -0.0536801646975, -0.105109365666, -0.553358112213, -0.00518030191328, -0.00801186882178, 0.026662047528, -0.0707252848644, -0.00164194215674, -0.00696646258888, -0.0156368327091, -0.243055275894, -0.0265795567348, 0.0236131786527, 0.126044421422, 0.0661457097531, -0.110683337504, 0.0679989404147, 0.167213483369, -0.0287554883533, 0.0611867349866, -0.0321046144057, -0.255398058467, -0.290623015012, 0.0022774254305, -0.0126711530544, -0.0831671026105, -0.0064248106726, -0.170417879777, 9.04696649847)
[…]

### 2015-12-08 Project/regression/regression.py

@@ -1,6 +1,8 @@
 import pandas as pd 
 import pickle 
 import numpy as np 
+import random
+import math
 
 def create_dummy_variables(df, columns):
     dummies = pd.get_dummies(articleData[columns], prefix=columns)
@@ -66,6 +68,22 @@ def get_author_popularity(author_array):
     return sum(popularities) / float(len(popularities))
 
 
+def scaleFeatures(X, mean = None, sigma= None):
+	if mean == None and sigma == None:
+		mean = np.mean(X, axis=0)
+		sigma = np.std(X, axis=0)
+
+	meanArray = np.repeat(mean.reshape(1, -1), X.shape[0], axis=0)
+	sigmaArray = np.repeat(sigma.reshape(1, -1), X.shape[0], axis=0)
+
+	scaledFeatures = (X - meanArray)/sigmaArray 
+	
+	return scaledFeatures, mean, sigma
+
+def scale_features(array):
+	scaled_array = (array - np.mean(array))/np.std(array)
+	return scaled_array
+
 articleData['author_gender'] = articleData.authors.apply(get_author_gender)
 articleData['popularity_pre_log'] = articleData.authors.apply(get_author_popularity)
 
@@ -76,29 +94,64 @@ articleDataDummies['flesch-kincaid_score'].fillna(articleDataDummies['flesch-kin
 
 articleDataDummies['log_count'] = articleDataDummies['count'].apply(np.log)
 articleDataDummies['log_popularity'] = articleDataDummies.popularity_pre_log.apply(np.log)
+articleDataDummies['log_wcount'] = articleDataDummies.wcount.apply(np.log)
 
-columnstoRemove = ['authors', 'desFacet', 'geoFacet', 'headline', 'id', 'orgFacet', 'perFacet', 'publishedDate', 'count', 'popularity_pre_log', 'log_count']
+columnstoRemove = ['authors', 'desFacet', 'geoFacet', 'headline', 'id', 'orgFacet', 'perFacet', 'publishedDate', 
+'count', 'popularity_pre_log', 'log_count', 'author_gender_male', 'desk_Business', 'section_Arts', 'type_Article', 
+'typeOfMaterial_Interview', 'wcount']
 articleDataDummiesRegression = articleDataDummies.ix[:, articleDataDummies.columns - columnstoRemove]
+intercept = pd.Series(np.ones(len(articleDataDummiesRegression)))
[…]

### 2015-12-08 Project/regression/regression.py
[project] mse plots.

@@ -3,6 +3,7 @@ import pickle
 import numpy as np 
 import random
 import math
+import csv
 
 def create_dummy_variables(df, columns):
     dummies = pd.get_dummies(articleData[columns], prefix=columns)
@@ -141,7 +142,7 @@ for l in [0, .1, 1, 2, 3, 4, 5, 10, 25, 50, 100]:
 
     training_MSES += list(MSE(yPredict, y)[0])
     holdout_MSES += list(MSE(yPredictHoldout, holdout_labels.values)[0])
-    ls += l 
+    ls += [l] 
 
     if MSE(yPredictHoldout, holdout_labels.values) < best_holdout_MSE:
     	best_weights = beta
@@ -155,3 +156,7 @@ print holdout_MSES
 print best_holdout_MSE
 print best_l 
 
+MSE_performance = pd.Series(training_MSES).to_frame(name='training_mse')
+MSE_performance['holdout_mse'] = pd.Series(holdout_MSES)
+MSE_performance['lambdas'] = pd.Series(ls)
+MSE_performance.to_csv('mse_performance.csv', index=False, quotechar='', quoting=csv.QUOTE_NONE, escapechar=' ', sep=',', encoding='utf8')

### 2015-12-08 Project/gender_guessing/add_genders.R
Update add_genders.R

@@ -27,3 +27,55 @@ authors$prop_female <- as.numeric(authors$prop_female)
 
 write.table(authors, file='authors_with_genders.tsv', row.names=FALSE, sep="\t", 
             quote=FALSE, fileEncoding='UTF-8')
+
+
+# split the name string to get multiple author names 
+
+n <- str_split(authors$names,'AND ')
+
+authors_number <- vector()
+
+for (i in 1:7367){
+  print (i)
+  authors_number[i] <- length(n[[i]])
+}
+max(authors_number)
+
+author_1_first_name <- vector()
+author_2_first_name <- vector()
+
+for (i in 1:7367){
+  print (i)
+  if(n[[i]]==''){
+    author_1_first_name[i] = 'NA'
+    author_2_first_name[i] = 'NA'
+  } else if (length(n[[i]])==1){
+      author_1_first_name[i] = first_name(n[[i]])
+      author_2_first_name[i] = 'NA'
+    } else {
+    author_1_first_name[i] = first_name(n[[i]][1])
+    author_2_first_name[i] = first_name(n[[i]][2])
+  }
+}
+
+authors <- cbind(authors,author_1_first_name,author_2_first_name)
+
+colnames(authors)[5] <- 'prop_male_1'
+colnames(authors)[6] <- 'prop_female_1'
+
+critical <- .5
+authors$gender_1 <- rep('NA',nrow(authors))
+authors$gender_2 <- rep('NA',nrow(authors))
+
+authors$gender_1[which(authors$prop_male_1=='NA')] <- 'NA'
+authors$gender_1[which(authors$prop_male_1!='NA' & authors$prop_male_1>=critical)] <- '0'
+authors$gender_1[which(authors$prop_male_1!='NA' & authors$prop_female_1 >= critical)] <- '1'
+
+authors$gender_2[which(authors$prop_male_2=='NA')] <- 'NA'
+authors$gender_2[which(authors$prop_male_2!='NA' & authors$prop_male_2>=critical)] <- '0'
+authors$gender_2[which(authors$prop_male_2!='NA' & authors$prop_female_2 >= critical)] <- '1'
+
+table(authors$gender_1)
+table(authors$gender_2)
+
+write.csv(authors,'multiple_authors.csv',row.names = F)

### 2015-12-09 Project/regression/regression.py
[project] some tidying up.

@@ -133,7 +133,7 @@ ls = []
 
 for l in [0, .1, 1, 2, 3, 4, 5, 10, 25, 50, 100]:
     XtX = np.dot(np.transpose(X), X) + l*np.identity(X.shape[1])
-    XtXinv = np.linalg.pinv(XtX)
+    XtXinv = np.linalg.inv(XtX)
 
     beta = np.dot(np.dot(XtXinv, np.transpose(X)), y)
 

### 2015-12-09 Project/gender_guessing/parse_names.py
[project] point at new pickle file.

@@ -22,7 +22,7 @@ def get_popularity(author):
 def remove_quotes(string):
 	return string.replace('"', '')
 
-df = pandas.read_pickle('/Users/dholtz/Downloads/articleData.pkl')
+df = pandas.read_pickle('/Users/dholtz/Downloads/articleData-withWords.pkl')
 
 authors = list(itertools.chain(*df['authors']))
 unique_authors = set(authors)

### 2015-12-09 Project/regression/regression.py
[project] build time of day and day of week features.

@@ -4,6 +4,8 @@ import numpy as np
 import random
 import math
 import csv
+import datetime
+from pytz import timezone
 
 def create_dummy_variables(df, columns):
     dummies = pd.get_dummies(articleData[columns], prefix=columns)
@@ -14,9 +16,22 @@ def create_dummy_variables(df, columns):
 def clean_up_whitespace(string):
     while '  ' in string:
         string = string.replace('  ', ' ')
-    return string    .encode('utf-8', 'ignore')
+    return string.encode('utf-8', 'ignore')
 
+def get_weekday(time_stamp):
+    return datetime.datetime.fromtimestamp(int(time_stamp/1000), tz=timezone('UTC')).strftime('%A')
 
+def get_hour_chunk(time_stamp):
+    utc = datetime.datetime.fromtimestamp(int(time_stamp/1000), tz=timezone('UTC'))
+    hour = int(utc.astimezone(timezone('US/Eastern')).strftime('%H'))
+    if hour < 6:
+        return '0-5'
+    elif hour < 12:
+        return '6-11'
+    elif hour < 18:
+        return '12-17'
+    else:
+        return '18-23'
 
 articleData = pickle.load(open('articleData-withWords.pkl', 'rb'))
 
@@ -87,8 +102,10 @@ def scale_features(array):
 
 articleData['author_gender'] = articleData.authors.apply(get_author_gender)
 articleData['popularity_pre_log'] = articleData.authors.apply(get_author_popularity)
+articleData['weekday'] = articleData.publishedDate.apply(get_weekday)
+articleData['timeofday'] = articleData.publishedDate.apply(get_hour_chunk)
 
-dummyColumns = ['typeOfMaterial', 'desk', 'type', 'section', 'author_gender']
+dummyColumns = ['typeOfMaterial', 'desk', 'type', 'section', 'author_gender', 'weekday', 'timeofday']
 articleDataDummies = create_dummy_variables(articleData, dummyColumns)
 articleDataDummies = articleDataDummies.merge(fleschKincaid, on=['id'])
 articleDataDummies['flesch-kincaid_score'].fillna(articleDataDummies['flesch-kincaid_score'].median(), inplace=True)
@@ -99,7 +116,7 @@ articleDataDummies['log_wcount'] = articleDataDummies.wcount.apply(np.log)
 
[…]

### 2015-12-09 Project/Latex/Paper.tex
[project] mse plots.
[project] some tidying up.
[project] some new plots.

--- 2015-12-08 Latex/Paper.tex
@@
 
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{author_popularity_histogram.png}
+\caption{Histogram of Bing Search Results}
+\label{fig:ap_hist}
+\end{figure}
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{author_popularity_log_histogram.png}
+\caption{Histogram of log(Bing Search Results)}
+\label{fig:apl_hist}
+\end{figure}
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{flesch_data_histogram.png}
+\caption{Histogram of Flesch Reading Ease}
+\label{fig:fre_hist}
+\end{figure}
+
 \end{description}
@@
 
+\begin{figure*}[ht]\centering
+\includegraphics[width=\linewidth]{feature_weights.png}
+\caption{Linear Regression Feature Weights (excluding intercept term)}
+\label{fig:lin_reg_weights}
+\end{figure*}
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{mse_plot.png}
+\caption{The effect of regularization on MSE}
+\label{fig:mse_plot}
+\end{figure}
+
+\begin{center}
+\begin{tabular}{lr}
+\toprule
+Feature  &  Weight\\
+\midrule
+intercept & 9.114\\
+log(Word Count) & 0.736\\
+Desk: Foreign & 0.561\\
+Desk: Travel & -0.478\\
+Section: World & -0.410\\
+Type: BlogPost & -0.394\\
+Type of Material: Schedule & -0.330\\
+Type of Material: Review & -0.324\\
+Type of Material: Letter & -0.311\\
+Section: Fashion \& Style & 0.275\\
+Section: Movies & -0.267\\
+Desk: National & 0.245\\
+Section: Opinion & 0.245\\
+Time of Day: 12-17 & 0.239\\
+Section: Books & 0.219\\
+Section: Health & 0.200\\
+Section: Sports & -0.191\\
+Desk: Society & -0.178\\
+Section: Corrections & -0.167\\
+Desk: BookReview & -0.167\\
+\end{tabular}
+\end{center}
 %------------------------------------------------

### 2015-12-09 Project/regression/plots.R

@@ -1,14 +1,68 @@
 library(ggplot2)
 library(knitr)
+library(dplyr)
+library(tidyr)
+library(stringr)
+setwd('/Users/dholtz/Desktop/mit_6867/Project/Latex/')
 
-weights <- c(-0.0155673547965, -0.032512577208, -0.0183928993822, 0.0436023817917, -0.168735260358, -0.0548502431781, 0.0400864539071, 0.0477745936425, 0.0482044874247, 0.656851318914, 0.101863103466, 0.0787925572846, 0.0772113800894, -0.109357780179, 0.00259863923261, 0.253263211162, 0.00423288154355, -0.151785209635, -0.086932735094, 0.0642537764487, 0.0594296289471, -0.0784431104261, 0.0100914683538, -0.0455614111585, -0.150498327579, 0.0370920476908, -0.471571091079, 0.00575529953804, -0.174645309934, -0.00122143569646, -0.120772053981, 0.729336915393, 0.0191174119751, -0.0296600316436, 0.0414784388088, 0.192408869835, 0.0525969201364, -0.0430805415879, -0.179258076024, -0.13951006772, 0.0400864539072, -0.199291009302, 0.125170423372, -0.00107423696482, 0.1722575493, -0.103366129896, -0.0142565167554, -0.0481040689734, -0.277144753309, -0.00275229839806, 0.123937726235, 0.111792685733, 0.261123875719, -0.0202035470324, -0.0667582303632, 0.0161341447383, -0.214593015947, 0.0525950326542, 0.0457459855028, 0.0809072398456, 0.0841238574083, -0.0947854358489, -0.0536801646975, -0.105109365666, -0.553358112213, -0.00518030191328, -0.00801186882178, 0.026662047528, -0.0707252848644, -0.00164194215674, -0.00696646258888, -0.0156368327091, -0.243055275894, -0.0265795567348, 0.0236131786527, 0.126044421422, 0.0661457097531, -0.110683337504, 0.0679989404147, 0.167213483369, -0.0287554883533, 0.0611867349866, -0.0321046144057, -0.255398058467, -0.290623015012, 0.0022774254305, -0.0126711530544, -0.0831671026105, -0.0064248106726, -0.170417879777, 9.04696649847)
[…]

### 2015-12-09 Project/regression/plots.R
[project] scales.

@@ -3,6 +3,7 @@ library(knitr)
 library(dplyr)
 library(tidyr)
 library(stringr)
+library(scales)
 setwd('/Users/dholtz/Desktop/mit_6867/Project/Latex/')
 
 weights <- c(-0.00191432502654, -0.0245399612398, -0.0150996164294, 0.0440296750968, -0.167166767298, -0.0396022800434, 0.0513236665953, 0.0376881505235, 0.034420292747, 0.560583168704, 0.135760449819, 0.119415176685, 0.0976386412178, -0.109717531939, 0.000263172714949, 0.245491206079, 0.00520118309151, -0.165878067649, -0.0892505580431, 0.0776715694698, 0.0461934763289, -0.177856267015, 0.0250636545508, -0.123728689152, -0.146765225723, 0.0466043122084, -0.477516203847, 0.00122058106983, -0.124056938997, 0.00101097778077, -0.0958951336039, 0.735969046607, 0.0167414635067, -0.025959165804, 0.040895009896, 0.219165180421, 0.108392843047, -0.0230247100238, -0.167360937146, -0.143143446175, 0.0513236665944, -0.13857883042, 0.274650293308, -0.00203664259554, 0.20002254626, -0.110278613843, -0.015642686809, -0.0503790999032, -0.267093153066, 0.00630394482605, 0.13920691866, 0.142020776932, 0.244810158572, -0.0206936510913, -0.0641111765411, 0.0342515812982, -0.190922375506, 0.0507730247675, 0.0552535755133, 0.074599895454, 0.0932262989123, -0.0826574211138, -0.0541942203966, -0.0675520953361, -0.409982572288, -0.00532750280566, 0.238705410021, 0.156883824326, 0.163951257457, -0.0177637990337, 0.0234011367299, -0.153357482844, -0.0168654857936, 0.00441377186855, -0.0265800308445, -0.310952228957, -0.0486760700795, 0.0211400997113, 0.00150676823874, 0.0435852536984, -0.00317176562678, 0.0389422324357, 0.15528305309, -0.0603885589553, 0.0301666928586, -0.0452872719984, -0.323507092898, -0.329889081099, -0.00411448761229, -0.0133846358151, -0.0934607743805, -0.00825110975912, -0.394127027181, 0.0491173950399, 0.0599957920327, 0.0289952809853, 0.00144497100566, 0.0541630652262, 0.0187797252149, 9.11388341686)

### 2015-12-10 Project/Latex/Paper.tex
[project] first pass at regression discussion.

@@
 
+Once we have constructed our full set of features, we aim to predict log(article pageviews), $\mathbf{y}$, using our full set of variables, which we include in our design matrix, $\mathbf{X}$.
+We estimate the feature weights using the closed form solution for both OLS and ridge regression,
+
+\begin{center}
+\begin{equation}
+\beta = (\mathbf{X}^T\mathbf{X} + \lambda \mathbf{I})^{-1}\mathbf{X}^T\mathbf{y}
+\end{equation}
+\end{center}
+
+\noindent where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter.
+Setting $\lambda = 0$ corresponds to OLS, whereas non-zero values of $\lambda$ correspond to ridge regression.
+The motivation for performing ridge regression as opposed to OLS is to not overfit on our data.
+
+In order to choose an appropriate value of $\lambda$, we split our data into a training and validation set.
+90\% of the data is allocated to the training set, and 10\% of the data is allocated to the test set.
+We cross-validate the estimated $\beta$ values on our validation set for each $\lambda$, and choose the value of $\lambda$ that produces the lowest MSE on the validation data.
+
+The table below displays the 20 feature weights with the largest magnitudes given our model.
+A bar chart showing the magnitude of all of the weights (excluding the intercept weight) can be found in Figure \ref{fig:lin_reg_weights}.
+
 \begin{figure*}[ht]\centering
@@
 \end{figure*}
+
+Figure \ref{fig:mse_plot} shows the training and holdout MSE for various values of Lambda.
+There are a few things in this plot worth discussing.
+First, note that the holdout MSE is consistently lower than the training data MSE.
+Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training and validation data.
+However, this shouldn't effect the validity of our cross validation.
+
+Another thing worth noting is that even on the training dataset, the MSE goes down once we choose a non-zero value of $\lambda$.
+At first, this was surprising to our group, as conceptually OLS is often thought of as the linear regression method that minimizes MSE.
+However, it is important to note that OLS only holds this distinction amongst unbiased estimators.
+Hoerl and Kennard (1970) \cite{hoerl1970ridge} prove the existence theorem for ridge regression, which proves that there exists a value of $\lambda$ for which $\beta_{ridge}$ produces a lower MSE than $\beta_{OLS}$.
+
+Another way of framing this finding is through bias-variance tradeoff. Recall that the MSE can be written as a function of the bias and variance:
+
+\begin{equation}
+MSE = (\textrm{Bias})^2 + \textrm{Var}.
+\end{equation}
+
+\noindent For some values of $\lambda$, ridge regression is able to lower the MSE by decreasing variance, but adding a non-zero bias term.
+We believe this is what is happening in our data when cycling through different $\lambda$ values.
 

### 2015-12-10 Project/Latex/Paper.tex
[project] sentiment.

@@
 
+In order to measure article sentiment, we use a Naives Bayes text classification algorithm, as described in Rennie et al (2003) \cite{rennie2003tackling}.
+We assume that each article in our corpus can belong to one of three classes - `negative' sentiment, `neutral' sentiment, and `positive' sentiment, which we will denote as $C_k$.
+The Naive Bayes model assumes that the likelihood of observing a given article $\mathbf{x} = (x_1, ..., 1_n)$, where $x_i$ is the number of times that word $i$ appears, is
+
+\begin{equation}
+p(\mathbf{x}|C_k) = \frac{\left ( \sum_i x_i \right )!}{\prod_i x_i !} \prod_i p_{ki}^{x_i},
+\end{equation}
+
+\noindent where $p_{ki}$ is the probability of a document of class $K$ including word $i$. Given that we transform this problem into log space, we can compute $\log(p(\mathbf{x}|C_k))$ as
+
+\begin{equation}
+\log(p(\mathbf{x}|C_k)) = \log ( p(C_k)) + \sum_{i=1}^n x_i \cdot \log(p_{ki}).
+\end{equation}
+
+We coded up a basic implementation of the Naive Bayes algorithm, drawing heavy inspiration from Greg Lamp's 2014 python tutorial on Naive Bayes \cite{NaiveBayes}.
+In order to get the probabilities $p(C_k)$ and $p_{ki}$, we developed a training set of NY times data by providing 200 articles to workers on Amazon Mechanical Turk.
+Each turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive.
+We recorded 5 scores for every article, and calculated the average sentiment reported by the Turkers.
+We classified any article having an average score greater than 0.5 as `positive'.
+Any article with an average sentiment less than -0.5 was `negative.' Any other articles were classified as `neutral.' Ultimately, our labels were 66\% neutral, 14.5\% negative, and 19.5\% positive.
+This is unsurprising, as a newspaper such as the New York Times likely strives for neutrality when reporting on most topics.
+
 \subsection{Article Perplexity}

### 2015-12-10 Project/feature_extraction/sentiment/sentiment_nltk_evaluate.py

@@ -1,0 +1,64 @@
+import re
+import string
+from prettytable import PrettyTable
+import pandas as pd 
+import pickle 
+import nltk
+import csv
+
+classifier = pickle.load(open('naive_bayes_nltk.pkl', 'rb'))
+
+def is_string(text):
+    return isinstance(text, basestring)
+
+def remove_punctuation(s):
+    "see http://stackoverflow.com/questions/265960/best-way-to-strip-punctuation-from-a-string-in-python"
+    table = string.maketrans("","")
+    return s.translate(table, string.punctuation)
+
+def tokenize(text):
+    dictionary = {}
+    text = remove_punctuation(text)
+    text = text.lower()
+    listy = re.split("\W+", text)
+    for i in listy:
+        dictionary[i] = True 
+    return dictionary
+
+def tokenize_2(text):
+    text = remove_punctuation(text)
+    text = text.lower()
+    return re.split("\W+", text)
+
+def count_words(words):
+    wc = {}
+    for word in words:
+        wc[word] = wc.get(word, 0.0) + 1.0
+    return wc    
+
+MTurk_results = pd.read_csv('MTurk_Results.csv')
+MTurk_results_with_indices = pd.read_csv('sample_articles_with_indices.csv')
+MTurk_results_with_indices_merged = MTurk_results_with_indices.merge(MTurk_results, on=['text'])
+MTurk_results_with_indices_merged['is_string'] = MTurk_results_with_indices_merged.text.apply(is_string)
+MTurk_clean = MTurk_results_with_indices_merged[MTurk_results_with_indices_merged['is_string']]
+
+all_words = set(word.lower() for passage in MTurk_clean['text'] for word in tokenize_2(passage))
+articles = pd.read_csv('articles.csv')
+
+articles_for_validation = articles.head(n=1000)
+
+k = 0
+
+def naive_bayes_nltk(text):
+	global k 
+	k += 1 
+	print k 
+	if isinstance(text, basestring):
+		text_for_classification = {word: (word in tokenize_2(text)) for word in all_words}
+		return classifier.classify(text_for_classification)
+	else:
+		return 'neutral'
+
+articles_for_validation['naive_bayes_nltk'] = articles_for_validation.text.apply(naive_bayes_nltk)
+
+articles_for_validation[['index', 'naive_bayes_nltk']].to_csv('naive_bayes_nltk.csv', index=False, quotechar='', quoting=csv.QUOTE_NONE, escapechar=' ', sep=',', encoding='utf8')

### 2015-12-10 Project/regression/regression.py

@@ -45,6 +45,10 @@ fleschKincaid = fleschKincaid.ix[:, fleschKincaid.columns - ['index_value']]
 bigramPerplexity = pickle.load(open('bigram-perplexity.pkl', 'rb'))
 articleData = pd.concat([articleData, bigramPerplexity], axis=1)
 
+sentiment_labels = pd.read_csv('../feature_extraction/sentiment/naive_bayes.csv')
+sentiment_labels.columns = ['id', 'sentiment']
+articleData = articleData.merge(sentiment_labels, on=['id'])
+
 def get_author_gender(author_array):
     genders = []
 
@@ -105,7 +109,7 @@ articleData['popularity_pre_log'] = articleData.authors.apply(get_author_popular
 articleData['weekday'] = articleData.publishedDate.apply(get_weekday)
 articleData['timeofday'] = articleData.publishedDate.apply(get_hour_chunk)
 
-dummyColumns = ['typeOfMaterial', 'desk', 'type', 'section', 'author_gender', 'weekday', 'timeofday']
+dummyColumns = ['typeOfMaterial', 'desk', 'type', 'section', 'author_gender', 'weekday', 'timeofday', 'sentiment']
 articleDataDummies = create_dummy_variables(articleData, dummyColumns)
 articleDataDummies = articleDataDummies.merge(fleschKincaid, on=['id'])
 articleDataDummies['flesch-kincaid_score'].fillna(articleDataDummies['flesch-kincaid_score'].median(), inplace=True)
@@ -116,7 +120,7 @@ articleDataDummies['log_wcount'] = articleDataDummies.wcount.apply(np.log)
 
 columnstoRemove = ['authors', 'desFacet', 'geoFacet', 'headline', 'id', 'orgFacet', 'perFacet', 'publishedDate', 
 'count', 'popularity_pre_log', 'log_count', 'author_gender_male', 'desk_Business', 'section_Arts', 'type_Article', 
-'typeOfMaterial_Interview', 'wcount', 'weekday_Friday', 'timeofday_0-5']
+'typeOfMaterial_Interview', 'wcount', 'weekday_Friday', 'timeofday_0-5', 'sentiment_neutral']
 articleDataDummiesRegression = articleDataDummies.ix[:, articleDataDummies.columns - columnstoRemove]
 intercept = pd.Series(np.ones(len(articleDataDummiesRegression)))
 articleDataDummiesRegression = articleDataDummiesRegression.apply(scale_features, axis=0)

### 2015-12-10 Project/Latex/Paper.tex
[project] update some plots.

@@
 
+We wanted to determine how our Naive Bayes implementation did compared to an off-the-shelf implementation of the same algorithm.
+In order to do so, we trained NLTK's multinomial Naive Bayes classifier on the same training data, and then evaluated the sentiment of 1,000 articles.
+A comparison of the two implementations is found below:
+
 \subsection{Article Perplexity}

### 2015-12-10 Project/regression/plots.R

@@ -6,13 +6,13 @@ library(stringr)
 library(scales)
 setwd('/Users/dholtz/Desktop/mit_6867/Project/Latex/')
 
-weights <- c(-0.00191432502654, -0.0245399612398, -0.0150996164294, 0.0440296750968, -0.167166767298, -0.0396022800434, 0.0513236665953, 0.0376881505235, 0.034420292747, 0.560583168704, 0.135760449819, 0.119415176685, 0.0976386412178, -0.109717531939, 0.000263172714949, 0.245491206079, 0.00520118309151, -0.165878067649, -0.0892505580431, 0.0776715694698, 0.0461934763289, -0.177856267015, 0.0250636545508, -0.123728689152, -0.146765225723, 0.0466043122084, -0.477516203847, 0.00122058106983, -0.124056938997, 0.00101097778077, -0.0958951336039, 0.735969046607, 0.0167414635067, -0.025959165804, 0.040895009896, 0.219165180421, 0.108392843047, -0.0230247100238, -0.167360937146, -0.143143446175, 0.0513236665944, -0.13857883042, 0.274650293308, -0.00203664259554, 0.20002254626, -0.110278613843, -0.015642686809, -0.0503790999032, -0.267093153066, 0.00630394482605, 0.13920691866, 0.142020776932, 0.244810158572, -0.0206936510913, -0.0641111765411, 0.0342515812982, -0.190922375506, 0.0507730247675, 0.0552535755133, 0.074599895454, 0.0932262989123, -0.0826574211138, -0.0541942203966, -0.0675520953361, -0.409982572288, -0.00532750280566, 0.238705410021, 0.156883824326, 0.163951257457, -0.0177637990337, 0.0234011367299, -0.153357482844, -0.0168654857936, 0.00441377186855, -0.0265800308445, -0.310952228957, -0.0486760700795, 0.0211400997113, 0.00150676823874, 0.0435852536984, -0.00317176562678, 0.0389422324357, 0.15528305309, -0.0603885589553, 0.0301666928586, -0.0452872719984, -0.323507092898, -0.329889081099, -0.00411448761229, -0.0133846358151, -0.0934607743805, -0.00825110975912, -0.394127027181, 0.0491173950399, 0.0599957920327, 0.0289952809853, 0.00144497100566, 0.0541630652262, 0.0187797252149, 9.11388341686)
[…]

### 2015-12-10 Project/Latex/Paper.tex
[project] more writing.

@@
 In order to do so, we trained NLTK's multinomial Naive Bayes classifier on the same training data, and then evaluated the sentiment of 1,000 articles.
-A comparison of the two implementations is found below:
+A comparison of the two implementations is found below, where the columns indicate the prediction by the NLTK Naive Bayes implementation and the rows indicate the prediction by our implementation:
+
+\vspace{10 mm}
+
+\begin{tabular}{l | r r r}
+\toprule
+  & negative & neutral & positive\\
+\midrule
+negative & 0 & 54 & 0\\
+neutral & 2 & 894 & 0\\
+positive & 0 & 50 & 0\\
+\end{tabular}
+
+\vspace{10 mm}
+
+Overall, we find 89.4\% agreement between the two algorithms.
+However, alarmingly, the NLTK implementation seems to predict neutral an overwhelming percentage of the time (99.8\% of the time).
+This warrants further investigation, and may be due to small differences in implementation, or peculiarities in the sample of 1,000 articles we chose to compare the two algorithms across.
+In any case, our algorithm seems to be performing in the same neighborhood as the NLTK implementation (if not better), so we feel relatively comfortable moving forward using our sentiment labels.
 

### 2015-12-10 Project/regression/plots.R

@@ -6,8 +6,8 @@ library(stringr)
 library(scales)
 setwd('/Users/dholtz/Desktop/mit_6867/Project/Latex/')
 
-weights <- c(-0.00623197060315, -0.0242271706056, -0.014351445239, 0.0441530957983, -0.183141438435, -0.070792093083, 0.0479733002858, 0.0356982720475, 0.0162439138474, 0.579854386616, 0.120247858437, 0.1171156196, 0.0942094479819, -0.105564058985, -0.00421410048242, 0.278003363208, 0.0363435856959, -0.0859070347175, -0.109571640119, 0.0640607591967, 0.0346550859402, -0.240917647691, 0.0276850189048, -0.186985947534, -0.204141702973, 0.0448847885443, -0.482790157831, 0.00272621354153, -0.106190380173, 0.00398383468196, -0.101871591649, 0.733401300913, 0.0193655004814, -0.0327953544567, 0.0354846836801, 0.235394877441, 0.135712802758, -0.0316455913835, -0.125844772176, -0.144438638638, 0.047973300281, -0.14405519998, 0.349880569091, 0.00528367806774, 0.192871661011, -0.103270596605, -0.0124195747141, -0.0493038049738, -0.268649476666, 0.00678743426196, 0.135766399505, 0.127683656849, 0.238839384074, -0.0196807004655, -0.0568528106611, 0.0224381468046, -0.209821773103, 0.0440742363626, 0.051248749088, 0.0710445438536, 0.092186731318, -0.0869283836069, -0.0537041550833, -0.115784992925, -0.452823135391, -0.00771210916169, 0.0103434730221, -0.0249938994228, 0.260697275868, 0.152143644462, 0.19517491076, -0.0110564008367, 0.0267366730432, -0.11583385892, -0.000500866474, 0.00600399969973, 0.0040386595581, -0.294996109746, -0.0368505591042, 0.022672483996, 0.0541801991396, 0.0592330487156, -0.000601681234961, 0.0593623261442, 0.190843349729, -0.0418335251941, 0.0533460621188, -0.0318690621793, -0.302743426906, -0.319668694839, -0.0148220651828, -0.0130394903733, -0.0917688101034, -0.00721212406133, -0.516239251888, 0.0266169956416, 0.0631139305705, 0.00850334672107, -0.00545365435006, 0.0555282923315, 0.0129282301541, 9.12145353103)
[…]

### 2015-12-10 Project/regression/regression.py

@@ -42,7 +42,12 @@ fleschKincaid = pd.DataFrame.from_csv('../feature_extraction/readability/flesch_
 fleschKincaid['id'] = fleschKincaid['index_value']
 fleschKincaid = fleschKincaid.ix[:, fleschKincaid.columns - ['index_value']]
 
+trigramPerplexity = pickle.load(open('trigram-nn-perplexity.pkl', 'rb'))
+trigramPerplexity.columns = ['Trigram Perplexity']
+articleData = pd.concat([articleData, trigramPerplexity], axis=1)
+
 bigramPerplexity = pickle.load(open('bigram-perplexity.pkl', 'rb'))
+bigramPerplexity.columns = ['Bigram Perplexity']
 articleData = pd.concat([articleData, bigramPerplexity], axis=1)
 
 sentiment_labels = pd.read_csv('../feature_extraction/sentiment/naive_bayes.csv')
@@ -104,6 +109,9 @@ def scale_features(array):
 	scaled_array = (array - np.mean(array))/np.std(array)
 	return scaled_array
 
+def square(n):
+    return n*n
+
 articleData['author_gender'] = articleData.authors.apply(get_author_gender)
 articleData['popularity_pre_log'] = articleData.authors.apply(get_author_popularity)
 articleData['weekday'] = articleData.publishedDate.apply(get_weekday)
@@ -117,10 +125,12 @@ articleDataDummies['flesch-kincaid_score'].fillna(articleDataDummies['flesch-kin
 articleDataDummies['log_count'] = articleDataDummies['count'].apply(np.log)
 articleDataDummies['log_popularity'] = articleDataDummies.popularity_pre_log.apply(np.log)
 articleDataDummies['log_wcount'] = articleDataDummies.wcount.apply(np.log)
+articleDataDummies['Squared Trigram Perplexity'] = articleDataDummies['Trigram Perplexity'].apply(square)
+articleDataDummies['Squared Bigram Perplexity'] = articleDataDummies['Bigram Perplexity'].apply(square)
 
 columnstoRemove = ['authors', 'desFacet', 'geoFacet', 'headline', 'id', 'orgFacet', 'perFacet', 'publishedDate', 
 'count', 'popularity_pre_log', 'log_count', 'author_gender_male', 'desk_Business', 'section_Arts', 'type_Article', 
-'typeOfMaterial_Interview', 'wcount', 'weekday_Friday', 'timeofday_0-5', 'sentiment_neutral']
[…]

### 2015-12-10 Project/Latex/Paper.tex
Latex additions

--- 2015-12-10 Project/Latex/Paper.tex
@@
 
-In order to measure article sentiment, we use a Naives Bayes text classification algorithm, as described in Rennie et al (2003) \cite{rennie2003tackling}.
-We assume that each article in our corpus can belong to one of three classes - `negative' sentiment, `neutral' sentiment, and `positive' sentiment, which we will denote as $C_k$.
-The Naive Bayes model assumes that the likelihood of observing a given article $\mathbf{x} = (x_1, ..., 1_n)$, where $x_i$ is the number of times that word $i$ appears, is
+\subsection{Article Perplexity}
+In order to determine the perplexity score, we first need to build some language model that gives us the probabilities of each word.
+While perplexity typically is a measure of how well a probability distribution can predict a sample, in our context, we interpret perplexity essentially as a measure of article “uniqueness”.
+The argument here is that if our language model can’t predict the language used in article very well, then the language used in the article is atypical relative to the corpus used to build the language model.
+Hence, given some language model, an article's perplexity is given by:
+\begin{equation}
+	2^{-n\cdot\sum_{i=1}^n \log p(w_i)}
+\end{equation}
+where $n$ is the length of the article, and $p(w_i)$ is the probability of the $i$-th word in the article.
+As for out paper, we consider a couple of choices for our language model.
+First, we build a simple bi-gram language to establish a baseline.
+We also build more sophisticated word vector based n-gram neural network language models.
 
+We split our articles into training (~70\%), validation (~15\%), and test sets (~15\%).
+We build our corpus only from the text of the articles in the training dataset.
+To build our bigram model, we simply need to compute the counts in our training corpus.
+Specifically, the probability of some word $w_i$ conditional on its preceding word $w_{i-1}$ is given by:
 \begin{equation}
-p(\mathbf{x}|C_k) = \frac{\left ( \sum_i x_i \right )!}{\prod_i x_i !} \prod_i p_{ki}^{x_i},
+p(w_i|w_{i-1}) = \frac{\textnormal{count}(w_{i-1},w_i)}{\textnormal{count}}
 \end{equation}
+However, its reasonable to expect that there might be bigrams in the development or test corpora that aren't observed in the training corpus.
+If this is the case, then any article with an unobserved bigram would be assigned a predicted likelihood of 0.
+Needless to say, this is very bad.
+In order to avoid this issue, we apply a technique called add-$\alpha$ smoothing.
+Essentially what add-$\alpha$ smoothing does is it adds counts to every single possible bigram so that no bigram, given a fixed vocabulary $V$, will ever have 0 probability.
+This changes our our word probabilities to:
+\begin{equation}
+p(w_i|w_{i-1}) = \frac{\textnormal{count}(w_{i-1},w_i) + \alpha}{\textnormal{count} + \alpha |V|}
+\end{equation}
+where $|V|$ is the size of the vocabulary.
+In the bigram case, we can consider that words follow a multinomial distribution conditional on its preceding word.
+Essentially, what's happening here is we're applying a symmetric dirichlet prior with parameter $\alpha$ to each of these multinomial distributions.
 
-\noindent where $p_{ki}$ is the probability of a document of class $K$ including word $i$. Given that we transform this problem into log space, we can compute $\log(p(\mathbf{x}|C_k))$ as
+We use our validation set to determine the optimal value of of $\alpha$ by seeing what value of $\alpha$ minimizes the negative average log-likelihood per word (NALL) of our validation corpus:
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{bigram_add_a.png}
+\caption{Average Negative Log-Likelihood per Word as a Function of $\alpha$}
+\label{fig:fre_hist}
+\end{figure}
+Here we see exactly what we should expect.
+The NALL on the training corpus is lower than the NALL on the validation corpus.
+Furthermore, the NALL on our training corpus monotonically increases as we increase $alpha$ while the NALL on the validation corpus initially decreases hits a minimum, and starts to increase.
+For our language model, the optimal $\alpha$ is 0.004.
+This produces the following NALLs for our 3 corpora:
 
-\begin{equation}
-\log(p(\mathbf{x}|C_k)) = \log ( p(C_k)) + \sum_{i=1}^n x_i \cdot \log(p_{ki}).
-\end{equation}
+\begin{table}[hbt]
+\caption{NALL for Bigram Language Model with Smoothing}
+\centering
+\begin{tabular}{ccc}
+\toprule
+Training &  Validation & Test\\
+\midrule
+4.98487 & 6.42916 &  6.41888\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-We coded up a basic implementation of the Naive Bayes algorithm, drawing heavy inspiration from Greg Lamp's 2014 python tutorial on Naive Bayes \cite{NaiveBayes}.
-In order to get the probabilities $p(C_k)$ and $p_{ki}$, we developed a training set of NY times data by providing 200 articles to workers on Amazon Mechanical Turk.
-Each turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive.
-We recorded 5 scores for every article, and calculated the average sentiment reported by the Turkers.
-We classified any article having an average score greater than 0.5 as `positive'.
-Any article with an average sentiment less than -0.5 was `negative.' Any other articles were classified as `neutral.' Ultimately, our labels were 66\% neutral, 14.5\% negative, and 19.5\% positive.
-This is unsurprising, as a newspaper such as the New York Times likely strives for neutrality when reporting on most topics.
+One criticism of standard n-gram language models is that they are rather sensitive to the training data.
+Suppose that two words generally have fairly interchangeable use cases (meaning that the same or similar words tend to appear before these words), for example, ``coffee" and ``tea".
+Further suppose that for whatever reason, the phrase ``drink coffee" is predominantly featured in the training data but ``drink tea" is not.
+Then a standard bigram language model would assign a very low probability to ``drink tea" even if the in the training data contains many instances in which they are fairly interchangeable (``buy coffee/tea", ``brew coffee/tea", etc.).
 
-\subsection{Article Perplexity}
+Hence, we build word vector n-gram neural network models to see if we can achieve better performance.
+By using dense word vectors rather than one-hot encodings to represent words, we're able to more accurately represent the ``similarity" of words.
+For example, ``coffee" is more similar to ``tea" that it is to ``car".
+Rather astonishingly, these word vector values can be trained through back propagation just like the weights in a neural network just as long as they're randomly initiated!
+
+ In particular, we train both a bigram neural network language model and a trigram neural network language model.
+In both cases, we use word vectors of length 10 (for the bigram NN, our input layer is size 10 and for the trigram NN, the input layer is size 20) that map to a single hidden layer with 10 hidden units with a tanh activation function.
+The hidden layer is then connected to the output layer which is the size of our vocabulary with a softmax activation function.
+Given that we want to also train our word vectors, we use stochastic adaptive gradient descent.
+We purposefully keep our network relatively small since our data set is rather large and training even this small neural network takes considerable time.
+
+ Unfortunately, our neural networks are still in the process of training.
+Due to various complications, we were not able to train our model for as long as we would've liked.
+Couple this with the fact that our training corpus contains over 3,000,000 observations and each complete pass through it takes approximately 4.5 hours, means that as of writing this paper, we've only completed 4 full passes.
+However, our neural network is working as intended and with each pass through the training data, we see decreasing NALL:
+
+ %% GRAPH AND TABLE HERE
+
+Despite the fact that our training isn't finished, both our neural network-based language models are already performing better than a standard bi-gram model with smoothing.
+To build our article perplexity scores, we use the most recently completed pass of our trigram NN language models.
+
+Its important to note however, that we don't necessarily care exactly how good our language model is since we are the language model is an intermediate step to construct a feature for a regression task.
+It could very well be the case features generated by a worse language model can work just as well as features generated from a better language model if the relative variation of articles is maintained.
+To test whether a better language model can give us more predictive power, we compute two sets of perplexity scores, one based on the smoothed bigram language model and one based on the trigram neural network language model.
 

### 2015-12-10 Project/Latex/Paper.tex

@@
 
+In order to measure article sentiment, we use a Naives Bayes text classification algorithm, as described in Rennie et al (2003) \cite{rennie2003tackling}.
+We assume that each article in our corpus can belong to one of three classes - `negative' sentiment, `neutral' sentiment, and `positive' sentiment, which we will denote as $C_k$.
+The Naive Bayes model assumes that the likelihood of observing a given article $\mathbf{x} = (x_1, ..., 1_n)$, where $x_i$ is the number of times that word $i$ appears, is
+
+\begin{equation}
+p(\mathbf{x}|C_k) = \frac{\left ( \sum_i x_i \right )!}{\prod_i x_i !} \prod_i p_{ki}^{x_i},
+\end{equation}
+
+\noindent where $p_{ki}$ is the probability of a document of class $K$ including word $i$. Given that we transform this problem into log space, we can compute $\log(p(\mathbf{x}|C_k))$ as
+
+\begin{equation}
+\log(p(\mathbf{x}|C_k)) = \log ( p(C_k)) + \sum_{i=1}^n x_i \cdot \log(p_{ki}).
+\end{equation}
+
+We coded up a basic implementation of the Naive Bayes algorithm, drawing heavy inspiration from Greg Lamp's 2014 python tutorial on Naive Bayes \cite{NaiveBayes}.
+In order to get the probabilities $p(C_k)$ and $p_{ki}$, we developed a training set of NY times data by providing 200 articles to workers on Amazon Mechanical Turk.
+Each turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive.
+We recorded 5 scores for every article, and calculated the average sentiment reported by the Turkers.
+We classified any article having an average score greater than 0.5 as `positive'.
+Any article with an average sentiment less than -0.5 was `negative.' Any other articles were classified as `neutral.' Ultimately, our labels were 66\% neutral, 14.5\% negative, and 19.5\% positive.
+This is unsurprising, as a newspaper such as the New York Times likely strives for neutrality when reporting on most topics.
+
+We wanted to determine how our Naive Bayes implementation did compared to an off-the-shelf implementation of the same algorithm.
+In order to do so, we trained NLTK's multinomial Naive Bayes classifier on the same training data, and then evaluated the sentiment of 1,000 articles.
+A comparison of the two implementations is found below, where the columns indicate the prediction by the NLTK Naive Bayes implementation and the rows indicate the prediction by our implementation:
+
+\vspace{10 mm}
+
+\begin{tabular}{l | r r r}
+\toprule
+  & negative & neutral & positive\\
+\midrule
+negative & 0 & 54 & 0\\
+neutral & 2 & 894 & 0\\
+positive & 0 & 50 & 0\\
+\end{tabular}
+
+\vspace{10 mm}
+
+Overall, we find 89.4\% agreement between the two algorithms.
+However, alarmingly, the NLTK implementation seems to predict neutral an overwhelming percentage of the time (99.8\% of the time).
+This warrants further investigation, and may be due to small differences in implementation, or peculiarities in the sample of 1,000 articles we chose to compare the two algorithms across.
+In any case, our algorithm seems to be performing in the same neighborhood as the NLTK implementation (if not better), so we feel relatively comfortable moving forward using our sentiment labels.
+
 \subsection{Article Perplexity}

### 2015-12-10 Project/Latex/Paper.tex
[project] update some text for the paper.

@@
 \begin{description}
-	\item[Flesch Reading Ease] The Flesch reading ease is a metric developed by Flesch in 1948 \cite{flesch1948new}.
+	\item[Flesch Reading Ease] One can conceive of a few hypotheses that relate the readership of content to the ease with which people can read it.
+Maybe more complicated pieces of text are more engaging, and are more likely to be read.
+On the other hand, perhaps pieces of text that are easier to read will be consumed by more people.
+In order to capture relationships such as these in our data, we include the Flesch reading ease.
+The Flesch reading ease is a metric developed by Flesch in 1948 \cite{flesch1948new}.
 The score indicates how difficult a piece of English text is to understand.
@@
 
-	\item[Author Gender] We also attempt to construct a feature that indicates the most likely gender of the article author(s).
+	\item[Author Gender] We certainly hope that an author's gender would not have a causal impact on whether or not a particular article gets readership.
+However, we want to allow for this fact and measure this variable.
+We do so by constructing a feature that indicates the most likely gender of the article author(s).
 In cases where the gender of the author is unclear (e.g., Robin) or there are likely multiple authors with different genders (e.g., The New York Times Staff), we record a third gender value, ``ambiguous / unknown."
@@
 	\item[Material Type, Section, Desk, and Article Type] We also include dummy variables including the material type (e.g., `News' or `Obituary'), desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), and section (e.g., `Movies' or `World').
+The hypothesis driving the decision to include these variables is that certain types of content (e.g., political news or international affairs) may be more widely read than local material (such as real estate) or less popular sections of the NY Times (e.g., the sports section).
 
@@
 \end{center}
+
+\subsection{The Effect of Textual Features}
+
+Given the large amount of work we put into building numerous text-based features (such as the trigram neural network perplexity, the sentiment, and the Flesch reading ease), we may want to evaluate how impactful these various features actually are in our regression.
+How much incremental reduction in MSE are we getting by including them?
+We first re-run the exact same regression specification, but instead of using the trigram perplexity calculated from our neural network, we instead use the simple bigram perplexity discussed earlier in this paper.
+We find that this actually causes the MSE to decrease, from 2.392 to 2.389.
+This small, but modest change suggests that there is almost no incremental value from the neural network feature.
+
+Now, we might ask if any text features are adding much value.
+Particularly given that, in general, their magnitude is dwarved by the magnitudes of the contextual features when looking at feature weights, we might expect that text features do not add much.
+When removing perplexity, sentiment, reading ease, and word count features from our regression, we find that the MSE increases, from 2.392 to 2.502.
+This is result is encouraging, as it suggests that even if text features aren't doing much, they're doing something!
+Removing them from our model leads to a 4.6\% increase in the MSE.
+
+\begin{table}[hbt]
+\caption{Comparison of MSEs for different model specifications}
+\centering
+\begin{tabular}{ccc}
+\toprule
+NN Trigram & Bigram & No Text\\
+\midrule
+2.392 & 2.389 &  2.502 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 %------------------------------------------------

### 2015-12-10 Project/Latex/Paper.tex
[project] some more info on regression.
[merge]

@@
 where $n$ is the length of the article, and $p(w_i)$ is the probability of the $i$-th word in the article.
+We think that perplexity might have some predictive power since people generally have a preference for novelty.
+If many news articles about the same story are all using highly similar language, an article that covers the story using atypical language is likely unique in some way or another which may drive people to read it more or less.
+
 As for out paper, we consider a couple of choices for our language model.
@@
 We split our articles into training (~70\%), validation (~15\%), and test sets (~15\%).
-We build our corpus only from the text of the articles in the training dataset.
-To build our bigram model, we simply need to compute the counts in our training corpus.
+We build our language model corpus only from the text of the articles in the training dataset.
+To estimate a bigram model, we simply need to compute the counts in our training corpus.
 Specifically, the probability of some word $w_i$ conditional on its preceding word $w_{i-1}$ is given by:
@@
 In order to avoid this issue, we apply a technique called add-$\alpha$ smoothing.
-Essentially what add-$\alpha$ smoothing does is it adds counts to every single possible bigram so that no bigram, given a fixed vocabulary $V$, will ever have 0 probability.
-This changes our our word probabilities to:
+Essentially what add-$\alpha$ smoothing does is it adds (possibly fractional) counts to every single possible bigram.
+After smoothing, no bigram, given a fixed vocabulary $V$, will ever have 0 probability.
+This changes our our estimated word probabilities to:
 \begin{equation}
@@
 where $|V|$ is the size of the vocabulary.
-In the bigram case, we can consider that words follow a multinomial distribution conditional on its preceding word.
-Essentially, what's happening here is we're applying a symmetric dirichlet prior with parameter $\alpha$ to each of these multinomial distributions.
+This smoothing actually has quite a nice Bayesian interpretation as well.
+In the bigram case, we can say that words follow a multinomial distribution conditional on its preceding word.
+The smoothed language model
+
+Essentially, is we're applying a symmetric dirichlet prior with parameter $\alpha$ to each of these multinomial distributions.
 
@@
 Now, we might ask if any text features are adding much value.
-Particularly given that, in general, their magnitude is dwarved by the magnitudes of the contextual features when looking at feature weights, we might expect that text features do not add much.
+Particularly given that, in general, their magnitude is dwarfed by the magnitudes of the contextual features when looking at feature weights, we might expect that text features do not add much.
 When removing perplexity, sentiment, reading ease, and word count features from our regression, we find that the MSE increases, from 2.392 to 2.502.
@@
 
+\section{Future Work}
+Unfortunately due to time constraints as well as various unanticipated difficulties and technical errors, we were unable to implement all of several important features we had initially wanted to include in our regression.
+Most notably, this paper in its current state lacks some form of topic modeling which we expect to be a strong predictor of content viewership at least in our context.
+
+There is also much room for improvement with our existing features as well.
+For example, the perplexity score feature ended providing very little predictive power might be a result of the underlying language model not being good enough, especially considering that the neural network model still hasn't completed its training.
+Though we did observe that there was virtually no difference moving from the bigram perplexity scores to the trigram neural network perplexity scores, this again might be because the neural network wasn't given sufficient time to train.
+We can imagine that if we continued training our neural network for several more cycles, we might start to see visible performance gains.
+
+Similarly, the sentiment analysis has much room for improvement...
+
+Certainly one thing we definitely plan on doing if we end up trying to turn this project into a publishable paper is to include all of the articles in our dataset.
+With the much greater sample size and the much more expansive corpus from the additional articles, we might expect to see drastically different
 %------------------------------------------------

### 2015-12-10 Project/Latex/Paper.tex
[project] update that table.

--- 2015-12-10 Project/Latex/Paper.tex
@@
 
+\begin{figure*}[ht]\centering
+\includegraphics[width=\linewidth]{feature_weights_contd.png}
+\caption{Linear Regression Feature Weights (excluding intercept term)}
+\label{fig:lin_reg_weights}
+\end{figure*}
+
 Figure \ref{fig:mse_plot} shows the training and holdout MSE for various values of Lambda.
@@
 \midrule
-intercept & 9.114\\
-log(Word Count) & 0.736\\
-Desk: Foreign & 0.561\\
-Desk: Travel & -0.478\\
-Section: World & -0.410\\
-Type: BlogPost & -0.394\\
-Type of Material: Schedule & -0.330\\
-Type of Material: Review & -0.324\\
-Type of Material: Letter & -0.311\\
-Section: Fashion \& Style & 0.275\\
-Section: Movies & -0.267\\
-Desk: National & 0.245\\
-Section: Opinion & 0.245\\
-Time of Day: 12-17 & 0.239\\
-Section: Books & 0.219\\
-Section: Health & 0.200\\
-Section: Sports & -0.191\\
-Desk: Society & -0.178\\
-Section: Corrections & -0.167\\
-Desk: BookReview & -0.167\\
+intercept & 8.964\\
+log(Word Count) & 0.741\\
+Desk: Foreign & 0.550\\
+Desk: Travel & -0.465\\
+Section: World & -0.402\\
+Section: Opinion & 0.283\\
+Type: BlogPost & -0.282\\
+Type of Material: Schedule & -0.267\\
+Desk: None & -0.262\\
+Section: Movies & -0.239\\
+Time of Day: 12-17 & 0.232\\
+Type of Material: Review & -0.218\\
+Desk: National & 0.211\\
+Section: Health & 0.210\\
+Section: Books & 0.201\\
+Type of Material: Letter & -0.200\\
+Type of Material: News & 0.187\\
+Section: Sports & -0.175\\
+Type of Material: Op-Ed & 0.170\\
+Desk: BookReview & -0.154\\
 \end{tabular}
@@
 \centering
-\begin{tabular}{ccc}
+\begin{tabular}{cccc}
 \toprule
-NN Trigram & Bigram & No Text\\
+  & NN Trigram & Bigram & No Text\\
 \midrule
-2.392 & 2.389 &  2.502 \\
+Best Val. MSE & 2.392 & 2.389 &  2.502 \\
+Train. MSE & 2.011 & 2.011 &  2.274 \\
+OLS Val. MSE & 2.407 & 2.405 & 2.506 \\ 
 \bottomrule

### 2015-12-10 Project/Latex/Paper.tex

--- 2015-12-10 Project/Latex/Paper.tex
@@
 
+Another way we could improve our model is to apply a basis expansion to our variables to include polynomial and various interaction terms.
+In many cases, theres no good justification to why a feature is linearly related to the output feature.
+Hence, applying a polynomial expansion might uncover an entirely different relationship between our dependent variable and its covariates.
+We also think that interaction terms might provide some significant predictive power.
+
 Certainly one thing we definitely plan on doing if we end up trying to turn this project into a publishable paper is to include all of the articles in our dataset.
-With the much greater sample size and the much more expansive corpus from the additional articles, we might expect to see drastically different
+With the much greater sample size and the much more expansive corpus from the additional articles, we might expect to see drastically different optimal weights.
 %------------------------------------------------

### 2015-12-10 Project/Latex/Paper.tex
[project] add more future work discussion, division of labor.

@@
 
-\Authors{Dave Holtz\textsuperscript{1}, Jeremy Yang\textsuperscript{2}, Michael Zhao\textsuperscript{3}} % Authors
+\Authors{Dave Holtz\textsuperscript{1}, Jeremy Yang\textsuperscript{2*}, Michael Zhao\textsuperscript{3}} % Authors
 \affiliation{\textsuperscript{1}\textit{dholtz@mit.edu}}
@@
 
-Similarly, the sentiment analysis has much room for improvement...
+Similarly, there is ample room to improve our sentiment analysis methodologies.
+The training dataset that we generated using mechanical turk is relatively small (200 articles).
+Because of this, there are currently many words that may have strong conditional probabilities of appearing given a particular article sentiment that simply do not appear in our training data.
+With more time (and money!), we might be able to expand the size of our training data and consequently include the accuracy of our model.
+Furthermore, the model currently skips words it hasn't seen in the training corpus - it might make sense to augment our Naive Bayes implementation to incorporate a method such as Laplace smoothing so as to not simply ignore words we haven't seen in our training data.
+
+Furthermore, the current discrepancies between our implementation of Naive Bayes and NLTK's, while not large in magnitude, are alarming.
+Given more time in the future, we hope to dig deeper into what is happening.
+Our current hypothesis is that NLTK's implementation of Naive Bayes does not take into account the number of times a particular word appears in a document.
+Because of this, more distinct informative words are required to move a given article's prediction away from the neutral class, which has an extremely strong prior.
 
@@
 With the much greater sample size and the much more expansive corpus from the additional articles, we might expect to see drastically different optimal weights.
+
+\section{Division of Labor}
+
+The work for this project was divided as follows.
+Michael extracted the New York Times contextual data and page view counts.
+Dave and Jeremy built out many of the content-driven features and additional contextual features, such as word count, author gender, popularity, and Flesch reading ease.
+Dave and Jeremy built the implementation and validation of the Naive Bayes classifier, and Michael focused his efforts on the construction and validation of the language model.
+Dave, along with assistance from Michael, focused on the implementation of OLS and ridge regression, along with validation and discussion of results.
+Extenuating personal circumstances unfortunately prevented Jeremy from contributing to the project as much as originally anticipated.
+
 %------------------------------------------------
@@
 \section*{Acknowledgments} % The \section*{} command stops section numbering
+
+We would like to thank Professor Sinan Aral for his work procuring this dataset, as well as his limitless advice, mentorship, and patience. We'd also like to thank Paramveer Dhillon for his help.
 

### 2015-12-10 Project/regression/perplexity_prep.py
[project] perplexity histogram.

@@ -1,0 +1,8 @@
+import pickle 
+import pandas as pd 
+
+trigramPerplexity = pickle.load(open('trigram-nn-perplexity.pkl', 'rb'))
+trigramPerplexity.to_csv('trigram_perplexity.csv')
+
+bigramPerplexity = pickle.load(open('bigram-perplexity.pkl', 'rb'))
+bigramPerplexity.to_csv('bigram_perplexity.csv')

### 2015-12-10 Project/Latex/Paper.tex
[merge]

--- 2015-12-10 Project/Latex/Paper.tex
@@
 This activity data is stored as individual lines of json and includes who (if available) accessed what page at what time.
-Overall, it is over 20 terabytes in size and contains over 3 billion page views\footnote{Not all page views are content views, for example, some events that are also tracked are searches, or user account settings.} Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013.
+Overall, it is over 20 terabytes in size and contains over 3 billion page views.
+It's important to note that not all page views are content views, for example, some events that are also tracked are searches, or user account settings.
+Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of including the entire dataset for future work.
 
-We limit our dataset to consist of pages that only contain articles or blogposts published during the month of August.
-We parse the data to obtain a list of urls, which need to be stripped of potential garbage.
-After cleaning up the url data, we are left with 6682 unique pieces of content.
+We further constrain our dataset to only include articles or blogposts published during the month of August since these are the pieces of content that have machine parsable text data.
+For the sake of brevity, we use articles to refer to both articles and blogposts unless we explicitly state otherwise.
+We make a basic first pass through the data to simply to obtain a list of urls.
+After cleaning up the url data to make sure we had each url mapped to a unique piece of content, we were left with a total of 6682 urls.
 We then parse our dataset and aggregate the number of counts each url receives.
-In order to make the comparison between articles fairer since an article that's been out longer will have more page views on average, we only count the number of page views received up to 7 days after publication\footnote{Given that page views tend to sharply drop off soon after publication since recency is quite important to the News, the number of page views obtained during the 7 days after an article is published represents the vast majority (usually well over 90\%) of total page views an article receives.}.
-In total, our data consists of over 250 million page views.
+In order to make the comparison between articles fairer since an article that's been out longer will have more page views on average, we only count the number of page views received up to 7 days after publication.
+Given that page views tend to sharply drop off soon after publication since recency is quite important to the News, the number of page views obtained during the 7 days after an article is published represents the vast majority (usually well over 90\%) of total page views an article receives.
+In total, our data consists of 248,161,455 page views.
 As seen in Figure 1 below, the distribution of page views is highly skewed with very heavy tails.
@@
 \toprule
-Total Page Views  &  248161455\\
-\midrule
 Min               &  1\\
@@
 Median            &  10298.5\\
-Std. Dev.         &  88972.9\\
-Skewness          &  9.52191\\
-Kurtosis          &  173.061\\
+StDev.         &  88972.9\\
+Skew          &  9.52191\\
+Kurt.          &  173.061\\
 \midrule
-Observations &  6682\\
+Obs &  6682\\
 \bottomrule
@@
 Median            &  9.239754\\
-Std. Dev.         &  2.028668\\
-Skewness          &  -1.270368\\
-Kurtosis          &  3.800911\\
+StDev.            &  2.028668\\
+Skew              &  -1.270368\\
+Kurt.             &  3.800911\\
 \midrule
-Observations &  6682\\
+Obs &  6682\\
 \bottomrule
@@
 
-In addition to aggregating the counts, when
+In addition to aggregating the counts, when parser the internal web traffic data, we made sure to store various relevant article meta-data features such as the headline, time of publishing, authors, section, desk, and the NYT's internal article content description tags (if available\footnote{by this we mean that the NYT only produces these content tags only for some articles, its not a question of whether the data is available to us}).
 
 \subsection{Parsed NYT Article Content Data}
+Unfortunately, the NYT internal web traffic data does not the actual content displayed on each webpage, which is a very important aspect of our project.
+To resolve this issue, we used the python library ``newspaper".
+This library helped to scrape the html from the NYT website and extract all the raw textual data from the html.
+We further used some regular expressions to clean up what the library missed.
 
@@
 
+\begin{table}[hbt]
+\caption{Log Page Views Distribution Summary Statistics}
+\centering
+\begin{tabular}{rr}
+\toprule
+Min               &  7\\
+Max               &  8941\\  
+Mean              &  701.1405\\
+Median            &  625\\
+StDev.            &  591.9535\\
+Skew              &  3.335127\\
+Kurt.             &  24.35949\\
+\midrule
+Obs &  6682\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \begin{figure}[ht]\centering
@@
 \end{figure}
+
+\begin{table}[hbt]
+\caption{Log Page Views Distribution Summary Statistics}
+\centering
+\begin{tabular}{rr}
+\toprule
+Min               &  1.945910\\
+Max               &  9.098403\\  
+Mean              &  6.181044\\
+Median            &  6.437751\\
+StDev.            &  1.004414\\
+Skew              &  -1.178993\\
+Kurt.             &  1.944611\\
+\midrule
+Obs &  6682\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 Most notably, this paper in its current state lacks some form of topic modeling which we expect to be a strong predictor of content viewership at least in our context.
+We also think the constructing some features using only the headlines might also provide some improvement, especially considering that an article's headline usually ends up being fairly important in a user's decision to click through or not.
 
@@
 With the much greater sample size and the much more expansive corpus from the additional articles, we might expect to see drastically different optimal weights.
+
+Since we also have Twitter data that tracks every tweet and retweet involving a NYT article, we would like to extend this project to also include our Twitter data.
+We can then explore many interesting potential cross effects, such as how does the article content predict the amount of Twitter sharing and alternately, how does content of a Tweet predict viewership.
 %------------------------------------------------

### 2015-12-10 Project/Latex/Paper.tex

@@
 
-\JournalInfo{6.867 Final Paper} % Journal information
+\JournalInfo{6.864/6.867 Final Paper} % Journal information
 \Archive{} % Additional notes (e.g. copyright, DOI, review/research article)
@@
 
-\Authors{Dave Holtz\textsuperscript{1}, Jeremy Yang\textsuperscript{2}, Michael Zhao\textsuperscript{3}} % Authors
+\Authors{Dave Holtz\textsuperscript{1}, Jeremy Yang\textsuperscript{2}, Michael Zhao\textsuperscript{3*}} % Authors
 \affiliation{\textsuperscript{1}\textit{dholtz@mit.edu}}
@@
 \affiliation{\textsuperscript{3}\textit{mfzhao@mit.edu}}
+\affiliation{\textsuperscript{*}\textit{Michael is also using this project to fulfill Final Project requirements for both 6.864 as well as 6.867.
+Dave and Jeremy are only using this project for 6.867.}}
 
@@
 We feed all of our constructed features to into a predictive regression model.
-We find [MAJOR RESULTS HERE].
-}
+We find that our textual features modestly improve our predictive power.
+However, there are still many textual features we've not yet explored due to time constraints.
+We believe that adding these additional features may offer more significant improvements in performance.}
 
@@
 We further used some regular expressions to clean up what the library missed.
+In total, our 6682 articles contain 4,685,021 words of text.
+We find that the distribution of article length, much like the distribution of page views, is highly skewed.
+We again apply a log transformation.
+Histograms and various summary statistics of for the distribution of article length and log article length are provided below:
 
@@
 \end{table}
+
+In addition to extracting the raw text data, we made sure to check if the articles contained non-textual content such as pictures or videos.
+We created indicator variables that denote the presence of such content within an article.
 
@@
 positive & 0 & 50 & 0\\
+\bottomrule
 \end{tabular}
@@
 We build our language model corpus only from the text of the articles in the training dataset.
+In order to keep the size of our vocabulary relatively manageable, we ignore any case sensitivities (so ``Cat" is the same as ``cat").
+Furthermore, we only include a word if it appears at least 5 times.
+Words that don't make this cutoff are mapped to a generic ``rare word" indicator.
+Lastly, we also map any numbers (that is actual numbers with digits, not number written with words) to a generic ``number" indicator.
+This leaves us with a vocabulary size $|V|$ of 29359.
 To estimate a bigram model, we simply need to compute the counts in our training corpus.
@@
 \end{equation}
-where $|V|$ is the size of the vocabulary.
 This smoothing actually has quite a nice Bayesian interpretation as well.
@@
 \includegraphics[width=\linewidth]{bigram_add_a.png}
-\caption{Average Negative Log-Likelihood per Word as a Function of $\alpha$}
+\caption{Negative Average Log-Likelihood per Word as a Function of $\alpha$}
 \label{fig:fre_hist}
@@
 \begin{table}[hbt]
-\caption{NALL for Bigram Language Model with Smoothing}
+\caption{Bigram Language Model with Smoothing Negative Average Log-Likelihood}
 \centering
@@
  In particular, we train both a bigram neural network language model and a trigram neural network language model.
-In both cases, we use word vectors of length 10 (for the bigram NN, our input layer is size 10 and for the trigram NN, the input layer is size 20) that map to a single hidden layer with 10 hidden units with a tanh activation function.
+In both cases, we use word vectors of length 10 (so for the bigram NN, our input layer is size 10 and for the trigram NN, the input layer is size 20) that map to a single hidden layer with 10 hidden units with a tanh activation function.
 The hidden layer is then connected to the output layer which is the size of our vocabulary with a softmax activation function.
@@
 Couple this with the fact that our training corpus contains over 3,000,000 observations and each complete pass through it takes approximately 4.5 hours, means that as of writing this paper, we've only completed 4 full passes.
-However, our neural network is working as intended and with each pass through the training data, we see decreasing NALL:
+Unfortunately, as a result of an oversight in the code, we forgot to store the NALL of training data as we complete each full loop.
+However, we are highly confident our neural network is working as intended since we see we that the NALL decreases every pass on the validation corpus:
 
@@
 
+We fully expect the validation corpus NALL to eventually start increasing once the neural network begins to overfit the training corpus, however we simply haven't reached that point yet.
+
 Despite the fact that our training isn't finished, both our neural network-based language models are already performing better than a standard bi-gram model with smoothing.
-To build our article perplexity scores, we use the most recently completed pass of our trigram NN language models.
+We build 2 sets of article perplexity scores.
+One set is derived from the smoothed bigram model to serve as a base comparison.
+Our other set is derived from the most recently completed pass of our trigram NN language model.
+We report the following article perplexity scores distribution:
 
-Its important to note however, that we don't necessarily care exactly how good our language model is since we are the language model is an intermediate step to construct a feature for a regression task.
-It could very well be the case features generated by a worse language model can work just as well as features generated from a better language model if the relative variation of articles is maintained.
-To test whether a better language model can give us more predictive power, we compute two sets of perplexity scores, one based on the smoothed bigram language model and one based on the trigram neural network language model.
+%
+
+%% PUT PERPLEXITY DISTRIBUTIONS HERE
+
+%
+
+Its important to note however, that we don't necessarily care exactly how good our language model.
+Since we only use the language model as an intermediate step to construct a feature for a regression task.
+It could very well be the case features generated by a worse language model can work just as well as well in the regression task as features generated by a better language model if the relative variation of articles is maintained.
+To test whether a better language model can give us more predictive power, we compute two sets of perplexity scores, one based on the smoothed bigram language model and one based on the current best trigram neural network language model.
 
@@
 Once we have constructed our full set of features, we aim to predict log(article pageviews), $\mathbf{y}$, using our full set of variables, which we include in our design matrix, $\mathbf{X}$.
-We estimate the feature weights using the closed form solution for both OLS and ridge regression,
-
-\begin{center}
+We estimate the feature weights using the closed form solution for both OLS and ridge regression:
 \begin{equation}
@@
 \end{equation}
-\end{center}
-
-\noindent where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter.
+where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter.
 Setting $\lambda = 0$ corresponds to OLS, whereas non-zero values of $\lambda$ correspond to ridge regression.
@@
 The table below displays the 20 feature weights with the largest magnitudes given our model.
-A bar chart showing the magnitude of all of the weights (excluding the intercept weight) can be found in Figure \ref{fig:lin_reg_weights}.
+A bar chart showing the magnitude of all of the weights (excluding the intercept weight) can be found in Figures \ref{fig:lin_reg_weights1} and \ref{fig:lin_reg_weights2}.
+
+\begin{table}[hbt]
+\caption{20 Most Significant Weights}
+\centering
+\begin{tabular}{lr}
+\toprule
+Feature  &  Weight\\
+\midrule
+intercept & 8.964\\
+log(Word Count) & 0.741\\
+Desk: Foreign & 0.550\\
+Desk: Travel & -0.465\\
+Section: World & -0.402\\
+Section: Opinion & 0.283\\
+Type: BlogPost & -0.282\\
+Type of Material: Schedule & -0.267\\
+Desk: None & -0.262\\
+Section: Movies & -0.239\\
+Time of Day: 12-17 & 0.232\\
+Type of Material: Review & -0.218\\
+Desk: National & 0.211\\
+Section: Health & 0.210\\
+Section: Books & 0.201\\
+Type of Material: Letter & -0.200\\
+Type of Material: News & 0.187\\
+Section: Sports & -0.175\\
+Type of Material: Op-Ed & 0.170\\
+Desk: BookReview & -0.154\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 \caption{Linear Regression Feature Weights (excluding intercept term)}
-\label{fig:lin_reg_weights}
+\label{fig:lin_reg_weights1}
 \end{figure*}
@@
 \caption{Linear Regression Feature Weights (excluding intercept term)}
-\label{fig:lin_reg_weights}
+\label{fig:lin_reg_weights2}
 \end{figure*}
@@
 
-\begin{center}
-\begin{tabular}{lr}
-\toprule
-Feature  &  Weight\\
-\midrule
-intercept & 8.964\\
-log(Word Count) & 0.741\\
-Desk: Foreign & 0.550\\
-Desk: Travel & -0.465\\
-Section: World & -0.402\\
-Section: Opinion & 0.283\\
-Type: BlogPost & -0.282\\
-Type of Material: Schedule & -0.267\\
-Desk: None & -0.262\\
-Section: Movies & -0.239\\
-Time of Day: 12-17 & 0.232\\
-Type of Material: Review & -0.218\\
-Desk: National & 0.211\\
-Section: Health & 0.210\\
-Section: Books & 0.201\\
-Type of Material: Letter & -0.200\\
-Type of Material: News & 0.187\\
-Section: Sports & -0.175\\
-Type of Material: Op-Ed & 0.170\\
-Desk: BookReview & -0.154\\
-\end{tabular}
-\end{center}
-
 \subsection{The Effect of Textual Features}
@@
 Now, we might ask if any text features are adding much value.
-Particularly given that, in general, their magnitude is dwarfed by the magnitudes of the contextual features when looking at feature weights, we might expect that text features do not add much.
+Particularly given that, in general, the magnitude of their weights is dwarfed by the magnitude of the weights of the contextual features when looking at feature weights, we might expect that text features do not add much.
 When removing perplexity, sentiment, reading ease, and word count features from our regression, we find that the MSE increases, from 2.392 to 2.502.
@@
 We can imagine that if we continued training our neural network for several more cycles, we might start to see visible performance gains.
+Alternatively, we think its likely the its not the perplexity of an article relative to the entire corpus that matters, but rather perplexity relative to other articles that cover the same event or topic.
 
-Similarly, the sentiment analysis has much room for improvement...
+Similarly, there is ample room to improve our sentiment analysis methodologies.
+The training dataset that we generated using mechanical turk is relatively small (200 articles).
+Because of this, there are currently many words that may have strong conditional probabilities of appearing given a particular article sentiment that simply do not appear in our training data.
+With more time (and money!), we might be able to expand the size of our training data and consequently include the accuracy of our model.
+Furthermore, the model currently skips words it hasn't seen in the training corpus - it might make sense to augment our Naive Bayes implementation to incorporate a method such as Laplace smoothing so as to not simply ignore words we haven't seen in our training data.
+
+Furthermore, the current discrepancies between our implementation of Naive Bayes and NLTK's, while not large in magnitude, are alarming.
+Given more time in the future, we hope to dig deeper into what is happening.
+Our current hypothesis is that NLTK's implementation of Naive Bayes does not take into account the number of times a particular word appears in a document.
+Because of this, more distinct informative words are required to move a given article's prediction away from the neutral class, which has an extremely strong prior.
 
@@
 We can then explore many interesting potential cross effects, such as how does the article content predict the amount of Twitter sharing and alternately, how does content of a Tweet predict viewership.
+=======
+\section{Division of Labor}
+
+The work for this project was divided as follows.
+Michael extracted the New York Times contextual data and page view counts and wrote the NYT website scraping script.
+Dave and Jeremy built out many of the content-driven features and additional contextual features, such as word count, author gender, popularity, and Flesch reading ease.
+Dave and Jeremy built the implementation and validation of the Naive Bayes classifier, while Michael focused his efforts on the construction and validation of the language model.
+Dave, along with assistance from Michael, focused on the implementation of OLS and ridge regression, along with validation and discussion of results.
+Plot generation, table creation, writing, and editing was evenly split between Michael and Extenuating personal circumstances unfortunately prevented Jeremy from contributing to the project as much as originally anticipated.
+
 %------------------------------------------------
@@
 \section*{Acknowledgments} % The \section*{} command stops section numbering
+
+We would like to thank Professor Sinan Aral for his work procuring this dataset, as well as his limitless advice, mentorship, and patience.
+We'd also like to thank Paramveer Dhillon for his useful comments and his work pre-pre-processing our dataset.
 

### 2015-12-10 Project/Latex/Paper.tex
[project] perplexity stuff.
[project] add graphs to latex
[project] remove equal sign.

@@
 
- %% GRAPH AND TABLE HERE
+\begin{figure}[ht]
+\includegraphics[width=\linewidth]{nall.png}
+\caption{The negative average log-likelihood of the bigram and trigram models}
+\label{fig:perp_hist}
+\end{figure}
 
@@
 
-%
-
-%% PUT PERPLEXITY DISTRIBUTIONS HERE
-
-%
+\begin{figure}[ht]
+\includegraphics[width=\linewidth]{perplexity.png}
+\caption{The distribution of bigram and trigram perplexities}
+\label{fig:perp_hist}
+\end{figure}
 
@@
 Setting $\lambda = 0$ corresponds to OLS, whereas non-zero values of $\lambda$ correspond to ridge regression.
-The motivation for performing ridge regression as opposed to OLS is to not overfit on our data.
+The motivation for performing ridge regression as opposed to OLS is to not overfit on our data \cite{ridgeregressionNotes}.
 
@@
 We can then explore many interesting potential cross effects, such as how does the article content predict the amount of Twitter sharing and alternately, how does content of a Tweet predict viewership.
-=======
+
 \section{Division of Labor}

### 2015-12-10 Project/Latex/Paper.tex
[project] dave revise second half.

@@
 
-Once we have constructed our full set of features, we aim to predict log(article pageviews), $\mathbf{y}$, using our full set of variables, which we include in our design matrix, $\mathbf{X}$.
-We estimate the feature weights using the closed form solution for both OLS and ridge regression:
+Using our full set of features, we are now able to perform the regression task we originally had in mind, and attempt to determine how (if at all) content drives viewership.
+We regress log(article pageviews), $\mathbf{y}$, on our design matrix, $\Phi$, which includes entries for each of our $k-1$ features, plus an intercept term (in this case, $k$ = 102)..
+We estimate the feature weights using the closed form solution for OLS and ridge regression:
 \begin{equation}
-\beta = (\mathbf{X}^T\mathbf{X} + \lambda \mathbf{I})^{-1}\mathbf{X}^T\mathbf{y}
+\beta = (\Phi^T\Phi + \lambda \mathbf{I})^{-1}\Phi^T\mathbf{y},
 \end{equation}
-where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter.
-Setting $\lambda = 0$ corresponds to OLS, whereas non-zero values of $\lambda$ correspond to ridge regression.
-The motivation for performing ridge regression as opposed to OLS is to not overfit on our data \cite{ridgeregressionNotes}.
+\noindent where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter.
+Setting $\lambda = 0$ corresponds to OLS, whereas a non-zero value of $\lambda$ corresponds to ridge regression.
+The motivation for performing ridge regression as opposed to OLS is to not overfit on our data, and the value of $\lambda$ can be interpreted as the strength of our Bayesian prior on the feature weights being equal to 0 \cite{ridgeregressionNotes}.
 
-In order to choose an appropriate value of $\lambda$, we split our data into a training and validation set.
-90\% of the data is allocated to the training set, and 10\% of the data is allocated to the test set.
-We cross-validate the estimated $\beta$ values on our validation set for each $\lambda$, and choose the value of $\lambda$ that produces the lowest MSE on the validation data.
+In order to choose an appropriate value of $\lambda$, we split our data into training, validation, and test sets.
+90\% of the data is allocated to the training set, 10\% to the validation set, and 10\% to the test set.
+Although we estimate $\beta$ on the training data using the above closed-form solution, we cross-validate on our validation set for each $\lambda$, and choose the value of $\lambda$ that produces the lowest MSE on our validation dataset.
+To ensure that we have not overfit $\beta$ to our validation dataset, we also calculate the MSE on the test set as a final step.
+A comparison of the training, validation, and test MSEs for various values of $\lambda$ is found in Figure \ref{fig:mse_plot}.
 
-The table below displays the 20 feature weights with the largest magnitudes given our model.
-A bar chart showing the magnitude of all of the weights (excluding the intercept weight) can be found in Figures \ref{fig:lin_reg_weights1} and \ref{fig:lin_reg_weights2}.
+We find that $\lambda = 100$ minimizes the MSE on our validation set.
+Table 6 displays the 20 feature weights with the largest magnitudes $\lambda = 100$.
+Two charts showing weights for the full set of features (excluding the intercept term) can be found in Figures \ref{fig:lin_reg_weights1} and \ref{fig:lin_reg_weights2}.
 
@@
 \includegraphics[width=\linewidth]{feature_weights.png}
-\caption{Linear Regression Feature Weights (excluding intercept term)}
+\caption{Linear Regression Feature Weights (excluding intercept term) (1 of 2)}
 \label{fig:lin_reg_weights1}
@@
 \includegraphics[width=\linewidth]{feature_weights_contd.png}
-\caption{Linear Regression Feature Weights (excluding intercept term)}
+\caption{Linear Regression Feature Weights (excluding intercept term) (2 of 2)}
 \label{fig:lin_reg_weights2}
@@
 
-Figure \ref{fig:mse_plot} shows the training and holdout MSE for various values of Lambda.
+It's worth taking the time to discuss Figure \ref{fig:mse_plot}, which shows the training and holdout MSE for various values of $\lambda$, in slightly more depth.
 There are a few things in this plot worth discussing.
-First, note that the holdout MSE is consistently lower than the training data MSE.
-Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training and validation data.
-However, this shouldn't effect the validity of our cross validation.
+First, note that the validation MSE is consistently higher than the training data MSE, which is consistently higher than the test data MSE.
+Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training, validation, and test data.
+However, we don't expect this effect the validity of our cross validation.
 
-Another thing worth noting is that even on the training dataset, the MSE goes down once we choose a non-zero value of $\lambda$.
+Another thing worth noting is that even on the training dataset, there exist non-zero values of $\lambda$ that achieve a lower MSE than the OLS estimate of $\beta$.
 At first, this was surprising to our group, as conceptually OLS is often thought of as the linear regression method that minimizes MSE.
 However, it is important to note that OLS only holds this distinction amongst unbiased estimators.
-Hoerl and Kennard (1970) \cite{hoerl1970ridge} prove the existence theorem for ridge regression, which proves that there exists a value of $\lambda$ for which $\beta_{ridge}$ produces a lower MSE than $\beta_{OLS}$.
+Hoerl and Kennard (1970) \cite{hoerl1970ridge} prove the existence theorem for ridge regression, which claims the existence of some $\lambda$ such that $\beta_{ridge}$ produces a lower MSE than $\beta_{OLS}$.
 
@@
 
-\noindent For some values of $\lambda$, ridge regression is able to lower the MSE by decreasing variance, but adding a non-zero bias term.
-We believe this is what is happening in our data when cycling through different $\lambda$ values.
+\noindent For some values of $\lambda$, ridge regression is able to lower the MSE by decreasing variance, but increase the bias from zero to some non-zero value.
+We believe the changes in MSE we observe in our dataset as we vary $\lambda$ can be explained by this phenomenon.
 
@@
 
+In general, we can now interpret the strength of the feature weights produced by our regression to determine how predictive a particular feature is of readership.
+Note from Table 6 that most of the features with the most predictive power are not the text-based features.
+The intercept term in our regression is orders of magnitude larger than any other feature, implying that most of the NYTimes articles receive many pageviews in the baseline case.
+The strongest coefficients tend to be those that indicate the publication desk, section, and time of publication of the article.
+This suggests that the content of an article itself may not be as important as the context in which it is published.
+The one exception we see is that a higher word count is predictive of higher viewership.
+We suspect that what's going on here is correlative, rather than causal - the quality of longer pieces (e.g., NYTimes Magazine articles) is likely higher, thus driving more readership.
+But we don't expect that a website full of low-quality, 5,000 word pieces would be successful.
+
 \subsection{The Effect of Textual Features}
 
-Given the large amount of work we put into building numerous text-based features (such as the trigram neural network perplexity, the sentiment, and the Flesch reading ease), we may want to evaluate how impactful these various features actually are in our regression.
-How much incremental reduction in MSE are we getting by including them?
-We first re-run the exact same regression specification, but instead of using the trigram perplexity calculated from our neural network, we instead use the simple bigram perplexity discussed earlier in this paper.
-We find that this actually causes the MSE to decrease, from 2.392 to 2.389.
-This small, but modest change suggests that there is almost no incremental value from the neural network feature.
+Given the large amount of work we put into building numerous text-based features (such as the trigram neural network perplexity, the sentiment labels, and the Flesch reading ease) and the relatively low impact they seem to have had on our regression (based on feature weights), we want to specifically evaluate the impact of these features on our regression.
+Specifically, how much incremental reduction in MSE are we getting by including them?
+We first re-run the exact same regression specification, but instead of using the trigram perplexity calculated from a neural network, we instead use the simple bigram perplexity discussed earlier in this paper.
+We find that this actually leads to a reduction in MSE, from 2.392 to 2.389.
+This small, but modest change suggests that there is currently almost no incremental value from using a neural network language model as opposed to a more straightforward language model.
 
-Now, we might ask if any text features are adding much value.
-Particularly given that, in general, the magnitude of their weights is dwarfed by the magnitude of the weights of the contextual features when looking at feature weights, we might expect that text features do not add much.
-When removing perplexity, sentiment, reading ease, and word count features from our regression, we find that the MSE increases, from 2.392 to 2.502.
-This is result is encouraging, as it suggests that even if text features aren't doing much, they're doing something!
+As a next step, we ask if text features in general add much value to our model.
+Given that, in general, the magnitude of text feature weights is dwarfed by the magnitude of contextual feature weights, we might expect that text features do not add much.
+We find that removing the perplexity, sentiment, reading ease, and word count features from our regression leads to an increase in MSE, from 2.392 to 2.502.
+This result is encouraging, as it suggests that even if text features currently aren't contributing as much as we had hoped, they are doing something!
 Removing them from our model leads to a 4.6\% increase in the MSE.
@@
 \section{Future Work}
-Unfortunately due to time constraints as well as various unanticipated difficulties and technical errors, we were unable to implement all of several important features we had initially wanted to include in our regression.
-Most notably, this paper in its current state lacks some form of topic modeling which we expect to be a strong predictor of content viewership at least in our context.
-We also think the constructing some features using only the headlines might also provide some improvement, especially considering that an article's headline usually ends up being fairly important in a user's decision to click through or not.
+Unfortunately, we were unable to implement all of several important features we initially wanted to include in our regression, due to a combination of time constraints and various unanticipated technical difficulties and errors.
+Most notably, this paper currently lacks some form of topic modeling, which we expect to be a strong predictor of content viewership (at least in our context).
+We also suspect that the addition of features containing information about article headlines may provide some improvement.
+It seems reasonable to assume that an article's headline is a fairly important factor in a user's decision to click through or not.
 
-There is also much room for improvement with our existing features as well.
-For example, the perplexity score feature ended providing very little predictive power might be a result of the underlying language model not being good enough, especially considering that the neural network model still hasn't completed its training.
-Though we did observe that there was virtually no difference moving from the bigram perplexity scores to the trigram neural network perplexity scores, this again might be because the neural network wasn't given sufficient time to train.
-We can imagine that if we continued training our neural network for several more cycles, we might start to see visible performance gains.
-Alternatively, we think its likely the its not the perplexity of an article relative to the entire corpus that matters, but rather perplexity relative to other articles that cover the same event or topic.
+There is also room for improvement in the design and extraction of our existing features.
+For example, the perplexity score feature ultimately provided very little predictive power.
+This may be due to the issues in the quality of the underlying language model.
+This is particularly likely given that the neural network model was not able to complete many iterations, and has not yet converged.
+Although we observed virtually no difference in performance moving from the bigram perplexity scores to the trigram neural network perplexity scores, this again may be due to insufficient training time for the neural network.
+We find it conceivable that, given many more cycles to train, our neural network would start to yield visible performance gains over the bigram perplexity score.
+Another hypothesis is that rather than measuring the the perplexity of an article relative to the entire corpus of articles, it would be more effective to measure a given article's perplexity relative to other articles that cover the same event or topic.
 
-Similarly, there is ample room to improve our sentiment analysis methodologies.
-The training dataset that we generated using mechanical turk is relatively small (200 articles).
-Because of this, there are currently many words that may have strong conditional probabilities of appearing given a particular article sentiment that simply do not appear in our training data.
-With more time (and money!), we might be able to expand the size of our training data and consequently include the accuracy of our model.
-Furthermore, the model currently skips words it hasn't seen in the training corpus - it might make sense to augment our Naive Bayes implementation to incorporate a method such as Laplace smoothing so as to not simply ignore words we haven't seen in our training data.
+Tshere is ample room to improve our sentiment analysis methodology.
+The training dataset generated using mechanical turk is relatively small (200 training articles).
+Because of this, there may exist many words that have a strong probability of appearing in an article conditional on sentiment that simply do not appear in our training set.
+With more time (and money!), we could label more data and improve the accuracy of our model.
+Furthermore, the model currently skips words that do not occur in the training corpus.
+An extension of our Naive Bayes implementation could use a method such as Laplace smoothing, so as to not simply ignore words we haven't seen in our training data.
 
-Furthermore, the current discrepancies between our implementation of Naive Bayes and NLTK's, while not large in magnitude, are alarming.
-Given more time in the future, we hope to dig deeper into what is happening.
+In addition, the current discrepancies between our implementation of Naive Bayes and NLTK implementation of the algorithm, while not large in magnitude, are alarming.
+In the future, we hope to dig deeper into this discrepancy and identify the root cause.
 Our current hypothesis is that NLTK's implementation of Naive Bayes does not take into account the number of times a particular word appears in a document.
-Because of this, more distinct informative words are required to move a given article's prediction away from the neutral class, which has an extremely strong prior.
+Because of this, a larger number of distinct informative words are required in order to overcome the strong prior belief that a given article's sentiment is neutral.
 
-Another way we could improve our model is to apply a basis expansion to our variables to include polynomial and various interaction terms.
-In many cases, theres no good justification to why a feature is linearly related to the output feature.
+Another avenue for potential improvement to our model is the application of basis expansion to our variables.
+This would allow us to include polynomial and interaction terms.
+In many cases, there is no good justification for assuming a feature is linearly related to the output feature.
 Hence, applying a polynomial expansion might uncover an entirely different relationship between our dependent variable and its covariates.
-We also think that interaction terms might provide some significant predictive power.
+By a similar line of reasoning, we suspect that interaction terms may have significant predictive power.
 
-Certainly one thing we definitely plan on doing if we end up trying to turn this project into a publishable paper is to include all of the articles in our dataset.
-With the much greater sample size and the much more expansive corpus from the additional articles, we might expect to see drastically different optimal weights.
-
-Since we also have Twitter data that tracks every tweet and retweet involving a NYT article, we would like to extend this project to also include our Twitter data.
-We can then explore many interesting potential cross effects, such as how does the article content predict the amount of Twitter sharing and alternately, how does content of a Tweet predict viewership.
+Given that we have access to six full months of NYTimes data, we also hope to expand our dataset to include the full body of NYTimes articles we have available to us.
+With a larger sample size and more expansive corpus, we might expect to see drastically different optimal weights, as our results will be less sensitive to time-dependent trends in viewership and content.
+Since we also have access to Twitter data tracking every tweet and retweet involving a NYT article, we also hope incorporate this data into a future version of our model.
+This would enable us to explore potential cross effects, such as the relationship between article content and Twitter sharing.
 
@@
 Dave and Jeremy built the implementation and validation of the Naive Bayes classifier, while Michael focused his efforts on the construction and validation of the language model.
-Dave, along with assistance from Michael, focused on the implementation of OLS and ridge regression, along with validation and discussion of results.
-Plot generation, table creation, writing, and editing was evenly split between Michael and Extenuating personal circumstances unfortunately prevented Jeremy from contributing to the project as much as originally anticipated.
+Dave, with assistance from Michael, focused on the implementation of OLS and ridge regression, along with validation and discussion.
+Plot generation, table creation, writing, and editing was evenly split between Michael and Dave.
+Extenuating personal circumstances unfortunately prevented Jeremy from contributing to the project as much as originally anticipated - he will likely be in touch soon with the instructors regarding this issue.
 

### 2015-12-10 Project/Latex/Paper.tex
edits

--- 2015-12-10 Project/Latex/Paper.tex
@@
 \affiliation{\textsuperscript{3}\textit{mfzhao@mit.edu}}
-\affiliation{\textsuperscript{*}\textit{Michael is also using this project to fulfill Final Project requirements for both 6.864 as well as 6.867.
-Dave and Jeremy are only using this project for 6.867.}}
+\affiliation{\textsuperscript{*}\textit{Michael is using this project to fulfill Final Project requirements for both 6.864 as well as 6.867. Dave and Jeremy are only using this project for 6.867.}}
 
@@
 To determine how content drives viewership, we construct a unique dataset of all articles published by the New York Times (NYT) in August 2013.
-Our dataset is built from 2 major components, the NYT's internal web traffic data and article content data parsed from the NYT website.
+Our dataset is built from 2 major components, the NYT's internal web traffic data and web content data extracted from the NYT website.
 We use the internal web traffic data to accurately track the number of page views of each article as well as construct a set of robust control variables such as the desk and section of each article.
-To build content features, we use various machine learning and statistical natural language processing techniques on our parsed article content data and construct features such as article perplexity, sentiment, reading difficulty, and indicators that denote the presence of pictures, videos, etc.
-Additionally, we have access to the NYT's internal website traffic data.
-We feed all of our constructed features to into a predictive regression model.
-We find that our textual features modestly improve our predictive power.
-However, there are still many textual features we've not yet explored due to time constraints.
+To build textual content features, we use various machine learning and statistical natural language processing techniques on our extracted content data to produce perplexity scores and to determine the sentiment and reading difficulty of each article.
+We also generate indicators that denote the presence of pictures, videos, etc.
+Lastly, we use some secondary data sources to build some additional control features such as estimated author gender and author popularity.
+We take all these features and build a predictive regression model.
+Overall, we do find that our textual features improve our predictive power, though rather modestly.
+However, due to time constraints, there remain many textual features we haven't yet implemented.
 We believe that adding these additional features may offer more significant improvements in performance.}
@@
 
-In today's digital economy, many companies are very interested in attracting users to visit their websites in order to earn ad revenue.
-While many factors might motivate a user to visit a particular page, certainly one important factor is the content in that webpage.
-This paper explores the relationship between the content of a webpage and the number of page views it ultimately ends up receiving by constructing a unique dataset of all articles published by the New York Times (NYT) during August 2013.
-This dataset is built from two major components: the NYT's internal web traffic data and parsed NYT article content data.
+In today's digital economy, both individuals and companies are very interested in attracting users to visit their websites in order to earn ad revenue.
+While many factors might motivate a user to visit a particular page, certainly one important factor is the content contained within that webpage.
+This paper explores the relationship between the content of a webpage and the number of page views it receives by constructing a unique dataset of all articles published by the New York Times (NYT) during August 2013.
+This dataset is built from two major components: the NYT's internal web traffic data and webpage content data parsed from the NYT website.
 
-Typically, a study such as ours tends to be very difficult to conduct as either accurate measures of viewership are unavailable\footnote{While oftentimes precise viewership data tends to be not available openly, oftentimes researchers use related observables, such as Facebook likes} or the feature extraction of the content is too challenging (for example Youtube), or or both.
+Typically, a study such as ours tends to be very difficult to conduct  accurate measures of viewership are either private of unavailable\footnote{While precise viewership data tends to be not available openly, oftentimes researchers use related observables, such as Facebook likes, or number of Tweets and Retweets}.
+Even in cases where page views are publically, for example Youtube or various other video sharing sights, feature extraction of the content is far too challenging given the tools we have available to us today\footnote{Though this is quickly changing with advances in computing power and development of better and better machine learning methods}.
 Fortunately, our access to the the NYT's internal web traffic data allows us to exactly measure the number of page views an article receives.
@@
 We build on this study in two ways: first, we relate an article's content back to the number of page views it receives rather than its virality\footnote{Which companies arguably care more about since word-of-mouth virality is usually a means to increase page views}.
-Second, we employ more sophisticated machine learning feature extraction techniques to see if they work any better over their simple measures.
+Second, we employ more sophisticated machine learning feature extraction techniques.
 %------------------------------------------------
@@
 Our NYT internal web traffic dataset is a record of all individual user activity on the NYT website covering the period of April 3rd, 2013 to October 31st, 2013.
-This activity data is stored as individual lines of json and includes who (if available) accessed what page at what time.
-Overall, it is over 20 terabytes in size and contains over 3 billion page views.
-It's important to note that not all page views are content views, for example, some events that are also tracked are searches, or user account settings.
-Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of including the entire dataset for future work.
+Each time a user\footnote{in this case, a ``user" is uniquely defined by a device/browser id.
+So while the same person might have multiple devices or may use multiple browsers, the NYT backend treats each device/browser combination as a unique ``user" even though in reality its all the same person.
+In some cases, we are able to link various id's together if the person happens to register an official user account on the NYT website and then logs into her account from multiple devices/browsers.} moves from one page to another on the NYT website, this activity is captured as an indivudal line json.
+This data is incredibly detailed and can potentially tell us who accessed what page when and from what location.
+In addition it tracks the duration a user stays on a webpage as well as the path a user took to get to her current webpage.\footnote{Big brother is watching you and we are big brother.
+Welcome to the age of Big Data} Overall, this data is over 20 terabytes in size and contains over 3 billion page views.
+It's important to note that not all page views are page views of content we care about.
+For example, some events that are also tracked in the data are searches or changes in user account settings.
+Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of extending our analysis to our entire dataset for future work.
+We further restrict our dataset to only include articles or blogposts since these are the pieces of content that have contain readily extractable text data.
+For the sake of brevity, we use the term ``articles" to refer to both articles and blogposts unless we explicitly state otherwise.
 
-We further constrain our dataset to only include articles or blogposts published during the month of August since these are the pieces of content that have machine parsable text data.
-For the sake of brevity, we use articles to refer to both articles and blogposts unless we explicitly state otherwise.
 We make a basic first pass through the data to simply to obtain a list of urls.
-After cleaning up the url data to make sure we had each url mapped to a unique piece of content, we were left with a total of 6682 urls.
-We then parse our dataset and aggregate the number of counts each url receives.
-In order to make the comparison between articles fairer since an article that's been out longer will have more page views on average, we only count the number of page views received up to 7 days after publication.
-Given that page views tend to sharply drop off soon after publication since recency is quite important to the News, the number of page views obtained during the 7 days after an article is published represents the vast majority (usually well over 90\%) of total page views an article receives.
-In total, our data consists of 248,161,455 page views.
+After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6682 urls.
+We then parse all the web data for the month of August and the first week of September count the number of impressions a url receives.
+In order to make an apples to apples comparison between articles we only count the number of page views received up to 7 days after publication since an article that's been out longer should have more page views in expectation.
+Given the tendency for the viewership of an article to drop off sharply soon after publication (as recency is important factor in determining of news readership), our 7-day measure generally represents the vast majority (well over 90\%) of total page views that an article receives\footnote{at least for a reasonable stretch of time}.
+Even after all this subsampling, our data still consists of 248,161,455 page views\footnote{Though one video by PSY completely crushes this number}.
 As seen in Figure 1 below, the distribution of page views is highly skewed with very heavy tails.
@@
 
-In addition to aggregating the counts, when parser the internal web traffic data, we made sure to store various relevant article meta-data features such as the headline, time of publishing, authors, section, desk, and the NYT's internal article content description tags (if available\footnote{by this we mean that the NYT only produces these content tags only for some articles, its not a question of whether the data is available to us}).
+In addition to aggregating the counts, when we parse the internal web traffic data, we make sure to extract various relevant article meta-data features such as the headline, time of publishing, authors, section, desk, and the NYT's internal article content description tags (if available).
 
-\subsection{Parsed NYT Article Content Data}
-Unfortunately, the NYT internal web traffic data does not the actual content displayed on each webpage, which is a very important aspect of our project.
-To resolve this issue, we used the python library ``newspaper".
-This library helped to scrape the html from the NYT website and extract all the raw textual data from the html.
-We further used some regular expressions to clean up what the library missed.
+\subsection{Parsed NYT Webpage Content Data}
+Unfortunately, the NYT internal web traffic data does not contain the actual content displayed on each webpage, which is a very important aspect of our project.
+Luckily, all this content is freely hosted on the NYT website!
+Although the NYT limits the number of free articles you can access per month, the tracking system is cookie based.
+This means that scraping the raw html content via wget or something similar does not contribute to your article count limit\footnote{Alternatively, you can just keep clearing your cookies}.
+Specifically, we used the python library ``newspaper" to download the html content from the NYT website and then extract all the raw textual data from the html.
+We applied some additional regular expressions filters to clean up what the library missed.
 In total, our 6682 articles contain 4,685,021 words of text.
-We find that the distribution of article length, much like the distribution of page views, is highly skewed.
+We find that the distribution of article length, much like the distribution of page views, is highly skewed and heavy tailed.
 We again apply a log transformation.
@@
 
-In addition to extracting the raw text data, we made sure to check if the articles contained non-textual content such as pictures or videos.
+In addition to extracting the raw text data, we made sure to check if the articles also contained any addtional non-textual content such as pictures or videos.
 We created indicator variables that denote the presence of such content within an article.
@@
 
-In preparing to perform our regression, we construct a number of features, using both the parsed NYT article data and secondary sources.
-These features include the Flesch reading ease, the (guessed) gender of the author(s), the popularity of the author(s), the sentiment of the article text, the perplexity of the article text, and variables indicating the section the article appeared in and the article's content type.
-Found below is a full list of these features, as well as the methodology used to construct them and validation of the features, where appropriate.
+Using our collected this data, as well as some data from some additional secondary sources we construct the features that make up our predictive regression model.
+These features include the Flesch reading ease, the estimated gender of the author(s), the popularity of the author(s), variables indicating the section the article appeared in and the article's content type, the sentiment of the article text, and the perplexity of the article text.
+We provide a full list of these features below, as well as the methodology used to construct them.
+Where appropriate, we include discussion of testing and validation of our features and our algorithms.
 
 \begin{description}
-	\item[Flesch Reading Ease] One can conceive of a few hypotheses that relate the readership of content to the ease with which people can read it.
+	\item[Flesch Reading Ease] One can conceive of a few competing hypotheses that relate the readership of content to the ease with which people can read it.
 Maybe more complicated pieces of text are more engaging, and are more likely to be read.
@@
 
-	In order to measure something that will serve as a decent proxy for popularity, we programmatically searched for every distinct author in our dataset using Bing, and recorded the number of search results that were returned by the query.
-In cases where a particular article has more than one distinct author, we calculate an ``effective" popularity by simply calculating the average number of search results over all article authors.
+	In order to measure something that will serve as a decent proxy for popularity, we programmatically searched for every distinct author in our dataset on Bing and recorded the number of search results that were returned by the query.
+In cases where a particular article has more than one distinct author, we calculate an ``effective" popularity by simply averaging number of search results over all article authors.
 
-	\item[Author Gender] We certainly hope that an author's gender would not have a causal impact on whether or not a particular article gets readership.
-However, we want to allow for this fact and measure this variable.
-We do so by constructing a feature that indicates the most likely gender of the article author(s).
+	\item[Author Gender] While we certainly don't think that an author's gender has a causal impact on the readership on an article.
+We believe that this feature allows us to control for some latent unobserved heterogeneity.
+For example, its not a stretch to think that the experiences of women authors are drastically different male authors and these differences are reflected in their writing.
+We construct this feature that indicates the most likely gender of the article author(s).
 In cases where the gender of the author is unclear (e.g., Robin) or there are likely multiple authors with different genders (e.g., The New York Times Staff), we record a third gender value, ``ambiguous / unknown."
 
-	We construct our gender data by cross-referencing the first names of all of the authors in our dataset against U.S.
+	Our gender data is gathered by cross-referencing the first names of all of the authors in our dataset against U.S.
 Social Security Administration baby name data from 1935 to 1997.
@@
 
-	\item[Material Type, Section, Desk, and Article Type] We also include dummy variables including the material type (e.g., `News' or `Obituary'), desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), and section (e.g., `Movies' or `World').
+	\item[Material Type, Section, Desk, and Article Type] We also include dummy variables including the material type (e.g., `News' or `Obituary'), publishing desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), and section (e.g., `Movies' or `World').
 The hypothesis driving the decision to include these variables is that certain types of content (e.g., political news or international affairs) may be more widely read than local material (such as real estate) or less popular sections of the NY Times (e.g., the sports section).
@@
 
-We also include features that attempt to capture the article sentiment and the article text perplexity.
-Since the construction of these features was more involved and involved validation of our algorithms, we discuss these two features in separate subsections.
+We also build features that attempt to capture the article sentiment and the article text perplexity.
+Since the construction of these features was considerably more complex and involved validation of our algorithms, we discuss these two features in separate subsections.
 
@@
 
-\noindent where $p_{ki}$ is the probability of a document of class $K$ including word $i$. Given that we transform this problem into log space, we can compute $\log(p(\mathbf{x}|C_k))$ as
+\noindent where $p_{ki}$ is the probability a word $w_i$ conditional on a document of being of class $K$. Applying a log transformation, we can compute $\log(p(\mathbf{x}|C_k))$ as:
 
@@
 We coded up a basic implementation of the Naive Bayes algorithm, drawing heavy inspiration from Greg Lamp's 2014 python tutorial on Naive Bayes \cite{NaiveBayes}.
-In order to get the probabilities $p(C_k)$ and $p_{ki}$, we developed a training set of NY times data by providing 200 articles to workers on Amazon Mechanical Turk.
-Each turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive.
-We recorded 5 scores for every article, and calculated the average sentiment reported by the Turkers.
+In order to get the probabilities $p(C_k)$ and $p_{ki}$, we needed some labeled training data.
+In order to obtain these labels, we selected a random subset of 200 articles from our dataset and created a task on Amazon Mechanical Turk.
+Each Turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive.
+In order to make sure these scores were relatively robust, we recorded 5 scores for every article from 5 different Turkers and calculated the average sentiment score.
 We classified any article having an average score greater than 0.5 as `positive'.

### 2015-12-10 Project/Latex/Paper.tex
merge

@@
 If many news articles about the same story are all using highly similar language, an article that covers the story using atypical language is likely unique in some way or another which may drive people to read it more or less.
+For this section we generally follow the 6.864 Lecture Notes \cite{6864Lec23}.
 
-As for out paper, we consider a couple of choices for our language model.
-First, we build a simple bi-gram language to establish a baseline.
+As for out paper, we construct a couple of different language models.
+First, we build a simple bi-gram language to use as a baseline.
 We also build more sophisticated word vector based n-gram neural network language models.
 
-We split our articles into training (~70\%), validation (~15\%), and test sets (~15\%).
-We build our language model corpus only from the text of the articles in the training dataset.
+We split our articles into training (~70\%), validation (~15\%), and test corpora (~15\%).
 In order to keep the size of our vocabulary relatively manageable, we ignore any case sensitivities (so ``Cat" is the same as ``cat").
@@
 Words that don't make this cutoff are mapped to a generic ``rare word" indicator.
-Lastly, we also map any numbers (that is actual numbers with digits, not number written with words) to a generic ``number" indicator.
-This leaves us with a vocabulary size $|V|$ of 29359.
+Lastly, we also map any numbers (that is numbers comprised of digits, not numbers written with words) to a generic ``number" indicator.
+Ultimately, this leaves us with a vocabulary size $|V|$ of 29359.
 To estimate a bigram model, we simply need to compute the counts in our training corpus.
@@
 \end{equation}
-However, its reasonable to expect that there might be bigrams in the development or test corpora that aren't observed in the training corpus.
-If this is the case, then any article with an unobserved bigram would be assigned a predicted likelihood of 0.
+However, its reasonable to expect that there might be bigrams in the development or test corpora that aren't observed in the training corpus considering that for our vocabulary size, there are nearly 900 million unique bigrams and our training data contains just over 3 million observations.
+If this is the case, then any article with a bigram unobserved in the training corpus would be assigned a predicted likelihood of 0.
 Needless to say, this is very bad.
 In order to avoid this issue, we apply a technique called add-$\alpha$ smoothing.
-Essentially what add-$\alpha$ smoothing does is it adds (possibly fractional) counts to every single possible bigram.
-After smoothing, no bigram, given a fixed vocabulary $V$, will ever have 0 probability.
+Add-$\alpha$ smoothing just adds $\alpha$ to each cell in the probability table.
+Hence, after smoothing, no bigram, given a fixed vocabulary $V$, will ever have 0 probability.
 This changes our our estimated word probabilities to:
@@
 \end{equation}
-This smoothing actually has quite a nice Bayesian interpretation as well.
-In the bigram case, we can say that words follow a multinomial distribution conditional on its preceding word.
-The smoothed language model
-
-Essentially, is we're applying a symmetric dirichlet prior with parameter $\alpha$ to each of these multinomial distributions.
+Essentially, this smoothing is acting like a probablistic Robin Hood, it takes probability mass from the ``rich" and distributes it to the ``poor".
+Additionally, this smoothing actually has a nice Bayesian interpretation as well.
+For the bigram language model case, we can say that words follow a multinomial distribution conditional on its preceding word.
+The smoothed language model is equivalent to the posterior of horizontally stacked multinomial likelihoods with a symmetric Dirichlet prior with a parameter value of $\alpha$.
 
@@
 \end{figure}
+
 Here we see exactly what we should expect.
@@
 For our language model, the optimal $\alpha$ is 0.004.
-This produces the following NALLs for our 3 corpora:
+This level of smoothingproduces the following NALLs for our 3 corpora:
 
 \begin{table}[hbt]
-\caption{Bigram Language Model with Smoothing Negative Average Log-Likelihood}
+\caption{Negative Average Log-Likelihood for Smoothed Bigram Language Model}
 \centering
@@
 One criticism of standard n-gram language models is that they are rather sensitive to the training data.
-Suppose that two words generally have fairly interchangeable use cases (meaning that the same or similar words tend to appear before these words), for example, ``coffee" and ``tea".
+Suppose that two words generally have fairly interchangeable use cases (that is the sets of words that tend to precede these words tend to be highly similar).
+For example, consider ``coffee" and ``tea".
 Further suppose that for whatever reason, the phrase ``drink coffee" is predominantly featured in the training data but ``drink tea" is not.
-Then a standard bigram language model would assign a very low probability to ``drink tea" even if the in the training data contains many instances in which they are fairly interchangeable (``buy coffee/tea", ``brew coffee/tea", etc.).
+Then a standard bigram language model would assign a very low probability to ``tea" if it were preceded by the word ``drink" even if the in the training data contains many instances in which they are fairly interchangeable (``buy coffee/tea", ``brew coffee/tea", etc.).
 
 Hence, we build word vector n-gram neural network models to see if we can achieve better performance.
-By using dense word vectors rather than one-hot encodings to represent words, we're able to more accurately represent the ``similarity" of words.
+By using dense word vectors rather than one-hot encodings to represent words, we're able capture the underlying similarity of words and their meanings.
 For example, ``coffee" is more similar to ``tea" that it is to ``car".
@@
  In particular, we train both a bigram neural network language model and a trigram neural network language model.
-In both cases, we use word vectors of length 10 (so for the bigram NN, our input layer is size 10 and for the trigram NN, the input layer is size 20) that map to a single hidden layer with 10 hidden units with a tanh activation function.
-The hidden layer is then connected to the output layer which is the size of our vocabulary with a softmax activation function.
-Given that we want to also train our word vectors, we use stochastic adaptive gradient descent.
+In both cases, we use word vectors of length 10 meaning that the input layer of the bigram NN is a 10 (a single word) and the input layer of the trigram is 20 (2 words concatenated together).
+This input layer then maps to the single hidden layer with 10 hidden units through a tanh activation function.
+Finally, the hidden layer is then connected to the output layer which is has size $|V|$ through a softmax activation function.
+We use stochastic adaptive gradient descent so that the errors can be properly backporgated to the word vectors.
 We purposefully keep our network relatively small since our data set is rather large and training even this small neural network takes considerable time.
@@
 Due to various complications, we were not able to train our model for as long as we would've liked.
-Couple this with the fact that our training corpus contains over 3,000,000 observations and each complete pass through it takes approximately 4.5 hours, means that as of writing this paper, we've only completed 4 full passes.
-Unfortunately, as a result of an oversight in the code, we forgot to store the NALL of training data as we complete each full loop.
-However, we are highly confident our neural network is working as intended since we see we that the NALL decreases every pass on the validation corpus:
+Since our training corpus contains over 3,000,000 observations, each complete pass through it takes approximately 4.5 hours.
+At the time of writing this section, we've only completed 4 full passes.
+
+ We adapted our code so it saved a useable snapshot of the network after each full pass through the data.
+Unfortunately, as a result of an oversight in the code, we forgot to store the NALL of training data in this snapshot.
+All is not lost since we remembered to at least store the NALL on the validation corpus.
+We are highly confident our neural network is working as intended since we see we that the NALL decreases every pass on the validation corpus in Figure \ref{fig:perp_hist} below.
 
@@
 
-Despite the fact that our training isn't finished, both our neural network-based language models are already performing better than a standard bi-gram model with smoothing.
+Despite the fact that our training isn't finished, both our neural network-based language models have already achieved lower negative average log-likelihood compared to the smoothed standard bi-gram model.
 We build 2 sets of article perplexity scores.
@@
 \includegraphics[width=\linewidth]{perplexity.png}
-\caption{The distribution of bigram and trigram perplexities}
+\caption{The Distribution of Bigram and trigram Neural Network Perplexities}
 \label{fig:perp_hist}
@@
 
-Its important to note however, that we don't necessarily care exactly how good our language model.
-Since we only use the language model as an intermediate step to construct a feature for a regression task.
-It could very well be the case features generated by a worse language model can work just as well as well in the regression task as features generated by a better language model if the relative variation of articles is maintained.
-To test whether a better language model can give us more predictive power, we compute two sets of perplexity scores, one based on the smoothed bigram language model and one based on the current best trigram neural network language model.
+Its important to note however, that we only care about the goodness of the underlying language model insofar that a better language will produce features with better predictive power.
+However, it could very well be the case features generated by a worse language model can work just as well as well as features generated by a better language model in the regression model if the relative variation of articles is maintained.
+Hence our reasoning to produce the two sets of perplexity scores.
 

### 2015-12-10 Project/Latex/Paper.tex
[tex] edits.

--- 2015-12-10 Project/Latex/Paper.tex
@@
 Our dataset is built from 2 major components, the NYT's internal web traffic data and web content data extracted from the NYT website.
-We use the internal web traffic data to accurately track the number of page views of each article as well as construct a set of robust control variables such as the desk and section of each article.
+We use the internal web traffic data to accurately track the number of page views for each article, and to construct a set of robust control variables such as the desk and section of each article.
 To build textual content features, we use various machine learning and statistical natural language processing techniques on our extracted content data to produce perplexity scores and to determine the sentiment and reading difficulty of each article.
-We also generate indicators that denote the presence of pictures, videos, etc.
-Lastly, we use some secondary data sources to build some additional control features such as estimated author gender and author popularity.
-We take all these features and build a predictive regression model.
+Lastly, we use secondary data sources to build additional control features such as estimated author gender and author popularity.
+We combine all these features and build a predictive regression model.
 Overall, we do find that our textual features improve our predictive power, though rather modestly.
-However, due to time constraints, there remain many textual features we haven't yet implemented.
-We believe that adding these additional features may offer more significant improvements in performance.}
+However, there remain many textual features we haven't yet implemented due to time constrains.
+We believe that adding these additional features may offer significant improvements in performance.}
 
@@
 
-In today's digital economy, both individuals and companies are very interested in attracting users to visit their websites in order to earn ad revenue.
-While many factors might motivate a user to visit a particular page, certainly one important factor is the content contained within that webpage.
+In today's digital economy, both individuals and companies are interested in attracting users to their websites in order to earn ad revenue.
+While many factors might motivate a user to visit a particular page, certainly one important factor is that webpage's content.
 This paper explores the relationship between the content of a webpage and the number of page views it receives by constructing a unique dataset of all articles published by the New York Times (NYT) during August 2013.
@@
 
-Typically, a study such as ours tends to be very difficult to conduct  accurate measures of viewership are either private of unavailable\footnote{While precise viewership data tends to be not available openly, oftentimes researchers use related observables, such as Facebook likes, or number of Tweets and Retweets}.
-Even in cases where page views are publically, for example Youtube or various other video sharing sights, feature extraction of the content is far too challenging given the tools we have available to us today\footnote{Though this is quickly changing with advances in computing power and development of better and better machine learning methods}.
+Typically, a study such as ours tends to be very difficult to conduct.
+Accurate measures of viewership are either private of unavailable\footnote{While precise viewership data tends to be not available openly, oftentimes researchers use related observables, such as Facebook likes, or number of Tweets and Retweets}.
+Even in cases where page views are publicly, for example Youtube or various other video sharing sites, feature extraction of the content is far too challenging given the tools available to us today\footnote{Though this is quickly changing with advances in computing power and development of better and better machine learning methods}.
 Fortunately, our access to the the NYT's internal web traffic data allows us to exactly measure the number of page views an article receives.
@@
 They also show that articles that evoke high physiological positive or negative arousal (such as awe or anger) tend to be more viral than articles that evoke deactivating emotions (sadness).
-We build on this study in two ways: first, we relate an article's content back to the number of page views it receives rather than its virality\footnote{Which companies arguably care more about since word-of-mouth virality is usually a means to increase page views}.
-Second, we employ more sophisticated machine learning feature extraction techniques.
-%------------------------------------------------
+We build on this study in two ways: first, we relate an article's content back to the number of page views it receives rather than its virality\footnote{Which is arguably more important to companies, since word-of-mouth virality is usually a means to increase page views}.
+Second, we employ more sophisticated machine learning feature extraction techniques than those used by Berger and Milkman.
 
@@
 So while the same person might have multiple devices or may use multiple browsers, the NYT backend treats each device/browser combination as a unique ``user" even though in reality its all the same person.
-In some cases, we are able to link various id's together if the person happens to register an official user account on the NYT website and then logs into her account from multiple devices/browsers.} moves from one page to another on the NYT website, this activity is captured as an indivudal line json.
+In some cases, we are able to link various id's together if the person happens to register an official user account on the NYT website and then logs into her account from multiple devices/browsers.} moves from one page to another on the NYT website, this activity is captured as an individual JSON object.
 This data is incredibly detailed and can potentially tell us who accessed what page when and from what location.
-In addition it tracks the duration a user stays on a webpage as well as the path a user took to get to her current webpage.\footnote{Big brother is watching you and we are big brother.
-Welcome to the age of Big Data} Overall, this data is over 20 terabytes in size and contains over 3 billion page views.
+In addition, it tracks the amount of time a user stays on a webpage and path a user took to get to her current webpage.
+Overall, this dataset is over 20 terabytes in size and contains over 3 billion page views.
 It's important to note that not all page views are page views of content we care about.
-For example, some events that are also tracked in the data are searches or changes in user account settings.
-Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of extending our analysis to our entire dataset for future work.
-We further restrict our dataset to only include articles or blogposts since these are the pieces of content that have contain readily extractable text data.
+For example, searches or changes in user account settings are also recorded in our dataset.
+Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of extending our analysis to our entire dataset in the future.
+We further restrict our dataset to only include articles or blogposts, since these are the pieces of content that contain readily extractable text data.
 For the sake of brevity, we use the term ``articles" to refer to both articles and blogposts unless we explicitly state otherwise.
 
-We make a basic first pass through the data to simply to obtain a list of urls.
-After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6682 urls.
-We then parse all the web data for the month of August and the first week of September count the number of impressions a url receives.
-In order to make an apples to apples comparison between articles we only count the number of page views received up to 7 days after publication since an article that's been out longer should have more page views in expectation.
-Given the tendency for the viewership of an article to drop off sharply soon after publication (as recency is important factor in determining of news readership), our 7-day measure generally represents the vast majority (well over 90\%) of total page views that an article receives\footnote{at least for a reasonable stretch of time}.
+We conducted a first pass through the data to simply to obtain a list of URLs.
+After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6682 URLs.
+We then parses all the web data for the month of August and the first week of September, counting the number of impressions each URL received.
+In order to make an apples to apples comparison between articles, we only count the number of page views received in the 7 days immediately following publication, since an article that's been out longer should have more page views in expectation.
+Given the tendency for viewership of an article to drop off sharply soon after publication (as recency is an important factor in news readership), our 7-day measure generally represents the vast majority (well over 90\%) of total page views that an article receives\footnote{at least for a reasonable stretch of time}.
 Even after all this subsampling, our data still consists of 248,161,455 page views\footnote{Though one video by PSY completely crushes this number}.
@@
 
-In addition to aggregating the counts, when we parse the internal web traffic data, we make sure to extract various relevant article meta-data features such as the headline, time of publishing, authors, section, desk, and the NYT's internal article content description tags (if available).
+In addition to counting page views, when parsing the internal web traffic data we extract various relevant article meta-data features, such as the headline, time of publishing, authors, section, desk, and the NYT's internal article content description tags (if available).
 
@@
 Although the NYT limits the number of free articles you can access per month, the tracking system is cookie based.
-This means that scraping the raw html content via wget or something similar does not contribute to your article count limit\footnote{Alternatively, you can just keep clearing your cookies}.
+This means that scraping the raw HTML content via wget or something similar does not contribute to your article count limit\footnote{Alternatively, you can just keep clearing your cookies}.
 Specifically, we used the python library ``newspaper" to download the html content from the NYT website and then extract all the raw textual data from the html.
@@
 We again apply a log transformation.
-Histograms and various summary statistics of for the distribution of article length and log article length are provided below:
+Histograms and various summary statistics for the distribution of article length and log article length are provided below:
 
@@
 
-In addition to extracting the raw text data, we made sure to check if the articles also contained any addtional non-textual content such as pictures or videos.
+In addition to extracting the raw text data, we checked for the presence of additional non-textual content such as pictures or videos in each articles HTML content.
 We created indicator variables that denote the presence of such content within an article.
-
-%\begin{figure*}[ht]\centering % Using \begin{figure*} makes the figure take up the entire width of the page
-%\includegraphics[width=\linewidth]{view}
-%\caption{Wide Picture}
-%\label{fig:view}
-%\end{figure*}
-
-%------------------------------------------------
 
@@
 
-Using our collected this data, as well as some data from some additional secondary sources we construct the features that make up our predictive regression model.
+Using our collected NYT article data, as well as some additional data from secondary sources, we construct the features that will be fed into our predictive regression model.
 These features include the Flesch reading ease, the estimated gender of the author(s), the popularity of the author(s), variables indicating the section the article appeared in and the article's content type, the sentiment of the article text, and the perplexity of the article text.
-We provide a full list of these features below, as well as the methodology used to construct them.
+We provide a full list of these features below, as well as the methodology used to extract them.
 Where appropriate, we include discussion of testing and validation of our features and our algorithms.
@@
 On the other hand, perhaps pieces of text that are easier to read will be consumed by more people.
-In order to capture relationships such as these in our data, we include the Flesch reading ease.
+In order to capture relationships such as these in our data, we calculate the Flesch reading ease.
 The Flesch reading ease is a metric developed by Flesch in 1948 \cite{flesch1948new}.
 The score indicates how difficult a piece of English text is to understand.
-Lower scores correspond to more difficult passages, and the highest score attainable is 120.0.
+Lower scores correspond to more difficult passages, with 120.0 being the highest attainable score.
 The formula for calculating a passage's Flesch reading ease is
@@
 
-	\item[Author Popularity] We attempt to include some measure of a particular author's popularity.
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{flesch_data_histogram.png}
+\caption{Histogram of Articles by Flesch Reading Ease}
+\label{fig:fre_hist}
+\end{figure}
+
+	\item[Author Popularity] We also want to include some measure of a particular author's popularity.
 It stands to reason that a new article by Paul Krugman or A.O.
-Scott should garner more readership than a new article by an unknown graduate student enrolled in 6.867 at MIT!
+Scott should garner more readership than a blog post by an unknown graduate student enrolled in 6.867 at MIT!
 
@@
 In cases where a particular article has more than one distinct author, we calculate an ``effective" popularity by simply averaging number of search results over all article authors.
+A histogram showing the distribution of the number of bing search results can be found in Figure \ref{fig:ap_hist}.
+We note that this distribution, like many of those we consider, is highly skew.
+We correct for this by transforming this variable with a log transformation.
+The distribution of log(number of Bing search results) is found in Figure \ref{fig:apl_hist}.
 
-	\item[Author Gender] While we certainly don't think that an author's gender has a causal impact on the readership on an article.
-We believe that this feature allows us to control for some latent unobserved heterogeneity.
-For example, its not a stretch to think that the experiences of women authors are drastically different male authors and these differences are reflected in their writing.
-We construct this feature that indicates the most likely gender of the article author(s).
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{author_popularity_histogram.png}
+\caption{Histogram of Articles by Bing Search Results}
+\label{fig:ap_hist}
+\end{figure}
+
+\begin{figure}[ht]\centering
+\includegraphics[width=\linewidth]{author_popularity_log_histogram.png}
+\caption{Histogram of Articles by log(Bing Search Results)}
+\label{fig:apl_hist}
+\end{figure}
+
+	\item[Author Gender] While we certainly don't think that an author's gender has a causal impact on the readership on an article, we believe that this feature allows us to control for some latent unobserved heterogeneity.
+For example, its not a stretch to think that the experiences of women authors are drastically different male authors and these differences are reflected in their writing style and the topics they write on.
+For each (set of) author(s), we record the most likely gender of the author.
 In cases where the gender of the author is unclear (e.g., Robin) or there are likely multiple authors with different genders (e.g., The New York Times Staff), we record a third gender value, ``ambiguous / unknown."
@@
 If over 90\% of the babies with a given name have been female, we assume a given author is female.
-Otherwise, we record ``ambiguous / unknown."
+Otherwise, we record ``ambiguous / unknown." We find that this methodology is robust to changes in our threshold (chosen as 90\%).
+Changing this threshold to other reasonable values does not have very much impact on our predicted genders.
 
-	\item[Material Type, Section, Desk, and Article Type] We also include dummy variables including the material type (e.g., `News' or `Obituary'), publishing desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), and section (e.g., `Movies' or `World').
+	\item[Material Type, Section, Desk, and Article Type] We also include a number of dummy variables, indicating the material type (e.g., `News' or `Obituary'), publishing desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), section (e.g., `Movies' or `World'), and the day of week and time of day that the article was published.
 The hypothesis driving the decision to include these variables is that certain types of content (e.g., political news or international affairs) may be more widely read than local material (such as real estate) or less popular sections of the NY Times (e.g., the sports section).
-
-\begin{figure}[ht]\centering
-\includegraphics[width=\linewidth]{author_popularity_histogram.png}
-\caption{Histogram of Bing Search Results}
-\label{fig:ap_hist}
-\end{figure}
-
-\begin{figure}[ht]\centering
-\includegraphics[width=\linewidth]{author_popularity_log_histogram.png}
-\caption{Histogram of log(Bing Search Results)}
-\label{fig:apl_hist}
-\end{figure}
-
-\begin{figure}[ht]\centering
-\includegraphics[width=\linewidth]{flesch_data_histogram.png}
-\caption{Histogram of Flesch Reading Ease}
-\label{fig:fre_hist}
-\end{figure}
+We also suspect that publishing an article on certain days of the week (for example, weekends) or at particular times of day (such as lunch hour) may correspond to higher levels of readership.
 
@@
 We also build features that attempt to capture the article sentiment and the article text perplexity.
-Since the construction of these features was considerably more complex and involved validation of our algorithms, we discuss these two features in separate subsections.
+Since the design and computation of these features was considerably more complex and our algorithms required some amount of validation, we discuss these two features in separate subsections.
 
@@
 In order to measure article sentiment, we use a Naives Bayes text classification algorithm, as described in Rennie et al (2003) \cite{rennie2003tackling}.
-We assume that each article in our corpus can belong to one of three classes - `negative' sentiment, `neutral' sentiment, and `positive' sentiment, which we will denote as $C_k$.
-The Naive Bayes model assumes that the likelihood of observing a given article $\mathbf{x} = (x_1, ..., 1_n)$, where $x_i$ is the number of times that word $i$ appears, is
+We assume that each article in our corpus can belong to one of three classes - `negative' sentiment, `neutral' sentiment, or `positive' sentiment, which we will denote as $C_k$.
+The Naive Bayes model assumes that the likelihood of observing a given article $\mathbf{x} = (x_1, ..., x_n)$, where $x_i$ is the number of times that word $i$ appears in the article, is
 
@@
 
-\noindent where $p_{ki}$ is the probability a word $w_i$ conditional on a document of being of class $K$. Applying a log transformation, we can compute $\log(p(\mathbf{x}|C_k))$ as:
+\noindent where $p_{ki}$ is the probability of word $w_i$ conditional on a document belonging to class $K$.
+Applying a log transformation to this expression, we can compute $\log(p(\mathbf{x}|C_k))$ as:
 
@@
 
+\noindent to classify a given article, we simply compute $p(\mathbf{x}|C_k))$ for each class, and select the class with the highest log-likelihood.
+
 We coded up a basic implementation of the Naive Bayes algorithm, drawing heavy inspiration from Greg Lamp's 2014 python tutorial on Naive Bayes \cite{NaiveBayes}.
@@
 In order to obtain these labels, we selected a random subset of 200 articles from our dataset and created a task on Amazon Mechanical Turk.
-Each Turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive.
+Each Turker was asked to score the sentiment toward the subject of the article in question.
+Scoring was done on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive.
 In order to make sure these scores were relatively robust, we recorded 5 scores for every article from 5 different Turkers and calculated the average sentiment score.
 We classified any article having an average score greater than 0.5 as `positive'.
-Any article with an average sentiment less than -0.5 was `negative.' Any other articles were classified as `neutral.' Ultimately, our labels were 66\% neutral, 14.5\% negative, and 19.5\% positive.
+Any article with an average sentiment less than -0.5 was classified as `negative.' Any other articles were classified as `neutral.' Ultimately, our labels were 66\% neutral, 14.5\% negative, and 19.5\% positive.
 This is unsurprising, as a newspaper such as the New York Times likely strives for neutrality when reporting on most topics.
 
-We wanted to determine how our Naive Bayes implementation did compared to an off-the-shelf implementation of the same algorithm.
-In order to do so, we trained NLTK's multinomial Naive Bayes classifier on the same training data, and then evaluated the sentiment of 1,000 articles.
-A comparison of the two implementations is found below, where the columns indicate the prediction by the NLTK Naive Bayes implementation and the rows indicate the prediction by our implementation:
+We wanted to measure how our Naive Bayes implementation did compared to an off-the-shelf implementation of the same algorithm.
+In order to do so, we trained NLTK's multinomial Naive Bayes classifier \cite{BirdKleinLoper09} on the same training data, and then compared the predicted sentiment between the two articles on a 1,000 article subset of our data.
+We limited the comparison to 1,000 articles because the NLTK implementation of multinomial Naive Bayes actually took an incredibly long time to run.
+A comparison of the two implementations' classifications is found below, where the columns indicate the prediction by the NLTK Naive Bayes implementation and rows indicate the prediction by our implementation:
 
@@
 Overall, we find 89.4\% agreement between the two algorithms.
-However, alarmingly, the NLTK implementation seems to predict neutral an overwhelming percentage of the time (99.8\% of the time).
-This warrants further investigation, and may be due to small differences in implementation, or peculiarities in the sample of 1,000 articles we chose to compare the two algorithms across.
-In any case, our algorithm seems to be performing in the same neighborhood as the NLTK implementation (if not better), so we feel relatively comfortable moving forward using our sentiment labels.
+Alarmingly, however, the NLTK implementation seems to predict neutral an overwhelming percentage of the time (99.8\%).
+This warrants further investigation, and may be due to small differences in implementation, or peculiarities in the sample of 1,000 articles we chose to compare the two algorithms.
+In any case, the predictions of our algorithm are in the same neighborhood as the NLTK implementation and are of comparable, if not better, quality.
+As a result, we feel relatively comfortable moving forward using our sentiment labels.
 
@@
 \includegraphics[width=\linewidth]{perplexity.png}
-\caption{The distribution of bigram and trigram perplexities}
+\caption{Histogram of Articles by bigram and trigram perplexities}
 \label{fig:perp_hist}

### 2015-12-10 Project/Latex/Paper.tex
[project] final paper r evisions.

@@
 Our NYT internal web traffic dataset is a record of all individual user activity on the NYT website covering the period of April 3rd, 2013 to October 31st, 2013.
-Each time a user\footnote{in this case, a ``user" is uniquely defined by a device/browser id.
+Each time a user\footnote{In this case, a ``user" is uniquely defined by a device/browser id.
 So while the same person might have multiple devices or may use multiple browsers, the NYT backend treats each device/browser combination as a unique ``user" even though in reality its all the same person.
@@
 We conducted a first pass through the data to simply to obtain a list of URLs.
-After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6682 URLs.
+After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6,682\footnote{If we had just included a few more URLs, we could have had 6,867 observations!} URLs.
 We then parses all the web data for the month of August and the first week of September, counting the number of impressions each URL received.
@@
 If many news articles about the same story are all using highly similar language, an article that covers the story using atypical language is likely unique in some way or another which may drive people to read it more or less.
-For this section we generally follow the 6.864 Lecture Notes \cite{6864Lec23}.
+For this section we generally follow the 6.864 Lecture Notes 2 and 3 \cite{6864Lec23}.
 
-As for out paper, we construct a couple of different language models.
-First, we build a simple bi-gram language to use as a baseline.
+As for our paper, we construct a couple of different language models.
+First, we build a simple bigram language to use as a baseline.
 We also build more sophisticated word vector based n-gram neural network language models.
@@
 Lastly, we also map any numbers (that is numbers comprised of digits, not numbers written with words) to a generic ``number" indicator.
-Ultimately, this leaves us with a vocabulary size $|V|$ of 29359.
+Ultimately, this leaves us with a vocabulary size $|V|$ of 29,359.
 To estimate a bigram model, we simply need to compute the counts in our training corpus.
@@
 \end{equation}
-However, its reasonable to expect that there might be bigrams in the development or test corpora that aren't observed in the training corpus considering that for our vocabulary size, there are nearly 900 million unique bigrams and our training data contains just over 3 million observations.
+However, its reasonable to expect that there might be bigrams in the development or test corpora that aren't observed in the training corpus.
+This is particularly likely given that for our vocabulary size, there are nearly 900 million unique bigrams and our training data contains just over 3 million observations.
 If this is the case, then any article with a bigram unobserved in the training corpus would be assigned a predicted likelihood of 0.
@@
 In order to avoid this issue, we apply a technique called add-$\alpha$ smoothing.
-Add-$\alpha$ smoothing just adds $\alpha$ to each cell in the probability table.
+Add-$\alpha$ smoothing adds $\alpha$ to each cell in the probability table.
 Hence, after smoothing, no bigram, given a fixed vocabulary $V$, will ever have 0 probability.
@@
 \end{equation}
-Essentially, this smoothing is acting like a probablistic Robin Hood, it takes probability mass from the ``rich" and distributes it to the ``poor".
-Additionally, this smoothing actually has a nice Bayesian interpretation as well.
+Essentially, this smoothing is acting like a probablistic Robin Hood.
+It takes probability mass from the ``rich" and distributes it to the ``poor".
+Additionally, this smoothing actually has a nice Bayesian interpretation.
 For the bigram language model case, we can say that words follow a multinomial distribution conditional on its preceding word.
@@
 The NALL on the training corpus is lower than the NALL on the validation corpus.
-Furthermore, the NALL on our training corpus monotonically increases as we increase $alpha$ while the NALL on the validation corpus initially decreases hits a minimum, and starts to increase.
+Furthermore, the NALL on our training corpus monotonically increases as we increase $\alpha$ while the NALL on the validation corpus initially decreases, hits a minimum, and starts to increase.
 For our language model, the optimal $\alpha$ is 0.004.
-This level of smoothingproduces the following NALLs for our 3 corpora:
+This level of smoothing produces the following NALLs for our 3 corpora:
 
@@
 For example, consider ``coffee" and ``tea".
-Further suppose that for whatever reason, the phrase ``drink coffee" is predominantly featured in the training data but ``drink tea" is not.
-Then a standard bigram language model would assign a very low probability to ``tea" if it were preceded by the word ``drink" even if the in the training data contains many instances in which they are fairly interchangeable (``buy coffee/tea", ``brew coffee/tea", etc.).
+Further, suppose that for whatever reason, the phrase ``drink coffee" is predominantly featured in the training data but ``drink tea" is not.
+Then a standard bigram language model would assign a very low probability to ``tea" if it were preceded by the word ``drink" even if the training data contains many instances in which they are fairly interchangeable (``buy coffee/tea", ``brew coffee/tea", etc.).
 
 Hence, we build word vector n-gram neural network models to see if we can achieve better performance.
-By using dense word vectors rather than one-hot encodings to represent words, we're able capture the underlying similarity of words and their meanings.
+By using dense word vectors rather than one-hot encodings to represent words, we're able to capture the underlying similarity of words and their meanings.
 For example, ``coffee" is more similar to ``tea" that it is to ``car".
-Rather astonishingly, these word vector values can be trained through back propagation just like the weights in a neural network just as long as they're randomly initiated!
+Rather astonishingly, these word vector values can be trained through back propagation just like the weights in a neural network as long as they're randomly initiated!
 
@@
 This input layer then maps to the single hidden layer with 10 hidden units through a tanh activation function.
-Finally, the hidden layer is then connected to the output layer which is has size $|V|$ through a softmax activation function.
-We use stochastic adaptive gradient descent so that the errors can be properly backporgated to the word vectors.
+Finally, the hidden layer is then connected to the output layer which has size $|V|$ through a softmax activation function.
+We use stochastic adaptive gradient descent so that the errors can be properly backpropagated to the word vectors.
 We purposefully keep our network relatively small since our data set is rather large and training even this small neural network takes considerable time.
@@
 All is not lost since we remembered to at least store the NALL on the validation corpus.
-We are highly confident our neural network is working as intended since we see we that the NALL decreases every pass on the validation corpus in Figure \ref{fig:perp_hist} below.
+We are highly confident our neural network is working as intended since we see in Figure \ref{fig:perp_hist} that the NALL decreases on every pass through the validation corpus.
 
@@
 First, note that the validation MSE is consistently higher than the training data MSE, which is consistently higher than the test data MSE.
-Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training, validation, and test data.
+Given the (relatively) small size of our dataset (6,682 observations), this is likely due to the sampling we used to separate our data into training, validation, and test data.
 However, we don't expect this effect the validity of our cross validation.
@@
 
-Tshere is ample room to improve our sentiment analysis methodology.
+There is ample room to improve our sentiment analysis methodology.
 The training dataset generated using mechanical turk is relatively small (200 training articles).
